### 2012-12-10 icse-chase-2013/paper.tex
Set up the template for the chase 2013 paper.

% This paper is written for CHASE 2013 and features the concept and
% creation of "Impact" the awareness tool.

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
\usepackage{amsmath}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

%Paper title
\title{Impact: A Dependency Awareness Tool For Software Development}

%Author block
\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria\\
Victoria, British Columbia \\ jell@uvic.ca}
\and
\IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria\\
Victoria, British Columbia \\ danielad@cs.uvic.ca}}

\maketitle

% Paper abstract
\begin{abstract}
Write some abstract here.
\end{abstract}

\section{Introduction}

\section{Related Work}

\section{Design and Implementation of Impact}

\subsection{Extracting Developer Dependencies}

\subsection{Information Visualization}

\section{Evaluation}

\subsection{Participants}

\subsubsection{Evaluation Strategy}

\section{Results}

\section{Discussion and Future Work}

\section{Threats to Validity}

\section{Conclusion}

\section{Acknowledgements}

\bibliographystyle{IEEEtran}
\bibliography{paper}

%End of paper
\end{document}

### 2012-12-10 icse-chase-2013/paper.tex
Wrote first draft of extracting developer dependencies

@@
 
+\subsection{Studies of Awareness Practises}
+
+\subsection{Existing Awareness Tools}
+
 \section{Design and Implementation of Impact}
+The goal of this tool, \textit{Impact}, is to create an information space which
+developers can use to determine how other developer's source code changes are
+effecting their own source code. Impact will provide information to the 
+developer at a method or function level by utilizing method ownership and
+method invocation hierarchies effected by source code changes. 
+Impact is built around the source code change (changeset) level of development
+as opposed to real time. Impact is broken
+into two segments: extracting developer dependencies and information
+visualization, both of which will now be described.\\
 
 \subsection{Extracting Developer Dependencies}
+In order to provide developers with awareness information regarding changing 
+technical dependencies within a software project, the project's source code
+repository must be queried in order to obtain information regarding technical
+dependencies (For the purposes of this paper, the Git source control management
+system will be used for implementation descriptions). 
+These dependencies are built by extracting source code ownership,
+differences in changesets (diff), and method invocation hierarchies.
+
+The project's method invocation hierarchies must be extracted to form the 
+technical dependencies that are present in the project at a given 
+changeset. These method invocation hierarchies form \textit{call graphs} 
+which are graphs using method declarations as
+nodes and method invocations as directed edges between these nodes. To
+construct a method call graph as a given changeset, the project's source
+files must first be compiled into abstract syntax trees (AST). Unlike Bodden's
+et al.~\cite{Bodden:2003:HVJ} approach of using byte code, our method only
+requires source code and the creation of ASTs which does not have the
+assumption of being able to fully compile to byte code. The ASTs contain
+all information needed to create the method call graphs. The ASTs are now
+iterated over looking for method declarations to create the nodes of the
+method call graph and method invocations inside of method declarations
+to create the directed edges of the call graph which represents one method
+invoking another. This call graph now forms the technical dependencies for
+which Impact will determine the awareness information for developers.
+
+In order to transform technical dependencies into developer dependencies,
+Impact must query the source code repository in order to determine ownership
+of technical components which are the methods in the call graph. In our case,
+Git stores developer ownership of a project on a per line basis and can be
+queried using the command \textit{git blame}. The blame command can be used
+to obtain ownership of methods inside a project. Since the ASTs created above
+contain the start and end line numbers of each method declaration, the blame
+command can be used to determine which developer(s) have contributed to
+its creation. Ownership is divided among contributing developers into a 
+percentage of total lines of the method. If developer A has written 3 lines
+of method foo() and developer B has written 7 lines of this 10 line method,
+developer A owns 30\% and developer B owns 70\% of method foo(). Now that
+the developer ownership of methods have has been determined, the technical
+dependencies can now be transformed into developer dependencies.
+
+Now that developer dependencies have been obtained, dependency awareness
+information can be extracted for a given changeset by making use of
+a changeset's diff information. Git stores information regarding which 
+lines of a software project have been changed, added, or deleted by a developer
+at each changeset of the project. This information can be extracted 
+using the \textit{git diff} command. Once again, since the computed
+ASTs contain the start and end lines of each method declaration, the information
+of the diff can be compared against the ASTs to determine which methods
+have been changed by a given changeset. These newly found changed method
+form the basis of awareness information to be provided. To determine which
+developers need to become aware of these newly changed methods, the
+previously constructed method call graphs with ownership are to be queried.
+Methods which invoke those that have been changed are the result of this 
+query; specifically the owners of these invoking methods. These owners
+must be alerted to the source code changes of the given changeset because
+their code is being impacted by the changeset author's actions. These 
+dependency awareness alerts will be given weights to determine the
+total impact of the change. The weight is calculated by multiplying the
+invoking method's ownership percent by the percent modified of the changed
+method. This weight will allow developers to filter through the more
+impacting changes that are effecting their source code.
+
+The process just described can be seen in Figure~\ref{fig:network}. In
+this example, methods foo(), owned solely by Bret, and bar(), owned
+partially by Daniel and Anne, invoke method getX(). Carl has now
+committed a changeset to the source code repository in which he has
+changed method getX(). The resulting dependency awareness information of 
+getX() being changed will be provided to Bret, Daniel, and Anne with 
+appropriate information weights based on their method ownership.
+
+% Technical network figure
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.5\textwidth]{images/CallGraph}
+\caption{A method call graph with ownership.\label{fig:network}}
+\end{figure}
 

### 2012-12-15 icse-chase-2013/paper.tex
Re-initialized the paper for a better layout than before.
Rough draft of introduction is done.

--- 2012-12-10 icse-chase-2013/paper.tex
@@
 %Paper title
-\title{Impact: A Dependency Awareness Tool For Software Development}
+\title{Impact: An Indirect Dependency Awareness Tool For Software Development}
 
@@
 \begin{abstract}
-Write some abstract here.
+Awareness has been a largely studied field of awareness research within
+software engineering. Many tools and techniques been been proposed and built
+in order to provide software developers and stakeholders with a greater
+sense of workspace and task awareness within their software projects. These
+techniques and tools have been largely focused on detecting \textit{direct} 
+conflicts which arise over a project's life time or have created 
+an exploratory ground for stakeholders to use as a means of resolving self
+discovered direct or \textit{indirect} conflicts. However, detecting 
+and providing pertinent information regarding indirect conflicts has been
+a largely ignored field of research partially due to its inherently largely
+complexity than direct conflicts. Indirect conflicts arise from changes
+in one software artefact affecting another. In this paper, we present
+Impact, a new task awareness tool directly aimed at both detecting
+and presenting indirect conflicts which arise inside of a software project.
+We introduce previous indirect conflict awareness attempts, our design 
+and implementation as well as describe Impact's potential through
+two evaluation studies. 
 \end{abstract}
@@
 
+Awareness is characterized as "an understanding of the activities of others
+which provides a context for one's own activities"~\cite{Dourish:1992:ACS}.
+The study of awareness and its tools has become an important topic of
+research in software engineering especially with the new importance of
+distributed work and collaboration. Awareness is generally associated with
+both technical and social dependencies that are created and evolve over
+a software project's life time. The study of these dependencies has become
+the primary focus of most awareness related research. Task awareness has
+become the most prevalent field of research to understand how developers and
+project managers cope with these technical and social dependencies.\\
+
+Tools have been created to attempt to solve task awareness related issues
+with moderate success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, 
+Khurana:2009:PFC}. However, these tools have been designed 
+to solve task awareness related issues at the direct conflict level. 
+Examples of direct task awareness include knowing when two or more 
+developers are  editing same artefact, finding expert knowledge of a
+particular file, and knowing which developers are working in which files.
+Meanwhile, task awareness related issues at the indirect conflict level
+continue to be an issue which is largely unsolved by most coordination
+mechanisms~\cite{Khurana:2009:PFC}.\\
+
+Previous interviews and surveys conducted with software developers show a 
+consensus that developers of a software project view indirect task awareness 
+as a high priority issue in their developement~\cite{Damian:2007:GSE, 
+Halverson:2006:DTV, Begel:2010:CDE}. Examples of indirect task awareness
+conflicts include have one's own code effect by another's source
+code change and finding out who might be indirectly effect by one's
+own code change. Indirect conflicts arising in source code are inherantly
+difficult to resolve as most of the time, source code analysis must
+be preformed in order to find the relationships harmed by these conflicts.
+While some awareness tools have been created with these indirect conflicts
+primarily in mind~\cite{Begel:2010:CDE, Trainer:2005:BGT}, most have only 
+created an exploratory environment which is used by developers to
+solve conflicts which may arise. These tools lack the ability to detect
+indirect conflicts that arise and alert developers to their presence 
+inside of the software system. Some tools have started to work directly
+with solving indirect conflicts~\cite{Sarma:2007:TSA}, but continue
+to be limited by their usefulness.\\
+
+Despite software developer's need for indirect conflict awareness tools
+and existing exploratory indirect conflict awareness tools, detecting
+and alerting developers to arising indirect conflicts is still a major
+problem in the field of task awareness. Impact, a web based awareness
+tool, aims to solve this issue. In this paper, we describe Impact's
+design and implementation in order to both detect newly created
+indirect conflicts among software developers as well as alerting developers
+to these conflicts. By leveraging technical relationships inherent of 
+software projects with method call graphs as well as detecting changes
+to these technical relationship through software configuration management
+(SCM) systems, Impact is able to detect indirect conflicts as well as
+alert developers involved in such conflicts in task awareness.\\
+
+The rest of this paper is organized as follows. First, we begin by discussing
+similar task awareness related tools which have partially solved the
+issues presented by this paper and how their workings can be applied to 
+Impact. In the following section we describe the design and implementation
+of Impact as an awareness tool. We then discuss a preliminary evaluation of
+Impact followed by a discussion of future work and conclusions.\\
+
 \section{Related Work}
 
-\section{Design and Implementation of Impact}
+\section{Impact}
 
-\subsection{Extracting Developer Dependencies}
+\subsection{Design}
 
-\subsection{Information Visualization}
+\subsection{Implementation}
 
@@
 
-\subsection{Participants}
-
-\subsubsection{Evaluation Strategy}
-
 \section{Results}
 
-\section{Discussion and Future Work}
-
-\section{Threats to Validity}
-
-\section{Conclusion}
-
-\section{Acknowledgements}
+\section{Conclusion and Future Work}
 

### 2012-12-16 icse-chase-2013/paper.tex
Write the related work, design, and conclusion for chase 2013.

@@
 \section{Introduction}
-
 Awareness is characterized as "an understanding of the activities of others
@@
 \section{Related Work}
+Although there is an abundance of Awareness tools developed in research
+today, only a handful have made an attempt to examine indirect conflicts.
+Here, we will outline three of the forefront projects in indirect conflicts
+and how these projects have influenced the decision making process in
+the design and implementation of Impact.\\
+
+We first start with both Codebook~\cite{Begel:2010:CDE} and 
+Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
+environment for developers to handle indirect conflicts. Exploratory
+pertains to the ability to solve self determined conflicts. Meaning that
+once a developer discovers a conflict, they can use the tool as a type of
+lookup table to solve their issue. Codebook is a type of social developer
+network that is relating developers to source code, issue repositories and
+other social media while Ariadne only looks at source code for developer
+to source code association. In this, developers become
+owners of source code artifacts by querying the source code repository
+for ownership information. Both projects also use directed graphs
+in order to relate technical artifacts to each other. This creates technical
+dependencies within the source code which are used for indirect
+conflicts. These projects make use of method call graphs in order to 
+determine which methods invoke others. This is the basis for their 
+linkage of source code artifacts. While these projects can be great tools 
+for solving indirect conflicts which may arise, they lack the ability to
+detect conflicts on their own.\\
+
+A serious attempt at both detecting and informing developers of
+indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
+monitors developers activities in files with regards to class signatures.
+Once a developer changes the signature of a class by modifying changes
+in name, parameters, return values of public methods etc., any workspace
+of other developers which is using that file will be notified. Palantir utilizes
+a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
+a favored collection system among Awareness tools. Sarma et al. also
+develops a generic design for future indirect conflict awareness tools. 
+However, where Palantir falls short is in its collection and distribution
+mechanisms. Palantir only considers "outside" appearance of technical
+objects, being their return types, parameters, etc., and only delivers
+detected conflicts to developers who are presently viewing or editing
+the indirect object.\\
+
+Impact is designed to handle the holes in the aforementioned projects.
+Impact will focus on the detection of indirect conflicts at an internal level
+of technical objects as opposed from object signatures and the distribution
+of these conflicts to all appropriate developers regardless of their current
+workspace activities. Impact is also designed around the successes these
+projects have had in the past with directed technical artifact graphs as 
+well as elements of collection, ownership and distribution functionality.\\
 
 \section{Impact}
+This section will proceed to give an outlined detail of Impact in both its
+design and implementation. The design of Impact was created to be both
+a generic construct which can be applied to other indirect conflict 
+awareness tools while the implementation is specific to the technical
+goals of Impact.
 
 \subsection{Design}
+Compared to tool design for direct conflicts, the major concern of 
+indirect conflict tools is to relate technical objects to one another
+with a "uses" mentality. To say that object 1 uses object 2 is to infer
+a technical relationship between the two objects which can be used
+in part to detect indirect conflict that arise from modifying object
+2. This kind of relationship is modeled after a directed graph. Each
+technical object and represented by node while each "uses"
+relationship is represented by a directed edge. This representation
+is used to indicate all indirect relationships within a software project.\\
+
+While technical object relationships form the basis of indirect conflicts,
+social communication is the ultimate goal of resolving such conflicts.
+This being the case, developer ownership must be placed on the 
+identified technical objects. With this ownership, we can now infer
+relationships among developers based on their technical objects
+"uses" relationship. Developer A who owns object 1 which uses 
+object 2 owned by developer B may be notified by a change to
+object 2's internal workings. Most, if not all, ownership information
+of technical objects can be extracted from a project's source code
+repository (CVS, Git, SVN, etc.).\\
+
+Finally, the indirect conflict tool must be able to detect changes to
+to the technical objects defined above. Two approaches have been 
+proposed for change gathering techniques: real time and commit time.
+Real time indicates that all changes made to source code will be
+reported even without being committed to a project's central 
+repository where as push time indicates only committed changes
+will be subject to conflict analysis. We propose the use of push time
+information gathering. Commit time avoids the issue of developers 
+overwriting previous work or deleting modifications which would 
+produce information for changes that no longer exist. However, the
+trade off is that indirect conflicts must be committed before detected
+which results in conflicts being apart of the system before being able
+to be dealt with as opposed to catching conflicts before they happen.
+Commit time also avoids the need for a client side application to 
+monitor the development activity of project developers.\\
+
+With this three step design of: (i) creating directed graphs of technical
+objects, (ii) assigning ownership to those technical objects, and (iii)
+detecting changes within commit time, we believe a wide variety of
+indirect conflict awareness tools can be created or extended. The
+implementation of Impact in the following section will follow these
+three design guidelines.
 
@@
 \section{Conclusion and Future Work}
+In this paper, we have presented the issues that arise from indirect 
+conflicts in present awareness tools. We have proposed a generic 
+design for the future development of awareness tools in regards to
+handling indirect conflicts. We have finally presented a prototype 
+awareness tool Impact which was designed around our generic 
+awareness approach. Impact was evaluated on a small scale, showing
+its future potential as well as highlighting its current weaknesses.\\
+
+In future work we are planning to conduct interviews and surveys
+with software developers to confirm that indirect conflicts pose a
+threat to their projects as well as discovering what constitutes a 
+valid change inside of a given technical object. With these two
+improvements, we plan to justify our generic design of indirect
+conflict tools further as well as improve the ideas of detecting 
+valid and significant changes to technical objects.\\
 

### 2012-12-17 icse-chase-2013/paper.tex
All sections are now written in very rough form

@@
 complexity than direct conflicts. Indirect conflicts arise from changes
-in one software artefact affecting another. In this paper, we present
+in one software artifact affecting another. In this paper, we present
 Impact, a new task awareness tool directly aimed at both detecting
@@
 Examples of direct task awareness include knowing when two or more 
-developers are  editing same artefact, finding expert knowledge of a
+developers are  editing same artifact, finding expert knowledge of a
 particular file, and knowing which developers are working in which files.
@@
 consensus that developers of a software project view indirect task awareness 
-as a high priority issue in their developement~\cite{Damian:2007:GSE, 
+as a high priority issue in their development~\cite{Damian:2007:GSE, 
 Halverson:2006:DTV, Begel:2010:CDE}. Examples of indirect task awareness
@@
 code change and finding out who might be indirectly effect by one's
-own code change. Indirect conflicts arising in source code are inherantly
+own code change. Indirect conflicts arising in source code are inherently
 difficult to resolve as most of the time, source code analysis must
@@
 \subsection{Implementation}
+For Impact's implementation, we decided to focus on methods as our
+selected technical objects to infer a directed graph from. The "uses" 
+relationship described above form methods is method invocation.
+Thus, in our constructed directed graph, methods represent nodes
+and method invocations represent our directed edges. In order to 
+construct this directed graph, abstract syntax trees (ASTs) are 
+constructed from source files in the project. The ASTs allow us
+to construct method call graphs~\cite{} from which the directed 
+graphs can be constructed.\\
+
+Once the directed graph is constructed, we must now assign
+ownership to our technical objects (methods) as per our design.
+To do this, we simply query the source code repository. In our case
+we used Git as the source code repository, so the command "git blame"
+is used for querying ownership information (most source code 
+repositories have similar commands and functionality). This command 
+gives authors of source code per line which can be used to assign
+ownership to methods. If a method has 10 lines and developer A
+has written 3 while developer B has written 7, then ownership is
+assigned 30\% and 70\% respectively.\\
+
+To detect changes to our technical objects (methods), we simply 
+use a commit's \textit{diff} which is a representation of all changes
+made inside a commit. We can use the lines changed in the diff to 
+find methods that have been changed. This gives cause of potential
+indirect conflicts. We now find all methods in our directed graphs
+which invoke these changed methods. These are the final indirect
+conflicts.\\
+
+Now that the indirect conflicts have been found, we use the
+ownership information of our technical objects to send alerts to
+those developers involved in the indirect conflict. All owners
+of methods which invoke those that have been changed are alerted
+to the newly changed method. This can been seen in
+Figure~\ref{}, the user interface of Impact. Here, in an RSS type
+feed, the changing developer, time of change, changed method,
+invoking methods, and commit message are all displayed. This 
+list is auto updating as new commits are push to the central
+git repository. Developers are now free to use this information
+to be aware of and solve indirect conflicts which arise at the
+method level.\\
 
 \section{Evaluation}
+To fully evaluate both the generic design of detecting and resolving
+indirect conflicts as well as Impact, extensive testing and evaluation
+must be preformed. However, we feel that a simple evaluation is
+first needed to assess the foundation of our design and claims:
+does Impact improve developer awareness with indirect conflicts
+and help the resolving of such issues?.\\
 
-\section{Results}
+We preformed two user case studies to address our question where
+we gave Impact to two small development teams. Each team was
+free to use Impact at their leisure during their development process,
+after which interviews were conducted with leads developers from 
+each development team. The interviews were conducted after each
+team had used Impact for three weeks.\\
+
+We asked lead developers to address two main concerns: do indirect
+conflicts pose a threat at the method level, and did Impact help raise
+awareness and promote quicker conflict resolution for indirect
+conflicts. Our two interviews largely supported our initial claims of
+indirect conflicts posing a serious threat to developers, especially
+in medium to large teams or projects. Our interviews also pointed
+out that method use can be a particularly large area for indirect
+conflicts to arise. However, both interviewees also pointed out that
+any technical object which is used as an interface to some data
+construct or methodology, database access for instance, can be 
+a large potential issue for indirect conflicts.  Interview response to
+Impact was also largely positive, both interviewees said that Impact
+helped raise awareness among their teams with what other developers
+are doing as well as the influence it has on their own work. However,
+both interviews showed Impact to have information overload. Both
+interviews showed that while all method changes were being detected,
+not all are alert worthy. One lead developer suggested to only alert
+developers to indirect conflicts if the internal structure of a method
+changes due to modification to input parameters or output parameters.
+In other words, the boundaries of the technical objects (changing
+how a parameter is used inside the method, modifying the return
+result inside the method) seem to be more of interest than other 
+internal workings.\\
+
+These two studies have show that our design and approach to
+detecting and alerting developers to indirect conflicts appear
+to be on the correct path. Impact as a tool has laid the foundations
+for future work in detecting indirect conflicts as well as alerting
+developers although more thought must be given as to 
+what constitutes a meaningful change inside our selected 
+technical objects.\\
 

### 2012-12-18 icse-chase-2013/paper.tex
First rough now has images.

@@
 in part to detect indirect conflict that arise from modifying object
-2. This kind of relationship is modeled after a directed graph. Each
+2. This kind of relationship is modelled after a directed graph. Each
 technical object and represented by node while each "uses"
@@
 repository (CVS, Git, SVN, etc.).\\
+
+\begin{figure}[t!]
+\includegraphics{images/CallGraph}
+\caption{Technical object directed graph with ownership\label{fig:graph}}
+\end{figure}
 
@@
 to the newly changed method. This can been seen in
-Figure~\ref{}, the user interface of Impact. Here, in an RSS type
+Figure~\ref{fig:impact}, the user interface of Impact. Here, in an RSS type
 feed, the changing developer, time of change, changed method,
@@
 method level.\\
+
+\begin{figure}[t!]
+\includegraphics{images/ImpactDemo}
+\caption{Impact's RSS type information feed.\label{fig:impact}}
+\end{figure}
 

### 2012-12-19 icse-chase-2013/paper.tex
Cleaned all but the eval and conclusions.

@@
 \begin{abstract}
-Awareness has been a largely studied field of awareness research within
+Awareness has been largely studied in the field of awareness research within
 software engineering. Many tools and techniques been been proposed and built
@@
 and providing pertinent information regarding indirect conflicts has been
-a largely ignored field of research partially due to its inherently largely
-complexity than direct conflicts. Indirect conflicts arise from changes
+largely ignored partially due to its inherently larger
+complexity than direct conflicts. Indirect conflicts arise when changes
 in one software artifact affecting another. In this paper, we present
@@
 We introduce previous indirect conflict awareness attempts, our design 
-and implementation as well as describe Impact's potential through
+and implementation, as well as describe Impact's potential through
 two evaluation studies. 
@@
 the primary focus of most awareness related research. Task awareness has
-become the most prevalent field of research to understand how developers and
-project managers cope with these technical and social dependencies.\\
+become the most prevalent field of awareness research to understand 
+how developers cope with these technical and social dependencies.\\
 
@@
 to solve task awareness related issues at the direct conflict level. 
-Examples of direct task awareness include knowing when two or more 
+Examples of direct conflict awareness include knowing when two or more 
 developers are  editing same artifact, finding expert knowledge of a
@@
 
-Previous interviews and surveys conducted with software developers show a 
-consensus that developers of a software project view indirect task awareness 
-as a high priority issue in their development~\cite{Damian:2007:GSE, 
-Halverson:2006:DTV, Begel:2010:CDE}. Examples of indirect task awareness
-conflicts include have one's own code effect by another's source
-code change and finding out who might be indirectly effect by one's
+Previous interviews and surveys conducted with software developers have 
+show a pattern that developers of a software project view indirect conflict 
+awareness  as a high priority issue in their development~\cite{Damian:2007:GSE, 
+Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}. Examples 
+of indirect conflict awareness
+include having one's own code effect by another developer's source
+code change or finding out who might be indirectly effected by one's
 own code change. Indirect conflicts arising in source code are inherently
 difficult to resolve as most of the time, source code analysis must
-be preformed in order to find the relationships harmed by these conflicts.
+be preformed in order to find relationships which are harmed by changes.
 While some awareness tools have been created with these indirect conflicts
@@
 with solving indirect conflicts~\cite{Sarma:2007:TSA}, but continue
-to be limited by their usefulness.\\
+to be limited by their defined technical relationships.\\
 
 Despite software developer's need for indirect conflict awareness tools
-and existing exploratory indirect conflict awareness tools, detecting
-and alerting developers to arising indirect conflicts is still a major
-problem in the field of task awareness. Impact, a web based awareness
-tool, aims to solve this issue. In this paper, we describe Impact's
+detecting and alerting developers to arising indirect conflicts is still a major
+problem. Impact, a web based awareness
+tool, aims to solve this issue. In this paper, we describe Impact's generic
 design and implementation in order to both detect newly created
@@
 to these conflicts. By leveraging technical relationships inherent of 
-software projects with method call graphs as well as detecting changes
+software projects with method call graphs~\cite{Lakhotia:1993:CCM}
+ as well as detecting changes
 to these technical relationship through software configuration management
 (SCM) systems, Impact is able to detect indirect conflicts as well as
-alert developers involved in such conflicts in task awareness.\\
+alert developers involved in such conflicts in task awareness. This process
+fulfills the technical and social dependency congruence found by 
+Grinter  et al.~\cite{Grinter:2003:RCW} which suggests to offer a smoother
+work flow.\\
 
 The rest of this paper is organized as follows. First, we begin by discussing
-similar task awareness related tools which have partially solved the
+similar indirect conflict awareness tools which have partially solved the
 issues presented by this paper and how their workings can be applied to 
-Impact. In the following section we describe the design and implementation
+Impact. In the following section we describe a generic design and implementation
 of Impact as an awareness tool. We then discuss a preliminary evaluation of
-Impact followed by a discussion of future work and conclusions.\\
+Impact followed by conclusions and future work.\\
 
@@
 to source code association. In this, developers become
-owners of source code artifacts by querying the source code repository
-for ownership information. Both projects also use directed graphs
-in order to relate technical artifacts to each other. This creates technical
-dependencies within the source code which are used for indirect
-conflicts. These projects make use of method call graphs in order to 
-determine which methods invoke others. This is the basis for their 
-linkage of source code artifacts. While these projects can be great tools 
-for solving indirect conflicts which may arise, they lack the ability to
+owners of source code artifacts. Both projects also use program 
+dependency graphs~\cite{Horwitz:1992:UPD}
+in order to relate technical artifacts to each other. These projects make 
+use of method call graphs in order to 
+determine which methods invoke others which forms the basis for 
+linking source code artifacts; creating a directed graph. While these 
+projects can be great tools 
+for solving indirect conflicts which may arise, by querying such directed
+graphs to view impacts of conflict creating code, they lack the ability to
 detect conflicts on their own.\\
@@
 in name, parameters, return values of public methods etc., any workspace
-of other developers which is using that file will be notified. Palantir utilizes
+of other developers which are using that class will be notified. Palantir utilizes
 a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
@@
 develops a generic design for future indirect conflict awareness tools. 
-However, where Palantir falls short is in its collection and distribution
+However, where Palantir falls short are in its collection and distribution
 mechanisms. Palantir only considers "outside" appearance of technical
@@
 detected conflicts to developers who are presently viewing or editing
-the indirect object.\\
+the indirect object while other developers who have used the modified 
+class previously are not notified.\\
 
@@
 workspace activities. Impact is also designed around the successes these
-projects have had in the past with directed technical artifact graphs as 
+projects have had in the past with directed graphs as 
 well as elements of collection, ownership and distribution functionality.\\
@@
 This section will proceed to give an outlined detail of Impact in both its
-design and implementation. The design of Impact was created to be both
+design and implementation. The design of Impact was created to be
 a generic construct which can be applied to other indirect conflict 
@@
 indirect conflict tools is to relate technical objects to one another
-with a "uses" mentality. To say that object 1 uses object 2 is to infer
+with a "uses" relationship. To say that object 1 uses object 2 is to infer
 a technical relationship between the two objects which can be used
 in part to detect indirect conflict that arise from modifying object
-2. This kind of relationship is modelled after a directed graph. Each
-technical object and represented by node while each "uses"
+2. This kind of relationship is modeled after a directed graph ~\cite{Horwitz:1992:UPD}. 
+Each technical object and represented by node while each "uses"
 relationship is represented by a directed edge. This representation
@@
 relationships among developers based on their technical objects
-"uses" relationship. Developer A who owns object 1 which uses 
-object 2 owned by developer B may be notified by a change to
+"uses" relationship. Developer A, who owns object 1 which uses 
+object 2 owned by developer B, may be notified by a change to
 object 2's internal workings. Most, if not all, ownership information
@@
 Finally, the indirect conflict tool must be able to detect changes to
-to the technical objects defined above. Two approaches have been 
-proposed for change gathering techniques: real time and commit time.
-Real time indicates that all changes made to source code will be
-reported even without being committed to a project's central 
-repository where as push time indicates only committed changes
-will be subject to conflict analysis. We propose the use of push time
-information gathering. Commit time avoids the issue of developers 
+to the technical objects defined above and distribute information 
+to the appropriate owners. Two approaches have been 
+proposed for change gathering techniques: real time and commit 
+time~\cite{Fitzpatrick:2002:SPA}.
+We propose the use of commit time
+information gathering as it avoids the issue of developers 
 overwriting previous work or deleting modifications which would 
@@
 to be dealt with as opposed to catching conflicts before they happen.
-Commit time also avoids the need for a client side application to 
-monitor the development activity of project developers.\\
+At commit time, the tool must parse changed source code in relation
+to technical artifacts in the created directed graph detailed above.
+Once technical objects are found to be changed, appropriate owners
+of objects which use the changed object should be notified.
+In Figure~\ref{fig:graph}, X changes method (technical object) 1
+and thus effects methods 2 and 3 resulting in the alerting of
+developers Y and Z.
 
@@
 objects, (ii) assigning ownership to those technical objects, and (iii)
-detecting changes within commit time, we believe a wide variety of
+detecting changes within commit time and the distribution of information
+to appropriate owners, we believe a wide variety of
 indirect conflict awareness tools can be created or extended. The
@@
 selected technical objects to infer a directed graph from. The "uses" 
-relationship described above form methods is method invocation.
+relationship described above for methods is method invocation.
 Thus, in our constructed directed graph, methods represent nodes
@@
 constructed from source files in the project. The ASTs allow us
-to construct method call graphs~\cite{} from which the directed 
+to construct method call graphs from which the directed 
 graphs can be constructed.\\
@@
 To do this, we simply query the source code repository. In our case
-we used Git as the source code repository, so the command "git blame"
+we used Git as the source code repository, so the command \textit{git blame}
 is used for querying ownership information (most source code 
 repositories have similar commands and functionality). This command 
-gives authors of source code per line which can be used to assign
+gives authors of source code per line which can is used to assign
 ownership to methods. If a method has 10 lines and developer A
@@
 invoking methods, and commit message are all displayed. This 
-list is auto updating as new commits are push to the central
-git repository. Developers are now free to use this information
-to be aware of and solve indirect conflicts which arise at the
-method level.\\
+list is auto updating as new commits are pushed to the central
+git repository. The information provided also comes with a simple
+weight of percent changed of changed method multiplied by 
+ownership of the invoking method. This allows developers to filter
+through high and low changes effecting their own source code.
+Developers are now fable to be alerted of detected indirect
+conflicts and solve them by communicating the the developer
+who authored the change.\\
 

### 2012-12-19 icse-chase-2013/paper.tex
First draft is done. Need to clean bib file and get pdf ready for edits.

@@
 indirect conflicts posing a serious threat to developers, especially
-in medium to large teams or projects. Our interviews also pointed
+in medium to large teams or projects as opposed to the small
+teams which they were apart of. Our interviews also pointed
 out that method use can be a particularly large area for indirect
@@
 for future work in detecting indirect conflicts as well as alerting
-developers although more thought must be given as to 
+developers, although more thought must be given as to 
 what constitutes a meaningful change inside our selected 

### 2013-01-09 icse-chase-2013/paper.tex
Final first rough draft.
First edits to chase impact paper.

@@
 complexity than direct conflicts. Indirect conflicts arise when changes
-in one software artifact affecting another. In this paper, we present
+in one software artifact affect another. In this paper, we present
 Impact, a new task awareness tool directly aimed at both detecting
@@
 of indirect conflict awareness
-include having one's own code effect by another developer's source
+include having one's own code effected by another developer's source
 code change or finding out who might be indirectly effected by one's
@@
 inside of the software system. Some tools have started to work directly
-with solving indirect conflicts~\cite{Sarma:2007:TSA}, but continue
+with solving indirect conflicts~\cite{Sarma:2007:TSA} but continue
 to be limited by their defined technical relationships.\\
@@
 environment for developers to handle indirect conflicts. Exploratory
-pertains to the ability to solve self determined conflicts. Meaning that
+pertains to the ability to solve self determined conflicts, meaning that
 once a developer discovers a conflict, they can use the tool as a type of
 lookup table to solve their issue. Codebook is a type of social developer
-network that is relating developers to source code, issue repositories and
+network that relates developers to source code, issue repositories and
 other social media while Ariadne only looks at source code for developer
@@
 determine which methods invoke others which forms the basis for 
-linking source code artifacts; creating a directed graph. While these 
+linking source code artifacts creating a directed graph. While these 
 projects can be great tools 
@@
 a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
-a favored collection system among Awareness tools. Sarma et al. also
+a favored collection system among Awareness tools. Sarma et al.
+~\cite{Sarma:2007:TSA} also
 develops a generic design for future indirect conflict awareness tools. 
-However, where Palantir falls short are in its collection and distribution
-mechanisms. Palantir only considers "outside" appearance of technical
-objects, being their return types, parameters, etc., and only delivers
+However, Palantir falls short in its collection and distribution
+mechanisms. First, Palantir only considers "outside" appearance of technical
+objects, being their return types, parameters, etc.. Secondly, Palantir 
+only delivers
 detected conflicts to developers who are presently viewing or editing
@@
 projects have had in the past with directed graphs as 
-well as elements of collection, ownership and distribution functionality.\\
+well as elements of collection, ownership, and distribution functionality.\\
 
@@
 indirect conflict tools is to relate technical objects to one another
-with a "uses" relationship. To say that object 1 uses object 2 is to infer
+with a ``uses'' relationship. To say that object 1 uses object 2 is to infer
 a technical relationship between the two objects which can be used
@@
 2. This kind of relationship is modeled after a directed graph ~\cite{Horwitz:1992:UPD}. 
-Each technical object and represented by node while each "uses"
+Each technical object is represented by node while each ``uses''
 relationship is represented by a directed edge. This representation
 is used to indicate all indirect relationships within a software project.\\
-
-While technical object relationships form the basis of indirect conflicts,
-social communication is the ultimate goal of resolving such conflicts.
-This being the case, developer ownership must be placed on the 
-identified technical objects. With this ownership, we can now infer
-relationships among developers based on their technical objects
-"uses" relationship. Developer A, who owns object 1 which uses 
-object 2 owned by developer B, may be notified by a change to
-object 2's internal workings. Most, if not all, ownership information
-of technical objects can be extracted from a project's source code
-repository (CVS, Git, SVN, etc.).\\
 
@@
 \end{figure}
+
+While technical object relationships form the basis of indirect conflicts,
+social communication is the ultimate goal of resolving such conflicts.
+This being the case, developer ownership must be placed on the 
+identified technical objects. With this ownership, we now infer
+relationships among developers based on their technical objects
+``uses'' relationship. Developer A, who owns object 1, which uses 
+object 2 owned by developer B, may be notified by a change to
+object 2's internal workings. Most, if not all, ownership information
+of technical objects can be extracted from a project's source code
+repository (CVS, Git, SVN, etc.).\\
 
@@
 produce information for changes that no longer exist. However, the
-trade off is that indirect conflicts must be committed before detected
+trade off is that indirect conflicts must be committed before detected,
 which results in conflicts being apart of the system before being able
@@
 of objects which use the changed object should be notified.
-In Figure~\ref{fig:graph}, X changes method (technical object) 1
-and thus effects methods 2 and 3 resulting in the alerting of
+In Figure~\ref{fig:graph}, X changes method (technical object) 1,
+which effects methods 2 and 3 resulting in the alerting of
 developers Y and Z.
@@
 For Impact's implementation, we decided to focus on methods as our
-selected technical objects to infer a directed graph from. The "uses" 
+selected technical objects to infer a directed graph from. The ``uses'' 
 relationship described above for methods is method invocation.
@@
 we used Git as the source code repository, so the command \textit{git blame}
-is used for querying ownership information (most source code 
-repositories have similar commands and functionality). This command 
+is used for querying ownership information. (Most source code 
+repositories have similar commands and functionality.) This command 
 gives authors of source code per line which can is used to assign
@@
 use a commit's \textit{diff} which is a representation of all changes
-made inside a commit. We can use the lines changed in the diff to 
+made inside a commit. We can use the lines changed in the \textit{diff} to 
 find methods that have been changed. This gives cause of potential
@@
 git repository. The information provided also comes with a simple
-weight of percent changed of changed method multiplied by 
+weight of the percent changed of changed method multiplied by 
 ownership of the invoking method. This allows developers to filter
-through high and low changes effecting their own source code.
-Developers are now fable to be alerted of detected indirect
+through high and low changes affecting their own source code.
+Developers are now able to be alerted of detected indirect
 conflicts and solve them by communicating the the developer
@@
 indirect conflicts as well as Impact, extensive testing and evaluation
-must be preformed. However, we feel that a simple evaluation is
+must be performed. However, we feel that a simple evaluation is
 first needed to assess the foundation of our design and claims:
 does Impact improve developer awareness with indirect conflicts
-and help the resolving of such issues?.\\
+and help the resolving of such issues?\\
 
-We preformed two user case studies to address our question where
+We performed two user case studies to address our question where
 we gave Impact to two small development teams. Each team was
 free to use Impact at their leisure during their development process,
-after which interviews were conducted with leads developers from 
+after which interviews were conducted with lead developers from 
 each development team. The interviews were conducted after each
@@
 a large potential issue for indirect conflicts.  Interview response to
-Impact was also largely positive, both interviewees said that Impact
+Impact was also largely positive, as both interviewees said that Impact
 helped raise awareness among their teams with what other developers
@@
 design for the future development of awareness tools in regards to
-handling indirect conflicts. We have finally presented a prototype 
-awareness tool Impact which was designed around our generic 
+handling indirect conflicts. We have presented a prototype 
+awareness tool, Impact, which was designed around our generic 
 awareness approach. Impact was evaluated on a small scale, showing

### 2013-01-19 icse-chase-2013/paper.tex
Working on images.
First edit (CHASE)
Updated for first edit and fixed images to be cleaner. (CHASE)
Danas syntax changes for rev 3

@@
 in one software artifact affect another. In this paper, we present
-Impact, a new task awareness tool directly aimed at both detecting
+\textit{Impact}, a new task awareness tool directly aimed at both detecting
 and presenting indirect conflicts which arise inside of a software project.
 We introduce previous indirect conflict awareness attempts, our design 
-and implementation, as well as describe Impact's potential through
+and implementation, as well as describe \textit{Impact's} potential through
 two evaluation studies. 
@@
 become the most prevalent field of awareness research to understand 
-how developers cope with these technical and social dependencies.\\
+how developers cope with these technical and social dependencies.
 
 Tools have been created to attempt to solve task awareness related issues
-with moderate success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, 
-Khurana:2009:PFC}. However, these tools have been designed 
+with moderate success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
 to solve task awareness related issues at the direct conflict level. 
@@
 continue to be an issue which is largely unsolved by most coordination
-mechanisms~\cite{Khurana:2009:PFC}.\\
+mechanisms. Examples of indirect conflict awareness
+include having one's own code effected by another developer's source
+code change or finding out who might be indirectly effected by one's
+own code change.
 
@@
 awareness  as a high priority issue in their development~\cite{Damian:2007:GSE, 
-Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}. Examples 
-of indirect conflict awareness
-include having one's own code effected by another developer's source
-code change or finding out who might be indirectly effected by one's
-own code change. Indirect conflicts arising in source code are inherently
+Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}.
+Indirect conflicts arising in source code are inherently
 difficult to resolve as most of the time, source code analysis must
-be preformed in order to find relationships which are harmed by changes.
+be performed in order to find relationships between technical objects
+which are harmed by changes.
 While some awareness tools have been created with these indirect conflicts
@@
 indirect conflicts that arise and alert developers to their presence 
-inside of the software system. Some tools have started to work directly
+inside the software system. Some tools have started to work directly
 with solving indirect conflicts~\cite{Sarma:2007:TSA} but continue
-to be limited by their defined technical relationships.\\
+to be limited by their defined technical relationships.
 
@@
 detecting and alerting developers to arising indirect conflicts is still a major
-problem. Impact, a web based awareness
-tool, aims to solve this issue. In this paper, we describe Impact's generic
+problem. \textit{Impact}, a web based awareness
+tool, aims to solve this issue. In this paper, we describe \textit{Impact's}
 design and implementation in order to both detect newly created
@@
 to these technical relationship through software configuration management
-(SCM) systems, Impact is able to detect indirect conflicts as well as
-alert developers involved in such conflicts in task awareness. This process
+(SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
+alert developers involved in such conflicts in task awareness. While this paper
+outlines \textit{Impact's} specific implementation, the design used can be put
+towards many other similar indirect conflict awareness tools. This design
 fulfills the technical and social dependency congruence found by 
-Grinter  et al.~\cite{Grinter:2003:RCW} which suggests to offer a smoother
-work flow.\\
+Grinter  et al.~\cite{Grinter:2003:RCW} which suggests to offer a better 
+communication patterns inside the project.
 
@@
 issues presented by this paper and how their workings can be applied to 
-Impact. In the following section we describe a generic design and implementation
-of Impact as an awareness tool. We then discuss a preliminary evaluation of
-Impact followed by conclusions and future work.\\
+\textit{Impact}. In the following section we describe a generic design and implementation
+of \textit{Impact} as an awareness tool. We then discuss a preliminary evaluation of
+\textit{Impact} followed by conclusions and future work.
 
@@
 and how these projects have influenced the decision making process in
-the design and implementation of Impact.\\
+the design and implementation of \textit{Impact}.
 
@@
 other social media while Ariadne only looks at source code for developer
-to source code association. In this, developers become
+to source code association. through Codebook, developers become
 owners of source code artifacts. Both projects also use program 
@@
 graphs to view impacts of conflict creating code, they lack the ability to
-detect conflicts on their own.\\
+detect conflicts on their own.
 
@@
 monitors developers activities in files with regards to class signatures.
-Once a developer changes the signature of a class by modifying changes
-in name, parameters, return values of public methods etc., any workspace
+Once a developer changes the signature of a class, such as by modifying changes
+in name, parameters, or return values of public methods, any workspace
 of other developers which are using that class will be notified. Palantir utilizes
@@
 the indirect object while other developers who have used the modified 
-class previously are not notified.\\
+class previously are not notified.
 
-Impact is designed to handle the holes in the aforementioned projects.
-Impact will focus on the detection of indirect conflicts at an internal level
-of technical objects as opposed from object signatures and the distribution
-of these conflicts to all appropriate developers regardless of their current
-workspace activities. Impact is also designed around the successes these
+\textit{Impact} is designed to address the limitations in the aforementioned projects.
+\textit{Impact} will focus on (1) the detection of indirect conflicts at an internal level
+of technical objects as opposed from object signatures and (2) the dissemination of
+awareness information to all appropriate developers regardless of their current
+workspace activities. \textit{Impact} is also designed around the successes these
 projects have had in the past with directed graphs as 
-well as elements of collection, ownership, and distribution functionality.\\
+well as elements of collection, ownership, and distribution functionality.
 
 \section{Impact}
-This section will proceed to give an outlined detail of Impact in both its
-design and implementation. The design of Impact was created to be
+This section will proceed to give an outlined detail of \textit{Impact} in both its
+design and implementation. The design of \textit{Impact} was created to be
 a generic construct which can be applied to other indirect conflict 
 awareness tools while the implementation is specific to the technical
-goals of Impact.
+goals of \textit{Impact}.
 
@@
 relationship is represented by a directed edge. This representation
-is used to indicate all indirect relationships within a software project.\\
+is used to indicate all indirect relationships within a software project.
 
 \begin{figure}[t!]
-\includegraphics{images/CallGraph}
+\includegraphics[width=\columnwidth]{images/CallGraph}
 \caption{Technical object directed graph with ownership\label{fig:graph}}
@@
 While technical object relationships form the basis of indirect conflicts,
-social communication is the ultimate goal of resolving such conflicts.
+communication between developers is the ultimate goal of resolving such conflicts.
 This being the case, developer ownership must be placed on the 
@@
 of technical objects can be extracted from a project's source code
-repository (CVS, Git, SVN, etc.).\\
+repository (CVS, Git, SVN, etc.).
 
-Finally, the indirect conflict tool must be able to detect changes to
-to the technical objects defined above and distribute information 
-to the appropriate owners. Two approaches have been 
+Finally, the indirect conflict tool must be able to detect changes
+to the technical objects defined above and notify the appropriate owners
+to the conflict. 
+Two approaches have been 
 proposed for change gathering techniques: real time and commit 
@@
 to technical artifacts in the created directed graph detailed above.
+Where \textit{Impact's} design differs from that of Palantir's is that
+the object's entire body (method definition and internal body) is 
+parsed at commit time to detect changes anywhere in the technical object.
 Once technical objects are found to be changed, appropriate owners
@@
 objects, (ii) assigning ownership to those technical objects, and (iii)
-detecting changes within commit time and the distribution of information
+detecting changes within commit time and the notification of conflict information
 to appropriate owners, we believe a wide variety of
 indirect conflict awareness tools can be created or extended. The
-implementation of Impact in the following section will follow these
+implementation of \textit{Impact} in the following section follows these
 three design guidelines.
@@
 \subsection{Implementation}
-For Impact's implementation, we decided to focus on methods as our
+For \textit{Impact's} implementation, we decided to focus on methods as our
 selected technical objects to infer a directed graph from. The ``uses'' 
@@
 to construct method call graphs from which the directed 
-graphs can be constructed.\\
+graphs can be constructed.
 
@@
 has written 3 while developer B has written 7, then ownership is
-assigned 30\% and 70\% respectively.\\
+assigned 30\% and 70\% respectively.
 
@@
 which invoke these changed methods. These are the final indirect
-conflicts.\\
+conflicts.
 
 Now that the indirect conflicts have been found, we use the
-ownership information of our technical objects to send alerts to
+ownership information of our technical objects to send notifications to
 those developers involved in the indirect conflict. All owners
@@
 to the newly changed method. This can been seen in
-Figure~\ref{fig:impact}, the user interface of Impact. Here, in an RSS type
+Figure~\ref{fig:impact}, the user interface of \textit{Impact}. Here, in an RSS type
 feed, the changing developer, time of change, changed method,
@@
 Developers are now able to be alerted of detected indirect
-conflicts and solve them by communicating the the developer
-who authored the change.\\
+conflicts and solve them by communicating the developer
+who authored the change.
 
 \begin{figure}[t!]
-\includegraphics{images/ImpactDemo}
-\caption{Impact's RSS type information feed.\label{fig:impact}}
+\includegraphics[width=\columnwidth]{images/ImpactDemo}
+\caption{\textit{Impact's} RSS type information feed.\label{fig:impact}}
 \end{figure}
@@
 To fully evaluate both the generic design of detecting and resolving
-indirect conflicts as well as Impact, extensive testing and evaluation
+indirect conflicts as well as \textit{Impact}, extensive testing and evaluation
 must be performed. However, we feel that a simple evaluation is
 first needed to assess the foundation of our design and claims:
-does Impact improve developer awareness with indirect conflicts
-and help the resolving of such issues?\\
+does \textit{Impact} improve developer awareness with indirect conflicts
+and help the resolving of such issues?
 
 We performed two user case studies to address our question where
-we gave Impact to two small development teams. Each team was
-free to use Impact at their leisure during their development process,
+we gave \textit{Impact} to two small development teams. Each team was
+free to use \textit{Impact} at their leisure during their development process,
 after which interviews were conducted with lead developers from 
 each development team. The interviews were conducted after each
-team had used Impact for three weeks.\\
+team had used \textit{Impact} for three weeks.
 
 We asked lead developers to address two main concerns: do indirect
-conflicts pose a threat at the method level, and did Impact help raise
+conflicts pose a threat at the method level, and did \textit{Impact} help raise
 awareness and promote quicker conflict resolution for indirect
@@
 a large potential issue for indirect conflicts.  Interview response to
-Impact was also largely positive, as both interviewees said that Impact
+\textit{Impact} was also largely positive, as both interviewees said that \textit{Impact}
 helped raise awareness among their teams with what other developers
 are doing as well as the influence it has on their own work. However,
-both interviews showed Impact to have information overload. Both
+both interviews showed \textit{Impact} to have information overload. Both
 interviews showed that while all method changes were being detected,
@@
 result inside the method) seem to be more of interest than other 
-internal workings.\\
+internal workings.
 
-These two studies have show that our design and approach to
+These two studies have shown that our design and approach to
 detecting and alerting developers to indirect conflicts appear
-to be on the correct path. Impact as a tool has laid the foundations
+to be on the correct path. \textit{Impact} as a tool has laid the foundations
 for future work in detecting indirect conflicts as well as alerting
@@
 what constitutes a meaningful change inside our selected 
-technical objects.\\
+technical objects.
 
@@
 handling indirect conflicts. We have presented a prototype 
-awareness tool, Impact, which was designed around our generic 
-awareness approach. Impact was evaluated on a small scale, showing
-its future potential as well as highlighting its current weaknesses.\\
+awareness tool, \textit{Impact}, which was designed around our generic 
+awareness approach. \textit{Impact} was evaluated on a small scale, showing
+its future potential as well as highlighting its current weaknesses.
 
@@
 conflict tools further as well as improve the ideas of detecting 
-valid and significant changes to technical objects.\\
+valid and significant changes to technical objects.
 

### 2013-01-19 icse-chase-2013/paper.tex
Danas revision 3 is now final.

@@
 and presenting indirect conflicts which arise inside of a software project.
-We introduce previous indirect conflict awareness attempts, our design 
-and implementation, as well as describe \textit{Impact's} potential through
+\textit{Impact} represents a first step towards the design and implementation
+of awareness tools that specifically address indirect conflicts in software
+projects. To describe \textit{Impact}, we introduce previous indirect 
+conflict awareness attempts, our design 
+and implementation, as well as \textit{Impact's} potential through
 two evaluation studies. 
@@
 outlines \textit{Impact's} specific implementation, the design used can be put
-towards many other similar indirect conflict awareness tools. This design
-fulfills the technical and social dependency congruence found by 
-Grinter  et al.~\cite{Grinter:2003:RCW} which suggests to offer a better 
-communication patterns inside the project.
+towards many other similar indirect conflict awareness tools. 
+\textit{Impact} represents a first step towards the design and implementation
+of awareness tools which address indirect conflicts in software development.
 
@@
 in medium to large teams or projects as opposed to the small
-teams which they were apart of. Our interviews also pointed
+teams which they were apart of. It was also pointed
 out that method use can be a particularly large area for indirect
-conflicts to arise. However, both interviewees also pointed out that
+conflicts to arise. However, it was noted that
 any technical object which is used as an interface to some data
@@
 a large potential issue for indirect conflicts.  Interview response to
-\textit{Impact} was also largely positive, as both interviewees said that \textit{Impact}
+\textit{Impact} was also largely positive, as interviewees stated that \textit{Impact}
 helped raise awareness among their teams with what other developers
 are doing as well as the influence it has on their own work. However,
-both interviews showed \textit{Impact} to have information overload. Both
-interviews showed that while all method changes were being detected,
-not all are alert worthy. One lead developer suggested to only alert
+\textit{Impact} was shown to have information overload. It was suggested
+that while all method changes were being detected,
+not all are notification worthy. One developer suggested to only notify
 developers to indirect conflicts if the internal structure of a method
@@
 result inside the method) seem to be more of interest than other 
-internal workings.
+internal workings. More complex inner workings of methods were also noted 
+to be of interest to developers
+such as cyclomatic complexity, or time and space requirements.
 
@@
 to be on the correct path. \textit{Impact} as a tool has laid the foundations
-for future work in detecting indirect conflicts as well as alerting
+for future work in detecting indirect conflicts as well as notifying
 developers, although more thought must be given as to 
@@
 awareness approach. \textit{Impact} was evaluated on a small scale, showing
-its future potential as well as highlighting its current weaknesses.
+its future potential as well as highlighting its current weaknesses. This paper
+has demonstrated our first step towards understanding awareness tools
+that specifically address indirect conflicts and what questions are 
+still to be answered.
 
-In future work we are planning to conduct interviews and surveys
-with software developers to confirm that indirect conflicts pose a
-threat to their projects as well as discovering what constitutes a 
-valid change inside of a given technical object. With these two
-improvements, we plan to justify our generic design of indirect
-conflict tools further as well as improve the ideas of detecting 
-valid and significant changes to technical objects.
+\section{Future Work For Indirect Conflicts}
+While \textit{Impact} has laid the foundations for awareness tool construction
+aimed at resolving indirect conflicts, it has also raised questions 
+in regards to levels of granularity in notification of indirect conflicts and the
+implied trade off between notifications and information overload. To address
+these problems, we hope to find industry expert opinion through interviews
+and surveys with experienced developers. From these interactions, we hope to
+answer some of the questions presented in this paper such as which technical
+objects concern developers with indirect conflicts, and what types of changes
+to these objects are cause for notification. We also hope, through these
+expert opinions, to further the 
+push for indirect conflict awareness tools by establishing that indirect
+conflicts are real world problems with harmful consequences. 
+Further more, we hope to bring these opinions back to the realm of 
+\textit{Impact}, in combination with other academic works on balancing
+between notifications and information overload~\cite{Wang:2007}, to
+continue our development
+on this tool to further its life both as a research tool and a possible
+industry application.
 

### 2013-02-02 icse-chase-2013/paper.tex
Final Edit
This is the final edit for CHASE 2013 submission. Still needs to be
compiled and checked for length.

@@
 %Paper title
-\title{Impact: An Indirect Dependency Awareness Tool For Software Development}
+\title{Supporting Awareness of Indirect Conflicts with Impact}
 
@@
 \begin{abstract}
-Awareness has been largely studied in the field of awareness research within
-software engineering. Many tools and techniques been been proposed and built
+Awareness has been largely studied in the field of CSCW research within
+software engineering. Many tools and techniques have been proposed and built
 in order to provide software developers and stakeholders with a greater
@@
 an exploratory ground for stakeholders to use as a means of resolving self
-discovered direct or \textit{indirect} conflicts. However, detecting 
+discovered direct or indirect conflicts. However, detecting 
 and providing pertinent information regarding indirect conflicts has been
-largely ignored partially due to its inherently larger
+largely ignored partially due to their inherently larger
 complexity than direct conflicts. Indirect conflicts arise when changes
@@
 Tools have been created to attempt to solve task awareness related issues
-with moderate success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
+with some success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
 to solve task awareness related issues at the direct conflict level. 
@@
 developers are  editing same artifact, finding expert knowledge of a
-particular file, and knowing which developers are working in which files.
+particular file, and knowing which developers are working on which files.
 Meanwhile, task awareness related issues at the indirect conflict level
@@
 code change or finding out who might be indirectly effected by one's
-own code change.
-
-Previous interviews and surveys conducted with software developers have 
-show a pattern that developers of a software project view indirect conflict 
+own code change. Previous interviews and surveys conducted with software developers have 
+shown a pattern that developers of a software project view indirect conflict 
 awareness  as a high priority issue in their development~\cite{Damian:2007:GSE, 
 Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}.
+
 Indirect conflicts arising in source code are inherently
@@
 created an exploratory environment which is used by developers to
-solve conflicts which may arise. These tools lack the ability to detect
+solve conflicts which may arise. These tools were not designed to detect
 indirect conflicts that arise and alert developers to their presence 
-inside the software system. Some tools have started to work directly
-with solving indirect conflicts~\cite{Sarma:2007:TSA} but continue
-to be limited by their defined technical relationships.
+inside the software system. Sarma et al.~\cite{Sarma:2007:TSA} has started to work directly
+on solving indirect conflicts, however, these products are not designed to handle
+internal structures of technical objects.
 
-Despite software developer's need for indirect conflict awareness tools
-detecting and alerting developers to arising indirect conflicts is still a major
-problem. \textit{Impact}, a web based awareness
-tool, aims to solve this issue. In this paper, we describe \textit{Impact's}
-design and implementation in order to both detect newly created
-indirect conflicts among software developers as well as alerting developers
-to these conflicts. By leveraging technical relationships inherent of 
+In this paper, we report on our research into supporting indirect conflicts
+and present the design, implementation, and evaluation of the tool \textit{Impact},
+a web based tool that aims at detecting indirect conflicts among developers
+and notifying the appropriate members involved in these conflicts.
+By leveraging technical relationships inherent of 
 software projects with method call graphs~\cite{Lakhotia:1993:CCM}
@@
 alert developers involved in such conflicts in task awareness. While this paper
-outlines \textit{Impact's} specific implementation, the design used can be put
-towards many other similar indirect conflict awareness tools. 
+outlines \textit{Impact's} specific implementation, its design is rather
+generic and can be implemented in similar indirect conflict awareness tools.
 \textit{Impact} represents a first step towards the design and implementation
@@
 similar indirect conflict awareness tools which have partially solved the
-issues presented by this paper and how their workings can be applied to 
+issues presented by this paper and how their designs can be applied to 
 \textit{Impact}. In the following section we describe a generic design and implementation
@@
 \section{Related Work}
-Although there is an abundance of Awareness tools developed in research
+Although there is an abundance of awareness tools developed in research
 today, only a handful have made an attempt to examine indirect conflicts.
@@
 other social media while Ariadne only looks at source code for developer
-to source code association. through Codebook, developers become
+to source code association. Through Codebook, developers become
 owners of source code artifacts. Both projects also use program 
@@
 a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
-a favored collection system among Awareness tools. Sarma et al.
+a favored collection system among awareness tools. Sarma et al.
 ~\cite{Sarma:2007:TSA} also
-develops a generic design for future indirect conflict awareness tools. 
+developed a generic design for future indirect conflict awareness tools. 
 However, Palantir falls short in its collection and distribution
 mechanisms. First, Palantir only considers "outside" appearance of technical
-objects, being their return types, parameters, etc.. Secondly, Palantir 
+objects, being their return types, parameters, etc. Secondly, Palantir 
 only delivers
@@
 
-\textit{Impact} is designed to address the limitations in the aforementioned projects.
-\textit{Impact} will focus on (1) the detection of indirect conflicts at an internal level
-of technical objects as opposed from object signatures and (2) the dissemination of
+\textit{Impact} is designed to address these limitations by extending this work in
+two directions: (1) the detection of indirect conflicts at an internal level
+of technical objects as opposed from object signatures, and (2) the dissemination of
 awareness information to all appropriate developers regardless of their current
@@
 in part to detect indirect conflict that arise from modifying object
-2. This kind of relationship is modeled after a directed graph ~\cite{Horwitz:1992:UPD}. 
+2. This kind of relationship is modeled based on directed graphs ~\cite{Horwitz:1992:UPD}. 
 Each technical object is represented by node while each ``uses''
@@
 of objects which use the changed object should be notified.
-In Figure~\ref{fig:graph}, X changes method (technical object) 1,
+In Figure~\ref{fig:graph}, Carl changes method (technical object) 1,
 which effects methods 2 and 3 resulting in the alerting of
-developers Y and Z.
+developers Bret, Daniel, and Anne.
 
-With this three step design of: (i) creating directed graphs of technical
+With this three step design: (i) creating directed graphs of technical
 objects, (ii) assigning ownership to those technical objects, and (iii)
@@
 indirect conflict awareness tools can be created or extended. The
-implementation of \textit{Impact} in the following section follows these
+implementation of \textit{Impact} described in the following section follows these
 three design guidelines.
@@
 repositories have similar commands and functionality.) This command 
-gives authors of source code per line which can is used to assign
+returns the source code authors per linee which can is used to assign
 ownership to methods. If a method has 10 lines and developer A
@@
 
-Now that the indirect conflicts have been found, we use the
+Once the indirect conflicts have been found, we use the
 ownership information of our technical objects to send notifications to
@@
 must be performed. However, we feel that a simple evaluation is
-first needed to assess the foundation of our design and claims:
-does \textit{Impact} improve developer awareness with indirect conflicts
-and help the resolving of such issues?
+first needed to assess the foundation of \textit{Impact}'s design and claims
+about indirect conflicts at the method level.
 
-We performed two user case studies to address our question where
-we gave \textit{Impact} to two small development teams. Each team was
+We performed a user case study where we gave \textit{Impact} to 
+two small development teams composed of three developers. Each team was
 free to use \textit{Impact} at their leisure during their development process,
@@
 We asked lead developers to address two main concerns: do indirect
-conflicts pose a threat at the method level, and did \textit{Impact} help raise
+conflicts pose a threat
+at the method level (e.g. method 1 has a bug because it invokes method
+2 which has has its implementation changed recently), 
+and did \textit{Impact} help raise
 awareness and promote quicker conflict resolution for indirect
-conflicts. Our two interviews largely supported our initial claims of
+conflicts. Our two interviews largely supported our expectation of
 indirect conflicts posing a serious threat to developers, especially
@@
 what constitutes a meaningful change inside our selected 
-technical objects.
+technical objects to avoid issues of information overload.
+
+\section{Next Steps in Research of Indirect Conflicts}
+While \textit{Impact} has laid the foundations for awareness tool construction
+aimed at resolving indirect conflicts, it has also raised questions 
+in regards to levels of granularity in notification of indirect conflicts and the
+implied trade off between notifications and information overload. To address
+these problems, we plan to find industry expert opinion through interviews
+and surveys with experienced developers. From these interactions, we plan to
+answer some of the questions presented in this paper such as which technical
+objects concern developers with indirect conflicts, and what types of changes
+to these objects are cause for notification. We are also planning, through these
+expert opinions, to further the 
+need for indirect conflict awareness tools by establishing that indirect
+conflicts are real world problems with harmful consequences. 
+Further more, we plan to bring these opinions back to the realm of 
+\textit{Impact}, in combination with other academic works on balancing
+between notifications and information overload~\cite{Wang:2007}, to
+continue our development
+on this tool to further its life both as a research tool and a possible
+industry application.
 
@@
 awareness tool, \textit{Impact}, which was designed around our generic 
-awareness approach. \textit{Impact} was evaluated on a small scale, showing
-its future potential as well as highlighting its current weaknesses. This paper
-has demonstrated our first step towards understanding awareness tools
-that specifically address indirect conflicts and what questions are 
-still to be answered.
-
-\section{Future Work For Indirect Conflicts}
-While \textit{Impact} has laid the foundations for awareness tool construction
-aimed at resolving indirect conflicts, it has also raised questions 
-in regards to levels of granularity in notification of indirect conflicts and the
-implied trade off between notifications and information overload. To address
-these problems, we hope to find industry expert opinion through interviews
-and surveys with experienced developers. From these interactions, we hope to
-answer some of the questions presented in this paper such as which technical
-objects concern developers with indirect conflicts, and what types of changes
-to these objects are cause for notification. We also hope, through these
-expert opinions, to further the 
-push for indirect conflict awareness tools by establishing that indirect
-conflicts are real world problems with harmful consequences. 
-Further more, we hope to bring these opinions back to the realm of 
-\textit{Impact}, in combination with other academic works on balancing
-between notifications and information overload~\cite{Wang:2007}, to
-continue our development
-on this tool to further its life both as a research tool and a possible
-industry application.
+technical object approach. \textit{Impact} was evaluated on a small scale, showing
+its future potential as well as highlighting its current weaknesses. 
+Our preliminary evaluation has given us concrete insights into 
+understanding awareness tools that specifically address indirect and has
+strengthened the need for further research towards notification and information
+overload (what quantifies a change that is notification
+worthy), which has become the goal of our next step.
 

### 2013-02-05 icse-chase-2013/paper.tex
Final Edit
Chase paper is ready for submit.

@@
 discovered direct or indirect conflicts. However, detecting 
-and providing pertinent information regarding indirect conflicts has been
+and providing pertinent information regarding \textit{indirect} conflicts has been
 largely ignored partially due to their inherently larger
@@
 Our preliminary evaluation has given us concrete insights into 
-understanding awareness tools that specifically address indirect and has
+understanding awareness tools that specifically address indirect conflicts and has
 strengthened the need for further research towards notification and information
-overload (what quantifies a change that is notification
+overload (what makes a change notification
 worthy), which has become the goal of our next step.

### 2013-03-02 icse-src-2013/paper.tex
SRC camera ready is finished. Need to submit pdf to conference now.

% This will be the main document for the Technical Networks paper to
% be written by the Eggnet team of Jordan Ell, Triet Huynh and Braden
% Simpson in association with Adrian Schroeter and Daniela Damian.

\documentclass[conference]{IEEEtran}

% Use of outside images
\usepackage{graphicx} 
% Use text inside euqations
\usepackage{amsmath}

% Correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% Begin the paper here
\begin{document}

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{Identifying Failure Inducing Developer Pairs Within Developer Networks}

% Authors names
\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria,
Victoria, British Columbia \\ jell@uvic.ca}
}

% Make the title area
\maketitle

\begin{abstract}
Software systems have not only become larger over time, but the amount of
technical contributors and dependencies have also increased. With these expansions also comes
the increasing risk of introducing a software failure into a pre-existing system.
Software failures are a multi-billion dollar problem in the industry today and while integration and
other forms of testing are helping to ensure a minimal number of failures, research to understand
full impacts of code changes and their social implications is still a major concern. This paper describes
how analysis of code changes and the technical relationships they infer can be used to detect pairs
of developers whose technical dependencies may induce software failures. These developer pairs may
also be used to predict future software failures as well as provide recommendations to contributors
to solve these failures caused by source code changes.
\end{abstract}

\section{Introduction}

% Setup the problem.
Large software projects are created using highly modular and 
reusable code. This creates technical dependencies between methods or functions
that can be used in a wide variety of locations throughout the project. 
This causes changes to any given method to have a rippling 
effect across the rest of the project~\cite{Acharya:2011:PCI}. The larger these effects are,
the more likely they are to cause a software failure inside the system during the project's
life span~\cite{Zimmermann:2008:PDU}. These observations of technical dependencies open
the door to types of analysis on the developer networks they infer and preventing software
failures by improving coordination amongst dependant developers.

% Explain technical network research that already exists
Technical dependencies in a project can be used to predict success or failure in
 builds or code changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most 
research in this area is based on identifying
central modules inside a large code base which are likely to cause software failures or
detecting frequently changed code that can be associated with previous failures
~\cite{Kim:2006:AIB}. 
This module-based method also results in predictions 
at the file or binary level of software development as opposed to a code change level
and often lack the ability to provide recommendations for improved coordination
other than test focus.

% Research question
With the power of technical dependencies in predicting software failures, the question I
investigated in this project is : 
"\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
statistically relate to bugs?}"

% Explain our brief intended approach
This paper explains the approach used to locate these pairs of developers in developer networks.
The process utilizes code changes and the call hierarchies effected  to find patterns of developer 
relationships in successful and failed code changes.  These developer relationships can be used
to also promote the idea of leveraging socio-technical congruence, a measure of coordination compared
to technical dependencies amoungst stakeholders, to provide coordination recommendations.

% Technical network figure
\begin{figure*}[tb!]
\centering
\includegraphics[width=0.9\textwidth]{images/TecNetwork}
\caption{A technical network for a code change. Carl has changed method getX() which is being
called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
\end{figure*}

\section{Technical Approach}

\subsection{Extracting Technical Networks}
% End goal
The basis of this approach is to create a technical network of developers based on method ownership
and those method's call hierarchies effected by code changes. These networks will provide
dependency edges between contributors caused by code changes which may be 
identified as possible failure inducing pairings (Figure~\ref{fig:network}). To achieve this goal,
developer owners of methods, method call hierarchies (technical
dependencies) and code change effects on these hierarchies must be identified.
This approach is described in detail by illustrating its application to mining the data in a Git
repository although it can be used with any software repository.

% Owners
To determine which developers own which methods at a given code change,
the Git repository is queried. Git stores developers of a file per line, which was used to extrapolate
a percentage of ownership given a method inside a file. If developer A has written 6/10 lines of 
method foo, then developer A owns 60\% of said method.

% Call Graph
A method call graph is then constructed to extract method call hierarchies in a project at a given code change. 
Unlike other approaches such as Bodden's et al.~\cite{Bodden:2003:HVJ} 
of using byte code and whole projects, call graphs are built directly from source code files
inside of a code change, 
which does not have the assumptions of being able to compile or have access to all project 
files. It is important to not require project compilation at each code change because it is
an expensive operation as well as code change effects may cause the project
to be unable to compile. Using source files also allowed an update to the call graph
with changed files as opposed to completely rebuilding at 
every code change. This creates a rolling call graph which 
is used to show method hierarchy at each code change inside a project opposed to
a static project view. As some method invocations may only be determined at run time, all
possible method invocations are considered for these types of method calls while constructing
the call graph.

% Changeset effects
The code change effect, if any, to the call hierarchy is now found. The Git
software repository is used to determine what changes were made to any give file inside a 
code change. Specifically, methods modified by a code change are searched for. The call graph 
is then used to determine which methods call those that have been changed, which
gives the code change technical dependencies.

%Resulting network
These procedures result in a technical network based on contributor method ownership 
inside a call hierarchy effected by a code change (Figure~\ref{fig:network} left hand side).
The network is then simplified by only using edges between developers, since the results
are only interested in discovering the failure inducing edges between developers and not the 
methods themselves (Figure~\ref{fig:network} right hand side). This is the final technical 
network.

\subsection{Identifying Failure Inducing Developer Pairs}
To identify failure inducing developer pairs (edges) inside technical networks, 
edges in relation to discovered code change failures are now analysed. To determine whether a code change 
was a success or failure (introduce a software failure), the approach of
Sliwerski et al.~\cite{Sliwerski:2005:CIF} is used. The following steps are then taken:

\begin{enumerate}
\item Identify all possible edges from the technical networks.
\item For each edge, count occurrences in technical networks of failed code changes.
\item For each edge, count occurrences in technical networks of successful code changes.
\item Determine if the edge is related to success or failure.
\end{enumerate}

To determine an edge's relation to success or failure,  the value FI (failure
index) which represents the normalized chance of a code change failure in the presence
of the edge, is created.

\begin{equation}
\text{FI} = \frac{ \text{edge}_{failed} / \text{total}_{failed}}{\text{edge}_{failed} / \text{total}_{failed} + \text{edge}_{success} / \text{total}_{success}}
\end{equation}

Developer pairs with the highest FI value are said to be failure inducing structures
inside a project. These edges are stored in Table~\ref{tab:ratio}. A Fisher Exact Value test 
is also preformed on edge appearance in successful and failed
code changes, and non-appearance in successful and failed code changes to only
consider statistically significant edges (Table ~\ref{tab:ratio}'s p-value).

\section{Results}
To Illustrate the use of the approach, I conducted a case study of
the Hibernate-ORM project, an open source Java 
application hosted on GitHub\footnote{https://github.com/} with issue tracking 
performed by Jira\footnote{http://www.atlassian.com/software/jira/overview}.

This project was chosen because the tool created only handles Java code and it is written in Java 
for all internal structures and control flow
and uses Git for version control. Hibernate-ORM also uses issue tracking software which 
is needed for determining code change success or failure~\cite{Sliwerski:2005:CIF}.

In Hibernate-ORM, 27 statistically significant failure inducing developer pairs (A FI value of 0.5 or higher) 
were found out of a total of 46 statistically significant pairs that existed over the project's lifetime.
The pairings are ranked by their respective FI values (Table~\ref{tab:ratio}). These pairs can be used
to create coordination reccomendations between the developers to catch potential software malfunctions.

\begin{table}[h]
\begin{center}
\begin{tabular}{@{\hspace{.2cm}}ccc@{\hspace{.75cm}}c@{\hspace{.2cm}}c@{\hspace{.2cm}}}
\hline
Pair & Successful & Failed & FI & P-Value\\
\hline
(Daniel, Anne)	&	0&	14&	1.0000& 0.0001249		\\
(Carl, Bret)	&	1&	12&	0.9190& 0.003468	\\
(Emily, Frank)	&	1&	9&	0.8948& 0.02165      \\
\hline
\end{tabular}
\end{center}
\caption{Top 3 failure inducing developer pairs found in the study.\label{tab:ratio}}
\end{table}

\section{Conclusion and Future Work}
Technical dependencies are often used to predict software failures
in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. 
This paper has presented a method for detecting failure inducing pairs of developers inside
of technical networks based on code changes. These developer pairs can be used in the prediction
of future bugs as well as provide coordination recommendations for developers within a project.

In future work, the adding of communication networks on a per code change basis is planned. In this,
the investigation of social and technical congruence on these networks and its effects on
software quality will be studied.

\bibliographystyle{IEEEtran}
\bibliography{paper}

% End of the paper
\end{document}

### 2013-03-04 icse-src-2013/paper.tex
FINAL camera ready of SRC 2013

@@
 investigated in this project is : 
-"\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
-statistically relate to bugs?}"
+``\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
+statistically relate to bugs?}''
 
@@
 The process utilizes code changes and the call hierarchies effected  to find patterns of developer 
-relationships in successful and failed code changes.  These developer relationships can be used
-to also promote the idea of leveraging socio-technical congruence, a measure of coordination compared
-to technical dependencies amoungst stakeholders, to provide coordination recommendations.
+relationships in successful and failed code changes.  These developer relationships can also be used
+to promote the idea of leveraging socio-technical congruence, a measure of coordination compared
+to technical dependencies amongst stakeholders, to provide coordination recommendations.
 
@@
 The basis of this approach is to create a technical network of developers based on method ownership
-and those method's call hierarchies effected by code changes. These networks will provide
+and those methods' call hierarchies effected by code changes. These networks will provide
 dependency edges between contributors caused by code changes which may be 
@@
 inside a call hierarchy effected by a code change (Figure~\ref{fig:network} left hand side).
-The network is then simplified by only using edges between developers, since the results
-are only interested in discovering the failure inducing edges between developers and not the 
+The network is then simplified by only using edges between developers, since I
+am only interested in discovering the failure inducing edges between developers and not the 
 methods themselves (Figure~\ref{fig:network} right hand side). This is the final technical 
@@
 \section{Results}
-To Illustrate the use of the approach, I conducted a case study of
+To illustrate the use of the approach, I conducted a case study of
 the Hibernate-ORM project, an open source Java 
@@
 
-In Hibernate-ORM, 27 statistically significant failure inducing developer pairs (A FI value of 0.5 or higher) 
+In Hibernate-ORM, 27 statistically significant failure inducing developer pairs (FI value of 0.5 or higher) 
 were found out of a total of 46 statistically significant pairs that existed over the project's lifetime.
-The pairings are ranked by their respective FI values (Table~\ref{tab:ratio}). These pairs can be used
-to create coordination reccomendations between the developers to catch potential software malfunctions.
+The pairings are ranked by their respective FI values (Table~\ref{tab:ratio}).
 

### 2013-03-15 icse-src-2013/paper.tex
Balance columns
SRC-2013 camera ready update.

@@
 
+\usepackage{balance}
+\usepackage{float}
+\floatstyle{plaintop}
+\restylefloat{table}
+
 % Correct bad hyphenation here
@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{Identifying Failure Inducing Developer Pairs Within Developer Networks}
+\title{Identifying Failure Inducing Developer Pairs within Developer Networks}
 
@@
 \IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia \\ jell@uvic.ca}
+Victoria, British Columbia, Canada \\ jell@uvic.ca}
 }
@@
 \end{center}
-\caption{Top 3 failure inducing developer pairs found in the study.\label{tab:ratio}}
+\caption{Top 3 failure inducing developer pairs found.\label{tab:ratio}}
 \end{table}
@@
 \bibliographystyle{IEEEtran}
+\balance
 \bibliography{paper}

### 2013-06-05 impact-position-2013/paper.tex
Added new version of impact position paper.

--- 2013-02-05 icse-chase-2013/paper.tex
@@
 projects. To describe \textit{Impact}, we introduce previous indirect 
-conflict awareness attempts, our design 
-and implementation, as well as \textit{Impact's} potential through
-two evaluation studies. 
+conflict awareness attempts, the existing problems in the field, our design 
+and implementation for a potential solution, as well as \textit{Impact's} future work
+on a proposed solution to information overload which arises with indirect
+conflict awareness.
 \end{abstract}
@@
 Indirect conflicts arising in source code are inherently
-difficult to resolve as most of the time, source code analysis must
+difficult to resolve as most of the time, source code analysis or
+program slicing~\cite{Weiser:1981:PS} must
 be performed in order to find relationships between technical objects
@@
 
+While indirect conflicts remain to be a large problem area in task awareness,
+some attempts have been made to both detect and provide developer awareness
+of indirect conflicts. Servant et al.~\cite{Servant:2010:CPI}, have devised a tool
+which can both detect and alert developers to indirect conflicts arising in 
+their software projects. However, a main issue of information overload continues
+to arise from this and many other indirect conflict tools. As software systems
+grow larger and more complex, indirect dependencies tend to become more numerous.
+This causes small changes in source code to have a large impact on the software system
+which these tools often report exhaustively. Information overload must be addressed
+in any feasible attempt made on indirect conflicts.
+
 In this paper, we report on our research into supporting indirect conflicts
-and present the design, implementation, and evaluation of the tool \textit{Impact},
+and present the design, implementation, and future evaluation of the tool \textit{Impact},
 a web based tool that aims at detecting indirect conflicts among developers
@@
 (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
-alert developers involved in such conflicts in task awareness. While this paper
+alert developers involved in such conflicts in task awareness while limiting information
+overload by using design by contract~\cite{} solutions to method design. While this paper
 outlines \textit{Impact's} specific implementation, its design is rather
@@
 \textit{Impact}. In the following section we describe a generic design and implementation
-of \textit{Impact} as an awareness tool. We then discuss a preliminary evaluation of
-\textit{Impact} followed by conclusions and future work.
+of \textit{Impact} as an awareness tool. We then discuss future work in terms
+of evaluation and further understanding of the problems at hand.
 
@@
 today, only a handful have made an attempt to examine indirect conflicts.
-Here, we will outline three of the forefront projects in indirect conflicts
+Here, we will outline four of the forefront projects in indirect conflicts
 and how these projects have influenced the decision making process in
@@
 graphs to view impacts of conflict creating code, they lack the ability to
-detect conflicts on their own.
+detect potential conflicts on their own.
 
@@
 
+We will lastly examine the tool CASI~\cite{Servant:2010:CPI} which uses
+a sphere of influence for each developer to determine how source code
+changes are indirectly related to other components of a software project.
+CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
+call graphs in Ariadne~\cite{Trainer:2005:BGT} which gives dependencies among
+all source code entities. This provides a verbose output of dependencies when
+source code is changed. CASI also implements a visualization where
+a developer can see what parts of a software projects he or she may be effecting
+with the source code change. This allows the developer themselves to go and fix
+potential issues elsewhere in the project before the code change is committed
+to the software repository. While CASI covers great ground in its approach,
+it still leaves the issue of information overload, although attempts were made
+to solve this by having severity levels of indirect conflicts presented to
+the user.
+
 \textit{Impact} is designed to address these limitations by extending this work in
-two directions: (1) the detection of indirect conflicts at an internal level
-of technical objects as opposed from object signatures, and (2) the dissemination of
+two directions: (1) the detection of high importance indirect conflicts at an internal level
+of technical objects, (2) the dissemination of
 awareness information to all appropriate developers regardless of their current
-workspace activities. \textit{Impact} is also designed around the successes these
-projects have had in the past with directed graphs as 
+workspace activities, and (3) potential solutions to information overload. 
+\textit{Impact} is also designed around the successes these
+projects have had in the past with dependency graphs as 
 well as elements of collection, ownership, and distribution functionality.
@@
 While technical object relationships form the basis of indirect conflicts,
-communication between developers is the ultimate goal of resolving such conflicts.
+communication between developers is our ultimate goal of resolving such conflicts.
 This being the case, developer ownership must be placed on the 
@@
 the object's entire body (method definition and internal body) is 
-parsed at commit time to detect changes anywhere in the technical object.
+parsed, similar to that of CASI~\cite{Servant:2010:CPI},
+at commit time, as opposed to real time, to detect changes anywhere in the technical object.
+This is a first design step towards avoiding information overload.
 Once technical objects are found to be changed, appropriate owners
-of objects which use the changed object should be notified.
+of objects which use the changed object should be notified. However, to
+avoid information overload at this step, analyses of the changed object
+must be preformed to determine whether or not it is a ``notification worthy''
+change. Our proposed solution to this is explained in Section~\ref{sec:ns}.
 In Figure~\ref{fig:graph}, Carl changes method (technical object) 1,
 which effects methods 2 and 3 resulting in the alerting of
-developers Bret, Daniel, and Anne.
+developers Bret, Daniel, and Anne. We have opted to alert the invoking
+developers rather than the developer making the change to potential
+solutions as our conflicts are detected at commit time and this supports
+the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} 
+from software structure to communication patterns in awareness systems.
 
@@
 objects, (ii) assigning ownership to those technical objects, and (iii)
-detecting changes within commit time and the notification of conflict information
+detecting ``notification worthy'' changes at commit time and the dissemination of conflict information
 to appropriate owners, we believe a wide variety of
 indirect conflict awareness tools can be created or extended. The
-implementation of \textit{Impact} described in the following section follows these
-three design guidelines.
+on going implementation of \textit{Impact} described in the following 
+section follows these three design guidelines.
 
@@
 relationship described above for methods is method invocation.
-Thus, in our constructed directed graph, methods represent nodes
+Thus, in our constructed dependency graph, methods represent nodes
 and method invocations represent our directed edges. In order to 
 construct this directed graph, abstract syntax trees (ASTs) are 
-constructed from source files in the project. The ASTs allow us
-to construct method call graphs from which the directed 
-graphs can be constructed.
+constructed from source files in the project.
 
@@
 repositories have similar commands and functionality.) This command 
-returns the source code authors per linee which can is used to assign
-ownership to methods. If a method has 10 lines and developer A
-has written 3 while developer B has written 7, then ownership is
-assigned 30\% and 70\% respectively.
+returns the source code authors per line which can be used to assign
+ownership to methods.
 
@@
 find methods that have been changed. This gives cause of potential
-indirect conflicts. We now find all methods in our directed graphs
-which invoke these changed methods. These are the final indirect
-conflicts.
+indirect conflicts. Here is where information overload can occur and must
+be dealt with. Impact's current state does not address this as our potential 
+solution is still under investigation, but is the
+main focus of our next steps as detailed in Section~\ref{sec:ns}.
+We now find all methods in our directed graphs which invoke these changed methods. 
+These are the final indirect conflicts.
 
@@
 of methods which invoke those that have been changed are alerted
-to the newly changed method. This can been seen in
+to the newly changed method. Impact can been seen in
 Figure~\ref{fig:impact}, the user interface of \textit{Impact}. Here, in an RSS type
 feed, the changing developer, time of change, changed method,
-invoking methods, and commit message are all displayed. This 
-list is auto updating as new commits are pushed to the central
-git repository. The information provided also comes with a simple
-weight of the percent changed of changed method multiplied by 
-ownership of the invoking method. This allows developers to filter
-through high and low changes affecting their own source code.
-Developers are now able to be alerted of detected indirect
-conflicts and solve them by communicating the developer
-who authored the change.
+invoking methods, and commit message are all displayed. 
+The weight provided will be an indication of importance based
+on the future work to be outline in Section~\ref{sec:ns}. This
+again helps solve the issue of information overload.
 
@@
 
-\section{Evaluation}
-To fully evaluate both the generic design of detecting and resolving
-indirect conflicts as well as \textit{Impact}, extensive testing and evaluation
-must be performed. However, we feel that a simple evaluation is
-first needed to assess the foundation of \textit{Impact}'s design and claims
-about indirect conflicts at the method level.
-
-We performed a user case study where we gave \textit{Impact} to 
-two small development teams composed of three developers. Each team was
-free to use \textit{Impact} at their leisure during their development process,
-after which interviews were conducted with lead developers from 
-each development team. The interviews were conducted after each
-team had used \textit{Impact} for three weeks.
-
-We asked lead developers to address two main concerns: do indirect
-conflicts pose a threat
-at the method level (e.g. method 1 has a bug because it invokes method
-2 which has has its implementation changed recently), 
-and did \textit{Impact} help raise
-awareness and promote quicker conflict resolution for indirect
-conflicts. Our two interviews largely supported our expectation of
-indirect conflicts posing a serious threat to developers, especially
-in medium to large teams or projects as opposed to the small
-teams which they were apart of. It was also pointed
-out that method use can be a particularly large area for indirect
-conflicts to arise. However, it was noted that
-any technical object which is used as an interface to some data
-construct or methodology, database access for instance, can be 
-a large potential issue for indirect conflicts.  Interview response to
-\textit{Impact} was also largely positive, as interviewees stated that \textit{Impact}
-helped raise awareness among their teams with what other developers
-are doing as well as the influence it has on their own work. However,
-\textit{Impact} was shown to have information overload. It was suggested
-that while all method changes were being detected,
-not all are notification worthy. One developer suggested to only notify
-developers to indirect conflicts if the internal structure of a method
-changes due to modification to input parameters or output parameters.
-In other words, the boundaries of the technical objects (changing
-how a parameter is used inside the method, modifying the return
-result inside the method) seem to be more of interest than other 
-internal workings. More complex inner workings of methods were also noted 
-to be of interest to developers
-such as cyclomatic complexity, or time and space requirements.
-
-These two studies have shown that our design and approach to
-detecting and alerting developers to indirect conflicts appear
-to be on the correct path. \textit{Impact} as a tool has laid the foundations
-for future work in detecting indirect conflicts as well as notifying
-developers, although more thought must be given as to 
-what constitutes a meaningful change inside our selected 
-technical objects to avoid issues of information overload.
-
-\section{Next Steps in Research of Indirect Conflicts}
-While \textit{Impact} has laid the foundations for awareness tool construction
-aimed at resolving indirect conflicts, it has also raised questions 
-in regards to levels of granularity in notification of indirect conflicts and the
-implied trade off between notifications and information overload. To address
-these problems, we plan to find industry expert opinion through interviews
-and surveys with experienced developers. From these interactions, we plan to
-answer some of the questions presented in this paper such as which technical
-objects concern developers with indirect conflicts, and what types of changes
-to these objects are cause for notification. We are also planning, through these
-expert opinions, to further the 
-need for indirect conflict awareness tools by establishing that indirect
-conflicts are real world problems with harmful consequences. 
-Further more, we plan to bring these opinions back to the realm of 
-\textit{Impact}, in combination with other academic works on balancing
-between notifications and information overload~\cite{Wang:2007}, to
-continue our development
-on this tool to further its life both as a research tool and a possible
-industry application.
+\section{Proposed Solution to Information Overload}
+\label{sec:ns}
+As was previously stated, the main issue to be addressed with Impact is
+information overload. Our potential solution comes from the ideas of
+Myers~\cite{} on ``design by contract''. In this methodology, changes to method
+preconditions and postconditions are considered to be the most harmful. 
+This includes adding conditions that must be met by both input
+and output parameters such as limitations to input and expected
+output. To achieve this level of source code analysis, the ideas of
+Fluri et al.~\cite{Fluri:2007:CDT} can be used on the previously generated
+ASTs for high granularity
+source code change extraction when determining if preconditions or
+postconditions have changed.
+We plan to bring these ideas into the realm of what makes a
+change ``notification worthy'', as to deal with information overload for
+developers using Impact. By only examining ``notification worthy'' changes,
+Impact would cut down the number of notifications 
+issued to just those which developers care about. To investigate this solution, we plan to
+interview and survey industry software developers at a wide range of
+experience to determine what makes a change in software important. We believe
+by only monitoring what developers and Myers~\cite{} find important to
+indirect conflict awareness, we will be able to eliminate information overload
+in Impact and other dependency awareness tools. We believe that these preconditions
+and postconditions will serve as the main tool for dealing with information overload.
 
@@
 awareness tool, \textit{Impact}, which was designed around our generic 
-technical object approach. \textit{Impact} was evaluated on a small scale, showing
-its future potential as well as highlighting its current weaknesses. 
-Our preliminary evaluation has given us concrete insights into 
-understanding awareness tools that specifically address indirect conflicts and has
-strengthened the need for further research towards notification and information
-overload (what makes a change notification
-worthy), which has become the goal of our next step.
+technical object approach. We have also proposed a solution to information
+overload which occurs in most indirect conflict awareness tools. We plan
+on investigating our solution further through developer interview and
+surveys, and by finally implementing the solution into Impact.
 

### 2013-06-05 impact-position-2013/paper.tex
Added missing Meyer reference.

@@
 alert developers involved in such conflicts in task awareness while limiting information
-overload by using design by contract~\cite{} solutions to method design. While this paper
+overload by using design by contract~\cite{Meyer:1988} solutions to method design. While this paper
 outlines \textit{Impact's} specific implementation, its design is rather
@@
 information overload. Our potential solution comes from the ideas of
-Myers~\cite{} on ``design by contract''. In this methodology, changes to method
+Meyer~\cite{Meyer:1988} on ``design by contract''. In this methodology, changes to method
 preconditions and postconditions are considered to be the most harmful. 
@@
 experience to determine what makes a change in software important. We believe
-by only monitoring what developers and Myers~\cite{} find important to
+by only monitoring what developers and Meyer~\cite{Meyer:1988} find important to
 indirect conflict awareness, we will be able to eliminate information overload

### 2013-08-11 apie-2013/paper.tex
Added skeletons for api and iced papers to be written.

--- 2013-03-15 icse-src-2013/paper.tex
@@
-% This will be the main document for the Technical Networks paper to
-% be written by the Eggnet team of Jordan Ell, Triet Huynh and Braden
-% Simpson in association with Adrian Schroeter and Daniela Damian.
-
 \documentclass[conference]{IEEEtran}
@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{Identifying Failure Inducing Developer Pairs within Developer Networks}
+\title{APIE Evolution: An Exploration into Software API Trends and Processes}
 
@@
 
+\author{\IEEEauthorblockN{Braden Simpson}
+\IEEEauthorblockA{University of Victoria,
+Victoria, British Columbia, Canada \\ braden@uvic.ca}
+}
+
+\author{\IEEEauthorblockN{Daniela Damian}
+\IEEEauthorblockA{University of Victoria,
+Victoria, British Columbia, Canada \\ danielad@csc.uvic.ca}
+}
+
 % Make the title area
@@
 \begin{abstract}
-Software systems have not only become larger over time, but the amount of
-technical contributors and dependencies have also increased. With these expansions also comes
-the increasing risk of introducing a software failure into a pre-existing system.
-Software failures are a multi-billion dollar problem in the industry today and while integration and
-other forms of testing are helping to ensure a minimal number of failures, research to understand
-full impacts of code changes and their social implications is still a major concern. This paper describes
-how analysis of code changes and the technical relationships they infer can be used to detect pairs
-of developers whose technical dependencies may induce software failures. These developer pairs may
-also be used to predict future software failures as well as provide recommendations to contributors
-to solve these failures caused by source code changes.
+
 \end{abstract}
@@
 
-% Setup the problem.
-Large software projects are created using highly modular and 
-reusable code. This creates technical dependencies between methods or functions
-that can be used in a wide variety of locations throughout the project. 
-This causes changes to any given method to have a rippling 
-effect across the rest of the project~\cite{Acharya:2011:PCI}. The larger these effects are,
-the more likely they are to cause a software failure inside the system during the project's
-life span~\cite{Zimmermann:2008:PDU}. These observations of technical dependencies open
-the door to types of analysis on the developer networks they infer and preventing software
-failures by improving coordination amongst dependant developers.
+\section{Related Work}
 
-% Explain technical network research that already exists
-Technical dependencies in a project can be used to predict success or failure in
- builds or code changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most 
-research in this area is based on identifying
-central modules inside a large code base which are likely to cause software failures or
-detecting frequently changed code that can be associated with previous failures
-~\cite{Kim:2006:AIB}. 
-This module-based method also results in predictions 
-at the file or binary level of software development as opposed to a code change level
-and often lack the ability to provide recommendations for improved coordination
-other than test focus.
+\section{Methodology}
 
-% Research question
-With the power of technical dependencies in predicting software failures, the question I
-investigated in this project is : 
-``\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
-statistically relate to bugs?}''
+\subsection{Results}
 
-% Explain our brief intended approach
-This paper explains the approach used to locate these pairs of developers in developer networks.
-The process utilizes code changes and the call hierarchies effected  to find patterns of developer 
-relationships in successful and failed code changes.  These developer relationships can also be used
-to promote the idea of leveraging socio-technical congruence, a measure of coordination compared
-to technical dependencies amongst stakeholders, to provide coordination recommendations.
+\section{Discussion}
 
-% Technical network figure
-\begin{figure*}[tb!]
-\centering
-\includegraphics[width=0.9\textwidth]{images/TecNetwork}
-\caption{A technical network for a code change. Carl has changed method getX() which is being
-called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
-\end{figure*}
+\section{Future Work}
 
-\section{Technical Approach}
-
-\subsection{Extracting Technical Networks}
-% End goal
-The basis of this approach is to create a technical network of developers based on method ownership
-and those methods' call hierarchies effected by code changes. These networks will provide
-dependency edges between contributors caused by code changes which may be 
-identified as possible failure inducing pairings (Figure~\ref{fig:network}). To achieve this goal,
-developer owners of methods, method call hierarchies (technical
-dependencies) and code change effects on these hierarchies must be identified.
-This approach is described in detail by illustrating its application to mining the data in a Git
-repository although it can be used with any software repository.
-
-% Owners
-To determine which developers own which methods at a given code change,
-the Git repository is queried. Git stores developers of a file per line, which was used to extrapolate
-a percentage of ownership given a method inside a file. If developer A has written 6/10 lines of 
-method foo, then developer A owns 60\% of said method.
-
-% Call Graph
-A method call graph is then constructed to extract method call hierarchies in a project at a given code change. 
-Unlike other approaches such as Bodden's et al.~\cite{Bodden:2003:HVJ} 
-of using byte code and whole projects, call graphs are built directly from source code files
-inside of a code change, 
-which does not have the assumptions of being able to compile or have access to all project 
-files. It is important to not require project compilation at each code change because it is
-an expensive operation as well as code change effects may cause the project
-to be unable to compile. Using source files also allowed an update to the call graph
-with changed files as opposed to completely rebuilding at 
-every code change. This creates a rolling call graph which 
-is used to show method hierarchy at each code change inside a project opposed to
-a static project view. As some method invocations may only be determined at run time, all
-possible method invocations are considered for these types of method calls while constructing
-the call graph.
-
-% Changeset effects
-The code change effect, if any, to the call hierarchy is now found. The Git
-software repository is used to determine what changes were made to any give file inside a 
-code change. Specifically, methods modified by a code change are searched for. The call graph 
-is then used to determine which methods call those that have been changed, which
-gives the code change technical dependencies.
-
-%Resulting network
-These procedures result in a technical network based on contributor method ownership 
-inside a call hierarchy effected by a code change (Figure~\ref{fig:network} left hand side).
-The network is then simplified by only using edges between developers, since I
-am only interested in discovering the failure inducing edges between developers and not the 
-methods themselves (Figure~\ref{fig:network} right hand side). This is the final technical 
-network.
-
-\subsection{Identifying Failure Inducing Developer Pairs}
-To identify failure inducing developer pairs (edges) inside technical networks, 
-edges in relation to discovered code change failures are now analysed. To determine whether a code change 
-was a success or failure (introduce a software failure), the approach of
-Sliwerski et al.~\cite{Sliwerski:2005:CIF} is used. The following steps are then taken:
-
-\begin{enumerate}
-\item Identify all possible edges from the technical networks.
-\item For each edge, count occurrences in technical networks of failed code changes.
-\item For each edge, count occurrences in technical networks of successful code changes.
-\item Determine if the edge is related to success or failure.
-\end{enumerate}
-
-To determine an edge's relation to success or failure,  the value FI (failure
-index) which represents the normalized chance of a code change failure in the presence
-of the edge, is created.
-
-\begin{equation}
-\text{FI} = \frac{ \text{edge}_{failed} / \text{total}_{failed}}{\text{edge}_{failed} / \text{total}_{failed} + \text{edge}_{success} / \text{total}_{success}}
-\end{equation}
-
-Developer pairs with the highest FI value are said to be failure inducing structures
-inside a project. These edges are stored in Table~\ref{tab:ratio}. A Fisher Exact Value test 
-is also preformed on edge appearance in successful and failed
-code changes, and non-appearance in successful and failed code changes to only
-consider statistically significant edges (Table ~\ref{tab:ratio}'s p-value).
-
-\section{Results}
-To illustrate the use of the approach, I conducted a case study of
-the Hibernate-ORM project, an open source Java 
-application hosted on GitHub\footnote{https://github.com/} with issue tracking 
-performed by Jira\footnote{http://www.atlassian.com/software/jira/overview}.
-
-This project was chosen because the tool created only handles Java code and it is written in Java 
-for all internal structures and control flow
-and uses Git for version control. Hibernate-ORM also uses issue tracking software which 
-is needed for determining code change success or failure~\cite{Sliwerski:2005:CIF}.
-
-In Hibernate-ORM, 27 statistically significant failure inducing developer pairs (FI value of 0.5 or higher) 
-were found out of a total of 46 statistically significant pairs that existed over the project's lifetime.
-The pairings are ranked by their respective FI values (Table~\ref{tab:ratio}).
-
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{@{\hspace{.2cm}}ccc@{\hspace{.75cm}}c@{\hspace{.2cm}}c@{\hspace{.2cm}}}
-\hline
-Pair & Successful & Failed & FI & P-Value\\
-\hline
-(Daniel, Anne)	&	0&	14&	1.0000& 0.0001249		\\
-(Carl, Bret)	&	1&	12&	0.9190& 0.003468	\\
-(Emily, Frank)	&	1&	9&	0.8948& 0.02165      \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Top 3 failure inducing developer pairs found.\label{tab:ratio}}
-\end{table}
-
-\section{Conclusion and Future Work}
-Technical dependencies are often used to predict software failures
-in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. 
-This paper has presented a method for detecting failure inducing pairs of developers inside
-of technical networks based on code changes. These developer pairs can be used in the prediction
-of future bugs as well as provide coordination recommendations for developers within a project.
-
-In future work, the adding of communication networks on a per code change basis is planned. In this,
-the investigation of social and technical congruence on these networks and its effects on
-software quality will be studied.
+\section{Conclusions}
 

### 2013-08-13 iced-2013/paper.tex
Added skeletons for api and iced papers to be written.
ICED interview methodology draft.

--- 2013-08-11 apie-2013/paper.tex
@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{APIE Evolution: An Exploration into Software API Trends and Processes}
+\title{Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight}
 
@@
 Victoria, British Columbia, Canada \\ jell@uvic.ca}
-}
-
-\author{\IEEEauthorblockN{Braden Simpson}
-\IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ braden@uvic.ca}
 }
@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 
-\subsection{Results}
+Our study was performed in two parts. First, a round of semi-structured interviews were conducted which 
+addressed the larger who, what, when, where, and how about indirect conflicts. Secondly, a survey was given
+out which was used to confirm what was learned from the interviews on a larger sample size as well as get
+a larger opinion base on general issues.
 
-\section{Discussion}
+\subsection{Interview Participants}
+
+We selected our participants from a large breadth of both open and closed source software development companies
+and projects. The population which participated in our interview came from the following software groups: 
+IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
+Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
+Our participants we chosen based on their direct involvement in the actual writing of software for
+their respective companies or projects. These participants' software development titles included: senior developer,
+lead developer, software developer, software developer in test, and junior developer.
+In addition to software development, some participants were also chosen based on their previous experience
+of software development as well as their current experience with project management at some capacity. These management
+type participants involved in project planning to some capacity between technical architecture, project maintainer,  
+chief technical officer, or program manager.
+
+\subsection{Interview Procedure}
+
+Participants were invited to participate in the interviews by email and were sent a single reminder email one week
+after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
+19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
+interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
+audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
+our breadth of participants was rather large spanning multiple open and closed source project and our interview 
+answer saturation was quite high (to be seen in Section~\ref{sec:rnd}). However, we were disappointed in our word
+of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
+their respective organizations but did not receive any additional participants through this method.
+
+Interview participants first answered a number of structured demographic items. Next, participants were 
+asked to describe various software development processes they use or encounter in their current working
+environment. These structured process questions were used for later analysis to provided context to
+any answer for questions that involve indirect conflicts. They next were engaged in a
+semi-structured round of questions pertaining to 12 categories of questions. Each of the categories was used as a
+point of aggregation for later analysis as a way of answering the X research questions laid out in 
+Section~\ref{sec:intro}. While each of the 12 question categories had a number of starter questions, interviews 
+largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
+In general, once the category was explained to the participant, little prompting was necessary as the participant
+would jump right into an experience, or issue he or she had with the question or category. However, not all
+participants had strong opinions or any experience on every category mentioned. For these participants, answers 
+to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
+either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
+being apart of any one software project for extended periods of time. We account for these non answers
+in our analysis as seen in Section~\ref{sec:rnd}. Interviews lasted 
+from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
+participant had in software development as they were able to give more experiences and more informed opinions.
+
+After the interviews, we extracted a list of answers, opinions, and experiences for each of the 12 categorical
+question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
+with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
+Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
+aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
+experience and opinion for the 12 categories. These hypotheses were then used as a template for creating the
+second step survey, in order to further test the answers given by interview participants and validate any possible
+conclusions to be drawn.
+
+\section{Results and Discussion}
+\label{sec:rnd}
 
@@
 \section{Conclusions}
+
+\section{Acknowledgments}
+
+We would sincerely like to thank all participants who were willing to be interviewed or who participated
+in completeing our survey. We thank these people for taking time out of their day to participate and for
+sharing their developer experience with us. Without these people our research could not have been possible.
 

### 2013-08-13 iced-2013/paper.tex
ICED survey methodology draft.

@@
 
-% Authors names
 \author{\IEEEauthorblockN{Jordan Ell}
 \IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ jell@uvic.ca}
-}
-
-\author{\IEEEauthorblockN{Daniela Damian}
+Victoria, Canada \\ jell@uvic.ca}
+IEEEauthorblockN{Daniela Damian}
 \IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ danielad@csc.uvic.ca}
+Victoria, Canada \\ danielad@csc.uvic.ca}
 }
@@
 
-We selected our participants from a large breadth of both open and closed source software development companies
+We selected our interview participants from a large breadth of both open and closed source software development companies
 and projects. The population which participated in our interview came from the following software groups: 
@@
 
+\subsection{Survey Participants}
+
+We selected our survey participants from a similar breadth of open and closed source software development 
+companies and projects as the interviews participants with two large exceptions. The software organizations
+that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
+Subnet Solutions, Ruboss Technology Corporation, Amazon, Kano Apps, Fireworks Design, and Frost Tree Games.
+However, participants who took part in the round of interviews were not asked to participate in the surveys but
+rather to act as a contact point for other developers in their team, project, or organization who may be interested
+in completing the survey. Aside from this aforementioned list, two large groups of developers were asked to
+participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
+users were selected based on their large amounts of development activity on GitHub and the Apache developers
+were selected based on their software development participation on specific projects known to be used heavily
+by other organizations and projects.
+
+\subsection{Survey Procedure}
+
+The created survey was based off of the 12 categorical hypotheses given by the round of interviews. The survey
+was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
+or different opinions from those already acquired from the interviews. The survey went through two rounds of
+piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
+with feedback given at the end. The previous interview participants were selected based on their domain knowledge
+expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
+in a more polished survey. The final survey consisted of 2 multiple choice, 3 level of agreement,
+6 level of use, and 9 short answer questions. Each non demographic question was made optional as it was shown 
+through the interviews that some questions require more experience from participants than may be provided.
+
+Survey participants were invited to participate in the survey by email. No reminder email was ever sent as the
+survey responses were not connected with the invitation email addresses and thus participants who did respond
+could not be removed from a reminder list. We directly emailed 1300 participants and ended with 78 responses
+giving a response rate of 6\%. We attribute the low response rate with a couple of factors. One, the surveys
+were conducted during the months of July and August while many participants may be away from their regular positions.
+Two, our GitHub and Apache participants could not be verified as to whether or not they actively support the
+email addresses used in the invitations. And finally, the survey was considered by some to be long and require
+more development experience than may have been typical of some of those invited to participate.
+
+After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
+as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
+answer question fields while combining commonly mentioned answers where possible. The results again from this
+preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.
+
 \section{Results and Discussion}
@@
 
-\bibliographystyle{IEEEtran}
-\balance
-\bibliography{paper}
+%\bibliographystyle{IEEEtran}
+%\balance
+%\bibliography{paper}
 

### 2013-08-13 iced-2013/paper.tex
ICED methodology draft except analysis.

@@
 any answer for questions that involve indirect conflicts. They next were engaged in a
-semi-structured round of questions pertaining to 12 categories of questions. Each of the categories was used as a
+semi-structured round of questions pertaining to 12 categories of questions.
+
+\begin{itemize}
+\item Software development tools for dependency tracking and awareness.
+\item Software development process for preventing indirect conflicts.
+\item Software artifact dependency levels and where conflicts can arise.
+\item How developers find internal or external software dependencies.
+\item Examples of indirect conflicts from real world experiences.
+\item How indirect conflicts are detected and found.
+\item How indirect conflict issues are solved or dealt with.
+\item Developer opinion of preemptive measures to prevent indirect conflicts.
+\item Developer opinion on what types of changes are worth a preemptive action.
+\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
+\end{itemize}
+
+Each of the categories was used as a
 point of aggregation for later analysis as a way of answering the X research questions laid out in 
@@
 expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
-in a more polished survey. The final survey consisted of 2 multiple choice, 3 level of agreement,
-6 level of use, and 9 short answer questions. Each non demographic question was made optional as it was shown 
+in a more polished survey. The final survey consisted of the following questions:
+2 multiple choice questions for developer experience and demographic information; 3 level of agreement questions for an indication of
+what types of environments indirect conflicts are more likely to occur;
+6 level of use questions to indicate what types of software changed developers find trouble some with respect to
+indirect conflicts; and 9 short answer questions for indication of when indirect conflicts occur, what types of
+processes are used to prevent or react to indirect conflict, and to provide an outlet for general opinion
+on the matter. Each non demographic question was made optional as it was shown 
 through the interviews that some questions require more experience from participants than may be provided.

### 2013-08-14 iced-2013/paper.tex
ICED first draft of ICs nature.

@@
 \usepackage{float}
+
+\usepackage{enumitem}
+
+\setlist[description]{leftmargin=*}
 \floatstyle{plaintop}
@@
 
+We now present our results of both the interviews and surveys conducted in regards to our research questions
+outlined in Section~\ref{sec:intro}. We first present and discuss the general nature of indirect conflicts in
+terms of when, what types, and how often indirect conflicts occur. We next show the findings for the process
+related research questions in regards to how developers react to, processes to prevent, and reactionary steps
+to indirect conflicts. Lastly, we present and discuss findings for what potential tools are wanted by developers
+and what researchers are currently developing.
+
+\subsection{The Nature of Indirect Conflicts}
+
+Although a detailed manual inspection of a project's lifetime and the indirect conflicts it experiences throughout
+its development is beyond the scope of this paper, we first wanted to explore the general nature of indirect 
+conflicts. This being the case we explored, through our interviews and surveys, what software developers believed
+to be the case when it came to how indirect conflicts occur, when and how often they occur, and the types of 
+software object's that incur indirect conflicts.
+
+\begin{description}[topsep=6pt]
+	\item[RQ1] \textit{How do indirect conflicts occur?}
+\end{description}
+
+From the interviewed participants, 63\% of developers believe that a large contributing factor to the cause
+of indirect conflicts comes from the changing of a software object's contract. One developer stated that ``design
+by contract was invented to solve this problem and it does it quite well'', while another noted that software 
+object contracts do solve the problem in theory, but that doesn't mean that problems don't occur in practice.
+In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
+which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
+towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples had to do 
+with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
+are afraid of the long range issues that may arise from that change''. 36\% of interviewed developers mentioned 
+that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
+integration fails noting that major releases of libraries can be a large issue with one developer saying 
+``their build never breaks, but it breaks ours''. Finally, 29\% of developers interviewed explained that changes to
+databases are often a large source of indirect conflicts saying ``[schema changes] break stuff all over the place''.
+
+\begin{description}[topsep=6pt]
+	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
+\end{description}
+
+From the interviewed participants, 58\% of developers explained that indirect conflicts occur all the time in
+their development life cycle with a minimum occurrence of once a week. Of those 58\%, developers noted that a
+lot of their development time is put into understanding code to avoid indirect conflicts, more serious issue tend
+to happen once a month, and that the conflicts that do occur tend to be quite unique from each other.
+
+In terms of when in a project indirect conflicts are more likely to occur 63\% of developers interviewed said that
+when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
+than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
+[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
+changes a lot, especially with agile''. 21\% if developers interviewed mentioned that the development process
+changes once a stable point of a project is met and the ``add and deprecate model solves a lot of these issues''.
+
+For those developers interviewed, we asked how often indirect conflicts occur in their software development experience,
+how often they occur before and after a project's first release, as well as how often indirect conflicts occur late
+in a project's lifetime. Table~\ref{tab:often} shows the results of these questions.
+
+\begin{table*}[tp]
+\begin{center}
+\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
+\hline
+Question & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
+\hline
+\hline
+How often do you experience indirect conflicts in your software development? & 17\% & 27\% & 21\% & 17\% & 3\% & 5\% & 10\% \\ \hline
+How often do indirect conflicts occur in the early stages of a project’s development? & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
+How often do indirect conflicts occur before the first release of a project? & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
+How often do indirect conflicts occur after the first release of a project? & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
+How often do indirect conflicts occur late in a project’s lifetime? & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
+of developers surveyed.\label{tab:often}}
+\end{table*}
+
+Of those surveyed that could not give direct answers to the questions in Table~\ref{tab:often}, the more common answers were 
+as follows: ``People rushing to meet [the release date] is where most of my indirect conflicts occur.'', ``Indirect conflicts
+after a release depends on how well the project was built at first'', ``They tend to slow down a bit after a major release, 
+unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
+with spikes during large revamps or the implementation of cross-cutting new features.''.
+
 \section{Future Work}
@@
 We would sincerely like to thank all participants who were willing to be interviewed or who participated
-in completeing our survey. We thank these people for taking time out of their day to participate and for
+in completing our survey. We thank these people for taking time out of their day to participate and for
 sharing their developer experience with us. Without these people our research could not have been possible.

### 2013-08-14 iced-2013/paper.tex
ICED skeleton for other research questions now in place.

@@
 
-\section{Results and Discussion}
+\section{Results}
 \label{sec:rnd}
@@
 
-\begin{table*}[tp]
+\begin{table*}[tb!]
 \begin{center}
@@
 
+\begin{description}[topsep=6pt]
+	\item[RQ3] \textit{What software artifacts are susceptible to indirect conflicts?}
+\end{description}
+
+From the developers interviews, 9 were currently working with large scale database applications and all 9 listed database
+schema as a large source of indirect conflicts. Out of the 5 developers interviewed that currently work on either software
+library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. Some
+other notable artifacts were user interface displays for web development, and full components in component base game
+architecture.
+
+\subsection{Processes of Indirect Conflicts}
+
+The results that follow pertain to software development process and its relation to indirect conflicts. We explore
+how developers react to indirect conflicts, which stakeholders are responsible for their fixes, and if any development
+processes can be used to prevent or catch indirect conflicts.
+
+\begin{description}[topsep=6pt]
+	\item[RQ4] \textit{How do developers react to and solve indirect conflicts?}
+\end{description}
+
+% Here use anything they said for tools and process
+
+\begin{description}[topsep=6pt]
+	\item[RQ5] \textit{Who is responsible for fixing indirect conflict issues?}
+\end{description}
+
+\begin{description}[topsep=6pt]
+	\item[RQ6] \textit{Who is responsible for fixing indirect conflict issues?}
+\end{description}
+
+\subsection{Tools for indirect Conflicts}
+
+\begin{description}[topsep=6pt]
+	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
+\end{description}
+
+\begin{description}[topsep=6pt]
+	\item[RQ8] \textit{When are reactionary indirect conflict tools best used and what would developers want from them?}
+\end{description}
+
+\section{Discussion}
+
 \section{Future Work}

### 2013-08-27 iced-2013/paper.tex
ICED
Up to RQ7 are now inserted into the paper.

@@
 \begin{description}[topsep=6pt]
-	\item[RQ3] \textit{What software artifacts are susceptible to indirect conflicts?}
+	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
 \end{description}
@@
 
+As per environments, Table~\ref{tab:env} shows which development team environment developers believe to be the most
+prone to indirect conflicts.
+
+\begin{table*}[tb!]
+\begin{center}
+\begin{tabular}{| p{7cm} | c | c | c | c | c |}
+\hline
+Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
+\hline
+\hline
+Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
+Developing in a team between 2 - 5 developers (Inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
+Developing in a multi team environment (Inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Results of survey questions to development environments in which indirect conflicts are likely to occur, in terms of percentage
+of developers surveyed.\label{tab:env}}
+\end{table*}
+
 \subsection{Processes of Indirect Conflicts}
@@
 \begin{description}[topsep=6pt]
-	\item[RQ4] \textit{How do developers react to and solve indirect conflicts?}
+	\item[RQ4] \textit{How do developers catch indirect conflicts?}
 \end{description}
 
-% Here use anything they said for tools and process
+Of the 19 developers interviewed, 69\% mentioned forms of testing (unit, integration, etc) as the major component of catching indirect
+conflict issues, subscribing to the idea of ``run the regression and integration tests and just see if anything breaks''. 31\% of developers 
+said build processes (either nightly builds or building the project themselves), and others mentioned code reviews
+while those dealing with a user interface mentioned user complaints from run time testing.
+
+Of the 78 developers survey, 49\% mentioned forms of testing as the major tool used to catch indirect conflicts, 33\% said build processes,
+while 31\% used work their IDE or IDE plugins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
+and project expertise (6\%) as factors of catching indirect conflicts.
 
 \begin{description}[topsep=6pt]
-	\item[RQ5] \textit{Who is responsible for fixing indirect conflict issues?}
+	\item[RQ5] \textit{How to developers react to and solve indirect conflicts?}
 \end{description}
 
-\begin{description}[topsep=6pt]
-	\item[RQ6] \textit{Who is responsible for fixing indirect conflict issues?}
-\end{description}
+Once an indirect conflict has occurred, usually alerted to the developer by test or build failures, 75\% of developers interviewed said
+they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
+``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
+code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
+frames of when events occurred. Developers also mentioned their personal expertise in the project as a large component of solving indirect
+conflicts as well as issue or work item trackers as many tests and builds can be linked to these management tools.
+
+Of the developers surveyed, 37\% confirmed that they used some type of source control management tools to help them solve their indirect
+conflicts while 23\% said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
+to solve indirect conflicts. Interestingly, only 13\% of developers mentioned a form of communication with other developers in aid to
+solving these conflicts and only 4\% mentioned the reading of formal documentation.
 
@@
 \begin{description}[topsep=6pt]
-	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
+	\item[RQ6] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
 \end{description}
 
+The developers who were interviewed had a major concern with any pre-emptive tool in that the amount of false positives
+provided by the tool may render the tool useless. Developers said ``this could relate to information overload'', ``this would
+be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
+good the results are in regards to false positives''. These concerns combined with RQ2 suggest that pre-emptive tools are
+best used in unstable phases on the project when indirect conflicts are more likely to happen, thus reducing the number of false
+positives. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
+Interviewed developers also suggested that their proper software development process are already in place to
+catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
+or communication within the project.
+
+As per what developers would want from such pre-emptive tools, Table~\ref{tab:pre} shows how developers feel towards being 
+alerted in certain types of functional changes within code which they are using.
+
+\begin{table*}[tb!]
+\begin{center}
+\begin{tabular}{| p{7cm} | c | c | c | c | c |}
+\hline
+Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
+\hline
+\hline
+Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
+Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
+Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
+User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
+Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
+Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Results of survey questions to source code changes that developers deem notification worthy, in terms of percentage
+of developers surveyed.\label{tab:pre}}
+\end{table*}
+
 \begin{description}[topsep=6pt]
-	\item[RQ8] \textit{When are reactionary indirect conflict tools best used and what would developers want from them?}
+	\item[RQ7] \textit{When are reactionary indirect conflict tools best used and what would developers want from them?}
 \end{description}

### 2013-08-28 iced-2013/paper.tex
ICED 2014
Just added results to the last research question. Need to start the
discussion now.

@@
 catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
-or communication within the project.
+or communication within the project. Finally, most interviewed developers mentioned that their current work flow does
+not subscribe to the ideas of pre-emptive prevention and that they just ``work until something breaks''.
 
@@
 \begin{description}[topsep=6pt]
-	\item[RQ7] \textit{When are reactionary indirect conflict tools best used and what would developers want from them?}
+	\item[RQ7] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
 \end{description}
+
+The developers who were surveyed, the following list represents the more common suggestions for how an indirect conflict
+resolution tool should be implemented.
+
+\begin{itemize}
+	\item Aid in debugging indirect conflicts by finding which recent code changes are breaking a particular area of code or a test.
+	\item Automated testing to show how code changes affect old tests and the ability to automatically write new tests to compensate for the change.
+	\item IDE plug-ins to show how current changes will affect other components of the current project.
+	\item Analysis of library releases to show how upgrading to a new release will affect your project.
+	\item Built in language features to either the source code architecture (i.e. Eiffle or Java Modeling Language) or the compile time tools to display
+				warning messages towards potential issues.
+	\item A code review time tool which allows deeper analysis of a new patch or pull request to the project allowing the reviewer to see potential 
+				indirect conflicts before actually merging the code in.
+	\item A tool which is non-obtrusive and integrates into their pre-existing development styles without them having to take extra steps.
+\end{itemize}
 

### 2013-08-29 iced-2013/paper.tex
Added abstract.

@@
 
-\usepackage{enumitem}
-
-\setlist[description]{leftmargin=*}
 \floatstyle{plaintop}
@@
 \begin{abstract}
+Developer awareness techniques have been proposed and studied to aid in developer
+understanding and  efficient developer output in terms of source code produced. These techniques
+involve the sharing of information regarding source code changes in a specific project in order
+to either catch potential issues before they are fully integrated or to help developers debug
+pre-existing issues. These techniques have largely focused on either \textit{direct} or 
+\textit{indirect conflict} issues. While the techniques and tools for direct conflicts have had
+large success, tools either proposed or studied which focus on indirect conflicts have had many common
+struggles and poor developer interest. To better understand why indirect conflict tools either fail or have little
+developer investment, we performed a grounded theory study with 97 developers involved in either
+interviews or a web survey. We identified the main features that characterize indirect conflict issues,
+what processes and techniques developers use to aid in preventing and solving indirect conflict issues,
+and what developers would look for in a tool to aid them in these conflicts.
 
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ1] \textit{How do indirect conflicts occur?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ4] \textit{How do developers catch indirect conflicts?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ5] \textit{How to developers react to and solve indirect conflicts?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ6] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
@@
 
-\begin{description}[topsep=6pt]
+\begin{description}
 	\item[RQ7] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}

### 2013-08-29 iced-2013/paper.tex
Introduction added large parts.

@@
 Victoria, Canada \\ jell@uvic.ca}
-IEEEauthorblockN{Daniela Damian}
+\IEEEauthorblockN{Daniela Damian}
 \IEEEauthorblockA{University of Victoria,
@@
 \begin{abstract}
+
 Developer awareness techniques have been proposed and studied to aid in developer
@@
 
-\section{Related Work}
+As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
+development has become the standard development process instead of the exception. With this parallel development
+comes the need for larger awareness among developers to have ``an understanding of the activities of others
+which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness in theory
+mitigates the downside of parallel development which is the cost of conflict prevention and resolution, however,
+in practice we see these mitigated losses appear quite frequently. Not only do these conflicts still occur,
+but at times they can prove to be a significant and time-consuming chore for developers~\cite{Perry:2001:PCL}.
+
+Two types of conflicts have attracted the attention of academics over the years, \textit{direct} and 
+\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
+artifact, finding expert knowledge about a particular file, or knowing when another developer is working on
+the same file as you. Tools have been created and studied for these types of issues
+~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
+positive developer feedback. However, indirect conflict tools have not had the same success. Indirect conflicts
+can arise in source code for such reasons as having one's own code negatively affected by a library upgrade or
+making a code change only to find out it had negative effects on unit or integration test or  having negative effects
+on the way other developer's have interpreted the use of your code. Tools that have attempted to either
+help developers prevent or solve indirect conflicts have either attracted little developer interest or have
+had some common struggles which remain to be studied or solved.
+
+% Another paragraph about indirect conflict tools.
+
+We are interested in exploring three main areas in the hopes of resolving some of the issues that are occurring
+with indirect conflict tools: the nature of these conflicts in the how, what, and where, the processes that
+developers use prevent, catch, and solve indirect conflicts, and what developer opinions are on preventative 
+indirect conflicts tools as well as what they would look for in any tool for indirect conflicts. We broke
+these categories down into the following 8 research questions:
+
+\begin{description}
+	\item[RQ1] \textit{How do indirect conflicts occur?}
+	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
+	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
+	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
+	\item[RQ5] \textit{How do developers catch indirect conflicts?}
+	\item[RQ6] \textit{How to developers react to and solve indirect conflicts?}
+	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
+	\item[RQ8] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+\end{description}
+
+We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
+developers in order to answer these questions.
+
+% Paragraph(s) which summarizes the key result points.
+
+We continue this paper by first presenting our motivations for this research in Section~\ref{sec:mot}.
+We then present the summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
+continue with the results of our study in regards to our 8 research questions in Section~\ref{sec:results} 
+followed by a discussion in Section~\ref{sec:disc} of what was found and how it can be applied to future study and implementation
+of indirect tools or processes. We next present how evaluation was performed in terms of the credibility
+and quality of our results in Section~\ref{sec:eval}, then cover related work in Section~\ref{sec:related}
+and conclude in Section~\ref{sec:conc}.
+
+\section{Motivation}
+\label{sec:mot}
 
 \section{Methodology}
+\label{sec:meth}
 
@@
 our breadth of participants was rather large spanning multiple open and closed source project and our interview 
-answer saturation was quite high (to be seen in Section~\ref{sec:rnd}). However, we were disappointed in our word
+answer saturation was quite high (to be seen in Section~\ref{sec:results}). However, we were disappointed in our word
 of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
@@
 being apart of any one software project for extended periods of time. We account for these non answers
-in our analysis as seen in Section~\ref{sec:rnd}. Interviews lasted 
+in our analysis as seen in Section~\ref{sec:results}. Interviews lasted 
 from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
@@
 \section{Results}
-\label{sec:rnd}
+\label{sec:results}
 
@@
 \begin{description}
+	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
+\end{description}
+
+% Design by contract, add and deprecate
+
+\begin{description}
 	\item[RQ4] \textit{How do developers catch indirect conflicts?}
@@
 \section{Discussion}
+\label{sec:disc}
 
-\section{Future Work}
+\section{Evaluation}
+\label{sec:eval}
+
+\section{Related Work}
+\label{sec:related}
 
 \section{Conclusions}
+\label{sec:conc}
 
@@
 
-%\bibliographystyle{IEEEtran}
-%\balance
-%\bibliography{paper}
+\bibliographystyle{IEEEtran}
+\balance
+\bibliography{paper}
 

### 2013-08-31 iced-2013/paper.tex
Added the dicussion of topics 1 and 2.

@@
 
-We continue this paper by first presenting our motivations for this research in Section~\ref{sec:mot}.
-We then present the summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
+We continue this paper by first presenting our summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
 continue with the results of our study in regards to our 8 research questions in Section~\ref{sec:results} 
@@
 and conclude in Section~\ref{sec:conc}.
-
-\section{Motivation}
-\label{sec:mot}
 
@@
 \begin{description}
-	\item[RQ4] \textit{How do developers catch indirect conflicts?}
+	\item[RQ5] \textit{How do developers catch indirect conflicts?}
 \end{description}
@@
 \begin{description}
-	\item[RQ5] \textit{How to developers react to and solve indirect conflicts?}
+	\item[RQ6] \textit{How to developers react to and solve indirect conflicts?}
 \end{description}
@@
 \begin{description}
-	\item[RQ6] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
+	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
 \end{description}
@@
 \begin{description}
-	\item[RQ7] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+	\item[RQ8] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
 \end{description}
@@
 \label{sec:disc}
+
+We discuss the results of Section~\ref{sec:results} as follows: first we look at and discuss the nature of indirect conflicts by addressing
+research questions 1 - 3, we proceed to talk about how developers structure their work flow around indirect conflicts currently by
+examining the results of research questions 4 - 6, and we lastly investigate why developers have not had much interest in tools built previously
+for indirect conflicts and what we can do moving forward to accommodate developer work flow with potential tools by looking at
+research questions 7 and 8.
+
+\subsection{The Nature of Indirect Conflicts}
+
+In order to build solution tools for indirect conflicts, we must first understand their nature; what causes them and how often do they 
+occur. From the results we see in RQ1, a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
+fully understand the ramifications of their source code change. This was especially apparent when developers were dealing with legacy code.
+Developers identified the changing of older code to be quite prone to indirect conflicts as they were not sure where the code was being used
+anymore. We would think this can be a simple fix with static analysis tools of source
+code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
+the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
+cross language dependencies which also breaks down any potential for static analysis on these software dependencies. 
+Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
+static analysis of the changing code is not always possible. The other side of upgrading libraries also posed a problem. Developers who were often
+tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
+how they may affect these outside projects with their changes.
+
+We see from RQ3 that developers have shown that the human factor involved with indirect conflicts cannot be ignored.
+From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
+that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
+do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
+that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
+67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. So even when developers are working alone,
+they do not have enough awareness of their own code as to how a software change may affect another portion of their project. This can be due to
+any of the reasons mentioned above. The problem also seems to be compounded once multiple developers and teams are involved in parallel development
+schemes. The lack of awareness as to other developers work in a project again comes to light. This lacking is increased as project size trends up.
+It would appear possible that the addition of software developers to a project, which seems to compound the likely hood of indirect
+conflicts, developers would be able to leverage other developer expertise of the project to help avoid or solve indirect conflicts.
+However, we will see in Section~\ref{sec:pd} that most developers do not leverage their social connections within a project to help
+solve indirect conflicts.
+
+From RQ2, we can make the deduction that indirect conflicts are more likely to occur early in a projects development life cycle as opposed
+to later. We used to notion of the first major release to make the distinction between early and late in a project but the same hypothesis
+could be applied to iterative development where during the earlier stages of the current phase could be considered more probable for indirect
+conflict to occur. In Table~\ref{tab:often}, we see that as the project becomes more mature. we see a drop off in the frequency of indirect
+conflicts. Developers surveyed answered that daily and weekly indirect conflict frequencies trail off past the project's first release and 
+those developers interviewed mentioned a point in the project, the ``mature point'', in which indirect conflicts become less frequent after.
+This would suggest that, as will be seen in Section~\ref{sec:td}, preemptive indirect conflict tools would be more useful before this point as the
+number of false positives will have a higher chance of being true positives. This ``mature point'' may be cause for one of two reasons. The
+first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
+of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
+development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
+would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
+to indirect conflicts. This two possibilities are not mutually exclusive. In the end, developers survey pointed out that before the ``mature
+point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
+said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).
+
+From the results of RQ3 we can also see evidence that the problemed areas of software architecture in regards to indirect conflicts are
+the interfaces between structures. As was seen, developers listed both methods/functions and database schema's as the largest contributors
+to indirect conflict structures. Since methods are often the interface which developers use most, it seems natural that they are identified
+as the main software structure in regards to indirect conflicts. It appears that when interfaces of methods change through natural software
+evolution, indirect conflicts are more prone to occurring. This corresponds well with where most academic research has taken place with indirect
+conflicts. Most tools built so far have focused on the method level, and those that attempt to prevent indirect conflicts often look for a change
+in interface to a method, either in signature or in internal body composition. The more interesting result was that of database schema based 
+indirect conflicts. There has been little to no focus of indirect conflicts at the schema level, possibly due to the inherit difficult nature
+of analysis a database schema in relation to source code structure. Project often use unique libraries and procedures when dealing with
+database operations which causes a generic tool for this analysis to be difficult at best to implement. However, developers have noted that this
+is often a very problemed area and can affect large portions of the software project.
+
+We theorize the nature of indirect conflicts with the following findings: 
+\begin{description}
+	\item[F1] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
+						by the natural evolution of software. This lack of awareness and understanding is compounded by having a multi developer
+						or multi team environment.
+	\item[F2] Indirect conflicts occur more often before a project reaches its ``mature point'' rather than after for each development
+						cycle.
+	\item[F3] Indirect conflicts are brought on by the changing of interfaces to software structures. The most common type of structures
+						to cause indirect conflicts are methods/functions and database schemas.
+\end{description}
+
+\subsection{Processes of Indirect Conflicts}
+\label{sec:pd}
+
+To be able to accommodate developers in the prevention and solving of indirect conflicts, it is important to understand the current
+work flows of developers in regards to how they prevent, catch, and solve indirect conflicts. RQ4 shows us that most developers rely on
+two items in order to prevent indirect conflicts. The first of these items is the development process known as add and deprecate. This
+is the process of instead of changing old code, the code is simply copied into a new method / function and then modified to make a new 
+version while the old version gets deprecated and phased out later in the project's life. Most project's which are past their ``mature point''
+use this development process as their number one way of preventing indirect conflicts. The second item is to have highly experienced team
+members be involved in the project planning in order to foresee where problems may arise and act accordingly. (Developers also mentioned
+their personal experience in preventing indirect conflicts.) These two tools are really the only measures that developers have right now
+in order to prevent indirect conflicts. This is probably due to the inherit difficulties in predicting indirect conflicts, as will be seen in
+Section~\ref{td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and conflicts 
+continue to occur. However, developers did not seem overly concerned that these two items do not totally prevent indirect conflicts. This
+is because most developers ascribe to the idea of ``I work until something breaks''. This shows that developers are somewhat content with
+the level of prevention that happens for indirect conflicts and would explain why preemptive indirect conflict tools have failed to catch on,
+as will be seen in Section~\ref{sec:td}. This again is probably due to the difficult nature of predicting indirect conflicts. RQ4 has mostly
+shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
+more effectively spent developing and fixing problems as needed than putting excess time into the prevention of potential errors.
+
+The results of RQ5 show us how developers currently catch indirect conflicts. It appears that most indirect conflicts are currently
+detected by using what some would call proper software practices. The use of unit and integration testing, and the use of nightly or weekly
+project builds are the main contributing factors of such discovery. The words ``use case coverage'' were constantly being used by developers
+when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
+not all indirect conflicts should be caught. This leads us to believe that as long as the contract of a software structure is understood,
+proper tests should be able to be written which will catch most if not all indirect conflicts. The problems occur when the contract is
+not properly understood and use cases are not tested for or examined as explained in previous discussion. Once an object is not properly
+tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. Aside from testing
+was the build process. Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
+the build process more to catch compile time errors witch prevent the product from being used. These types of build breaking indirect
+conflicts do occur but for the most part developers posses tools capable to build their solutions in order to catch these problems before
+the code change is added to the project. Therefore, unit and integration testing appears to be the largest factor in catching indirect conflicts.
+
+Finally, once the indirect conflict is caught, the results of RQ6 show us how developers tend to solve the issues. Here, the generic approach
+brought forth by developers was that of history checking. Developers would look into source code change logs and build or test failure
+logs in order to see what is broken and what might have changed recently that could be affecting the break. This gives the impression that
+developers already have the tools to solve indirect conflicts. However, this leads to the question: how quickly and
+efficiently are they able to solve them?
+While each developer gave some examples on solving indirect conflicts and the times needed, they all agreed that to be able to solve them
+quicker would be a huge benefit. A possible solutions as to solving indirect conflicts quicker are the automatic analysis of source code
+change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
+by several researchers in both delta debugging and code slicing. The interest that developer's have in fixing issues as they
+arise more quickly again shows that developer's prefer fixing real problems directly as opposed to trying to prevent potential problems.
+Their time is more effectively spent developing than speculating.
+
+We sum up the developer process with indirect conflicts with the following findings:
+\begin{description}
+	\item[F4] Developers use a combination of the add and deprecate model as well as expert knowledge in order to prevent indirect conflicts.
+	\item[F5] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
+	\item[F6] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
+\end{description}
+
+\subsection{Tools for indirect Conflicts}
+\label{sec:td}
 

### 2013-08-31 iced-2013/paper.tex
Added the discussion of topic 3.

@@
 not properly understood and use cases are not tested for or examined as explained in previous discussion. Once an object is not properly
-tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. Aside from testing
+tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. Some languages
+such as Eiffle and the Java Modeling Language enforce object contracts to be well understood through the ideas of pre and post conditions
+of structures. Aside from testing
 was the build process. Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
@@
 
+Ultimately, the goal of this research would be to produce any sort of tool that could aid developers in any stage (prevention, detection, or
+solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
+questions along with research questions 7 and 8. We first examined the prevention / detection side of indirect conflicts through the use of
+preemptive indirect conflict tools. Preemptive indirect conflict tools are used to find potential code changes that could result in indirect
+conflicts. As was found through RQ7, developers major concern with preemptive indirect conflict tools was the number of false positives put
+forth by such a tool and the information overload that was sure to follow. Developers were very quick to identify this problem which has
+been a common occurrence among any preemptive indirect conflict tool built. The main issue with false positives is the detection of changes
+which frequently cause indirect conflicts. Change types are very possible to detect through software static analysis, but even those that are
+prone to causing indirect conflicts do not cause them all the time. This is quite evident in Table~\ref{tab:pre} where developers identified
+that there is an even distribution of software changes that could cause indirect conflicts and as to if they would want to be made aware of
+those changes. The only two exceptions which stand out are that of method signature changes and main algorithm changes. The problem here however,
+is that method signature changes will be caught by IDE errors, nightly builds, unit testing, integration testing and so on, while main algorithm
+changes are very difficult to detect. This all leads to the hypothesis that preemptive indirect conflict tools cannot be built without a very
+large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
+changed pre-conditions in methods most times if not always, there is still a large chance that that changed pre-condition will actually not
+introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
+overload, and eventually to developers abandoning the tool.
+
+In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
+have an automatic understanding of how the project's code is suppose to work. From the halting problem, we know it is a rather difficult
+if not impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
+human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
+of the use cases of a particular code artifact, then they can write proper tests to ensure those use cases are always met and not broken which
+would be the best form of preventing and catching indirect conflicts.
+
+Given the above challenges, some indirect conflict tools have had some general success. The key difference with these tools is that they
+contain the idea of developer exploration. The idea of only presenting information to users on request is the complete opposite of the preemptive
+notion of notifying developers to potential upcoming problems. The tools which provide information on user request usually involve taking some
+artifact of interest and presenting information which is dependent on the artifact. Here, the idea that a user can provide some sort of
+input, some sort of context, is a powerful tool for reducing false positives. We have really just moved the false positives of the preemptive
+tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
+precision of results. These types of tools are in direct comparison with the results of RQ8 in that developers were more interested in
+the debugging of issues quickly and being able to narrow down where problems emanate from. This also sticks with the findings found in
+Section~\ref{sec:pd} being that developers would rather spend their time fixing and exploring a real issue when it does occur than trying
+to prevent potential issues moving forward. As researchers, we should be focusing our efforts on the solving as opposed to prevention and 
+the detection of indirect conflicts. Prevention does not appear to worth developer time as indirect conflicts and software failures will always
+occur, detection will always require some form of human intervention to determine the correctness of a program, so this only leaves us with
+the time and effort spent on debugging.
+
+We sum up our conclusions for future indirect conflict tools with the following findings:
+\begin{description}
+	\item[F7] It is extremely difficult, if not impossible, to create a preemptive indirect conflict tool which can detect indirect conflicts
+						with a low rate of false positives.
+	\item[F8] Indirect conflict tools are best used by providing exploratory environments to aid in the solving of indirect conflicts.
+	\item[F9] Indirect conflict tools should be reducing debugging time of indirect conflicts as opposed to preventative
+						time in order to better fit developer activities.
+\end{description}
+
 \section{Evaluation}

### 2013-09-03 iced-2013/paper.tex
Finished intro, added eval, and conclusions. Need related work.

@@
 
+\usepackage{hyperref}
+\usepackage{nameref}
+
 \floatstyle{plaintop}
@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\makeatletter
+\def\namedlabel#1#2{\begingroup
+   \def\@currentlabel{#2}%
+   \label{#1}\endgroup
+}
+\makeatother
 
@@
 
-% Another paragraph about indirect conflict tools.
+The inherent difficulty regarding indirect conflicts is one, to determine which software dependencies are
+causing indirect conflicts, and two, to be able to analyze these dependencies, specifically as to when they change
+and whether or not they will induce an indirect conflict or otherwise negative consequences of a software project.
+Sarma et al.~\cite{Sarma:2007:TSA} have devised a tool (Palantir) which can both detect potential indirect conflicts and alert
+developers to these conflicts arising in their software projects at the class signature level. Ariadne~\cite{Trainer:2005:BGT}, another indirect 
+conflict tool, creates an exploratory environment where developers can see what areas of a project they are affecting with
+source code changes at the method level and thus where indirect conflicts may occur. 
+Another tool developer for indirect conflicts is CASI~\cite{Servant:2010:CPI}
+which utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of the more traditional method call graphs to provide 
+again an environment to see what areas of a project are being affected by a source code change. 
+These tools have all shown to have the same common difficulty of scalability in the face of information overload. Too many
+dependencies with the potential for indirect conflicts are being reported to developers to the point where the number of false
+positives reported becomes unbearable. Our own unpublished tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered this same fate.
+
+Many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through the means of other 
+developers~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. Common issues arise here in that developers do not always
+utilize others to solve issues, socio-technical congruence is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problem with scalability in that many many developers may be linked to
+anyone particular persons work. Again the notion of confirmation overload becomes apparent and false positives become a large
+concern.
 
@@
 \begin{description}
-	\item[RQ1] \textit{How do indirect conflicts occur?}
-	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
-	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
-	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
-	\item[RQ5] \textit{How do developers catch indirect conflicts?}
-	\item[RQ6] \textit{How to developers react to and solve indirect conflicts?}
-	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
-	\item[RQ8] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{How do indirect conflicts occur?}
+	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
+	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
+	\item[RQ4\namedlabel{itm:rq4}{RQ4}] \textit{How do developer prevent indirect conflicts?}
+	\item[RQ5\namedlabel{itm:rq5}{RQ5}] \textit{How do developers catch indirect conflicts?}
+	\item[RQ6\namedlabel{itm:rq6}{RQ6}] \textit{How to developers react to and solve indirect conflicts?}
+	\item[RQ7\namedlabel{itm:rq7}{RQ7}] \textit{When are preemptive indirect conflict tools best used and what would developers want from them?}
+	\item[RQ8\namedlabel{itm:rq8}{RQ8}] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
 \end{description}
@@
 We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
-developers in order to answer these questions.
-
-% Paragraph(s) which summarizes the key result points.
+developers in order to answer these questions. Our findings indicate that while indirect conflicts are very much a real problem
+in software development, developers already poses the tools to prevent and catch indirect conflicts to a certain manageable degree and have
+the mindset of working until something breaks, then attempting a fix. Thus, developers would rather spend their time developing instead
+of taking or reading preventative measures, resulting in the need of tools to debug and solve indirect conflicts quickly
+(reactionary tools) as opposed to tools which attempt to prevent or catch indirect conflicts early (preemptive tools).
 
@@
 \begin{description}
-	\item[RQ1] \textit{How do indirect conflicts occur?}
+	\item[~\ref{itm:rq1}] \textit{How do indirect conflicts occur?}
 \end{description}
@@
 \begin{description}
-	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
+	\item[~\ref{itm:rq2}] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
 \end{description}
@@
 \begin{description}
-	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
+	\item[~\ref{itm:rq3}] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
 \end{description}
@@
 \begin{description}
-	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
+	\item[~\ref{itm:rq4}] \textit{How do developer prevent indirect conflicts?}
 \end{description}
@@
 \begin{description}
-	\item[RQ5] \textit{How do developers catch indirect conflicts?}
+	\item[~\ref{itm:rq5}] \textit{How do developers catch indirect conflicts?}
 \end{description}
@@
 Of the 78 developers survey, 49\% mentioned forms of testing as the major tool used to catch indirect conflicts, 33\% said build processes,
-while 31\% used work their IDE or IDE plugins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
+while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
 and project expertise (6\%) as factors of catching indirect conflicts.
@@
 \begin{description}
-	\item[RQ6] \textit{How to developers react to and solve indirect conflicts?}
+	\item[~\ref{itm:rq6}] \textit{How to developers react to and solve indirect conflicts?}
 \end{description}
@@
 \begin{description}
-	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
+	\item[~\ref{itm:rq7}] \textit{When are preemptive indirect conflict tools best used and what would developers want from them?}
 \end{description}
 
-The developers who were interviewed had a major concern with any pre-emptive tool in that the amount of false positives
+The developers who were interviewed had a major concern with any preemptive tool in that the amount of false positives
 provided by the tool may render the tool useless. Developers said ``this could relate to information overload'', ``this would
 be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
-good the results are in regards to false positives''. These concerns combined with RQ2 suggest that pre-emptive tools are
+good the results are in regards to false positives''. These concerns combined with RQ2 suggest that preemptive tools are
 best used in unstable phases on the project when indirect conflicts are more likely to happen, thus reducing the number of false
@@
 or communication within the project. Finally, most interviewed developers mentioned that their current work flow does
-not subscribe to the ideas of pre-emptive prevention and that they just ``work until something breaks''.
+not subscribe to the ideas of preemptive prevention and that they just ``work until something breaks''.
 
-As per what developers would want from such pre-emptive tools, Table~\ref{tab:pre} shows how developers feel towards being 
+As per what developers would want from such preemptive tools, Table~\ref{tab:pre} shows how developers feel towards being 
 alerted in certain types of functional changes within code which they are using.
@@
 \begin{description}
-	\item[RQ8] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+	\item[~\ref{itm:rq8}] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
 \end{description}
@@
 				indirect conflicts before actually merging the code in.
-	\item A tool which is non-obtrusive and integrates into their pre-existing development styles without them having to take extra steps.
+	\item A tool which is non-obtrusive and integrates into their preexisting development styles without them having to take extra steps.
 \end{itemize}
@@
 
-We discuss the results of Section~\ref{sec:results} as follows: first we look at and discuss the nature of indirect conflicts by addressing
-research questions 1 - 3, we proceed to talk about how developers structure their work flow around indirect conflicts currently by
-examining the results of research questions 4 - 6, and we lastly investigate why developers have not had much interest in tools built previously
-for indirect conflicts and what we can do moving forward to accommodate developer work flow with potential tools by looking at
-research questions 7 and 8.
+We discuss the results of Section~\ref{sec:results} as follows: first we look at and discuss the nature of indirect conflicts by addressing~\ref{itm:rq1} 
+-~\ref{itm:rq3}, we proceed to talk about how developers structure their work flow around indirect conflicts currently by
+examining the results of~\ref{itm:rq4} -~\ref{itm:rq6}, and we lastly investigate why developers have not had much interest in tools built previously
+for indirect conflicts and what we can do moving forward to accommodate developer work flow with potential tools by looking at~\ref{itm:rq7} 
+and~\ref{itm:rq8}.
 
@@
 In order to build solution tools for indirect conflicts, we must first understand their nature; what causes them and how often do they 
-occur. From the results we see in RQ1, a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
+occur. From the results we see in~\ref{itm:rq1}, a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
 fully understand the ramifications of their source code change. This was especially apparent when developers were dealing with legacy code.
@@
 
-We see from RQ3 that developers have shown that the human factor involved with indirect conflicts cannot be ignored.
+We see from~\ref{itm:rq3} that developers have shown that the human factor involved with indirect conflicts cannot be ignored.
 From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
@@
 
-From RQ2, we can make the deduction that indirect conflicts are more likely to occur early in a projects development life cycle as opposed
+From ~\ref{itm:rq2}, we can make the deduction that indirect conflicts are more likely to occur early in a projects development life cycle as opposed
 to later. We used to notion of the first major release to make the distinction between early and late in a project but the same hypothesis
@@
 
-From the results of RQ3 we can also see evidence that the problemed areas of software architecture in regards to indirect conflicts are
+From the results of~\ref{itm:rq3} we can also see evidence that the problemed areas of software architecture in regards to indirect conflicts are
 the interfaces between structures. As was seen, developers listed both methods/functions and database schema's as the largest contributors
@@
 \begin{description}
-	\item[F1] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
+	\item[F1\namedlabel{itm:f1}{F1}] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
 						by the natural evolution of software. This lack of awareness and understanding is compounded by having a multi developer
 						or multi team environment.
-	\item[F2] Indirect conflicts occur more often before a project reaches its ``mature point'' rather than after for each development
+	\item[F2\namedlabel{itm:f2}{F2}] Indirect conflicts occur more often before a project reaches its ``mature point'' rather than after for each development
 						cycle.
-	\item[F3] Indirect conflicts are brought on by the changing of interfaces to software structures. The most common type of structures
+	\item[F3\namedlabel{itm:f3}{F3}] Indirect conflicts are brought on by the changing of interfaces to software structures. The most common type of structures
 						to cause indirect conflicts are methods/functions and database schemas.
@@
 To be able to accommodate developers in the prevention and solving of indirect conflicts, it is important to understand the current
-work flows of developers in regards to how they prevent, catch, and solve indirect conflicts. RQ4 shows us that most developers rely on
+work flows of developers in regards to how they prevent, catch, and solve indirect conflicts. ~\ref{itm:rq4} shows us that most developers rely on
 two items in order to prevent indirect conflicts. The first of these items is the development process known as add and deprecate. This
-is the process of instead of changing old code, the code is simply copied into a new method / function and then modified to make a new 
+is the process of instead of changing old code, the code is simply copied into a new method/function and then modified to make a new 
 version while the old version gets deprecated and phased out later in the project's life. Most project's which are past their ``mature point''
@@
 in order to prevent indirect conflicts. This is probably due to the inherit difficulties in predicting indirect conflicts, as will be seen in
-Section~\ref{td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and conflicts 
+Section~\ref{sec:td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and conflicts 
 continue to occur. However, developers did not seem overly concerned that these two items do not totally prevent indirect conflicts. This
@@
 the level of prevention that happens for indirect conflicts and would explain why preemptive indirect conflict tools have failed to catch on,
-as will be seen in Section~\ref{sec:td}. This again is probably due to the difficult nature of predicting indirect conflicts. RQ4 has mostly
+as will be seen in Section~\ref{sec:td}. This again is probably due to the difficult nature of predicting indirect conflicts. ~\ref{itm:rq4} has mostly
 shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
@@
 
-The results of RQ5 show us how developers currently catch indirect conflicts. It appears that most indirect conflicts are currently
+The results of~\ref{itm:rq5} show us how developers currently catch indirect conflicts. It appears that most indirect conflicts are currently
 detected by using what some would call proper software practices. The use of unit and integration testing, and the use of nightly or weekly
@@
 
-Finally, once the indirect conflict is caught, the results of RQ6 show us how developers tend to solve the issues. Here, the generic approach
+Finally, once the indirect conflict is caught, the results of~\ref{itm:rq6} show us how developers tend to solve the issues. Here, the generic approach
 brought forth by developers was that of history checking. Developers would look into source code change logs and build or test failure
@@
 \begin{description}
-	\item[F4] Developers use a combination of the add and deprecate model as well as expert knowledge in order to prevent indirect conflicts.
-	\item[F5] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
-	\item[F6] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
+	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of the add and deprecate model as well as expert 
+																	 knowledge in order to prevent indirect conflicts.
+	\item[F5\namedlabel{itm:f5}{F5}] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
+	\item[F6\namedlabel{itm:f6}{F6}] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
 \end{description}
@@
 solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
-questions along with research questions 7 and 8. We first examined the prevention / detection side of indirect conflicts through the use of
+questions along with~\ref{itm:rq7} and~\ref{itm:rq8}. We first examined the prevention / detection side of indirect conflicts through the use of
 preemptive indirect conflict tools. Preemptive indirect conflict tools are used to find potential code changes that could result in indirect
-conflicts. As was found through RQ7, developers major concern with preemptive indirect conflict tools was the number of false positives put
+conflicts. As was found through~\ref{itm:rq7}, developers major concern with preemptive indirect conflict tools was the number of false positives put
 forth by such a tool and the information overload that was sure to follow. Developers were very quick to identify this problem which has
@@
 large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
-changed pre-conditions in methods most times if not always, there is still a large chance that that changed pre-condition will actually not
+changed preconditions in methods most times if not always, there is still a large chance that that changed precondition will actually not
 introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
@@
 tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
-precision of results. These types of tools are in direct comparison with the results of RQ8 in that developers were more interested in
+precision of results. These types of tools are in direct comparison with the results of~\ref{itm:rq8} in that developers were more interested in
 the debugging of issues quickly and being able to narrow down where problems emanate from. This also sticks with the findings found in
@@
 \begin{description}
-	\item[F7] It is extremely difficult, if not impossible, to create a preemptive indirect conflict tool which can detect indirect conflicts
-						with a low rate of false positives.
-	\item[F8] Indirect conflict tools are best used by providing exploratory environments to aid in the solving of indirect conflicts.
-	\item[F9] Indirect conflict tools should be reducing debugging time of indirect conflicts as opposed to preventative
+	\item[F7\namedlabel{itm:f7}{F7}] It is extremely difficult, if not impossible, to create a preemptive indirect 
+																	 conflict tool which can detect indirect conflicts with a low rate of false positives.
+	\item[F8\namedlabel{itm:f8}{F8}] Indirect conflict tools are best used by providing exploratory environments to aid in the solving of indirect conflicts.
+	\item[F9\namedlabel{itm:f9}{F9}] Indirect conflict tools should be reducing debugging time of indirect conflicts as opposed to preventative
 																	 time in order to better fit developer activities.
@@
 
+As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.
+We have chosen three of these criteria and explain how we fulfill them.
+
+{\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
+criterion is used to verify the correctness of our finding and to ensure they resonate and fit with participant opinion. It is also
+required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
+after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
+on and potentially agree with what the majority has been on the topic. Often the developers own opinions already matched those of
+the majority before them and did not necessarily have to directly verify it themselves.
+
+To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:disc} to either a majority of agreeing
+responses on a topic or to a large amount of direct quotes presented by participants.
+
+{\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
+Although our findings~\ref{itm:f1} -~\ref{itm:f6} may not be entirely novel or even surprising, the combination of these results allow us to discover the
+insightful findings of~\ref{itm:f7} -~\ref{itm:f9} regarding indirect conflict tools. Given how many indirect conflict tools are left with the same common
+issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
+been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.
+
+10 of the 78 participants who were surveyed sent direct responses to us asking for any results of the research to be sent directly to
+them in order to improve their indirect conflict work flows. 7 of the 19 participants survey expressed interest concerning any possible
+tools or plans for tools to come out of this research as well. The combination of academia relatability and direct industry interest 
+in our results help us fulfill this criterion.
+
+{\bfseries Variation.} ``Has variation been built into the findings?'' Variation shows that an event is complex and that any findings
+made accurately demonstrate that complexity. Since those participants interviewed came from such a diverse set of organizations, 
+software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
+expressed unique situations that did not fully meet our generalized findings or on going theories. In these cases, we worked in the specific
+cases which were presented as boundary cases and can be seen in quotations in Section~\ref{sec:results}. These quotations add to the variation
+to show how the complexity of the situation also resides in a significant number of unique boundary situations as well as the complexity
+in the generalized theories and findings.
+
 \section{Related Work}
@@
 \label{sec:conc}
+
+Indirect conflicts are a significant issue with real world developers, however, many proposed techniques and tools to mitigate
+losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
+19 interviewed developers from 12 organizations as well as 78 surveyed developers, we have provided characterization of indirect conflicts,
+current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
+in their work flow with indirect conflicts.
+
+We have shown through findings~\ref{itm:f1} -~\ref{itm:f3} why indirect conflicts occur, when indirect conflicts are more likely to
+occur, as well as what types of software objects are susceptible to these conflicts. Findings~\ref{itm:f4} -~\ref{itm:f6} have shown
+how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings~\ref{itm:f7} 
+-~\ref{itm:f9} have provided a foundation as to why past techniques and tools have had low adoption rates and where
+researchers should focus their current and future efforts in handling indirect conflicts. We hope that this study and its 
+findings will inspire future techniques and tools for dealing with indirect conflicts that 
+will aid developers in industry as well as test and validate our theories put forth in this paper.
 

### 2013-09-04 iced-2013/paper.tex
Added grounded theory and evaluation.

@@
 
+\newcommand{\totalCategories}{10}
+
 % Begin the paper here
@@
 
-\author{\IEEEauthorblockN{Jordan Ell}
+\author{\IEEEauthorblockN{Jordan Ell and Daniela Damian}
 \IEEEauthorblockA{University of Victoria,
-Victoria, Canada \\ jell@uvic.ca}
-\IEEEauthorblockN{Daniela Damian}
-\IEEEauthorblockA{University of Victoria,
-Victoria, Canada \\ danielad@csc.uvic.ca}
+Victoria, Canada \\ jell@uvic.ca, danielad@csc.uvic.ca}
 }
@@
 
-Our study was performed in two parts. First, a round of semi-structured interviews were conducted which 
+Our grounded theory study was performed in two parts. First, a round of semi-structured interviews were conducted which 
 addressed the larger who, what, when, where, and how about indirect conflicts. Secondly, a survey was given
-out which was used to confirm what was learned from the interviews on a larger sample size as well as get
-a larger opinion base on general issues.
+out which was used to confirm what was theorized from the interviews on a larger sample size as well as obtain
+larger developer opinion.
+
+\subsection{Grounded Theory}
+We approached our study based on grounded theory as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
+Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
+\textit{open code} to create a social theory ``grounded'' in the empirical data. For an exploratory study such as
+the ours, grounded theory is well suited because involves starting from very broad and abstract type questions and
+making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
+regining the sampling criteria throughout the course of the study to ensure that participants are able to answer new
+questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
+both interviews and surveys (when open ended questions were involved) was analyzed using open coding. Open coding involves
+assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
+codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
+link our defined codes to categories of grounded theory such as context and consequences. The ultimate goal of grounded 
+theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in
+findings~\ref{itm:f1} -~\ref{itm:f9}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+3 criteria that are commonly used in evaluating grounded theory studies.
 
@@
 our breadth of participants was rather large spanning multiple open and closed source project and our interview 
-answer saturation was quite high (to be seen in Section~\ref{sec:results}). However, we were disappointed in our word
-of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
-their respective organizations but did not receive any additional participants through this method.
+answer saturation~\cite{Guest:2006:SAP} was quite high (to be seen in Section~\ref{sec:results}).
 
 Interview participants first answered a number of structured demographic items. Next, participants were 
-asked to describe various software development processes they use or encounter in their current working
-environment. These structured process questions were used for later analysis to provided context to
-any answer for questions that involve indirect conflicts. They next were engaged in a
-semi-structured round of questions pertaining to 12 categories of questions.
+asked to describe various software development experiences regarding our three main areas of focus: the nature
+of indirect conflicts, current processes used regarding indirect conflicts, opinions on future and the use of
+current tools. processes they use or encounter in their current working environment. These three areas of focus
+were studied by having the interview participants talk about their experiences and opinions in the following
+\totalCategories{} semi-structured research topics which were asked of:
 
@@
 
-Each of the categories was used as a
-point of aggregation for later analysis as a way of answering the X research questions laid out in 
-Section~\ref{sec:intro}. While each of the 12 question categories had a number of starter questions, interviews 
+While each of the \totalCategories{} question categories had a number of starter questions, interviews 
 largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
@@
 being apart of any one software project for extended periods of time. We account for these non answers
-in our analysis as seen in Section~\ref{sec:results}. Interviews lasted 
+in our analysis and results as seen in Section~\ref{sec:results}. Interviews lasted 
 from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
@@
 
-After the interviews, we extracted a list of answers, opinions, and experiences for each of the 12 categorical
-question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
-with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
-Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
-aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
-experience and opinion for the 12 categories. These hypotheses were then used as a template for creating the
-second step survey, in order to further test the answers given by interview participants and validate any possible
-conclusions to be drawn.
+%After the interviews, we extracted a list of answers, opinions, and experiences for each of the \totalCategories{} categorical
+%question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
+%with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
+%Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
+%aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
+%experience and opinion for the \totalCategories{} categories. These hypotheses were then used as a template for creating the
+%second step survey, in order to further test the answers given by interview participants and validate any possible
+%conclusions to be drawn.
 
@@
 that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
-Subnet Solutions, Ruboss Technology Corporation, Amazon, Kano Apps, Fireworks Design, and Frost Tree Games.
+Subnet Solutions, and Amazon.
 However, participants who took part in the round of interviews were not asked to participate in the surveys but
@@
 participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
-users were selected based on their large amounts of development activity on GitHub and the Apache developers
+users were selected based on large amounts of development activity on GitHub and the Apache developers
 were selected based on their software development participation on specific projects known to be used heavily
-by other organizations and projects.
+utilized by other organizations and projects.
 
@@
 
-The created survey was based off of the 12 categorical hypotheses given by the round of interviews. The survey
+The created survey was based off of the \totalCategories{} categorical hypotheses given by the round of interviews. The survey
 was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
@@
 expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
-in a more polished survey. The final survey consisted of the following questions:
+in a more polished survey. This also allowed the previously interviewed developers to examine what hypotheses were
+formed and we would be moving forward with. The final survey consisted of the following:
 2 multiple choice questions for developer experience and demographic information; 3 level of agreement questions for an indication of
@@
 
-After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
-as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
-answer question fields while combining commonly mentioned answers where possible. The results again from this
-preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.
+%After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
+%as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
+%answer question fields while combining commonly mentioned answers where possible. The results again from this
+%preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.
+
+\subsection{Evaluation}
+
+Following our data collection and analysis, we re-interviewed some of our initial interview participants
+in order to validate our findings. We confirmed our hypotheses as to whether or not they resonate with 
+industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
+applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
+five participants. Those that were re-interviewed came from the range of 5-10 years of software development
+experience. Re-interviewed participants were given our 8 research questions along with results and 9 findings
+and asked open ended questions regarding their opinions and experiences to validate our findings. The five
+participants found FILL IN HERE WHEN COMPLETE.
 
@@
 related research questions in regards to how developers react to, processes to prevent, and reactionary steps
-to indirect conflicts. Lastly, we present and discuss findings for what potential tools are wanted by developers
-and what researchers are currently developing.
+to indirect conflicts.
 

### 2013-09-09 iced-2013/paper.tex

@@
 From the interviewed participants, 63\% of developers believe that a large contributing factor to the cause
-of indirect conflicts comes from the changing of a software object's contract. One developer stated that ``design
+of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
+what a software object guarantees, meaning how the input, output, or how any aspect of the object is garuntee
+to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\textregistered. 
+One developer stated that ``design
 by contract was invented to solve this problem and it does it quite well'', while another noted that software 
@@
 tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. Some languages
-such as Eiffle and the Java Modeling Language enforce object contracts to be well understood through the ideas of pre and post conditions
-of structures. Aside from testing
+such as Eiffle and the Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}
+enforce object contracts to be well understood through the ideas of pre and post conditions of structures. Aside from testing
 was the build process. Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
@@
 
+% insert here
+
 \section{Conclusions}

### 2013-09-10 iced-2013/paper.tex
iced 2014 draft one done.

@@
 of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
-what a software object guarantees, meaning how the input, output, or how any aspect of the object is garuntee
-to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\textregistered. 
-One developer stated that ``design
-by contract was invented to solve this problem and it does it quite well'', while another noted that software 
-object contracts do solve the problem in theory, but that doesn't mean that problems don't occur in practice.
+what a software object guarantees, meaning how the input, output, or how any aspect of the object is garunteed
+to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
 In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
-towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples had to do 
+towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples dealt
 with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
@@
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
-changes a lot, especially with agile''. 21\% if developers interviewed mentioned that the development process
-changes once a stable point of a project is met and the ``add and deprecate model solves a lot of these issues''.
+changes a lot, especially with agile''.
 
@@
 
-% Design by contract, add and deprecate
+Of the developers interviewed, 3 main points of interest were found in preventing indirect conflicts. The first was
+the use of ``Design by Contract''\texttrademark, meaning that software objects have certain guarantees about their behavior
+in terms of input, output, and internal workings. This methodology can be used as a type of documentation towards awareness of
+a software object. One developer stated that ``design by contract was invented to solve this problem and it does it quite well'', 
+while another noted that software  object contracts do solve the problem in theory, but that doesn't mean that problems don't 
+still occur in practice. The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
+that the development process changes once a stable point of a project is met and the ``add and deprecate model [prevents] a lot 
+of these issues''. Add and deprecate meaning instead of editing code, the developer simply clones code (if needed) and edits the clone
+while slowly phasing out the out code in subsequent releases or other development cycle needs. This allows a stoppage of indirect conflicts
+by allowing developers to use the older versions of software objects which remain unchanged. Lastly, pure developer experience was mentioned
+as a source to prevent indirect conflicts. 37\% of developers mentioned that when planning code changes, either a very experienced member
+of the project was involved in the planning and has duties to foresee any indirect conflicts that may arise, or that while implementing
+a code change developers must use their personal knowledge to predict where indirect conflicts will occur.
+
+Of the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
+total), 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of ``Design
+by Contract'' or the testing of a methods contract (which could be viewed as catching indirect conflicts), and 14\% said that add and
+deprecate was used in their projects to avoid indirect conflicts.
 
@@
 could be applied to iterative development where during the earlier stages of the current phase could be considered more probable for indirect
-conflict to occur. In Table~\ref{tab:often}, we see that as the project becomes more mature. we see a drop off in the frequency of indirect
-conflicts. Developers surveyed answered that daily and weekly indirect conflict frequencies trail off past the project's first release and 
+conflict to occur. In Table~\ref{tab:often}, we see that as the project becomes more mature, a drop off in the frequency of indirect
+conflicts occurs. Developers surveyed answered that daily and weekly indirect conflict frequencies trail off past the project's first release and 
 those developers interviewed mentioned a point in the project, the ``mature point'', in which indirect conflicts become less frequent after.
@@
 as the main software structure in regards to indirect conflicts. It appears that when interfaces of methods change through natural software
-evolution, indirect conflicts are more prone to occurring. This corresponds well with where most academic research has taken place with indirect
-conflicts. Most tools built so far have focused on the method level, and those that attempt to prevent indirect conflicts often look for a change
+evolution, indirect conflicts are more prone to occurring.
+Most tools built so far have focused on the method level, and those that attempt to prevent indirect conflicts often look for a change
 in interface to a method, either in signature or in internal body composition. The more interesting result was that of database schema based 
-indirect conflicts. There has been little to no focus of indirect conflicts at the schema level, possibly due to the inherit difficult nature
+indirect conflicts. There has been little focus of indirect conflicts at the relational database schema level~\cite{Maule:2008:IAD}, 
+possibly due to the inherit difficult nature
 of analysis a database schema in relation to source code structure. Project often use unique libraries and procedures when dealing with
@@
 work flows of developers in regards to how they prevent, catch, and solve indirect conflicts. ~\ref{itm:rq4} shows us that most developers rely on
-two items in order to prevent indirect conflicts. The first of these items is the development process known as add and deprecate. This
-is the process of instead of changing old code, the code is simply copied into a new method/function and then modified to make a new 
-version while the old version gets deprecated and phased out later in the project's life. Most project's which are past their ``mature point''
-use this development process as their number one way of preventing indirect conflicts. The second item is to have highly experienced team
+two items in order to prevent indirect conflicts. The first of these items are development processes known as
+``Design by Contract''\texttrademark and the add and deprecate methodology. These processes are implementation of software artifacts based
+on what guarantees should be made about their behavior, and instead of changing old code, the code is simply copied into a new method/function and then modified to make a new version while the old version gets deprecated and phased out later in the project's life, respectively.
+``Design by Contract''\texttrademark ensures that software artifacts, while they can be changed, should always behave in a way that is understood
+by other developers (the guaranteed aspects of the artifact), and should be documented in a way to reflect this behavior in order to avoid
+misinterpretations of the artifact. Most project's which are past their ``mature point''
+use add and deprecate development process as a way of preventing indirect conflicts of legacy code by not changing it. 
+The second item is to have highly experienced team
 members be involved in the project planning in order to foresee where problems may arise and act accordingly. (Developers also mentioned
-their personal experience in preventing indirect conflicts.) These two tools are really the only measures that developers have right now
+their personal experience in preventing indirect conflicts.) These three tools are really the only measures that developers use right now
 in order to prevent indirect conflicts. This is probably due to the inherit difficulties in predicting indirect conflicts, as will be seen in
-Section~\ref{sec:td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and conflicts 
-continue to occur. However, developers did not seem overly concerned that these two items do not totally prevent indirect conflicts. This
+Section~\ref{sec:td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and thus, conflicts 
+continue to occur. However, developers did not seem overly concerned that these three items do not totally prevent indirect conflicts. This
 is because most developers ascribe to the idea of ``I work until something breaks''. This shows that developers are somewhat content with
@@
 change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
-by several researchers in both delta debugging and code slicing. The interest that developer's have in fixing issues as they
+by several researchers in both delta debugging~\cite{Zeller:2002:ICC} and program slicing~\cite{Weiser:1982:PUS}. 
+The interest that developer's have in fixing issues as they
 arise more quickly again shows that developer's prefer fixing real problems directly as opposed to trying to prevent potential problems.
@@
 \begin{description}
-	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of the add and deprecate model as well as expert 
+	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of object contract understanding, the add and deprecate model, as well as expert 
 																	 knowledge in order to prevent indirect conflicts.
@@
 In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
-have an automatic understanding of how the project's code is suppose to work. From the halting problem, we know it is a rather difficult
-if not impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
+have an automatic understanding of how the project's code is suppose to work. From the halting problem~\cite{Burkholder:1987}, 
+we know it is an impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
 human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
@@
 
-% insert here
+Since this paper has covered a wide spectrum in regards to indirect conflicts (preventing, catching, solving, process, and tools), there
+exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
+in the notion of indirect conflicts, lessons learned from topic in awareness, preemptive direct conflict detection, and debugging can be used.
+
+Kim conducted several initial focus groups as well as web surveys to determine what developers are interested in with regards to software
+modifications~\cite{Kim:2011:ESA}. The top two interests found were that developers wanted to know whose recent code changes semantically
+interfere with their own code changes, and whether their code is impacted by a code change. These areas of interest resonate with what was
+found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way.  Kim
+also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
+were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
+noting ``I get a big laundry list... I see the email and I delete it''.
+
+In a case study of impact awareness de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
+preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
+development processes.
+
+In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards false positive output
+as well as developer process integration. Developers complained that static analysis tools usually produce a large amount of false
+positives to the point where output is ignored, as well as the tool not being designed properly to fit their current development work
+flow. In our case we found that the current development work flow is ``work until something breaks'' as opposed to try and find where things
+might break before making code changes.
+
+Hattori et. al.~\cite{Hattori:2012:ICG} found, through qualitative user studies, that developers tend to use the bare amount of communication
+towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
+is the same mentality found in this paper as ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
+sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
+This is the situation many indirect conflict tools have strived for.
+
+While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
+can be, and should be, applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
+changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
+with the idea that developers wait for something to go wrong, then wish to fix the issue as fast as possible. Program slicing~\cite{Xu:2005:BSP}
+has also been extensively studied in order to aid in debugging in manual or automated techniques.
 

### 2013-09-30 iced-2013/paper.tex
ICED new abs, intro, methodology.

@@
 
-Developer awareness techniques have been proposed and studied to aid in developer
-understanding and  efficient developer output in terms of source code produced. These techniques
-involve the sharing of information regarding source code changes in a specific project in order
-to either catch potential issues before they are fully integrated or to help developers debug
-pre-existing issues. These techniques have largely focused on either \textit{direct} or 
-\textit{indirect conflict} issues. While the techniques and tools for direct conflicts have had
-large success, tools either proposed or studied which focus on indirect conflicts have had many common
-struggles and poor developer interest. To better understand why indirect conflict tools either fail or have little
-developer investment, we performed a grounded theory study with 97 developers involved in either
-interviews or a web survey. We identified the main features that characterize indirect conflict issues,
-what processes and techniques developers use to aid in preventing and solving indirect conflict issues,
-and what developers would look for in a tool to aid them in these conflicts.
+Awareness techniques have been proposed and studied to aid in developer
+understanding, efficiency, and quality of software produced. Some of these techniques have focused 
+on either \textit{direct} or 
+\textit{indirect conflicts} in regards to prevent, catching, or debugging these conflicts as they arise
+through source code changes. While the techniques and tools for direct conflicts have had
+large success, tools either proposed or studied for indirect conflicts have had many common
+struggles and little developer interest. To better understand common indirect conflict tool issues, 
+we performed a grounded theory study with 97 developers involved in either
+interviews or a web survey. We found that: indirect conflicts are a real and frequent problem,
+developers already posses a large suite of tools and processes which can prevent and catch
+indirect conflicts at a manageable scale, and that developers have a work until something breaks philosophy
+which leads to the lesser want of prevention tool as to the want of quick and perhaps automatic debugging tools
+for indirect conflicts.
 
@@
 comes the need for larger awareness among developers to have ``an understanding of the activities of others
-which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness in theory
-mitigates the downside of parallel development which is the cost of conflict prevention and resolution, however,
-in practice we see these mitigated losses appear quite frequently. Not only do these conflicts still occur,
+which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
+mitigates some downsides of parallel development which include the cost of conflict prevention and resolution, however,
+in practice we see these mitigated losses continue to appear quite frequently. Not only do these conflicts still occur,
 but at times they can prove to be a significant and time-consuming chore for developers~\cite{Perry:2001:PCL}.
@@
 \textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
-artifact, finding expert knowledge about a particular file, or knowing when another developer is working on
-the same file as you. Tools have been created and studied for these types of issues
+artifact, or finding expert knowledge about a particular file. Tools have been created and studied for direct conflicts
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. However, indirect conflict tools have not had the same success. Indirect conflicts
-can arise in source code for such reasons as having one's own code negatively affected by a library upgrade or
-making a code change only to find out it had negative effects on unit or integration test or  having negative effects
+positive developer feedback. However, indirect conflict tools have not shared the same success. Indirect conflicts
+can arise in source code for such reasons as having one's own code negatively affected by a library upgrade,
+making a code change only to find out it had negative effects on unit or integration test, or having negative effects
 on the way other developer's have interpreted the use of your code. Tools that have attempted to either
@@
 
-The inherent difficulty regarding indirect conflicts is one, to determine which software dependencies are
-causing indirect conflicts, and two, to be able to analyze these dependencies, specifically as to when they change
-and whether or not they will induce an indirect conflict or otherwise negative consequences of a software project.
-Sarma et al.~\cite{Sarma:2007:TSA} have devised a tool (Palantir) which can both detect potential indirect conflicts and alert
-developers to these conflicts arising in their software projects at the class signature level. Ariadne~\cite{Trainer:2005:BGT}, another indirect 
-conflict tool, creates an exploratory environment where developers can see what areas of a project they are affecting with
-source code changes at the method level and thus where indirect conflicts may occur. 
-Another tool developer for indirect conflicts is CASI~\cite{Servant:2010:CPI}
-which utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of the more traditional method call graphs to provide 
-again an environment to see what areas of a project are being affected by a source code change. 
-These tools have all shown to have the same common difficulty of scalability in the face of information overload. Too many
-dependencies with the potential for indirect conflicts are being reported to developers to the point where the number of false
-positives reported becomes unbearable. Our own unpublished tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered this same fate.
+Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
+when those changes will create negative consequences in a software project.
+Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
+and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
+one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
+and create a taxonomy for different
+types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
+creates an environment where developers can see how source code changes will affect other areas of a project at the 
+method level and thus where indirect conflicts may occur. 
+Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
+utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
+an environment to see what areas of a project are being affected by a source code change. 
+Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
+or providing redundant information to preexisting tools. 
+Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
 
-Many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-developers solve their indirect conflict issues through the means of other 
-developers~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. Common issues arise here in that developers do not always
-utilize others to solve issues, socio-technical congruence is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problem with scalability in that many many developers may be linked to
-anyone particular persons work. Again the notion of confirmation overload becomes apparent and false positives become a large
-concern.
+Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
+However, socio-technical congruence is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
 
-We are interested in exploring three main areas in the hopes of resolving some of the issues that are occurring
-with indirect conflict tools: the nature of these conflicts in the how, what, and where, the processes that
-developers use prevent, catch, and solve indirect conflicts, and what developer opinions are on preventative 
-indirect conflicts tools as well as what they would look for in any tool for indirect conflicts. We broke
-these categories down into the following 8 research questions:
+We are interested in exploring three questions in the hopes of resolving some of the issues that are occurring
+with indirect conflict tools:
 
 \begin{description}
-	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{How do indirect conflicts occur?}
-	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
-	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
-	\item[RQ4\namedlabel{itm:rq4}{RQ4}] \textit{How do developer prevent indirect conflicts?}
-	\item[RQ5\namedlabel{itm:rq5}{RQ5}] \textit{How do developers catch indirect conflicts?}
-	\item[RQ6\namedlabel{itm:rq6}{RQ6}] \textit{How to developers react to and solve indirect conflicts?}
-	\item[RQ7\namedlabel{itm:rq7}{RQ7}] \textit{When are preemptive indirect conflict tools best used and what would developers want from them?}
-	\item[RQ8\namedlabel{itm:rq8}{RQ8}] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What is the nature of indirect conflicts?}
+	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What types of process and tools are being used by developers in regards to indirect conflicts?}
+	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{How should indirect conflict tools work to best benefit developers?}
 \end{description}
@@
 We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
-developers in order to answer these questions. Our findings indicate that while indirect conflicts are very much a real problem
-in software development, developers already poses the tools to prevent and catch indirect conflicts to a certain manageable degree and have
+developers in order to answer the aforementioned questions. Our findings indicate that while indirect conflicts are very much a real 
+and frequent problem
+in software development, developers already posses the tools to prevent and catch indirect conflicts to a manageable degree and have
 the mindset of working until something breaks, then attempting a fix. Thus, developers would rather spend their time developing instead
-of taking or reading preventative measures, resulting in the need of tools to debug and solve indirect conflicts quickly
-(reactionary tools) as opposed to tools which attempt to prevent or catch indirect conflicts early (preemptive tools).
+of taking or reading preventative measures, resulting in the need of tools to debug and solve indirect conflicts quickly as opposed to tools 
+which attempt to prevent or catch indirect conflicts early.
 
 We continue this paper by first presenting our summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
-continue with the results of our study in regards to our 8 research questions in Section~\ref{sec:results} 
-followed by a discussion in Section~\ref{sec:disc} of what was found and how it can be applied to future study and implementation
-of indirect tools or processes. We next present how evaluation was performed in terms of the credibility
-and quality of our results in Section~\ref{sec:eval}, then cover related work in Section~\ref{sec:related}
+then explain and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}.
+Next, an evaluation of how our study was performed in terms of the credibility
+and quality of our results is described in Section~\ref{sec:eval}. We then cover related work in Section~\ref{sec:related}
 and conclude in Section~\ref{sec:conc}.
@@
 Our grounded theory study was performed in two parts. First, a round of semi-structured interviews were conducted which 
-addressed the larger who, what, when, where, and how about indirect conflicts. Secondly, a survey was given
-out which was used to confirm what was theorized from the interviews on a larger sample size as well as obtain
-larger developer opinion.
+addressed the 3 research questions as mentioned above. Secondly, a survey was conducted
+which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
+larger developer opinion of the subject.
 
@@
 Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
-\textit{open code} to create a social theory ``grounded'' in the empirical data. For an exploratory study such as
-the ours, grounded theory is well suited because involves starting from very broad and abstract type questions and
+\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
+the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions and
 making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
-regining the sampling criteria throughout the course of the study to ensure that participants are able to answer new
+realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
 questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
@@
 codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
-link our defined codes to categories of grounded theory such as context and consequences. The ultimate goal of grounded 
+link our defined codes through categories of grounded theory such as context and consequences. The ultimate goal of grounded 
 theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in
@@
 Our participants we chosen based on their direct involvement in the actual writing of software for
-their respective companies or projects. These participants' software development titles included: senior developer,
-lead developer, software developer, software developer in test, and junior developer.
-In addition to software development, some participants were also chosen based on their previous experience
-of software development as well as their current experience with project management at some capacity. These management
-type participants involved in project planning to some capacity between technical architecture, project maintainer,  
-chief technical officer, or program manager.
+their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
+with an average of 8 years of experience.
+In addition to software development, some participants were also chosen based on both their experience of software development 
+as well as their experience with project management at some capacity.
 
@@
 interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
-audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
-our breadth of participants was rather large spanning multiple open and closed source project and our interview 
-answer saturation~\cite{Guest:2006:SAP} was quite high (to be seen in Section~\ref{sec:results}).
+audio being stored for future use. We were pleased with the response rate to our initial invitation emails as
+our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
+and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.
 
 Interview participants first answered a number of structured demographic items. Next, participants were 
-asked to describe various software development experiences regarding our three main areas of focus: the nature
-of indirect conflicts, current processes used regarding indirect conflicts, opinions on future and the use of
-current tools. processes they use or encounter in their current working environment. These three areas of focus
-were studied by having the interview participants talk about their experiences and opinions in the following
-\totalCategories{} semi-structured research topics which were asked of:
+asked to describe various software development experiences regarding our three research questions.
+Our three questions were studied by having the interview participants talk about their experiences and opinions in the following
+\totalCategories{} semi-structured research topics:
 
 \begin{itemize}
+\item Examples of indirect conflicts from real world experiences.
+\item Software artifact dependency levels and where conflicts can arise.
 \item Software development tools for dependency tracking and awareness.
 \item Software development process for preventing indirect conflicts.
-\item Software artifact dependency levels and where conflicts can arise.
 \item How developers find internal or external software dependencies.
-\item Examples of indirect conflicts from real world experiences.
 \item How indirect conflicts are detected and found.
@@
 largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
-In general, once the category was explained to the participant, little prompting was necessary as the participant
-would jump right into an experience, or issue he or she had with the question or category. However, not all
-participants had strong opinions or any experience on every category mentioned. For these participants, answers 
+However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
 to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
@@
 in our analysis and results as seen in Section~\ref{sec:results}. Interviews lasted 
-from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
-participant had in software development as they were able to give more experiences and more informed opinions.
+from 15 minutes up to 75 minutes and were very much dependent on the experience of the participant.
 
@@
 Subnet Solutions, and Amazon.
-However, participants who took part in the round of interviews were not asked to participate in the surveys but
-rather to act as a contact point for other developers in their team, project, or organization who may be interested
-in completing the survey. Aside from this aforementioned list, two large groups of developers were asked to
+However, participants who took part in the round of interviews were asked to act as a contact point for other 
+developers in their team, project, or organization who may be interested
+in completing the survey. Aside from this aforementioned list, two groups of developers were asked to
 participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
 users were selected based on large amounts of development activity on GitHub and the Apache developers
-were selected based on their software development participation on specific projects known to be used heavily
+were selected based on their software development contributions on specific projects known to be used heavily
 utilized by other organizations and projects.
@@
 
-The created survey was based off of the \totalCategories{} categorical hypotheses given by the round of interviews. The survey
+Survey participants were invited to participate in the survey by email. No reminder email was sent as the
+survey responses were not connected with the invitation email addresses and thus participants who did respond
+could not be identified. We directly emailed 1300 participants and ended with 78 responses
+giving a response rate of 6\%. We attribute the low response rate with: the surveys
+were conducted during the months of July and August while many participants may be away from their regular positions.
+and our GitHub and Apache participants could not be verified as to whether or not they actively support the
+email addresses used in the invitations. In addition, the survey was considered by some to be long and require
+more development experience than may have been typical of some of those invited to participate.
+
+The created the survey was based off of categorical hypotheses created by the round of interviews. The survey
 was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
@@
 piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
-with feedback given at the end. The previous interview participants were selected based on their domain knowledge
-expertise with indirect conflicts in order to provide what we think would be amount to better feedback and resulting
-in a more polished survey. This also allowed the previously interviewed developers to examine what hypotheses were
-formed and we would be moving forward with. The final survey consisted of the following:
-2 multiple choice questions for developer experience and demographic information; 3 level of agreement questions for an indication of
+with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed the previously interviewed developers to examine what hypotheses were
+formed and what we would be moving forward with. The final survey consisted of:
+2 multiple choice questions for demographic information; 3 level of agreement questions for an indication of
 what types of environments indirect conflicts are more likely to occur;
@@
 through the interviews that some questions require more experience from participants than may be provided.
-
-Survey participants were invited to participate in the survey by email. No reminder email was ever sent as the
-survey responses were not connected with the invitation email addresses and thus participants who did respond
-could not be removed from a reminder list. We directly emailed 1300 participants and ended with 78 responses
-giving a response rate of 6\%. We attribute the low response rate with a couple of factors. One, the surveys
-were conducted during the months of July and August while many participants may be away from their regular positions.
-Two, our GitHub and Apache participants could not be verified as to whether or not they actively support the
-email addresses used in the invitations. And finally, the survey was considered by some to be long and require
-more development experience than may have been typical of some of those invited to participate.
 
@@
 five participants. Those that were re-interviewed came from the range of 5-10 years of software development
-experience. Re-interviewed participants were given our 8 research questions along with results and 9 findings
+experience. Re-interviewed participants were given our 3 research questions along with results and 9 findings,
 and asked open ended questions regarding their opinions and experiences to validate our findings. The five

### 2013-10-03 iced-2013/paper.tex
ICED 2nd draft almost done. Needs one read.

@@
 
-\usepackage{balance}
+%\usepackage{balance}
 \usepackage{float}
@@
 
-\section{Results}
+\section{Results and Discussion}
 \label{sec:results}
 
-We now present our results of both the interviews and surveys conducted in regards to our research questions
-outlined in Section~\ref{sec:intro}. We first present and discuss the general nature of indirect conflicts in
-terms of when, what types, and how often indirect conflicts occur. We next show the findings for the process
-related research questions in regards to how developers react to, processes to prevent, and reactionary steps
-to indirect conflicts.
+We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
+outlined in Section~\ref{sec:intro}. We list each finding followed by our qualitative and quantitative results
+from which we draw the finding which are then followed by discussion.
 
-\subsection{The Nature of Indirect Conflicts}
+\subsection{What is the nature of indirect conflicts?}
 
@@
 its development is beyond the scope of this paper, we first wanted to explore the general nature of indirect 
-conflicts. This being the case we explored, through our interviews and surveys, what software developers believed
-to be the case when it came to how indirect conflicts occur, when and how often they occur, and the types of 
+conflicts. This being the case, we explored, through our interviews and surveys, what software developers believed
+towards to how indirect conflicts occur, when and how often they occur, and the types of 
 software object's that incur indirect conflicts.
@@
 \begin{description}
-	\item[~\ref{itm:rq1}] \textit{How do indirect conflicts occur?}
+	\item[F1\namedlabel{itm:f1}{F1}] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
+						by the manipulation of software contracts. This lack of awareness and understanding is compounded by having a multi developer
+						or multi team environment.
 \end{description}
@@
 of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
-what a software object guarantees, meaning how the input, output, or how any aspect of the object is garunteed
+what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
 to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
@@
 with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
-are afraid of the long range issues that may arise from that change''. 36\% of interviewed developers mentioned 
-that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
-integration fails noting that major releases of libraries can be a large issue with one developer saying 
-``their build never breaks, but it breaks ours''. Finally, 29\% of developers interviewed explained that changes to
-databases are often a large source of indirect conflicts saying ``[schema changes] break stuff all over the place''.
+are afraid of the long range issues that may arise from that change''.
+
+From these listed results, we see a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
+fully understand the ramifications of their source code change, especially when a software object's contract is being changed. 
+This was largely apparent when developers were dealing with legacy code,
+or another developer's code in general. The non-understanding of how implementation decisions were made is a large factor and that changing
+older code was quite prone to indirect conflicts as they were unsure of where the code was being used. This presents an immediate problem regarding
+indirect conflicts in software maintenance and we would think the simple fix would be static analysis tools for source
+code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
+the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
+cross language dependencies which also breaks down any potential for static analysis on these software dependencies. Another potential solution
+to this issue could be formal documentation such as UML. However, from Petre~\cite{Petre:2013:UP} we see that UML is not used so much in practice.
+These short comings leave the door open for other type of pro-developer documentation or tools to allow quick understanding of code changes and
+how they may impact cross team, cross project, or other dependencies. At any rate, it is clear that developers do not posses the tools needed
+to quickly evaluate a software change (especially when contracts change) from a wide spectrum of dependencies that may not have the ability to
+be statically analyzed.
 
 \begin{description}
-	\item[~\ref{itm:rq2}] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
+	\item[F2\namedlabel{itm:f2}{F2}] Indirect conflicts, while occurring quite frequently throughout a development cycle,
+						occur more often before a project reaches its ``mature point'' rather than after for each development cycle.
 \end{description}
 
-From the interviewed participants, 58\% of developers explained that indirect conflicts occur all the time in
+From the interviewed participants, 58\% of developers explained that indirect conflicts occur ``all the time'' in
 their development life cycle with a minimum occurrence of once a week. Of those 58\%, developers noted that a
-lot of their development time is put into understanding code to avoid indirect conflicts, more serious issue tend
-to happen once a month, and that the conflicts that do occur tend to be quite unique from each other.
+lot of their development time is put into understanding code to avoid indirect conflicts, with more serious issues tending
+to occur once a month, and that the conflicts that do occur tend to be quite unique from each other. Right away we can see a serious issue with the standard
+development work flow. Developers are not actually spending their time developing, rather they are spending large amounts of time trying to avoid indirect
+conflicts. This shows that indirect conflicts are a serious threat to developer productivity and are a problem worth study and investigation.
 
@@
 
-For those developers interviewed, we asked how often indirect conflicts occur in their software development experience,
-how often they occur before and after a project's first release, as well as how often indirect conflicts occur late
+For those developers surveyed, we used this notion of a mature or stable point to ask how often indirect conflicts occur in their software development experience,
+how often they occur before and after a project's first release (mature point), as well as how often indirect conflicts occur late
 in a project's lifetime. Table~\ref{tab:often} shows the results of these questions.
@@
 
+From the interviews and the surveys combined we can make the deduction that indirect conflicts are more likely to occur early 
+in a projects development cycle as opposed to later. This must be a major concern for any indirect conflict tools moving forward. Since many
+of the tools listen in Section~\ref{sec:intro} implement some sort of filtering system or taxonomy to avoid information overload, this
+finding should be included in future workings. Knowing that indirect conflicts occur more frequently early in a development cycle will let
+these tools know that those reported later in the development life cycle have a high probability (compared to without this knowledge) of
+being a true positive. This is a major concern when dealing with information overload and to provide developers with what they may find
+relevant to their work. The results in in Table~\ref{tab:often} also suggest that preemptive indirect conflict tools may be more accurate
+towards the mature point of a project, but have potentially more use earlier where the indirect conflicts are more likely to occur.
+While what the ``mature point'' of a project is could be speculated upon, we suggest it is caused for one of two reasons, The
+first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
+of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
+development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
+would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
+to indirect conflicts. In the end, developers survey pointed out that before the ``mature
+point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
+said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).
+
 Of those surveyed that could not give direct answers to the questions in Table~\ref{tab:often}, the more common answers were 
@@
 unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
-with spikes during large revamps or the implementation of cross-cutting new features.''.
-
-\begin{description}
-	\item[~\ref{itm:rq3}] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
-\end{description}
-
-From the developers interviews, 9 were currently working with large scale database applications and all 9 listed database
-schema as a large source of indirect conflicts. Out of the 5 developers interviewed that currently work on either software
-library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. Some
-other notable artifacts were user interface displays for web development, and full components in component base game
-architecture.
+with spikes during large revamps or the implementation of cross-cutting new features.''. These results can be argued to show that the occurance
+of indirect conflicts can largely depend on the management, and deadlines of projects. These are important factors to consider when trying
+to define a ``mature point''. A large refactoring or rewrite can reset the ``mature point'' of a project and should be considered for
+any taxonomies created for preemptive indirect conflict tools.
 
@@
 
-\subsection{Processes of Indirect Conflicts}
+From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
+that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
+do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
+that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
+67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. These results show that indirect conflicts
+seem to be compounded by a multiple human factor found in parallel development. The lack of awareness as to other developers work in a project 
+again comes to light. We would think that developers would leverage the other developer's knowledge in their projects to avoid indirect conflicts
+but as we will see most developers tend not leverage their social connections within a project to help solve indirect conflicts. This could be
+a potential issue of social structure in that developers simply do not have the means or will to talk to their colleagues, or it could be that
+developers heavily leverage the notion of software stigmergy as shown by Bolici et. al~\cite{Bolici}. Creating effective communication around
+technical dependencies between developers and teams through either explicit or implicit means appears to be a large task to accomplish but a
+carsickness task as shown by these result being that static analysis breaks down at these levels.
 
-The results that follow pertain to software development process and its relation to indirect conflicts. We explore
+\begin{description}
+	\item[F3\namedlabel{itm:f3}{F3}] Indirect conflicts are largely brought on by the changing of interfaces to software structures. The most common type of structures
+						to cause indirect conflicts are methods/functions and database schemas.
+\end{description}
+
+From the developers interviewed, 9 were currently working with large scale database applications and all 9 listed database
+schema as a large source of indirect conflicts with developers saying ``[schema changes] break stuff all over the place''.
+Out of the 5 developers interviewed that currently work on either software
+library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. 
+36\% of interviewed developers mentioned that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
+integration fails noting that major releases of libraries can be a large issue with one developer saying 
+``their build never breaks, but it breaks ours''. Some
+other notable artifacts were user interface displays for web development, and full components in component base game
+architecture.
+
+From these results, we can see the evidence that the problemed areas of software architecture in regards to indirect conflicts are
+the interfaces between structures, mainly being methods and database schemas. This is good news for preexisting indirect conflict tools as
+many of them have focused at the signature level of methods as a spot for technical dependency changes. However, there has been little focus on indirect
+conflicts at the relation database schema level~\cite{Maule:2008:IAD}, possibly due to the inherit difficult nature
+of analysis of a database schema in relation to source code structure. Developers have given researchers an area of indirect conflicts that have been
+over looked and we should be making an effort to provide answers in this area. Also, note here that developers while in F1 showed that changes to software
+object contracts were the concern, the focus here was shifted to the interface. While the interface does make up a large component of a software object contract,
+it does not fully encompass it. Precondition and postcondition checks, which are part of the object contract are performed in the body of a method and are largley overlooked
+by indirect conflict tools. This may be a source of many false negatives from these tools.
+
+Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
+static analysis of the changing code is not always possible. Developers who were also
+tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
+how they may affect these outside projects with their changes. However, it appears that when interfaces of methods change through software
+evolution, indirect conflicts are more prone to occur. This result again shows that even at the library or binary level, any type of support tools for indirect
+conflicts should focus at the interface level of change.
+
+\subsection{What types of process and tools are being used by developers in regards to indirect conflicts?}
+
+The findings that follow pertain to software development process and its relation to indirect conflicts. We explore
 how developers react to indirect conflicts, which stakeholders are responsible for their fixes, and if any development
@@
 \begin{description}
-	\item[~\ref{itm:rq4}] \textit{How do developer prevent indirect conflicts?}
+	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of object contract understanding, the add and deprecate model, as well as expert 
+																	 knowledge in order to prevent indirect conflicts.
 \end{description}
@@
 while another noted that software object contracts do solve the problem in theory, but that doesn't mean that problems don't 
-still occur in practice. The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
+still occur in practice.  Since ``Design by Contract''\texttrademark ensures that software artifacts, while they can be changed, 
+should always behave in a way that is understood by other developers, languages should just directly support ``Design by Contract''\texttrademark,
+however, this is not the case. One developer said that ``is it too much of a burden'' in regards to strictly making the contract
+apart of the source code. But why is it not such a burden is write separate unit tests that are normally designed around testing
+the contract? There is obviously a trade off between having rules built into the source versus having external tests to check if rules
+were implemented correctly. The question should become which is more effective for software development in terms of bugs produced 
+from contract breaks or bugs produced from convoluted code. This paper cannot begin to attempt to answer this question.
+
+The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
 that the development process changes once a stable point of a project is met and the ``add and deprecate model [prevents] a lot 
@@
 while slowly phasing out the out code in subsequent releases or other development cycle needs. This allows a stoppage of indirect conflicts
-by allowing developers to use the older versions of software objects which remain unchanged. Lastly, pure developer experience was mentioned
+by allowing developers to use the older versions of software objects which remain unchanged. From the interviews, it seems this is a largely
+effective way of preventing indirect conflicts, however, it was mostly used for libraries and methods only. In some cases such as development
+in a large database project, the add and deprecate model simply does not work as all code must run off the newest changes. So while the
+add and deprecate model is certainly a good fix for heavily version-ed software, especially when given to 3rd party developers, it does
+not necessarily work on internal projects which need to cooperate on which versions of methods or schemas they use.
+
+Lastly, pure developer experience was mentioned
 as a source to prevent indirect conflicts. 37\% of developers mentioned that when planning code changes, either a very experienced member
 of the project was involved in the planning and has duties to foresee any indirect conflicts that may arise, or that while implementing
-a code change developers must use their personal knowledge to predict where indirect conflicts will occur.
+a code change developers must use their personal knowledge to predict where indirect conflicts will occur. While most interviewed developers
+had such an expert in their software projects, they admitted to not being able to foresee all indirect conflicts that may occur, especially
+when talking about external teams or projects who may or may not use their final product. This raises the question of how much planning
+by expert developers should really go into avoiding indirect conflicts? Dromey~\cite{} raises the debate of prevention versus cure of software
+defects and how difficult a problem it is to measure. Should developers be using certain tools to plan their changes or should they be even
+planning at all? Which method provides the best software quality for the minimum effort. These are not easy questions to answer but may
+be key in improving developer productivity when concerning indirect conflicts.
 
-Of the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
+To confirm our findings, the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
 total), 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of ``Design
@@
 
+Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and thus, conflicts 
+continue to occur. However, developers did not seem overly concerned that these three items do not totally prevent indirect conflicts. This
+is because most developers ascribe to the idea of ``I work until something breaks'', or taking the curative approach. This shows that developers are somewhat content with
+the level of prevention that happens for indirect conflicts, however, this could be an issue of developers simply following old standards.
+It seems that the cure approach to indirect conflicts is heavily used in industry, but is it better than prevention?
+The lack of preventative process being used is probably due to the difficult nature of predicting indirect conflicts. These results have mostly
+shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
+more effectively spent, in the developer's eyes, developing and fixing problems as needed than putting excess time into the prevention of potential errors.
+This is an important consideration for moving forward with indirect conflict tools and research.
+
 \begin{description}
-	\item[~\ref{itm:rq5}] \textit{How do developers catch indirect conflicts?}
+	\item[F5\namedlabel{itm:f5}{F5}] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
 \end{description}
@@
 said build processes (either nightly builds or building the project themselves), and others mentioned code reviews
-while those dealing with a user interface mentioned user complaints from run time testing.
+while those dealing with a user interface mentioned user complaints from run time testing. It appears that most indirect conflicts are currently
+detected by using what some would call proper software development practices. The words ``use case coverage'' were constantly being used by developers
+when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
+not all indirect conflicts should be caught. This leads us to believe that as long as the contract of a software structure is understood,
+proper tests should be able to be written which will catch most if not all indirect conflicts. This is another example of developers subscribing
+to the work until something breaks philosophy in that if proper tests are written, the errors will be caught after they break something.
+However, problems occur when a software contract is
+not properly understood and use cases are not tested for or examined. Once an object is not properly
+tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. This is the major
+negative result of work until something breaks. The break detectors need to be properly written. Some languages
+such as Eiffle and the Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}
+enforce object contracts to be well understood through the ideas of preconditions and postconditions of structures which is an example of preemptive
+conflict detection. However, the trade off between external tests and in-line contracts is still in place as previously mentioned.
 
@@
 and project expertise (6\%) as factors of catching indirect conflicts. 
+Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
+the build process more to catch compile time errors witch prevent the product from being used. These types of build breaking indirect
+conflicts do occur but for the most part developers posses tools capable to build their solutions in order to catch these problems before
+the code change is added to the project. Therefore, unit and integration testing appears to be the largest factor in catching indirect conflicts
+as long as they are properly written and contracts are understood.
 
 \begin{description}
-	\item[~\ref{itm:rq6}] \textit{How to developers react to and solve indirect conflicts?}
+	\item[F6\namedlabel{itm:f6}{F6}] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
 \end{description}
@@
 
-Of the developers surveyed, 37\% confirmed that they used some type of source control management tools to help them solve their indirect
-conflicts while 23\% said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
+Here, the generic approach
+brought forth by developers was seen to be history checking. Developers would look into source code change logs and build or test failure
+logs in order to see what is broken and what might have changed recently that could be affecting the break. This gives the impression that
+developers already have the tools to solve indirect conflicts. However, this leads to the question: how quickly and
+efficiently are they able to solve them and is the time spent debugging outweighing the potential time spent preventing such errors?
+While each developer gave some examples on solving indirect conflicts and the times needed, they all agreed that to be able to solve them
+quicker would be a huge benefit. This again shows the developer preference towards work until something breaks as opposed to prevent the
+error in the first place. While the debate of prevention versus cure goes on, possible solutions as to solving
+indirect conflicts quicker are the automatic analysis of source code
+change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
+by several researchers in both delta debugging~\cite{Zeller:2002:ICC} and program slicing~\cite{Weiser:1982:PUS}. The fact that researchers
+have had such success in the cure or debugging side of indirect conflicts could be due to the lack of indirect conflict. Once a problem has
+occurred, there is no chance of it being a false positive. Academics do not have to create taxonomies to filter results and developers do
+not need to spend time working on only potential issues. Their time is more effectively, in their eyes, spent developing than speculating.
+
+Of the developers surveyed, 23\% also said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
 to solve indirect conflicts. Interestingly, only 13\% of developers mentioned a form of communication with other developers in aid to
-solving these conflicts and only 4\% mentioned the reading of formal documentation.
+solving these conflicts and only 4\% mentioned the reading of formal documentation. This shows is the serious lack of static analysis tools
+available for developers in debugging as many projects and languages simply do not support it (adding cause for non static analysis tools moving
+forward), but more importantly, that developers tend to not communicate or use formal documentation for debugging indirect conflicts. Petre~\cite{Petre:2013:UP},
+as already mentioned, showed that it is not industry practice to use UML, which could open to floor to more informal means such as online forums. The none
+developer communication, as again already mentioned, could be addressed by either more formal communication structures or by developer stigmergy 
+in their projects if needed.
 
-\subsection{Tools for indirect Conflicts}
+\subsection{How should indirect conflict tools work to best benefit developers?}
+Ultimately, the goal of this research would be to produce any sort of tool that could aid developers in any stage (prevention, detection, or
+solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
+questions along with the proceeding results.
 
 \begin{description}
-	\item[~\ref{itm:rq7}] \textit{When are preemptive indirect conflict tools best used and what would developers want from them?}
+	\item[F7\namedlabel{itm:f7}{F7}] It is extremely difficult, if not impossible, to create a preemptive indirect 
+																	 conflict tool which can detect indirect conflicts with a low rate of false positives. Preemptive tools 
+																	 should be exploratory in nature.
 \end{description}
@@
 be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
-good the results are in regards to false positives''. These concerns combined with RQ2 suggest that preemptive tools are
-best used in unstable phases on the project when indirect conflicts are more likely to happen, thus reducing the number of false
-positives. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
+good the results are in regards to false positives''. These concerns combined with RQ1 suggest that preemptive tools are
+best used in stable phases on the project when indirect conflicts are less likely to happen, thus reducing the number of false
+positives reported. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
 Interviewed developers also suggested that their proper software development process are already in place to
 catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
-or communication within the project. Finally, most interviewed developers mentioned that their current work flow does
-not subscribe to the ideas of preemptive prevention and that they just ``work until something breaks''.
+or communication within the project. All of this again shows that developers would prefer to subscribe to the 
+``work until something breaks'' philosophy. Again, it is up for debate whether or not this is the correct approach the developers should
+be taking (effectiveness of prevention versus cure), but it is the one they currently prefer.
 
@@
 
+Developers were very quick to identify this problem which has
+been a common occurrence among any preemptive indirect conflict tool built. The main issue with false positives is the detection of changes
+which frequently cause indirect conflicts. Change types are very possible to detect through software static analysis~\cite{Fluri:2007:CDT} (when possible), 
+but even those that are
+prone to causing indirect conflicts do not cause them all the time. This is quite evident in Table~\ref{tab:pre} where developers identified
+that there is an even distribution of software changes that could cause indirect conflicts and as to if they would want to be made aware of
+those changes. The only two exceptions which stand out are that of method signature changes and main algorithm changes. The problem here however,
+is that method signature changes will be caught by IDE errors, nightly builds, unit testing, integration testing and so on, while main algorithm
+changes are very difficult to detect. This all leads to the hypothesis that preemptive indirect conflict tools cannot be built without a very
+large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
+changed preconditions in methods most times if not always, there is still a large chance that that changed precondition will actually not
+introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
+overload, and eventually to developers abandoning the tool.
+
+%In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
+%have an automatic understanding of how the project's code is suppose to work. From the halting problem~\cite{Burkholder:1987}, 
+%we know it is an impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
+%human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
+%of the use cases of a particular code artifact, then they can write proper tests to ensure those use cases are always met and not broken which
+%would be the best form of preventing and catching indirect conflicts.
+
+Given the above challenges, some indirect conflict tools have had some general success. The key difference with these tools is that they
+contain the idea of developer exploration. The idea of only presenting information to users on request is the complete opposite of the preemptive
+notion of notifying developers to potential upcoming problems. Here, the idea that a user can provide some sort of
+input, a sort of context, is a powerful tool for reducing false positives. We have really just moved the false positives of the preemptive
+tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
+precision of results given the context. These types of tools stick to previous findings being developers working until something breaks, then
+requesting information to fix it.
+
 \begin{description}
-	\item[~\ref{itm:rq8}] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
+	\item[F8\namedlabel{itm:f8}{F8}] Indirect conflict tools should take a curative approach towards indirect conflicts as opposed to preventative
+																	 in order to better fit developer activities.
 \end{description}
@@
 The developers who were surveyed, the following list represents the more common suggestions for how an indirect conflict
-resolution tool should be implemented.
+curative tools should be implemented.
 
@@
 
-\section{Discussion}
-\label{sec:disc}
-
-We discuss the results of Section~\ref{sec:results} as follows: first we look at and discuss the nature of indirect conflicts by addressing~\ref{itm:rq1} 
--~\ref{itm:rq3}, we proceed to talk about how developers structure their work flow around indirect conflicts currently by
-examining the results of~\ref{itm:rq4} -~\ref{itm:rq6}, and we lastly investigate why developers have not had much interest in tools built previously
-for indirect conflicts and what we can do moving forward to accommodate developer work flow with potential tools by looking at~\ref{itm:rq7} 
-and~\ref{itm:rq8}.
-
-\subsection{The Nature of Indirect Conflicts}
-
-In order to build solution tools for indirect conflicts, we must first understand their nature; what causes them and how often do they 
-occur. From the results we see in~\ref{itm:rq1}, a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
-fully understand the ramifications of their source code change. This was especially apparent when developers were dealing with legacy code.
-Developers identified the changing of older code to be quite prone to indirect conflicts as they were not sure where the code was being used
-anymore. We would think this can be a simple fix with static analysis tools of source
-code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
-the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
-cross language dependencies which also breaks down any potential for static analysis on these software dependencies. 
-Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
-static analysis of the changing code is not always possible. The other side of upgrading libraries also posed a problem. Developers who were often
-tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
-how they may affect these outside projects with their changes.
-
-We see from~\ref{itm:rq3} that developers have shown that the human factor involved with indirect conflicts cannot be ignored.
-From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
-that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
-do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
-that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
-67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. So even when developers are working alone,
-they do not have enough awareness of their own code as to how a software change may affect another portion of their project. This can be due to
-any of the reasons mentioned above. The problem also seems to be compounded once multiple developers and teams are involved in parallel development
-schemes. The lack of awareness as to other developers work in a project again comes to light. This lacking is increased as project size trends up.
-It would appear possible that the addition of software developers to a project, which seems to compound the likely hood of indirect
-conflicts, developers would be able to leverage other developer expertise of the project to help avoid or solve indirect conflicts.
-However, we will see in Section~\ref{sec:pd} that most developers do not leverage their social connections within a project to help
-solve indirect conflicts.
-
-From ~\ref{itm:rq2}, we can make the deduction that indirect conflicts are more likely to occur early in a projects development life cycle as opposed
-to later. We used to notion of the first major release to make the distinction between early and late in a project but the same hypothesis
-could be applied to iterative development where during the earlier stages of the current phase could be considered more probable for indirect
-conflict to occur. In Table~\ref{tab:often}, we see that as the project becomes more mature, a drop off in the frequency of indirect
-conflicts occurs. Developers surveyed answered that daily and weekly indirect conflict frequencies trail off past the project's first release and 
-those developers interviewed mentioned a point in the project, the ``mature point'', in which indirect conflicts become less frequent after.
-This would suggest that, as will be seen in Section~\ref{sec:td}, preemptive indirect conflict tools would be more useful before this point as the
-number of false positives will have a higher chance of being true positives. This ``mature point'' may be cause for one of two reasons. The
-first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
-of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
-development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
-would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
-to indirect conflicts. This two possibilities are not mutually exclusive. In the end, developers survey pointed out that before the ``mature
-point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
-said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).
-
-From the results of~\ref{itm:rq3} we can also see evidence that the problemed areas of software architecture in regards to indirect conflicts are
-the interfaces between structures. As was seen, developers listed both methods/functions and database schema's as the largest contributors
-to indirect conflict structures. Since methods are often the interface which developers use most, it seems natural that they are identified
-as the main software structure in regards to indirect conflicts. It appears that when interfaces of methods change through natural software
-evolution, indirect conflicts are more prone to occurring.
-Most tools built so far have focused on the method level, and those that attempt to prevent indirect conflicts often look for a change
-in interface to a method, either in signature or in internal body composition. The more interesting result was that of database schema based 
-indirect conflicts. There has been little focus of indirect conflicts at the relational database schema level~\cite{Maule:2008:IAD}, 
-possibly due to the inherit difficult nature
-of analysis a database schema in relation to source code structure. Project often use unique libraries and procedures when dealing with
-database operations which causes a generic tool for this analysis to be difficult at best to implement. However, developers have noted that this
-is often a very problemed area and can affect large portions of the software project.
-
-We theorize the nature of indirect conflicts with the following findings: 
-\begin{description}
-	\item[F1\namedlabel{itm:f1}{F1}] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
-						by the natural evolution of software. This lack of awareness and understanding is compounded by having a multi developer
-						or multi team environment.
-	\item[F2\namedlabel{itm:f2}{F2}] Indirect conflicts occur more often before a project reaches its ``mature point'' rather than after for each development
-						cycle.
-	\item[F3\namedlabel{itm:f3}{F3}] Indirect conflicts are brought on by the changing of interfaces to software structures. The most common type of structures
-						to cause indirect conflicts are methods/functions and database schemas.
-\end{description}
-
-\subsection{Processes of Indirect Conflicts}
-\label{sec:pd}
-
-To be able to accommodate developers in the prevention and solving of indirect conflicts, it is important to understand the current
-work flows of developers in regards to how they prevent, catch, and solve indirect conflicts. ~\ref{itm:rq4} shows us that most developers rely on
-two items in order to prevent indirect conflicts. The first of these items are development processes known as
-``Design by Contract''\texttrademark and the add and deprecate methodology. These processes are implementation of software artifacts based
-on what guarantees should be made about their behavior, and instead of changing old code, the code is simply copied into a new method/function and then modified to make a new version while the old version gets deprecated and phased out later in the project's life, respectively.
-``Design by Contract''\texttrademark ensures that software artifacts, while they can be changed, should always behave in a way that is understood
-by other developers (the guaranteed aspects of the artifact), and should be documented in a way to reflect this behavior in order to avoid
-misinterpretations of the artifact. Most project's which are past their ``mature point''
-use add and deprecate development process as a way of preventing indirect conflicts of legacy code by not changing it. 
-The second item is to have highly experienced team
-members be involved in the project planning in order to foresee where problems may arise and act accordingly. (Developers also mentioned
-their personal experience in preventing indirect conflicts.) These three tools are really the only measures that developers use right now
-in order to prevent indirect conflicts. This is probably due to the inherit difficulties in predicting indirect conflicts, as will be seen in
-Section~\ref{sec:td}. Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and thus, conflicts 
-continue to occur. However, developers did not seem overly concerned that these three items do not totally prevent indirect conflicts. This
-is because most developers ascribe to the idea of ``I work until something breaks''. This shows that developers are somewhat content with
-the level of prevention that happens for indirect conflicts and would explain why preemptive indirect conflict tools have failed to catch on,
-as will be seen in Section~\ref{sec:td}. This again is probably due to the difficult nature of predicting indirect conflicts. ~\ref{itm:rq4} has mostly
-shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
-more effectively spent developing and fixing problems as needed than putting excess time into the prevention of potential errors.
-
-The results of~\ref{itm:rq5} show us how developers currently catch indirect conflicts. It appears that most indirect conflicts are currently
-detected by using what some would call proper software practices. The use of unit and integration testing, and the use of nightly or weekly
-project builds are the main contributing factors of such discovery. The words ``use case coverage'' were constantly being used by developers
-when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. This leads us to believe that as long as the contract of a software structure is understood,
-proper tests should be able to be written which will catch most if not all indirect conflicts. The problems occur when the contract is
-not properly understood and use cases are not tested for or examined as explained in previous discussion. Once an object is not properly
-tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. Some languages
-such as Eiffle and the Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}
-enforce object contracts to be well understood through the ideas of pre and post conditions of structures. Aside from testing
-was the build process. Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
-the build process more to catch compile time errors witch prevent the product from being used. These types of build breaking indirect
-conflicts do occur but for the most part developers posses tools capable to build their solutions in order to catch these problems before
-the code change is added to the project. Therefore, unit and integration testing appears to be the largest factor in catching indirect conflicts.
-
-Finally, once the indirect conflict is caught, the results of~\ref{itm:rq6} show us how developers tend to solve the issues. Here, the generic approach
-brought forth by developers was that of history checking. Developers would look into source code change logs and build or test failure
-logs in order to see what is broken and what might have changed recently that could be affecting the break. This gives the impression that
-developers already have the tools to solve indirect conflicts. However, this leads to the question: how quickly and
-efficiently are they able to solve them?
-While each developer gave some examples on solving indirect conflicts and the times needed, they all agreed that to be able to solve them
-quicker would be a huge benefit. A possible solutions as to solving indirect conflicts quicker are the automatic analysis of source code
-change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
-by several researchers in both delta debugging~\cite{Zeller:2002:ICC} and program slicing~\cite{Weiser:1982:PUS}. 
-The interest that developer's have in fixing issues as they
-arise more quickly again shows that developer's prefer fixing real problems directly as opposed to trying to prevent potential problems.
-Their time is more effectively spent developing than speculating.
-
-We sum up the developer process with indirect conflicts with the following findings:
-\begin{description}
-	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of object contract understanding, the add and deprecate model, as well as expert 
-																	 knowledge in order to prevent indirect conflicts.
-	\item[F5\namedlabel{itm:f5}{F5}] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
-	\item[F6\namedlabel{itm:f6}{F6}] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
-\end{description}
-
-\subsection{Tools for indirect Conflicts}
-\label{sec:td}
-
-Ultimately, the goal of this research would be to produce any sort of tool that could aid developers in any stage (prevention, detection, or
-solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
-questions along with~\ref{itm:rq7} and~\ref{itm:rq8}. We first examined the prevention / detection side of indirect conflicts through the use of
-preemptive indirect conflict tools. Preemptive indirect conflict tools are used to find potential code changes that could result in indirect
-conflicts. As was found through~\ref{itm:rq7}, developers major concern with preemptive indirect conflict tools was the number of false positives put
-forth by such a tool and the information overload that was sure to follow. Developers were very quick to identify this problem which has
-been a common occurrence among any preemptive indirect conflict tool built. The main issue with false positives is the detection of changes
-which frequently cause indirect conflicts. Change types are very possible to detect through software static analysis, but even those that are
-prone to causing indirect conflicts do not cause them all the time. This is quite evident in Table~\ref{tab:pre} where developers identified
-that there is an even distribution of software changes that could cause indirect conflicts and as to if they would want to be made aware of
-those changes. The only two exceptions which stand out are that of method signature changes and main algorithm changes. The problem here however,
-is that method signature changes will be caught by IDE errors, nightly builds, unit testing, integration testing and so on, while main algorithm
-changes are very difficult to detect. This all leads to the hypothesis that preemptive indirect conflict tools cannot be built without a very
-large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
-changed preconditions in methods most times if not always, there is still a large chance that that changed precondition will actually not
-introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
-overload, and eventually to developers abandoning the tool.
-
-In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
-have an automatic understanding of how the project's code is suppose to work. From the halting problem~\cite{Burkholder:1987}, 
-we know it is an impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
-human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
-of the use cases of a particular code artifact, then they can write proper tests to ensure those use cases are always met and not broken which
-would be the best form of preventing and catching indirect conflicts.
-
-Given the above challenges, some indirect conflict tools have had some general success. The key difference with these tools is that they
-contain the idea of developer exploration. The idea of only presenting information to users on request is the complete opposite of the preemptive
-notion of notifying developers to potential upcoming problems. The tools which provide information on user request usually involve taking some
-artifact of interest and presenting information which is dependent on the artifact. Here, the idea that a user can provide some sort of
-input, some sort of context, is a powerful tool for reducing false positives. We have really just moved the false positives of the preemptive
-tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
-precision of results. These types of tools are in direct comparison with the results of~\ref{itm:rq8} in that developers were more interested in
-the debugging of issues quickly and being able to narrow down where problems emanate from. This also sticks with the findings found in
-Section~\ref{sec:pd} being that developers would rather spend their time fixing and exploring a real issue when it does occur than trying
-to prevent potential issues moving forward. As researchers, we should be focusing our efforts on the solving as opposed to prevention and 
-the detection of indirect conflicts. Prevention does not appear to worth developer time as indirect conflicts and software failures will always
-occur, detection will always require some form of human intervention to determine the correctness of a program, so this only leaves us with
-the time and effort spent on debugging.
-
-We sum up our conclusions for future indirect conflict tools with the following findings:
-\begin{description}
-	\item[F7\namedlabel{itm:f7}{F7}] It is extremely difficult, if not impossible, to create a preemptive indirect 
-																	 conflict tool which can detect indirect conflicts with a low rate of false positives.
-	\item[F8\namedlabel{itm:f8}{F8}] Indirect conflict tools are best used by providing exploratory environments to aid in the solving of indirect conflicts.
-	\item[F9\namedlabel{itm:f9}{F9}] Indirect conflict tools should be reducing debugging time of indirect conflicts as opposed to preventative
-																	 time in order to better fit developer activities.
-\end{description}
+From the above list, we can see three main areas of focus that researchers of indirect conflicts should spend their time in. The first
+is the debugging process through utilizing logs. Showing which recent code changes affect an area of code or how a commit and break a specific
+unit test are key methodologies that developers want in order to aid their preferred curative approaches. Developers also want to be able to
+understand their software object contracts better or even automatically. Detections of contract changes will play a large part in indirect conflict
+and dependency awareness moving forward, especially at the method body and database schema levels. These contract detections also play a role in
+the automatic generation of unit tests in order to comply with use case coverage. Finally, and most importantly, developers want tools that
+integrate into their current development styles, that being ``work until something breaks''. While a complete evaluation of the curative approach
+is beyond the scope of this paper, it is how developers act today and  in the end, tools should be created to aid developers.
 
@@
 Although our findings~\ref{itm:f1} -~\ref{itm:f6} may not be entirely novel or even surprising, the combination of these results allow us to discover the
-insightful findings of~\ref{itm:f7} -~\ref{itm:f9} regarding indirect conflict tools. Given how many indirect conflict tools are left with the same common
+insightful findings of~\ref{itm:f7} and~\ref{itm:f8} regarding indirect conflict tools. Given how many indirect conflict tools are left with the same common
 issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
@@
 sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
-This is the situation many indirect conflict tools have strived for.
+This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici} give the 
+possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea
+that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
 
@@
 how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings~\ref{itm:f7} 
--~\ref{itm:f9} have provided a foundation as to why past techniques and tools have had low adoption rates and where
+and~\ref{itm:f8} have provided a foundation as to why past techniques and tools have had low adoption rates and where
 researchers should focus their current and future efforts in handling indirect conflicts. We hope that this study and its 
 findings will inspire future techniques and tools for dealing with indirect conflicts that 
-will aid developers in industry as well as test and validate our theories put forth in this paper.
+will aid developers in industry as well as test and validate our theories put forth in this paper as well as spark a potential
+investigation into preventative versus curative approaches to software defects.
 
@@
 \bibliographystyle{IEEEtran}
-\balance
+%\balance
 \bibliography{paper}

### 2013-10-03 iced-2013/paper.tex
ICED 2nd draft complete.

@@
 theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in
-findings~\ref{itm:f1} -~\ref{itm:f9}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+findings~\ref{itm:f1} -~\ref{itm:f8}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
 3 criteria that are commonly used in evaluating grounded theory studies.
@@
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
-towards the rest of the project. Of those 73\%, examples were broken down as follows. 21\% of examples dealt
+towards the rest of the project. Of those 73\%, 21\% dealt
 with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
@@
 This was largely apparent when developers were dealing with legacy code,
-or another developer's code in general. The non-understanding of how implementation decisions were made is a large factor and that changing
+or another developer's code in general. The non-understanding of how implementation decisions were previously made is a large factor and that changing
 older code was quite prone to indirect conflicts as they were unsure of where the code was being used. This presents an immediate problem regarding
@@
 code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
-the easy fix. For starters, not all programming languages are subject to strong static analysis if any, while some projects involve many
+the easy fix. Not all programming languages are subject to strong static analysis if any, while some projects involve many
 cross language dependencies which also breaks down any potential for static analysis on these software dependencies. Another potential solution
@@
 to quickly evaluate a software change (especially when contracts change) from a wide spectrum of dependencies that may not have the ability to
-be statically analyzed.
+be analyzed by traditional means. This finding lays the foundation for the reality of indirect conflicts occurring in industry,
 
@@
 development work flow. Developers are not actually spending their time developing, rather they are spending large amounts of time trying to avoid indirect
-conflicts. This shows that indirect conflicts are a serious threat to developer productivity and are a problem worth study and investigation.
+conflicts. This shows that indirect conflicts are a serious threat to developer productivity and are a problem worth study and investigation. This also
+raises the point of how developers should be spending their time: preventing, or solving indirect conflicts.
 
@@
 finding should be included in future workings. Knowing that indirect conflicts occur more frequently early in a development cycle will let
-these tools know that those reported later in the development life cycle have a high probability (compared to without this knowledge) of
+these tools know that those reported later in the development life cycle have a higher probability (compared to without this knowledge) of
 being a true positive. This is a major concern when dealing with information overload and to provide developers with what they may find
@@
 unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
-with spikes during large revamps or the implementation of cross-cutting new features.''. These results can be argued to show that the occurance
+with spikes during large revamps or the implementation of cross-cutting new features.''. These results can be argued to show that the occurrence
 of indirect conflicts can largely depend on the management, and deadlines of projects. These are important factors to consider when trying
 to define a ``mature point''. A large refactoring or rewrite can reset the ``mature point'' of a project and should be considered for
-any taxonomies created for preemptive indirect conflict tools.
+any taxonomies created for preemptive indirect conflict tools. The ``mature point'' of a project (or multiple mature points) should be
+considered when building a taxonomy for indirect conflict tools.
 
@@
 a potential issue of social structure in that developers simply do not have the means or will to talk to their colleagues, or it could be that
-developers heavily leverage the notion of software stigmergy as shown by Bolici et. al~\cite{Bolici}. Creating effective communication around
+developers heavily leverage the notion of software stigmergy as shown by Bolici et. al~\cite{Bolici:2009}. Creating effective communication around
 technical dependencies between developers and teams through either explicit or implicit means appears to be a large task to accomplish but a
 carsickness task as shown by these result being that static analysis breaks down at these levels.
+
+This finding characterizes that indirect conflicts are a frequent problem in industry. It also shows what environments are more prone to indirect
+conflicts and should be handled with more delicate care than others.
 
@@
 when talking about external teams or projects who may or may not use their final product. This raises the question of how much planning
-by expert developers should really go into avoiding indirect conflicts? Dromey~\cite{} raises the debate of prevention versus cure of software
+by expert developers should really go into avoiding indirect conflicts? Dromey~\cite{Dromey:2003} raises the debate of prevention versus cure of software
 defects and how difficult a problem it is to measure. Should developers be using certain tools to plan their changes or should they be even
@@
 
-To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:disc} to either a majority of agreeing
+To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:results} to either a majority of agreeing
 responses on a topic or to a large amount of direct quotes presented by participants.
@@
 sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
-This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici} give the 
+This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici:2009} give the 
 possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea

### 2013-10-08 apie-2013/paper.tex
APIE methodology, results, and conclusions are done first draft. Need intro / related work and future work still.

@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{APIE Evolution: An Exploration into Software API Trends and Processes}
+\title{API Evolution: A Study of Software Stability and Change Trends}
 
@@
 \author{\IEEEauthorblockN{Jordan Ell}
-\IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ jell@uvic.ca}
-}
-
-\author{\IEEEauthorblockN{Braden Simpson}
-\IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ braden@uvic.ca}
-}
-
-\author{\IEEEauthorblockN{Daniela Damian}
-\IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ danielad@csc.uvic.ca}
+\IEEEauthorblockA{University of Victoria\\
+British Columbia, Canada \\ jell@uvic.ca}
+\and
+\IEEEauthorblockN{Braden Simpson}
+\IEEEauthorblockA{University of Victoria\\
+British Columbia, Canada \\ braden@uvic.ca}
+\and
+\IEEEauthorblockN{Daniela Damian}
+\IEEEauthorblockA{University of Victoria\\
+British Columbia, Canada \\ danielad@csc.uvic.ca}
 }
@@
 \begin{abstract}
-
+Something
 \end{abstract}
@@
 \section{Introduction}
+Intorduction goes here.
 
 \section{Related Work}
+Related work goes here.
 
 \section{Methodology}
+In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{}. This tool allows us to detect fine grained
+source code changes in Java projects This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
+the smallest possible edit distance between the tree. This results in the source code change at a fine grained level performed in the commit.
 
-\subsection{Results}
+We took ChangeDistiller and applied it across 10 open source Java projects. For each of the projects, we obtained the software configuration management (SCM) system
+which is used to store all source code changes of a project. When it was necessary, we converted some forms of SCM system to Git in order to reduce implementation
+burdens of using multiple SCMs. Once the SCM was obtained, we used ChangeDistiller and iterated over every commit found in a project's git master branch. We stored
+34 of ChangeDistiller's built in source code change types for each commit. We noted how many of each change type was performed in each commit and stored that information
+in a PostgreSQL database. In order to filter and protect our results, we manually inspected the 10 Java projects studied in order to identify code built for test.
+We separated changes to this test code from all other code to ensure our results only focused on real implementation while allowing us to study changes to
+test based code separately.
 
-\section{Discussion}
+Once the ChangeDistiller information was collected, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
+have preconceived notions of software stability, we decided that by studying the types of changes surrounding these releases, we could get a better understanding of
+what types of source code changes or trends constitute software stability or maturity. In order to study the release points, we went to each of our 10 project's 
+home pages on-line and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total we identified 472 releases
+across our 10 studied projects.
+
+Once the release dates were collected we set about analyzing out data by creating change ratios surrounding the release dates of each project as a way to measure
+the trend of a particular change type. 
+To do this we used Equation~\ref{eq:norm}. Equation~\ref{eq:norm} works by summing a particular change type in commits ($T_c$)
+from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$), then dividing
+this by the sum of that same particular change type in commits($T_c$) from a given number of days ($d$) before the release date ($r$) to
+the release date divided by divided by the number of commits in this date range ($|c|$). 
+This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
+type occurs more frequently after the release and if it is below 1 then is occurs more frequently before the release. For the purposes of our study, we set the
+number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
+formed much of the basis for the results to come in Section~\ref{sec:results}
+
+\begin{equation}
+\text{Ratio}_{norm} = \frac{ \sum_{c=r}^{r+d} T_c / |c|} { \sum_{c=r}^{r-d} T_c / |c|}
+\label{eq:norm}
+\end{equation}
+
+Aside from generating quantitative data, we also created a web application for the visualization of the data called API Evolution (APIE). This visualizer allowed
+us to inspect a single project and a single change type metric at a time (see Figure~\ref{fig:apie}) for qualitative analysis of software evolution trends. We
+used this tool to manually inspect 4 specific change type trends surround release dates. To do this, we aggregated change types across 50 commits, meaning that
+each point in the graph represented the date of a commit and the sum of the particular change type's occurrences over the last 50 commits. This was used to smooth
+out the curves presented by the tool to allow easier manual inspection. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
+trending, and local minimum. Since the graphs were quite turbulent, best estimations were given by two judges at each release point to fit the graph into the aforementioned
+4 slope categories. The two judges used 1.5 months before and after the release date as start and end points for the graph trend line.
+
+% APIE graph figure
+\begin{figure}[tb!]
+\centering
+\includegraphics[width=0.5\textwidth]{images/apie.png}
+\caption{An screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed.\label{fig:apie}}
+\end{figure}
+
+We performed 1888 manual inspections across 10 projects, 472 release dates and 4 change types, and used this data to form the basis of our qualitative data.
+Quantitative data was cross validated with qualitative data to form some of the final results to be seen in the next section.
+
+\section{Results}
+\label{sec:results}
+
+To answer our research question, we conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
+eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
+because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.
+
+Due to the space requirements of this paper, we focus our results on Major releases of the 10 case study projects and a select few of the calculated change ratios. 
+There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm}
+can be seen in Table~\ref{tab:ratio}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{| l | c | c | c |}
+\hline
+Object & Added & Changed & Removed\\
+\hline
+Public Classes & 0.78 & 0.76 & 0.67 \\
+Public Methods (Signature) & 1.12 & 0.66 & 2.45 \\
+Public Methods (Bodies) & - & 0.86 & - \\
+Private Classes & 0.59 & 0.92 & 0.32 \\
+Private Methods (Signatures) & 0.86 & 0.91 & 1.49 \\
+Private Methods (Bodies) & - & 0.91 & - \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Implementation oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:ratio}}
+\end{table}
+
+As it can be seen in Table~\ref{tab:ratio}, there a few notable change type trends around major releases. We can see that any changes to public classes are
+more likely to occur before a major release than after, noting that the addition, changing, and deletion of public classes have change ratios of 0.78, 
+0.76 and 0.67 respectively. The two most drastic change type trends to note from this table however are the removing of public methods and the removing
+of private classes. We can see that public methods are more than twice as likely to occur after a major release than before it and the removal of 
+private classes is three times as likely to happen before a major release than after.
+
+In terms of added code, we found as per Table~\ref{tab:ratio} that additions to private classes are a large indicator of a project being at a stable
+release point with almost being twice as likely to occur before a major release than after. A large surprise to changed code indicators was that of the
+public method bodies. We can see that with a change ration of 0.86 there is not a major indication of these types of changes happening before or after a 
+release which leads us to believe that method implementation may not be able to be detected as stable through software metrics. Changes
+to method signatures also point towards changes before a release rather than after which may indicate that stable software only needs to be stable after the
+release and not prior, as changing method signature is considered a fairly unstable change~\cite{Fluri}.
+
+Our qualitative results from manual graph inspections can be seen in Table~\ref{tab:qual}. These results show that adding, changing signatures and bodies of, and
+removing public methods tend to all be at a local minimum of change type trends at major releases.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{| l | c | c | c | c |}
+\hline
+Change Type & Upward Trend & Local Maximum & Downward Trend & Local Minimum\\
+\hline
+Added Public Methods & 21.6\% & 17.2\% & 14.7\% & 33.6\% \\
+Changed Public Methods (Signature) & 6.0\% & 19.8\% & 19.0\% & 39.7\% \\
+Changed Public Methods (Bodies) & 9.2\% & 16.5\% & 26.6\% & 37.6\% \\
+Removed Public Methods & 7.8\% & 16.4\% & 12.9\% & 41.4\% \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Qualitative graph analysis results. \label{tab:qual}}
+\end{table}
+
+Lastly we found that software changes related to testing can be highly used as an indicator of a major release point within the projects studied. The change
+ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the three largest indicators of a project release and stability with regards to test based
+changes as the addition of test classes, the removal of test classes, and the removal of test methods.
+
+We summarize our key findings as to source code change type trends which can be found at points of stability as the following: the addition of private classes more often
+before a stable point, the changing of public method signatures more often before a stable point, the removal of public methods more than twice as often
+after a stable point, the removal of private classes three times as often after a stable point, and the removal of test classes more than twice as often after
+a stable point. There are of course other change ratios to consider as listed in this section, however we found the aforementioned to be the most indicative of 
+software stability around major releases.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{| l | c | c | c |}
+\hline
+Object & Added & Changed & Removed\\
+\hline
+Classes & 0.72 & 0.86 & 0.36 \\
+Methods & 1.55 & 0.98 & 1.93 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Test oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:test}}
+\end{table}
 
@@
 
-\bibliographystyle{IEEEtran}
-\balance
-\bibliography{paper}
+Software stability is an often used measure in determining some aspects of code quality. This paper has present some initial indications as to how
+software stability may be found, based on major release stabilities, by analyzing source code change type trends both before and after major software
+releases. The source code change trends found in this paper, in combination with future work and analysis, will allow for the detection of stable code
+points throughout a software projects life time. This paper has also shown the beginnings of a visualization for source code change trends which may
+be used as a visual cue towards project stability and potential areas of instability where action may need to be taken.
 
@@
 \end{document}
+
+\bibliographystyle{IEEEtran}
+%\balance
+\bibliography{paper}

### 2013-10-08 apie-2013/paper.tex
APIE new methodology formula for new ratio calculation.

@@
 
-Once the release dates were collected we set about analyzing out data by creating change ratios surrounding the release dates of each project as a way to measure
-the trend of a particular change type. 
-To do this we used Equation~\ref{eq:norm}. Equation~\ref{eq:norm} works by summing a particular change type in commits ($T_c$)
-from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$), then dividing
-this by the sum of that same particular change type in commits($T_c$) from a given number of days ($d$) before the release date ($r$) to
-the release date divided by divided by the number of commits in this date range ($|c|$). 
+Once the release dates were collected we set about analyzing out data by creating average change ratios surrounding the release dates of each 
+project as a way to measure the trend of a particular change type at a release type. 
+To do this we used Equation~\ref{eq:norm}. Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
+a given release type a sum of a particular change type in commits ($T_c$)
+from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
+is created by summing across all release of a given release type
+a sum of that same particular change type in commits($T_c$) from a given number of days ($d$) before the release date ($r$) to
+the release date divided by divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
 This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
@@
 \begin{equation}
-\text{Ratio}_{norm} = \frac{ \sum_{c=r}^{r+d} T_c / |c|} { \sum_{c=r}^{r-d} T_c / |c|}
+\text{ChangeRatio} = \frac{ \sum_{r_0}^{r_n}\sum_{c=r}^{r+d} T_c / |c|} { \sum_{r_0}\sum_{c=r}^{r-d} T_c / |c|}
 \label{eq:norm}

### 2013-10-08 apie-2013/paper.tex
APIE methodology, results, future work, and conclusion first draft done.

@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{API Evolution: A Study of Software Stability and Change Trends}
+\title{API Evolution: A Study of Software Releases and Change Trends Leading to Stability}
 
@@
 \section{Methodology}
-In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{}. This tool allows us to detect fine grained
+In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows us to detect fine grained
 source code changes in Java projects This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
@@
 Due to the space requirements of this paper, we focus our results on Major releases of the 10 case study projects and a select few of the calculated change ratios. 
-There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm}
+There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
 can be seen in Table~\ref{tab:ratio}.
@@
 \begin{center}
+\tabcolsep=0.11cm
 \begin{tabular}{| l | c | c | c |}
@@
 \hline
-Public Classes & 0.78 & 0.76 & 0.67 \\
-Public Methods (Signature) & 1.12 & 0.66 & 2.45 \\
-Public Methods (Bodies) & - & 0.86 & - \\
-Private Classes & 0.59 & 0.92 & 0.32 \\
-Private Methods (Signatures) & 0.86 & 0.91 & 1.49 \\
-Private Methods (Bodies) & - & 0.91 & - \\
+Public Classes & 1.14 & 0.86 & 1.16 \\
+Public Methods (Signature) & 1.07 & 0.92 & 1.34 \\
+Public Methods (Bodies) & - & 1.06 & - \\
+Private Classes & 0.81 & 1.18 & 1.44 \\
+Private Methods (Signatures) & 1.00 & 1.10 & 1.22 \\
+Private Methods (Bodies) & - & 1.08 & - \\
+Files & 1.12 & 0.96 & 1.14 \\
+Documentation & - & 0.99 & - \\
 \hline
@@
 As it can be seen in Table~\ref{tab:ratio}, there a few notable change type trends around major releases. We can see that any changes to public classes are
-more likely to occur before a major release than after, noting that the addition, changing, and deletion of public classes have change ratios of 0.78, 
-0.76 and 0.67 respectively. The two most drastic change type trends to note from this table however are the removing of public methods and the removing
-of private classes. We can see that public methods are more than twice as likely to occur after a major release than before it and the removal of 
-private classes is three times as likely to happen before a major release than after.
+more likely to occur before a major release than after, noting that the addition and deletion of public classes have change ratios of 1.14, 
+and 1.16 respectively. This means that on average classes are added and removed 10\% more often after major releases than before.
+The two most drastic change type trends to note from this table however are the adding and removing of private classes. We can see that private classes are
+44\% more likely to be added after a major release and that adding private classes to a project is roughly 24\% more likely to occur before a major release.
+All results in Table~\ref{tab:ratio} could be used as identified trends of major software releases, while we have just highlighted the larger ratios in
+this paper.
 
-In terms of added code, we found as per Table~\ref{tab:ratio} that additions to private classes are a large indicator of a project being at a stable
-release point with almost being twice as likely to occur before a major release than after. A large surprise to changed code indicators was that of the
-public method bodies. We can see that with a change ration of 0.86 there is not a major indication of these types of changes happening before or after a 
-release which leads us to believe that method implementation may not be able to be detected as stable through software metrics. Changes
-to method signatures also point towards changes before a release rather than after which may indicate that stable software only needs to be stable after the
-release and not prior, as changing method signature is considered a fairly unstable change~\cite{Fluri}.
+Some of the more interesting trends found around major release points from Table~\ref{tab:ratio} are private methods being removed and public method signatures
+being changed. We hypothesized that public method signatures being changed would be heavily favored towards before a major release indicating a larger sens
+of stability in the public API, however we found that only a change ratio of 0.92 exists meaning that public method signatures undergo changed almost through
+the project steadily. The private methods being removed is also suggestive of possible maintenance being done post release, meaning that code is being
+cleaned up in order to reduce the number of overall methods in a project which are not publicly available.
 
@@
 
-\begin{table}[h]
+\begin{table*}[tb!]
 \begin{center}
@@
 \caption{Qualitative graph analysis results. \label{tab:qual}}
-\end{table}
+\end{table*}
 
@@
 ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the three largest indicators of a project release and stability with regards to test based
-changes as the addition of test classes, the removal of test classes, and the removal of test methods.
-
-We summarize our key findings as to source code change type trends which can be found at points of stability as the following: the addition of private classes more often
-before a stable point, the changing of public method signatures more often before a stable point, the removal of public methods more than twice as often
-after a stable point, the removal of private classes three times as often after a stable point, and the removal of test classes more than twice as often after
-a stable point. There are of course other change ratios to consider as listed in this section, however we found the aforementioned to be the most indicative of 
-software stability around major releases.
+changes are the removal of test classes, the changing of method signatures (both occurring more before releases), and test classes being changed. An interesting
+side note is that documentation is changed heavily before releases as opposed to after.
 
@@
 \hline
-Classes & 0.72 & 0.86 & 0.36 \\
-Methods & 1.55 & 0.98 & 1.93 \\
+Classes & 1.07 & 1.21 & 0.76 \\
+Methods (Signatures) & 1.23 & 0.83 & 1.01 \\
+Methods (Bodies) & - & 0.90 & - \\
+Documentation & - & 0.72 & - \\
 \hline
@@
 
+We summarize our key findings as to source code change type trends which can be found at points of stability as the following: the addition of private classes more often
+before a release point, the removal of private classes more often after a release point, the removal of public methods more often after a release point, 
+the removal of private methods more often after a release point, the removal of test classes more often after a release point, and the changing of test method
+signatures more often before a release point. While all change ratios may need to be considered for a final analysis or taxonomy, we have offered the strongest
+change trends.
+
 \section{Future Work}
+
+In future work of change trend analysis, we have planned further statistical tests of the change trends identified in this paper. We would like to
+be able to determine if these patterns are unique to major releases versus other types of releases (minor, alpha, beta, release candidate) and to
+be able to determine if these pattern are unique to releases alone or can be spotted throughout the history of a project.
+
+If it is deemed that these change trends are present in no other releases, some, all, or in non release history of the project, we will
+compare the history of the projects as to where these trends are found against preexisting software stability and quality metrics. We will use this data to
+create a new method of discovering when software can be considered stable in reference to the types of software changes being made.
 
@@
 
-% End of the paper
-\end{document}
-
 \bibliographystyle{IEEEtran}
@@
 \bibliography{paper}
+
+% End of the paper
+\end{document}

### 2013-10-09 apie-2013/paper.tex
APIE abs, intro, and related work now have first draft. This completes the overall first draft of this paper.

@@
 \begin{abstract}
-Something
+As software evolves over the lifetime of a project, changes to software objects such as methods or classes occur which can have ripple effects throughout
+the rest of the system causing relationships to be re-engineered and causing instability. However, projects are sometimes deemed stable at certain release
+points by projects owners or maintainers. We know that studying a release history of a project can give clues as to a project's history, release process,
+and how processes change, but can it shed light on what causes a project to be released and deemed stable? This paper presents our findings from studying
+source code change trends surrounding 109 major release points of 10 open source projects. We found 6 large source code change trends at these release
+points that can be used to determine if a project is likely near a major release point and thus deemed stable. 
 \end{abstract}
@@
 \section{Introduction}
-Intorduction goes here.
+Release points are a vital milestone of software projects. From major releases of a Waterfall based project to minor iterations of an Agile development,
+releases form an interesting single point of a project's development history.
+Third party users of a system often only see a product at a release point either major or minor,
+and expect the system to come with a sense of reliability and stability at this point. However, the decision as to when a project is ready for public usage
+as to its reliability, quality and stability can be a difficult decision to make for most project owners or maintainers.
+
+While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
+the sub study of software stability and its implications on quality and reliability remains a difficult subject to understand. 
+The decision of what makes a project stable and ready
+for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
+the project~\cite{Conway:1968}. Code is an often looked to statistic for stability but can be grossly misleading
+in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC}. 
+Creating a different quantitative approach to determining software stability and release preparedness is still a large open area of
+interest in software engineering research.
+
+We turn our analysis to the notions of software change trends, specifically those trends around major releases. Change trends have been used to detect
+stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
+With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
+be leveraged to detect stability and the evolution of source code, the question we investigate in this paper is:
+``\textit{Is it possible to identify trends in source code changes surrounding major releases of open source projects as a notion towards a project
+stability measure?}''. By answering this question, we hope to provide industry with the ability to determine a project's stability and perhaps
+reliability based on its current change trends. We also hope to provide researchers with the ability to perform further analysis of projects in
+regards to software quality with the findings of this paper.
+
+In this paper, we perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
+throughout their history. We studied 26 quantitative and 16 qualitative change trends and identified a core group of 6 change trends which occur
+prominently at major release points of the projects studied.
+
+The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work of release study and stability. Sections~\ref{sec:meth}
+explains our methodology and tools used in solving our research problem.
+Section \ref{sec:results} gives a full representation of the results and trends found as well as a summarized results of the 6 change trends found
+to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.
 
 \section{Related Work}
-Related work goes here.
+\label{sec:rel}
+While very little has been published about release quality studies and stability, there have been a few studies which attempt to address the issues directly
+or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
+pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
+and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a the project's behavior
+around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
+indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be being performed.
+Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
+at major and minor releases.
+
+The study of open source projects revolving around release points has become more available by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
+a resource of historical release dates for open source software projects to be used for future studies by other researchers.
+
+In terms of software stability, many proposed development techniques have been proposed to increase software stability. 
+Fayad~\cite{Fayad:2001:TOI}~\cite{Fayad:2002:ASS} suggests that ``business objects'' (BOs) do not change in nature and that they are inherently stable. These
+objects only need to change to accommodate external modules at the interface. Some studies such as Chow et al.~\cite{Chow:2011:SJI} have investigated
+the stability of changes to interfaces which are considered a good indication of stability.
+Mockus et al.~\cite{Mockus:2008:IQR} used major and minor release points to compare industry process quality to customer-perceived quality of the software
+project. Mockus et al. found that defect density is lowest at major releases but at the same time software quality is at its lowest all when compared to minor
+releases. The low software quality here related to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
+showed that stable core architectures can be detected by using source code changes. Finally, Fayad et al.~\cite{Fayad:2010:SSM} have
+investigated the Software Stability Model (SSM) for Software Product Lines to show that the SSM's impact on architecture and design of a software product
+can help improve the life of the product line and make it more adaptable and applicable.
 
 \section{Methodology}
+\label{sec:meth}
 In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows us to detect fine grained
@@
 
-We summarize our key findings as to source code change type trends which can be found at points of stability as the following: the addition of private classes more often
+We summarize our key findings as to these 6 source code change type trends which can be found at points of stability as the following: the addition of private classes more often
 before a release point, the removal of private classes more often after a release point, the removal of public methods more often after a release point, 
@@
 signatures more often before a release point. While all change ratios may need to be considered for a final analysis or taxonomy, we have offered the strongest
-change trends.
+change trends here as a suggestion for future focus.
 
 \section{Future Work}
+\label{sec:fut}
 
@@
 \section{Conclusions}
+\label{sec:con}
 

### 2013-11-04 iced-2013/paper.tex
ICED new results section written.

@@
 	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What is the nature of indirect conflicts?}
-	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What types of process and tools are being used by developers in regards to indirect conflicts?}
-	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{How should indirect conflict tools work to best benefit developers?}
+	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What types of process or tools are being used by developers in regards to indirect conflicts?}
+	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
 \end{description}
@@
 
-\section{Results and Discussion}
+\section{Results}
 \label{sec:results}
@@
 We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
-outlined in Section~\ref{sec:intro}. We list each finding followed by our qualitative and quantitative results
-from which we draw the finding which are then followed by discussion.
+outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
+results from which we draw our discussion to be seen in Section~\ref{sec:disc}.
 
 \subsection{What is the nature of indirect conflicts?}
-
-Although a detailed manual inspection of a project's lifetime and the indirect conflicts it experiences throughout
-its development is beyond the scope of this paper, we first wanted to explore the general nature of indirect 
-conflicts. This being the case, we explored, through our interviews and surveys, what software developers believed
-towards to how indirect conflicts occur, when and how often they occur, and the types of 
-software object's that incur indirect conflicts.
-
-\begin{description}
-	\item[F1\namedlabel{itm:f1}{F1}] Indirect conflicts are caused by a lack of understanding and awareness of source code changes which are brought on
-						by the manipulation of software contracts. This lack of awareness and understanding is compounded by having a multi developer
-						or multi team environment.
-\end{description}
 
@@
 what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
-to work; made famous by Eiffle Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
+to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
 In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
@@
 with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
-are afraid of the long range issues that may arise from that change''.
+are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
+object contacts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
+Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
+follow common patterns.
 
-From these listed results, we see a common theme of awareness is obvious. Developers cause indirect conflicts because they do not
-fully understand the ramifications of their source code change, especially when a software object's contract is being changed. 
-This was largely apparent when developers were dealing with legacy code,
-or another developer's code in general. The non-understanding of how implementation decisions were previously made is a large factor and that changing
-older code was quite prone to indirect conflicts as they were unsure of where the code was being used. This presents an immediate problem regarding
-indirect conflicts in software maintenance and we would think the simple fix would be static analysis tools for source
-code to allow developers to identify which parts of the project they are affecting with a change. However, as we found out this is not always
-the easy fix. Not all programming languages are subject to strong static analysis if any, while some projects involve many
-cross language dependencies which also breaks down any potential for static analysis on these software dependencies. Another potential solution
-to this issue could be formal documentation such as UML. However, from Petre~\cite{Petre:2013:UP} we see that UML is not used so much in practice.
-These short comings leave the door open for other type of pro-developer documentation or tools to allow quick understanding of code changes and
-how they may impact cross team, cross project, or other dependencies. At any rate, it is clear that developers do not posses the tools needed
-to quickly evaluate a software change (especially when contracts change) from a wide spectrum of dependencies that may not have the ability to
-be analyzed by traditional means. This finding lays the foundation for the reality of indirect conflicts occurring in industry,
-
-\begin{description}
-	\item[F2\namedlabel{itm:f2}{F2}] Indirect conflicts, while occurring quite frequently throughout a development cycle,
-						occur more often before a project reaches its ``mature point'' rather than after for each development cycle.
-\end{description}
+In regards to object contract changes, 9 developers currently working with large scale database applications listed database
+schemas as a large source of indirect conflicts while 5 developers that work on either software
+library projects or are in test said that methods or functions were the root of their indirect conflict issues. 
+36\% of interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
+library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
+other notable indirect conflict artifacts were user interfaces in web development and full components in component 
+base game architecture.
 
 From the interviewed participants, 58\% of developers explained that indirect conflicts occur ``all the time'' in
-their development life cycle with a minimum occurrence of once a week. Of those 58\%, developers noted that a
-lot of their development time is put into understanding code to avoid indirect conflicts, with more serious issues tending
-to occur once a month, and that the conflicts that do occur tend to be quite unique from each other. Right away we can see a serious issue with the standard
-development work flow. Developers are not actually spending their time developing, rather they are spending large amounts of time trying to avoid indirect
-conflicts. This shows that indirect conflicts are a serious threat to developer productivity and are a problem worth study and investigation. This also
-raises the point of how developers should be spending their time: preventing, or solving indirect conflicts.
+their development life cycle with a minimum occurrence of once a week, with more serious issues tending
+to occur once a month. 65\% of surveyed developers answered that indirect conflicts occur on minimum bi-weekly,
+with the majority of developers saying that weekly occurrences are most common.
 
-In terms of when in a project indirect conflicts are more likely to occur 63\% of developers interviewed said that
+63\% of developers interviewed said that
 when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
@@
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
-changes a lot, especially with agile''.
-
-For those developers surveyed, we used this notion of a mature or stable point to ask how often indirect conflicts occur in their software development experience,
-how often they occur before and after a project's first release (mature point), as well as how often indirect conflicts occur late
-in a project's lifetime. Table~\ref{tab:often} shows the results of these questions.
+changes a lot, especially with agile''. Surveyed developers also added ``indirect conflicts after a release 
+depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a 
+major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during 
+large revamps or the implementation of cross-cutting new features.''. Surveyed participants also answered that 
+indirect conflicts are more likely to occur before the first major release rather than after at the daily
+and weekly occurrence intervals as seen in Table~\ref{tab:often}.
 
@@
 \hline
-Question & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
+Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
 \hline
 \hline
-How often do you experience indirect conflicts in your software development? & 17\% & 27\% & 21\% & 17\% & 3\% & 5\% & 10\% \\ \hline
-How often do indirect conflicts occur in the early stages of a project’s development? & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
-How often do indirect conflicts occur before the first release of a project? & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
-How often do indirect conflicts occur after the first release of a project? & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
-How often do indirect conflicts occur late in a project’s lifetime? & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
+Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
+Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
+After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
+Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
 \end{tabular}
@@
 
-From the interviews and the surveys combined we can make the deduction that indirect conflicts are more likely to occur early 
-in a projects development cycle as opposed to later. This must be a major concern for any indirect conflict tools moving forward. Since many
-of the tools listen in Section~\ref{sec:intro} implement some sort of filtering system or taxonomy to avoid information overload, this
-finding should be included in future workings. Knowing that indirect conflicts occur more frequently early in a development cycle will let
-these tools know that those reported later in the development life cycle have a higher probability (compared to without this knowledge) of
-being a true positive. This is a major concern when dealing with information overload and to provide developers with what they may find
-relevant to their work. The results in in Table~\ref{tab:often} also suggest that preemptive indirect conflict tools may be more accurate
-towards the mature point of a project, but have potentially more use earlier where the indirect conflicts are more likely to occur.
-While what the ``mature point'' of a project is could be speculated upon, we suggest it is caused for one of two reasons, The
-first would be that the project's source code has stabilized quite a bit in terms of large changes not occurring as much. This could be a result
-of a project being released to a customer or because it is now being used as part of many other projects. The second would be that a change in
-development process has occurred in that source code is not being changed so much as newer code is being added or old code is being deleted. This
-would cause a slow in indirect conflicts as developers interviewed said that changes to code, not the addition of or deletion of code is prone
-to indirect conflicts. In the end, developers survey pointed out that before the ``mature
-point'' indirect conflicts are like to occur daily and weekly (32\% and 18\%) as opposed to after the ``mature point'' when developers
-said they are more likely to occur on a monthly or even yearly frequency (18\% and 12\%).
+In terms of organizational structure, surveyed participants answered that as a project becomes larger and more 
+developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
+occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
+saying they are like to occur in single developer projects.
 
-Of those surveyed that could not give direct answers to the questions in Table~\ref{tab:often}, the more common answers were 
-as follows: ``People rushing to meet [the release date] is where most of my indirect conflicts occur.'', ``Indirect conflicts
-after a release depends on how well the project was built at first'', ``They tend to slow down a bit after a major release, 
-unless the next release is a major rework.'', and ``Conflicts continue to occur at roughly the same rate after the first release,
-with spikes during large revamps or the implementation of cross-cutting new features.''. These results can be argued to show that the occurrence
-of indirect conflicts can largely depend on the management, and deadlines of projects. These are important factors to consider when trying
-to define a ``mature point''. A large refactoring or rewrite can reset the ``mature point'' of a project and should be considered for
-any taxonomies created for preemptive indirect conflict tools. The ``mature point'' of a project (or multiple mature points) should be
-considered when building a taxonomy for indirect conflict tools.
+\subsection{What types of process or tools are being used by developers in regards to indirect conflicts?}
 
-As per environments, Table~\ref{tab:env} shows which development team environment developers believe to be the most
-prone to indirect conflicts.
+In terms of preventative processes used for indirect conflicts, 3 major components were found.
+First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or understand
+when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
+contract when possible, and that an object's contract is used as a type of documentation towards awareness of the 
+software object. One developer stated that ``design by contract was invented to solve this problem and it does it 
+quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't 
+mean that problems don't still occur in practice.
+Second, 21\% of interview developers mentioned that the add and deprecate 
+model is used to prevent indirect conflicts once
+the project, feature, or component has reached a stable or mature point.
+Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
+while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to 
+use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
+Lastly, pure developer experience was mentioned with 37\% of developers mentioning that when planning code changes, 
+either a very experienced member of the project was involved in the planning and has duties to foresee any 
+indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect 
+conflicts will occur while implementing.
 
-\begin{table*}[tb!]
-\begin{center}
-\begin{tabular}{| p{7cm} | c | c | c | c | c |}
-\hline
-Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
-\hline
-\hline
-Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
-Developing in a team between 2 - 5 developers (Inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
-Developing in a multi team environment (Inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
-\end{tabular}
-\end{center}
-\caption{Results of survey questions to development environments in which indirect conflicts are likely to occur, in terms of percentage
-of developers surveyed.\label{tab:env}}
-\end{table*}
-
-From Table~\ref{tab:env}, is can clearly be seen that as a project becomes larger and more and more developers are added, even to the point
-that multiple teams have to be formed, indirect conflicts become more likely to occur. However, this is not to say that indirect conflicts
-do not occur at the lower number of developer level. Even when developing individual projects, 43\% of developers agreed or strongly agreed
-that indirect conflicts are likely to occur. Once more than one developer is associated with a project, or a project has multiple teams, 
-67\% and 74\% of developers agreed or strongly agreed that indirect conflicts are likely to occur. These results show that indirect conflicts
-seem to be compounded by a multiple human factor found in parallel development. The lack of awareness as to other developers work in a project 
-again comes to light. We would think that developers would leverage the other developer's knowledge in their projects to avoid indirect conflicts
-but as we will see most developers tend not leverage their social connections within a project to help solve indirect conflicts. This could be
-a potential issue of social structure in that developers simply do not have the means or will to talk to their colleagues, or it could be that
-developers heavily leverage the notion of software stigmergy as shown by Bolici et. al~\cite{Bolici:2009}. Creating effective communication around
-technical dependencies between developers and teams through either explicit or implicit means appears to be a large task to accomplish but a
-carsickness task as shown by these result being that static analysis breaks down at these levels.
-
-This finding characterizes that indirect conflicts are a frequent problem in industry. It also shows what environments are more prone to indirect
-conflicts and should be handled with more delicate care than others.
-
-\begin{description}
-	\item[F3\namedlabel{itm:f3}{F3}] Indirect conflicts are largely brought on by the changing of interfaces to software structures. The most common type of structures
-						to cause indirect conflicts are methods/functions and database schemas.
-\end{description}
-
-From the developers interviewed, 9 were currently working with large scale database applications and all 9 listed database
-schema as a large source of indirect conflicts with developers saying ``[schema changes] break stuff all over the place''.
-Out of the 5 developers interviewed that currently work on either software
-library projects or in test, all 5 said that methods or functions were the root of their indirect conflict issues. 
-36\% of interviewed developers mentioned that indirect conflicts occur when an update to an external project, library, or service occurs and the resulting 
-integration fails noting that major releases of libraries can be a large issue with one developer saying 
-``their build never breaks, but it breaks ours''. Some
-other notable artifacts were user interface displays for web development, and full components in component base game
-architecture.
-
-From these results, we can see the evidence that the problemed areas of software architecture in regards to indirect conflicts are
-the interfaces between structures, mainly being methods and database schemas. This is good news for preexisting indirect conflict tools as
-many of them have focused at the signature level of methods as a spot for technical dependency changes. However, there has been little focus on indirect
-conflicts at the relation database schema level~\cite{Maule:2008:IAD}, possibly due to the inherit difficult nature
-of analysis of a database schema in relation to source code structure. Developers have given researchers an area of indirect conflicts that have been
-over looked and we should be making an effort to provide answers in this area. Also, note here that developers while in F1 showed that changes to software
-object contracts were the concern, the focus here was shifted to the interface. While the interface does make up a large component of a software object contract,
-it does not fully encompass it. Precondition and postcondition checks, which are part of the object contract are performed in the body of a method and are largley overlooked
-by indirect conflict tools. This may be a source of many false negatives from these tools.
-
-Instances of upgrading libraries to newer versions was a common cause of indirect conflicts, and since these may be binary files,
-static analysis of the changing code is not always possible. Developers who were also
-tasked with updating older libraries had no knowledge of what other outside projects used these libraries and were therefore unable to understand
-how they may affect these outside projects with their changes. However, it appears that when interfaces of methods change through software
-evolution, indirect conflicts are more prone to occur. This result again shows that even at the library or binary level, any type of support tools for indirect
-conflicts should focus at the interface level of change.
-
-\subsection{What types of process and tools are being used by developers in regards to indirect conflicts?}
-
-The findings that follow pertain to software development process and its relation to indirect conflicts. We explore
-how developers react to indirect conflicts, which stakeholders are responsible for their fixes, and if any development
-processes can be used to prevent or catch indirect conflicts.
-
-\begin{description}
-	\item[F4\namedlabel{itm:f4}{F4}] Developers use a combination of object contract understanding, the add and deprecate model, as well as expert 
-																	 knowledge in order to prevent indirect conflicts.
-\end{description}
-
-Of the developers interviewed, 3 main points of interest were found in preventing indirect conflicts. The first was
-the use of ``Design by Contract''\texttrademark, meaning that software objects have certain guarantees about their behavior
-in terms of input, output, and internal workings. This methodology can be used as a type of documentation towards awareness of
-a software object. One developer stated that ``design by contract was invented to solve this problem and it does it quite well'', 
-while another noted that software object contracts do solve the problem in theory, but that doesn't mean that problems don't 
-still occur in practice.  Since ``Design by Contract''\texttrademark ensures that software artifacts, while they can be changed, 
-should always behave in a way that is understood by other developers, languages should just directly support ``Design by Contract''\texttrademark,
-however, this is not the case. One developer said that ``is it too much of a burden'' in regards to strictly making the contract
-apart of the source code. But why is it not such a burden is write separate unit tests that are normally designed around testing
-the contract? There is obviously a trade off between having rules built into the source versus having external tests to check if rules
-were implemented correctly. The question should become which is more effective for software development in terms of bugs produced 
-from contract breaks or bugs produced from convoluted code. This paper cannot begin to attempt to answer this question.
-
-The second point was the add and deprecate methodology. 21\% if developers interviewed mentioned 
-that the development process changes once a stable point of a project is met and the ``add and deprecate model [prevents] a lot 
-of these issues''. Add and deprecate meaning instead of editing code, the developer simply clones code (if needed) and edits the clone
-while slowly phasing out the out code in subsequent releases or other development cycle needs. This allows a stoppage of indirect conflicts
-by allowing developers to use the older versions of software objects which remain unchanged. From the interviews, it seems this is a largely
-effective way of preventing indirect conflicts, however, it was mostly used for libraries and methods only. In some cases such as development
-in a large database project, the add and deprecate model simply does not work as all code must run off the newest changes. So while the
-add and deprecate model is certainly a good fix for heavily version-ed software, especially when given to 3rd party developers, it does
-not necessarily work on internal projects which need to cooperate on which versions of methods or schemas they use.
-
-Lastly, pure developer experience was mentioned
-as a source to prevent indirect conflicts. 37\% of developers mentioned that when planning code changes, either a very experienced member
-of the project was involved in the planning and has duties to foresee any indirect conflicts that may arise, or that while implementing
-a code change developers must use their personal knowledge to predict where indirect conflicts will occur. While most interviewed developers
-had such an expert in their software projects, they admitted to not being able to foresee all indirect conflicts that may occur, especially
-when talking about external teams or projects who may or may not use their final product. This raises the question of how much planning
-by expert developers should really go into avoiding indirect conflicts? Dromey~\cite{Dromey:2003} raises the debate of prevention versus cure of software
-defects and how difficult a problem it is to measure. Should developers be using certain tools to plan their changes or should they be even
-planning at all? Which method provides the best software quality for the minimum effort. These are not easy questions to answer but may
-be key in improving developer productivity when concerning indirect conflicts.
-
-To confirm our findings, the developers surveyed who could give positive identification to preventative measures taken to avoid indirect conflicts (37 developers
-total), 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of ``Design
-by Contract'' or the testing of a methods contract (which could be viewed as catching indirect conflicts), and 14\% said that add and
+Of the the 37 developers surveyed who could give positive identification of preventative processes for indirect conflicts,
+27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
+by contract or the testing of a methods contract, and 14\% said that add and
 deprecate was used in their projects to avoid indirect conflicts.
 
-Even with proper process and expert knowledge, it is still improbable to predict every indirect conflict, and thus, conflicts 
-continue to occur. However, developers did not seem overly concerned that these three items do not totally prevent indirect conflicts. This
-is because most developers ascribe to the idea of ``I work until something breaks'', or taking the curative approach. This shows that developers are somewhat content with
-the level of prevention that happens for indirect conflicts, however, this could be an issue of developers simply following old standards.
-It seems that the cure approach to indirect conflicts is heavily used in industry, but is it better than prevention?
-The lack of preventative process being used is probably due to the difficult nature of predicting indirect conflicts. These results have mostly
-shown that developers would rather spend their time actually writing the software than trying to prevent only potential issues. Time is
-more effectively spent, in the developer's eyes, developing and fixing problems as needed than putting excess time into the prevention of potential errors.
-This is an important consideration for moving forward with indirect conflict tools and research.
+In regards to catching indirect conflicts, 69\% of interviewed developers mentioned forms of testing (unit, and integration) 
+as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
+tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
+when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
+not all indirect conflicts should be caught. 31\% of developers said build processes (either nightly builds or building the project 
+themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
+time testing. The surveyed developers confirmed these results with 49\% mentioning forms of testing as the major tool used to 
+catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
+Surveyed developers also mentioned review process and personal expertise as factors of catching indirect conflicts.
 
-\begin{description}
-	\item[F5\namedlabel{itm:f5}{F5}] Developers use unit and integration testing to examine use case coverage in order to detect indirect conflicts.
-\end{description}
-
-Of the 19 developers interviewed, 69\% mentioned forms of testing (unit, integration, etc) as the major component of catching indirect
-conflict issues, subscribing to the idea of ``run the regression and integration tests and just see if anything breaks''. 31\% of developers 
-said build processes (either nightly builds or building the project themselves), and others mentioned code reviews
-while those dealing with a user interface mentioned user complaints from run time testing. It appears that most indirect conflicts are currently
-detected by using what some would call proper software development practices. The words ``use case coverage'' were constantly being used by developers
-when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. This leads us to believe that as long as the contract of a software structure is understood,
-proper tests should be able to be written which will catch most if not all indirect conflicts. This is another example of developers subscribing
-to the work until something breaks philosophy in that if proper tests are written, the errors will be caught after they break something.
-However, problems occur when a software contract is
-not properly understood and use cases are not tested for or examined. Once an object is not properly
-tested, the negative consequences may not be caught until later in a project with the potential for more damage to be caused. This is the major
-negative result of work until something breaks. The break detectors need to be properly written. Some languages
-such as Eiffle and the Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}
-enforce object contracts to be well understood through the ideas of preconditions and postconditions of structures which is an example of preemptive
-conflict detection. However, the trade off between external tests and in-line contracts is still in place as previously mentioned.
-
-Of the 78 developers survey, 49\% mentioned forms of testing as the major tool used to catch indirect conflicts, 33\% said build processes,
-while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts. Other developers surveyed also mention code review process (14\%)
-and project expertise (6\%) as factors of catching indirect conflicts. 
-Developers said that the build process is simply there to ensure a, at least, semi-functioning product. They use
-the build process more to catch compile time errors witch prevent the product from being used. These types of build breaking indirect
-conflicts do occur but for the most part developers posses tools capable to build their solutions in order to catch these problems before
-the code change is added to the project. Therefore, unit and integration testing appears to be the largest factor in catching indirect conflicts
-as long as they are properly written and contracts are understood.
-
-\begin{description}
-	\item[F6\namedlabel{itm:f6}{F6}] Developers prefer to spend time debugging real problems opposed to preventing potential problems.
-\end{description}
-
-Once an indirect conflict has occurred, usually alerted to the developer by test or build failures, 75\% of developers interviewed said
+Once an indirect conflict has occurred and developers need to resolve it, 75\% of developers interviewed said
 they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
@@
 code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
-frames of when events occurred. Developers also mentioned their personal expertise in the project as a large component of solving indirect
-conflicts as well as issue or work item trackers as many tests and builds can be linked to these management tools.
+frames of when events occurred. Of the developers surveyed, 23\% said they used native IDE tools and 21\% said they use 
+features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers 
+mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading 
+of formal documentation.
 
-Here, the generic approach
-brought forth by developers was seen to be history checking. Developers would look into source code change logs and build or test failure
-logs in order to see what is broken and what might have changed recently that could be affecting the break. This gives the impression that
-developers already have the tools to solve indirect conflicts. However, this leads to the question: how quickly and
-efficiently are they able to solve them and is the time spent debugging outweighing the potential time spent preventing such errors?
-While each developer gave some examples on solving indirect conflicts and the times needed, they all agreed that to be able to solve them
-quicker would be a huge benefit. This again shows the developer preference towards work until something breaks as opposed to prevent the
-error in the first place. While the debate of prevention versus cure goes on, possible solutions as to solving
-indirect conflicts quicker are the automatic analysis of source code
-change logs to detect what changes will affect a certain part of a project's code. Luckily, some of this solution has already been implemented
-by several researchers in both delta debugging~\cite{Zeller:2002:ICC} and program slicing~\cite{Weiser:1982:PUS}. The fact that researchers
-have had such success in the cure or debugging side of indirect conflicts could be due to the lack of indirect conflict. Once a problem has
-occurred, there is no chance of it being a false positive. Academics do not have to create taxonomies to filter results and developers do
-not need to spend time working on only potential issues. Their time is more effectively, in their eyes, spent developing than speculating.
+%Very important
+Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
+most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
+approach. This means that while developers do have processes and tool for prevention, they would rather spend their time
+at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
+while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.
 
-Of the developers surveyed, 23\% also said they used native IDE tools and 21\% said they use features of the language's compiler and debugger in order
-to solve indirect conflicts. Interestingly, only 13\% of developers mentioned a form of communication with other developers in aid to
-solving these conflicts and only 4\% mentioned the reading of formal documentation. This shows is the serious lack of static analysis tools
-available for developers in debugging as many projects and languages simply do not support it (adding cause for non static analysis tools moving
-forward), but more importantly, that developers tend to not communicate or use formal documentation for debugging indirect conflicts. Petre~\cite{Petre:2013:UP},
-as already mentioned, showed that it is not industry practice to use UML, which could open to floor to more informal means such as online forums. The none
-developer communication, as again already mentioned, could be addressed by either more formal communication structures or by developer stigmergy 
-in their projects if needed.
+\subsection{What do developers want from future indirect conflict tools?}
 
-\subsection{How should indirect conflict tools work to best benefit developers?}
-Ultimately, the goal of this research would be to produce any sort of tool that could aid developers in any stage (prevention, detection, or
-solving) of indirect conflicts. In order to address what types of tools would be best, we analyze the results from our previous research
-questions along with the proceeding results.
+When asked about preventative tools, interviewed developers had major concerns that the amount of false positives
+provided by the tool which may render the tool useless. Developers said ``this would
+be a real challenge with the number of dependencies'', ``it depends on how
+good the results are in regards to false positives'', and ``I only want to know if it will break me'', meaning that
+developers seem to care mostly about negative impacts of code changes as opposed to all impacts in order to reduce 
+false positives and to keep preventative measures focused on real resulting issues as opposed to preventing potential
+issues. Overall, developers had little interest in preventative tools or processes.
 
-\begin{description}
-	\item[F7\namedlabel{itm:f7}{F7}] It is extremely difficult, if not impossible, to create a preemptive indirect 
-																	 conflict tool which can detect indirect conflicts with a low rate of false positives. Preemptive tools 
-																	 should be exploratory in nature.
-\end{description}
+In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
+are already in place to catch potential issues such as testing, code review, 
+individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
+want to be notified about method signature changes as they have a high chance to break the code as opposed to only
+23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
+interface change. Other change types varied from never being notified to most times being notified, showcasing the
+complexity of change types which may or may not negatively affect a project.
 
-The developers who were interviewed had a major concern with any preemptive tool in that the amount of false positives
-provided by the tool may render the tool useless. Developers said ``this could relate to information overload'', ``this would
-be a real challenge with the number of dependencies'', ``I only want to know if it will break me'', and ``it depends on how
-good the results are in regards to false positives''. These concerns combined with RQ1 suggest that preemptive tools are
-best used in stable phases on the project when indirect conflicts are less likely to happen, thus reducing the number of false
-positives reported. One developer in particular noted ``You will spend all your time getting notified and reviewing instead of implementing''.
-Interviewed developers also suggested that their proper software development process are already in place to
-catch potential issues before they arise such as code review, individual knowledge, IDE call hierarchies (static language analysis),
-or communication within the project. All of this again shows that developers would prefer to subscribe to the 
-``work until something breaks'' philosophy. Again, it is up for debate whether or not this is the correct approach the developers should
-be taking (effectiveness of prevention versus cure), but it is the one they currently prefer.
-
-As per what developers would want from such preemptive tools, Table~\ref{tab:pre} shows how developers feel towards being 
-alerted in certain types of functional changes within code which they are using.
-
-\begin{table*}[tb!]
-\begin{center}
-\begin{tabular}{| p{7cm} | c | c | c | c | c |}
-\hline
-Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
-\hline
-\hline
-Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
-Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
-Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
-User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
-Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
-Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
-\end{tabular}
-\end{center}
-\caption{Results of survey questions to source code changes that developers deem notification worthy, in terms of percentage
-of developers surveyed.\label{tab:pre}}
-\end{table*}
-
-Developers were very quick to identify this problem which has
-been a common occurrence among any preemptive indirect conflict tool built. The main issue with false positives is the detection of changes
-which frequently cause indirect conflicts. Change types are very possible to detect through software static analysis~\cite{Fluri:2007:CDT} (when possible), 
-but even those that are
-prone to causing indirect conflicts do not cause them all the time. This is quite evident in Table~\ref{tab:pre} where developers identified
-that there is an even distribution of software changes that could cause indirect conflicts and as to if they would want to be made aware of
-those changes. The only two exceptions which stand out are that of method signature changes and main algorithm changes. The problem here however,
-is that method signature changes will be caught by IDE errors, nightly builds, unit testing, integration testing and so on, while main algorithm
-changes are very difficult to detect. This all leads to the hypothesis that preemptive indirect conflict tools cannot be built without a very
-large number of false positives being reported at any given time. Even if developers, as per Table~\ref{tab:pre}, want to be alerted to 
-changed preconditions in methods most times if not always, there is still a large chance that that changed precondition will actually not
-introduce an indirect conflict in the source code. This leads to a large number of false positive being reported, which leads to information
-overload, and eventually to developers abandoning the tool.
-
-%In order to prevent or catch indirect conflicts with an automated tool without introducing a large amount of false positives, would be to
-%have an automatic understanding of how the project's code is suppose to work. From the halting problem~\cite{Burkholder:1987}, 
-%we know it is an impossible task to write one program that can verify the correctness of another. This leaves us with the only option of some kind of
-%human intervention in the matter which leads us back to use case coverage. As was stated before, if the developer has a complete understand
-%of the use cases of a particular code artifact, then they can write proper tests to ensure those use cases are always met and not broken which
-%would be the best form of preventing and catching indirect conflicts.
-
-Given the above challenges, some indirect conflict tools have had some general success. The key difference with these tools is that they
-contain the idea of developer exploration. The idea of only presenting information to users on request is the complete opposite of the preemptive
-notion of notifying developers to potential upcoming problems. Here, the idea that a user can provide some sort of
-input, a sort of context, is a powerful tool for reducing false positives. We have really just moved the false positives of the preemptive
-tool to the ``search results'' of the exploratory tool, however, with the user input the results can be narrowed down to provide a larger
-precision of results given the context. These types of tools stick to previous findings being developers working until something breaks, then
-requesting information to fix it.
-
-\begin{description}
-	\item[F8\namedlabel{itm:f8}{F8}] Indirect conflict tools should take a curative approach towards indirect conflicts as opposed to preventative
-																	 in order to better fit developer activities.
-\end{description}
-
-The developers who were surveyed, the following list represents the more common suggestions for how an indirect conflict
-curative tools should be implemented.
+When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
+Surveyed developers suggested the following improvements to curative tools:
 
 \begin{itemize}
-	\item Aid in debugging indirect conflicts by finding which recent code changes are breaking a particular area of code or a test.
-	\item Automated testing to show how code changes affect old tests and the ability to automatically write new tests to compensate for the change.
+	\item Aid in debugging by finding which recent code changes are breaking a particular area of code or a test.
+	\item Automatically write new tests to compensate for changes.
 	\item IDE plug-ins to show how current changes will affect other components of the current project.
 	\item Analysis of library releases to show how upgrading to a new release will affect your project.
-	\item Built in language features to either the source code architecture (i.e. Eiffle or Java Modeling Language) or the compile time tools to display
-				warning messages towards potential issues.
-	\item A code review time tool which allows deeper analysis of a new patch or pull request to the project allowing the reviewer to see potential 
+	\item Built in language features to either the source code architecture (i.e. Eiffle or Java Modeling Language) or the compile 
+				time tools to display warning messages towards potential issues.
+	\item A code review time tool which allows deeper analysis of a new patch to the project allowing the reviewer to see potential 
 				indirect conflicts before actually merging the code in.
@@
 
-From the above list, we can see three main areas of focus that researchers of indirect conflicts should spend their time in. The first
-is the debugging process through utilizing logs. Showing which recent code changes affect an area of code or how a commit and break a specific
-unit test are key methodologies that developers want in order to aid their preferred curative approaches. Developers also want to be able to
-understand their software object contracts better or even automatically. Detections of contract changes will play a large part in indirect conflict
-and dependency awareness moving forward, especially at the method body and database schema levels. These contract detections also play a role in
-the automatic generation of unit tests in order to comply with use case coverage. Finally, and most importantly, developers want tools that
-integrate into their current development styles, that being ``work until something breaks''. While a complete evaluation of the curative approach
-is beyond the scope of this paper, it is how developers act today and  in the end, tools should be created to aid developers.
+\section{Discussion}
+\label{sec:disc}
 

### 2013-11-05 iced-2013/paper.tex
ICED new methodology.

@@
 
-Our grounded theory study was performed in two parts. First, a round of semi-structured interviews were conducted which 
-addressed the 3 research questions as mentioned above. Secondly, a survey was conducted
+We performed a grounded theory study in two parts. First, a round of semi-structured interviews were conducted which 
+addressed the 3 research questions as mentioned above. Second, a survey was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
@@
 \textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
-the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions and
+the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
 making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
@@
 questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
-both interviews and surveys (when open ended questions were involved) was analyzed using open coding. Open coding involves
+both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
 assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
@@
 theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in 
-findings~\ref{itm:f1} -~\ref{itm:f8}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+Section~\ref{sec:disc}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
 3 criteria that are commonly used in evaluating grounded theory studies.
@@
 We selected our interview participants from a large breadth of both open and closed source software development companies
-and projects. The population which participated in our interview came from the following software groups: 
+and projects as well as Agile and Waterfall based projects.
+The population which participated in our interview came from the following software groups: 
 IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
 Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
-Our participants we chosen based on their direct involvement in the actual writing of software for
+Our participants were chosen based on their direct involvement in the actual writing of software for
 their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
 with an average of 8 years of experience.
-In addition to software development, some participants were also chosen based on both their experience of software development 
-as well as their experience with project management at some capacity.
+In addition to software development, some participants were also chosen based on their experience with project management
+at some capacity.
 
@@
 
-Participants were invited to participate in the interviews by email and were sent a single reminder email one week
-after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
+Participants were invited to be interviewed by email and were sent a single reminder email one week
+after the initial invitation if no response was made. We directly emailed 22 participants and ended up conducting
 19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
@@
 
-Interview participants first answered a number of structured demographic items. Next, participants were 
+Interview participants first answered a number of structured demographic items, followed by being  
 asked to describe various software development experiences regarding our three research questions.
-Our three questions were studied by having the interview participants talk about their experiences and opinions in the following
-\totalCategories{} semi-structured research topics:
-
-\begin{itemize}
-\item Examples of indirect conflicts from real world experiences.
-\item Software artifact dependency levels and where conflicts can arise.
-\item Software development tools for dependency tracking and awareness.
-\item Software development process for preventing indirect conflicts.
-\item How developers find internal or external software dependencies.
-\item How indirect conflicts are detected and found.
-\item How indirect conflict issues are solved or dealt with.
-\item Developer opinion of preemptive measures to prevent indirect conflicts.
-\item Developer opinion on what types of changes are worth a preemptive action.
-\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
-\end{itemize}
+Our three questions were studied by having the interview participants talk about their experiences and opinions in
+\totalCategories{} semi-structured research topics.
 
@@
 being apart of any one software project for extended periods of time. We account for these non answers
-in our analysis and results as seen in Section~\ref{sec:results}. Interviews lasted 
-from 15 minutes up to 75 minutes and were very much dependent on the experience of the participant.
-
-%After the interviews, we extracted a list of answers, opinions, and experiences for each of the \totalCategories{} categorical
-%question fields. We combined commonly mentioned opinions and answers where possible, and noted where experiences
-%with indirect conflicts from participants were similar but generally kept the experiences tied to the individual.
-%Where more direct answers were given, we used the wording of our participants rather than our own descriptive 
-%aggregations. The results of this preliminary analysis was a list of short hypotheses supported by direct developer
-%experience and opinion for the \totalCategories{} categories. These hypotheses were then used as a template for creating the
-%second step survey, in order to further test the answers given by interview participants and validate any possible
-%conclusions to be drawn.
+in our analysis and results as seen in Section~\ref{sec:results}.
 
@@
 Subnet Solutions, and Amazon.
-However, participants who took part in the round of interviews were asked to act as a contact point for other 
+Participants who took part in the round of interviews were asked to act as a contact point for other 
 developers in their team, project, or organization who may be interested
-in completing the survey. Aside from this aforementioned list, two groups of developers were asked to
+in completing the survey. Aside from this aforementioned list, two other groups of developers were asked to
 participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
@@
 
-The created the survey was based off of categorical hypotheses created by the round of interviews. The survey
+The created survey~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
+was based off of hypotheses created by the round of interviews. The survey
 was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
@@
 piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
-with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed the previously interviewed developers to examine what hypotheses were
-formed and what we would be moving forward with. The final survey consisted of:
-2 multiple choice questions for demographic information; 3 level of agreement questions for an indication of
-what types of environments indirect conflicts are more likely to occur;
-6 level of use questions to indicate what types of software changed developers find trouble some with respect to
-indirect conflicts; and 9 short answer questions for indication of when indirect conflicts occur, what types of
-processes are used to prevent or react to indirect conflict, and to provide an outlet for general opinion
-on the matter. Each non demographic question was made optional as it was shown 
+with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
+the previously interviewed developers to examine what hypotheses were
+formed and what we would be moving forward with.
+Each non demographic question was made optional as it was shown 
 through the interviews that some questions require more experience from participants than may be provided.
-
-%After the survey was closed, preliminary analysis was applied to the short answer questions in the same manor
-%as the interview responses. We extracted a list of answers, opinions, and experiences for each of the 9 short
-%answer question fields while combining commonly mentioned answers where possible. The results again from this
-%preliminary analysis is a list of common and unique hypotheses to each of the 9 short answer questions.
 

### 2013-11-06 iced-2013/paper.tex
ICED new discussion and implication for tools.

@@
 	\item Analysis of library releases to show how upgrading to a new release will affect your project.
-	\item Built in language features to either the source code architecture (i.e. Eiffle or Java Modeling Language) or the compile 
+	\item Built in language features to either the source code architecture (i.e. Eiffle or 
+				Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}) or the compile 
 				time tools to display warning messages towards potential issues.
@@
 \label{sec:disc}
+
+\textbf{Unwanted Awareness} As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
+sporadically, and very greatly from case to case. These conditions pose very large issues for the creation of generalizable theories
+or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
+to proposed or implemented tools as discussed in Section~\ref{sec:related}. These complexities are what we believe to be the
+leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
+information being presented to them, rendering the system useless. These false positives are caused by a disjoint in what
+developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
+be seen, is cause by developers only considering events which cause some action to be taken on their part, to be true positives,
+where as awareness theory would state any event which is related to an individual's 
+work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
+
+Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
+their project awareness. To be able to fully understand a developer's awareness intuition, we can look to the common statement of
+developers only wanting to know about an event if it affects them by forcing some sort of action on their part. In a sense, developers
+don't care about what they don't need to account for. This sense of unwanted awareness is crucial to understand why generalized
+awareness techniques of difficult to generalize problems, such as indirect conflicts, often encounter difficulties. 
+Developers simply do not want to know about events which effect them, but require no action on their part.
+
+This unwanted awareness, or limited awareness, seems counter intuitive to classical awareness theory which would state that
+being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
+but as it was found in practice, this is an incorrect assumption. Since software developers have limits on their time, awareness of
+all events surrounding a developer's work or project is not possible. Developers prefer to spend their limited time dealing with
+the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
+of events occurring around their work which pose no direct threat to the consumption of their time.
+
+Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
+developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
+them initially, we must consider the pro and cons of adding this, for now, unwanted awareness. A pro of total, or near total
+awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
+with higher quality or understanding of the product. 
+A con would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
+producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
+however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
+be more productive.
+
+\textbf{Prevention versus Cure} We have seen through the results, that developers posses both tools and practices for the prevention,
+detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
+that developers tend to prefer workings on real issues that have already occurred as opposed to preventing issues of the future which
+may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
+is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
+
+Dromey~\cite{Dromey:2003} raises the debate of prevention versus cure in software development and how difficult a problem it is
+to measure. Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
+place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
+them in the first place.
+
+Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
+accomplished essentially by
+not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
+developers are spending little to no time in prevention. Developers do however spend a large amount of time in detection and cure through the
+writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
+conflicts occurred at either the detection or cure levels. Obviously, developers either prefer to spend their time in curative
+measures, or do not posses the proper tools to take better action in the preventative stages.
+
+The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
+Even with an experienced system architect, it was shown that identifying dependencies and notifying those involved is a daunting task 
+which is ignored more than dealt with. The second being the knowledge of when a dependency will fail also requires vast knowledge
+of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. The
+abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
+real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
+this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
+curative measures are more productive than preventative measures.
+
+The pros and cons of prevention versus cure are similar to those of unwanted awareness in that they result in a trade off 
+of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
+it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
+not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
+to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
+Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.
+
+A last interesting observation about current curative measures being taken in industry, is that developers view testing of software 
+are curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
+are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
+thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
+However, if a test never fails, could it not be said that it is preventing changes from causing it to fail rather than detecting
+when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
+measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative.
+
+This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
+designers, architects, or managerial stakeholders. 
+From the project manager's interviewed it was shown that they are heavily favored towards planning and 
+prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
+approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
+is being viewed in research.
+
+%mature point, static analysis
+\textbf{Gap in Evolution Analysis} As was see in Section~\ref{sec:results}, we know that indirect conflicts are more likely to occur,
+from developer opinion, before a mature point is reached in the project's evolution. However, this context as well as any other
+potential contextual attributes are rarely identified or accounted for in regards to indirect conflicts. Developers have given
+us a clue into the idea that different change types may occur at different rates throughout a project's life time and that this
+may drastically effect the outcomes of indirect conflict tools or processes.
+
+Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
+changes occur at a given point in time. This type of static analysis may be useful in regards to indirect conflicts in that we
+could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
+means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding. This sort
+of deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
+research than may have been previously thought.
+
+This added level of context, through the notion of a mature point, 
+would add to the prevention versus cure debate as previously discussed. Prevention may be a better
+choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
+to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
+causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
+and the mature point in order to fully understand the nature of indirect conflicts and their context.
+
+While this added sense of understanding is important to indirect conflicts, contextual attributes of a project's current progress
+or measures of performance are often more synonymous with project management
+rather than software developers themselves. An understanding of a project's evolution, a mature point being reached as an example,
+may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
+conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.
+These understandings may lead to the detection of indirect conflict issues quicker than traditional means.
+
+Aside from context, we have show that dependency identification and tracking is a key missing component of
+indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
+organizational structures of software teams. Between increased modularization (multiple sub projects or repositories), 
+cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
+become quite limited in identifying and tracking dependencies where they were once strong.
+
+Through our results, developers have said that relational database schemas are one of the highest sources of indirect
+conflicts found in their projects, yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
+been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
+dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.
+
+These increases of complexity in software products has left quite a gap in dependency identification and tracking which 
+has lead to some of the deficiencies of indirect conflict research.
+
+\subsection{Implication for Research}
+\label{sec:implr}
+
+\subsection{Implication for Tools}
+\label{sec:implt}
+
+The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
+developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
+said, if researchers wish for their tools to be largely adopted among current developers, a focus at these two stages
+is crucial for success. It should be noted here that detection must come with an almost zero rate of false positives as
+current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
+
+The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
+tools are used for indirect conflicts alone (automatic debugging, unit tests, etc), most of these tools
+have need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
+automatically written for a given software object's contract, the ability to find a change  in an external project
+which has broken a developer's own project, or any automation of the existing detection and resolution stages of indirect
+conflicts are what developers currently seek. But of course, most of these implications rely on the improvement of 
+static analysis tools.
+
+This research, in regards to tools, most importantly implies the further development of static analysis tools. Static 
+analysis lays at the heart of most if not all stages of indirect conflict research. We must be able to track and manage
+software dependencies across the new landscape of software development that is multiple projects, repositories, and cross
+language support. These improvements will allow the further development of both current and future indirect conflict
+tools.
 

### 2013-11-06 iced-2013/paper.tex
ICED new implications for research section.

@@
 	\item Analysis of library releases to show how upgrading to a new release will affect your project.
-	\item Built in language features to either the source code architecture (i.e. Eiffle or 
+	\item Built in language features to either the source code architecture (i.e. Eiffel or 
 				Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}) or the compile 
@@
 
+The largest implication for future research found in this paper is the need for continued study of the open question
+``Which of prevention or cure is more effective for software development?''. This simple question will undoubtedly 
+produce extremely complex answers. With software developers focusing their current efforts on curative measures suggests
+that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
+least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
+provide more insight into possibilities, as its answer may determine many future actions taken by the research community.
+
+Recommendations towards the study of prevention versus cure involve the examination of formal processes and tools used
+by industry professionals with measurements of efficiency and effectiveness, similar to the work of
+Tiwana~\cite{Tiwana:2008:ICD} in coordination tools. With these studies, we may find insight into the correct balance
+of prevention versus cure, thus being able to increase developer productivity as well as identify gaps in theory 
+where new discoveries may be made.
+
+A second implication for research derived by the paper, which is more directly related to awareness, is the continued
+understanding of the disjoint between theoretical awareness and awareness as perceived by industry practitioners.
+As was identified in the previous discussion, while theory suggests that awareness of all events related to ones work
+produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
+to the point where they only want to be aware of events which require action on their part. This disjoint, of what 
+theoretically should be and what physically is, should be further understood to combat future potential failures in
+tool development or theories attempted to be used in practice. These disjoints may be detected by conducting case 
+studies of tools built primarily on theoretical models, while having these tools tested in industry. We found developers
+to be quite honest when using a tool they dislike which should lend itself to the detection of the aforementioned
+disjoints.
+
 \subsection{Implication for Tools}
@@
 The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
-tools are used for indirect conflicts alone (automatic debugging, unit tests, etc), most of these tools
+tools are used for indirect conflicts alone (automatic debugging~\cite{Zeller:2005:WPF}, unit tests, etc), most of these tools
 have need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
@@
 
-Since this paper has covered a wide spectrum in regards to indirect conflicts (preventing, catching, solving, process, and tools), there
+Since this paper has covered a wide spectrum in regards to indirect conflicts (prevention, detection, resolution, process, and tools), there
 exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly

### 2013-11-06 iced-2013/paper.tex
ICED new conclusion and updated evaluation.

@@
 
+As added insurance, we conducted 5 post results interviews with developers to once again confirm our results, and discussions. These
+procedures can be seen in Section~\ref{sec:meth}.
+
 To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:results} to either a majority of agreeing
@@
 {\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
-Although our findings~\ref{itm:f1} -~\ref{itm:f6} may not be entirely novel or even surprising, the combination of these results allow us to discover the
-insightful findings of~\ref{itm:f7} and~\ref{itm:f8} regarding indirect conflict tools. Given how many indirect conflict tools are left with the same common
+Although our results may not be entirely novel or even surprising, the combination of said results allowed us to discover the
+the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
+insight into the debate of prevention versus cure in software development. Given how many indirect conflict tools are left with the same common
 issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
@@
 expressed unique situations that did not fully meet our generalized findings or on going theories. In these cases, we worked in the specific
-cases which were presented as boundary cases and can be seen in quotations in Section~\ref{sec:results}. These quotations add to the variation
+cases which were presented as boundary cases and can be seen in Section~\ref{sec:results} as some unique findings or highly
+specialized cases. These cases add to the variation
 to show how the complexity of the situation also resides in a significant number of unique boundary situations as well as the complexity
@@
 
-Indirect conflicts are a significant issue with real world developers, however, many proposed techniques and tools to mitigate
+Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
 losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
@@
 
-We have shown through findings~\ref{itm:f1} -~\ref{itm:f3} why indirect conflicts occur, when indirect conflicts are more likely to
-occur, as well as what types of software objects are susceptible to these conflicts. Findings~\ref{itm:f4} -~\ref{itm:f6} have shown
-how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings~\ref{itm:f7} 
-and~\ref{itm:f8} have provided a foundation as to why past techniques and tools have had low adoption rates and where
-researchers should focus their current and future efforts in handling indirect conflicts. We hope that this study and its 
-findings will inspire future techniques and tools for dealing with indirect conflicts that 
+We have shown through the findings of RQ1 why indirect conflicts occur, when indirect conflicts are more likely to
+occur, as well as what types of software objects are susceptible to these conflicts. Results of RQ 2 have shown
+how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings of RQ3
+have provided a foundation as to why past techniques and tools have had low adoption rates and where
+researchers should focus their current and future efforts in handling indirect conflicts.
+
+We have brought to attention through discussion the disjoint of theoretical awareness needs versus the practical
+awareness needs found in industry, the debate of prevention versus cure and industries tendency towards curative measures,
+and the gap of analytical evolution tools needed for dependency identification and indirect conflicts.
+
+We hope that this study and its 
+findings will inspire future processes and tools for dealing with indirect conflicts that 
 will aid developers in industry as well as test and validate our theories put forth in this paper as well as spark a potential

### 2013-11-06 iced-2013/paper.tex
ICED new introduction. It is limited to a single page now.

@@
 which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
-mitigates some downsides of parallel development which include the cost of conflict prevention and resolution, however,
-in practice we see these mitigated losses continue to appear quite frequently. Not only do these conflicts still occur,
-but at times they can prove to be a significant and time-consuming chore for developers~\cite{Perry:2001:PCL}.
+mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
+in practice we see these mitigated losses continue to appear quite frequently and can prove to be a significant
+and time-consuming chore for developers~\cite{Perry:2001:PCL}.
 
-Two types of conflicts have attracted the attention of academics over the years, \textit{direct} and 
+Two types of conflicts have attracted the attention of academics, \textit{direct} and 
 \textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
@@
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. However, indirect conflict tools have not shared the same success. Indirect conflicts
-can arise in source code for such reasons as having one's own code negatively affected by a library upgrade,
-making a code change only to find out it had negative effects on unit or integration test, or having negative effects
-on the way other developer's have interpreted the use of your code. Tools that have attempted to either
-help developers prevent or solve indirect conflicts have either attracted little developer interest or have
-had some common struggles which remain to be studied or solved.
+positive developer feedback. Indirect conflicts involve source code change influences such as making a change which
+causes an error in another location in the system, having code negatively affected by a library upgrade, or even
+breaking an integration test of a separate component than the one changed. Indirect conflict tools however, have
+not shared the same success as direct conflict 
+tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.
 
-Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
-when those changes will create negative consequences in a software project.
-Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
-and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
-one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
-and create a taxonomy for different
-types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
-creates an environment where developers can see how source code changes will affect other areas of a project at the 
-method level and thus where indirect conflicts may occur. 
-Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
-utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
-an environment to see what areas of a project are being affected by a source code change. 
-Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
-or providing redundant information to preexisting tools. 
-Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
+While indirect conflict tools have shown potential from developer studies, some of the same problems continue
+to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
+overload. Through case studies, developers have noted that current indirect conflict tools provide too many 
+false positive results leading to information overload and the tool eventually being
+ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
+tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
+invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
+identification of failure inducing changes, other than those which are already identifiable by other means such
+as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
+also compounded by the scale of many software development projects leading to further information overload.
 
-Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
-times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
-However, socio-technical congruence is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+Human factors have also shown to cause issues as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence has been leveraged in several indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
+Factors again of information overload, human dependencies, and scalability lay at the root of these problems.
 
-We are interested in exploring three questions in the hopes of resolving some of the issues that are occurring
-with indirect conflict tools:
+%Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
+%when those changes will create negative consequences in a software project.
+%Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
+%and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
+%one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
+%and create a taxonomy for different
+%types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
+%creates an environment where developers can see how source code changes will affect other areas of a project at the 
+%method level and thus where indirect conflicts may occur. 
+%Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
+%utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
+%an environment to see what areas of a project are being affected by a source code change. 
+%Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
+%or providing redundant information to preexisting tools. 
+%Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
+
+%Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+%times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+%developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
+%However, socio-technical congruence is largely unproven in regards to its correlation to
+%software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+
+In the hopes of resolving some of the aforementioned issues surrounding indirect conflicts and their tools, we 
+are interested in exploring the following three research questions:
 
@@
 We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
-developers in order to answer the aforementioned questions. Our findings indicate that while indirect conflicts are very much a real 
-and frequent problem
-in software development, developers already posses the tools to prevent and catch indirect conflicts to a manageable degree and have
-the mindset of working until something breaks, then attempting a fix. Thus, developers would rather spend their time developing instead
-of taking or reading preventative measures, resulting in the need of tools to debug and solve indirect conflicts quickly as opposed to tools 
-which attempt to prevent or catch indirect conflicts early.
+developers in order to answer the aforementioned questions. Our findings indicate that: indirect conflicts are very much a real 
+and frequent problem in software development, contextual patterns of indirect conflicts have been missed in previous tools through
+the lack of dependency identification, that developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, and that developers do not want awareness mechanisms which provide non actionable results.
 
-We continue this paper by first presenting our summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
-then explain and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}.
+We continue this paper by first presenting our methodology used to perform this study in Section~\ref{sec:meth}. We
+then present and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}
+and Section~\ref{sec:disc}.
 Next, an evaluation of how our study was performed in terms of the credibility
@@
 theoretically should be and what physically is, should be further understood to combat future potential failures in
-tool development or theories attempted to be used in practice. These disjoints may be detected by conducting case 
-studies of tools built primarily on theoretical models, while having these tools tested in industry. We found developers
+tool development or theories attempted to be used in practice.
+
+These disjoints may be detected by conducting case 
+studies of tools built primarily on theoretical models, while having said tools tested in industry. We found developers
 to be quite honest when using a tool they dislike which should lend itself to the detection of the aforementioned

### 2013-11-06 iced-2013/paper.tex
ICED cleaned up abstract to represent new changes.

@@
 on either \textit{direct} or 
-\textit{indirect conflicts} in regards to prevent, catching, or debugging these conflicts as they arise
+\textit{indirect conflicts} in regards to prevent, detect, or resolve these conflicts as they arise
 through source code changes. While the techniques and tools for direct conflicts have had
@@
 we performed a grounded theory study with 97 developers involved in either
-interviews or a web survey. We found that: indirect conflicts are a real and frequent problem,
-developers already posses a large suite of tools and processes which can prevent and catch
-indirect conflicts at a manageable scale, and that developers have a work until something breaks philosophy
-which leads to the lesser want of prevention tool as to the want of quick and perhaps automatic debugging tools
-for indirect conflicts.
+interviews or a web survey. We found that: indirect conflicts are a real and frequent problem in
+software development, that developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, that developers do not want awareness mechanisms which provide non actionable results,
+and that there exists a gap in software evolution analytical tools from the reliance on static analysis.
 
@@
 
-%Difficulties regarding indirect conflicts are to determine when software dependencies change and more crucially,
-%when those changes will create negative consequences in a software project.
-%Sarma et al.~\cite{Sarma:2007:TSA} built Palantir, which can both detect potential indirect conflicts, at the class signature level,
-%and alert developers to these conflicts. Holmes et al.~\cite{Holmes:2010:CAR} take it
-%one step further with their tool YooHoo by detecting fine grained source code changes, such as method return type changes, 
-%and create a taxonomy for different
-%types of changes and their proneness to cause indirect conflicts. The tool Ariadne~\cite{Trainer:2005:BGT} 
-%creates an environment where developers can see how source code changes will affect other areas of a project at the 
-%method level and thus where indirect conflicts may occur. 
-%Another indirect conflict tool, CASI~\cite{Servant:2010:CPI},
-%utilizes dependency slicing~\cite{Bajracharya:2009:SIS} instead of method call graphs to provide 
-%an environment to see what areas of a project are being affected by a source code change. 
-%Most of these tools have all shown to have common difficulties of: scalability, false positives, information overload, 
-%or providing redundant information to preexisting tools. 
-%Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
-
-%Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
-%times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-%developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
-%However, socio-technical congruence is largely unproven in regards to its correlation to
-%software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
-
 In the hopes of resolving some of the aforementioned issues surrounding indirect conflicts and their tools, we 
@@
 developers in order to answer the aforementioned questions. Our findings indicate that: indirect conflicts are very much a real 
-and frequent problem in software development, contextual patterns of indirect conflicts have been missed in previous tools through
-the lack of dependency identification, that developers tend to prefer to use detection and resolution processes or tools
-over those of prevention, and that developers do not want awareness mechanisms which provide non actionable results.
+and frequent problem in software development, that developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, that developers do not want awareness mechanisms which provide non actionable results, 
+and that there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
+context of indirect conflicts.
 

### 2013-11-07 iced-2013/paper.tex
ICED new related work.

@@
 exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
-in the notion of indirect conflicts, lessons learned from topic in awareness, preemptive direct conflict detection, and debugging can be used.
+in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging can be used.
+
+Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
+which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
+Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
+al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
+such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
+indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
+detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
+creates an environment where developers can see how source code changes will affect other areas of a project at the 
+method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
+been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
+needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
+instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
+Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
+which were explored in this paper. Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
+
+Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
+Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
+artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
+created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
+Socio-technical congruence however, is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
 
@@
 interfere with their own code changes, and whether their code is impacted by a code change. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way.  Kim
+found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
+require action on the developer's half.  Kim
 also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
@@
 
-In a case study of impact awareness de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
 code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
@@
 
-In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards false positive output
-as well as developer process integration. Developers complained that static analysis tools usually produce a large amount of false
-positives to the point where output is ignored, as well as the tool not being designed properly to fit their current development work
-flow. In our case we found that the current development work flow is ``work until something breaks'' as opposed to try and find where things
-might break before making code changes.
+In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
+ill fit of current static analysis to current development practices. Not being designed properly to fit current development work
+flow, and the over abundance of false positives were shown to be factors of developer non-use.
 
@@
 that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
+Blincoe et al.~\cite{Blincoe:2013:UCC} also found, through data mining techniques, a subset of events related to awareness 
+which occur that developers will communicate about as opposed to all awareness events. This again confirms our understanding that developers
+only care about certain events which require some action on their part.
 

### 2013-11-07 iced-2013/paper.tex
ICED evaluation is in. Not sure how much to write there.

@@
 Following our data collection and analysis, we re-interviewed some of our initial interview participants
-in order to validate our findings. We confirmed our hypotheses as to whether or not they resonate with 
+in order to validate our findings. We confirmed our findings as to whether or not they resonate with 
 industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
@@
 five participants. Those that were re-interviewed came from the range of 5-10 years of software development
-experience. Re-interviewed participants were given our 3 research questions along with results and 9 findings,
-and asked open ended questions regarding their opinions and experiences to validate our findings. The five
-participants found FILL IN HERE WHEN COMPLETE.
+experience. Re-interviewed participants were given our 3 research questions along with results and main
+discussion points, and asked open ended questions regarding their opinions and experiences to validate our 
+findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
+list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:eval}
 
@@
 
+From the re-interviewed participants, we found extremely positive feedback regarding both our results and major discussion
+points. Participants often had new stories and experiences to share once they had heard the results of this paper which
+confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
+as such but then realized the results held true to their daily actions for better or worse.
+
 As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.

### 2013-11-17 apie-2013/paper.tex
APIE added danas changes for R2.

@@
 \begin{abstract}
-As software evolves over the lifetime of a project, changes to software objects such as methods or classes occur which can have ripple effects throughout
+As software evolves over the lifetime of a project, changes to software objects such as methods or classes occur, which can have ripple effects throughout
 the rest of the system causing relationships to be re-engineered and causing instability. However, projects are sometimes deemed stable at certain release
@@
 and how processes change, but can it shed light on what causes a project to be released and deemed stable? This paper presents our findings from studying
-source code change trends surrounding 109 major release points of 10 open source projects. We found 6 large source code change trends at these release
+source code change trends surrounding 109 major release points of 10 open source projects. We found 9 large source code change trends at these release
 points that can be used to determine if a project is likely near a major release point and thus deemed stable. 
@@
 for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
-the project~\cite{Conway:1968}. Code is an often looked to statistic for stability but can be grossly misleading
+the project~\cite{Conway:1968}, or a reduction in code churn. Code is an often looked to statistic for stability but can be grossly misleading
 in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC}. 
-Creating a different quantitative approach to determining software stability and release preparedness is still a large open area of
+Creating an approach to determining software stability and release preparedness is still a large open area of
 interest in software engineering research.
 
-We turn our analysis to the notions of software change trends, specifically those trends around major releases. Change trends have been used to detect
+We turn our analysis to the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
+a likelihood for a change type to occur around a certain event. Change trends have been used to detect
 stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
@@
 be leveraged to detect stability and the evolution of source code, the question we investigate in this paper is:
-``\textit{Is it possible to identify trends in source code changes surrounding major releases of open source projects as a notion towards a project
+``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
 stability measure?}''. By answering this question, we hope to provide industry with the ability to determine a project's stability and perhaps
@@
 In this paper, we perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
-throughout their history. We studied 26 quantitative and 16 qualitative change trends and identified a core group of 6 change trends which occur
+throughout their history. We studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
 prominently at major release points of the projects studied.
@@
 explains our methodology and tools used in solving our research problem.
-Section \ref{sec:results} gives a full representation of the results and trends found as well as a summarized results of the 6 change trends found
+Section \ref{sec:results} gives a full representation of the results and trends found as well as a summarized results of the 9 change trends found
 to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.
@@
 In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows us to detect fine grained
-source code changes in Java projects This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
-the smallest possible edit distance between the tree. This results in the source code change at a fine grained level performed in the commit.
+source code changes in Java projects. This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
+the smallest possible edit distance between the trees. This results in the source code change at a fine grained level performed in the commit.
 
-We took ChangeDistiller and applied it across 10 open source Java projects. For each of the projects, we obtained the software configuration management (SCM) system
+We took ChangeDistiller and applied it across 10 open source Java projects. Java projects were nesissary because ChangeDistiller only works for Java source code.
+For each of the projects, we obtained the software configuration management (SCM) system
 which is used to store all source code changes of a project. When it was necessary, we converted some forms of SCM system to Git in order to reduce implementation
@@
 34 of ChangeDistiller's built in source code change types for each commit. We noted how many of each change type was performed in each commit and stored that information
-in a PostgreSQL database. In order to filter and protect our results, we manually inspected the 10 Java projects studied in order to identify code built for test.
+in a PostgreSQL database. In order to filter and protect our results, we manually inspected the 10 Java projects studied in order to identify code built for test
+purposes.
 We separated changes to this test code from all other code to ensure our results only focused on real implementation while allowing us to study changes to
@@
 what types of source code changes or trends constitute software stability or maturity. In order to study the release points, we went to each of our 10 project's 
-home pages on-line and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total we identified 472 releases
+web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total we identified 472 releases
 across our 10 studied projects.
@@
 Once the release dates were collected we set about analyzing out data by creating average change ratios surrounding the release dates of each 
-project as a way to measure the trend of a particular change type at a release type. 
-To do this we used Equation~\ref{eq:norm}. Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
+project as a way to measure the trend of a particular change type at a release type. This change ratio simply compares the number of change events (of a given
+change type) before a release to after the release. Both of the before and after event totals are divided by the number of commits on their respective side
+of the release to account for activity. We implemented this algorithm through Equation~\ref{eq:norm}.
+
+Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
 a given release type a sum of a particular change type in commits ($T_c$)
@@
 each point in the graph represented the date of a commit and the sum of the particular change type's occurrences over the last 50 commits. This was used to smooth
-out the curves presented by the tool to allow easier manual inspection. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
-trending, and local minimum. Since the graphs were quite turbulent, best estimations were given by two judges at each release point to fit the graph into the aforementioned
+out the curves presented by the tool to allow easier manual inspection. This method however does not take activity into account as seen in
+Equation~\ref{eq:norm}, so it represents the true activity and change types occuring. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
+trending, and local minimum. Since the graphs were quite turbulent, best estimations were conducted by two judges at each release point to fit the graph into the aforementioned
 4 slope categories. The two judges used 1.5 months before and after the release date as start and end points for the graph trend line.
@@
 \includegraphics[width=0.5\textwidth]{images/apie.png}
-\caption{An screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed.\label{fig:apie}}
+\caption{An screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
+major releases as vertical yellow lines.\label{fig:apie}}
 \end{figure}
@@
 We performed 1888 manual inspections across 10 projects, 472 release dates and 4 change types, and used this data to form the basis of our qualitative data.
-Quantitative data was cross validated with qualitative data to form some of the final results to be seen in the next section.
+Quantitative data was used to compliment the quantitative ratios found from the previous methodology.
 
@@
 
-As it can be seen in Table~\ref{tab:ratio}, there a few notable change type trends around major releases. We can see that any changes to public classes are
-more likely to occur before a major release than after, noting that the addition and deletion of public classes have change ratios of 1.14, 
-and 1.16 respectively. This means that on average classes are added and removed 10\% more often after major releases than before.
-The two most drastic change type trends to note from this table however are the adding and removing of private classes. We can see that private classes are
-44\% more likely to be added after a major release and that adding private classes to a project is roughly 24\% more likely to occur before a major release.
-All results in Table~\ref{tab:ratio} could be used as identified trends of major software releases, while we have just highlighted the larger ratios in
-this paper.
+To study the most prevalent change trends, we set a ratio threshold of greater than 1.2, or less than 0.83 (20\% greater trend of after the release date
+or 20\% greater trend of before the release date) to indicate the greatest trends.
 
-Some of the more interesting trends found around major release points from Table~\ref{tab:ratio} are private methods being removed and public method signatures
-being changed. We hypothesized that public method signatures being changed would be heavily favored towards before a major release indicating a larger sens
-of stability in the public API, however we found that only a change ratio of 0.92 exists meaning that public method signatures undergo changed almost through
-the project steadily. The private methods being removed is also suggestive of possible maintenance being done post release, meaning that code is being
-cleaned up in order to reduce the number of overall methods in a project which are not publicly available.
+As it can be seen in Table~\ref{tab:ratio}, there a few change type trends around major releases which pass the threshold. We can see that both public
+and private methods
+being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:ratio} also shows significance in the changes to private
+classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
+All results in Table~\ref{tab:ratio} could be used as identified trends of major software releases, while we have just highlighted the larger ratios
+which meet our threshold criteria in this paper.
 
-Our qualitative results from manual graph inspections can be seen in Table~\ref{tab:qual}. These results show that adding, changing signatures and bodies of, and
-removing public methods tend to all be at a local minimum of change type trends at major releases.
+Another interesting trend that can been seen in Table~\ref{tab:ratio} is that of changes to public objects. We can see for public classes and methods that
+5 out of 7 ratios indicate changes occur to these objects after major release rather than before. We hypothesize that these changes to the public API
+after a major release come from newly reported bugs from end users as well as having old features being deprecated while adding new features to the project
+after the stable build had been released.
+
+Our complementary qualitative results from manual graph inspections can be seen in Table~\ref{tab:qual}. These results show that adding, changing signatures
+and bodies of, and removing public methods tend to all be at a local minimum of change type trends at major releases when activity is not taken into
+account. These results confirm previous results of low code churn as an indication of stability.
 
@@
 
-Lastly we found that software changes related to testing can be highly used as an indicator of a major release point within the projects studied. The change
-ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the three largest indicators of a project release and stability with regards to test based
-changes are the removal of test classes, the changing of method signatures (both occurring more before releases), and test classes being changed. An interesting
-side note is that documentation is changed heavily before releases as opposed to after.
+Lastly we found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
+ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet our threshold and are indicators of stability with regards to test based
+changes are: the changing of test classes, the removal of test classes, the adding of methods, and the changing of method signatures, and test classes being changed.
+Changes to documentation also meets our threshold and occurs more before a major release.
 
@@
 
-We summarize our key findings as to these 6 source code change type trends which can be found at points of stability as the following: the addition of private classes more often
-before a release point, the removal of private classes more often after a release point, the removal of public methods more often after a release point, 
-the removal of private methods more often after a release point, the removal of test classes more often after a release point, and the changing of test method
-signatures more often before a release point. While all change ratios may need to be considered for a final analysis or taxonomy, we have offered the strongest
-change trends here as a suggestion for future focus.
+While all change ratios may need to be considered for continued analysis or a taxonomy of change trends, we have offered the strongest
+change trends in these results as a suggestion for future focus.
 
@@
 
-Software stability is an often used measure in determining some aspects of code quality. This paper has present some initial indications as to how
+Software stability is an often used measure in determining some aspects of code quality. This paper has presented some initial indications as to how
 software stability may be found, based on major release stabilities, by analyzing source code change type trends both before and after major software
-releases. The source code change trends found in this paper, in combination with future work and analysis, will allow for the detection of stable code
+releases.
+
+The 4 change trends found in this paper occurring before major releases are: added private classes, changed test method signatures, changed documentation,
+and removed test classes.
+The 5 change trends found in this paper occurring after major release are: added test methods, changed test classes, removed public methods, removed
+private classes, and removed private methods.
+
+The source code change trends found in this paper, in combination with future work and analysis, will allow for the detection of stable code
 points throughout a software projects life time. This paper has also shown the beginnings of a visualization for source code change trends which may

### 2013-11-18 apie-2013/paper.tex
APIE added Alessias changes. DRAFT 2 done.

@@
 for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
-the project~\cite{Conway:1968}, or a reduction in code churn. Code is an often looked to statistic for stability but can be grossly misleading
+the project~\cite{Conway:1968}. Code churn is an often looked to statistically for stability but can be grossly misleading
 in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC}. 
@@
 explains our methodology and tools used in solving our research problem.
-Section \ref{sec:results} gives a full representation of the results and trends found as well as a summarized results of the 9 change trends found
+Section \ref{sec:results} gives a full representation of the results and trends found as well as the 9 change trends found
 to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.
@@
 pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
-and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a the project's behavior
+and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a project's behavior
 around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
-indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be being performed.
+indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be performed.
 Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
-at major and minor releases.
+at major and minor releases, showing how test and production code can evolve at different rates and times. These results show that production
+and test code should be handled as different cases for a stability measure around major releases.
 
-The study of open source projects revolving around release points has become more available by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
+The study of open source projects revolving around release points has become more accessible by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
 a resource of historical release dates for open source software projects to be used for future studies by other researchers.
 
-In terms of software stability, many proposed development techniques have been proposed to increase software stability. 
+In terms of software stability, development techniques have been proposed to increase software stability. 
 Fayad~\cite{Fayad:2001:TOI}~\cite{Fayad:2002:ASS} suggests that ``business objects'' (BOs) do not change in nature and that they are inherently stable. These
@@
 project. Mockus et al. found that defect density is lowest at major releases but at the same time software quality is at its lowest all when compared to minor
-releases. The low software quality here related to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
+releases. The low software quality here relates to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
 showed that stable core architectures can be detected by using source code changes. Finally, Fayad et al.~\cite{Fayad:2010:SSM} have
@@
 Once the ChangeDistiller information was collected, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
-have preconceived notions of software stability, we decided that by studying the types of changes surrounding these releases, we could get a better understanding of
+have preconceived notions of software stability, we decided that by studying the change types surrounding these releases, we could get a better understanding of
 what types of source code changes or trends constitute software stability or maturity. In order to study the release points, we went to each of our 10 project's 
@@
 is created by summing across all release of a given release type
-a sum of that same particular change type in commits($T_c$) from a given number of days ($d$) before the release date ($r$) to
-the release date divided by divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
+a sum of that same particular change type in commits ($T_c$) from a given number of days ($d$) before the release date ($r$) to
+the release date divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
 This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
-type occurs more frequently after the release and if it is below 1 then is occurs more frequently before the release. For the purposes of our study, we set the
+type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of our study, we set the
 number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
@@
 \includegraphics[width=0.5\textwidth]{images/apie.png}
-\caption{An screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
+\caption{A screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
 major releases as vertical yellow lines.\label{fig:apie}}
@@
 
-Due to the space requirements of this paper, we focus our results on Major releases of the 10 case study projects and a select few of the calculated change ratios. 
+Due to space requirements of this paper, we focus our results on major releases of the 10 case study projects and select few of the calculated change ratios. 
 There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
@@
 
-As it can be seen in Table~\ref{tab:ratio}, there a few change type trends around major releases which pass the threshold. We can see that both public
+As it can be seen in Table~\ref{tab:ratio}, there are few change type trends around major releases which pass our threshold. We can see that both public
 and private methods

### 2013-11-20 iced-2013/paper.tex
ICED danas quick edits are in.

@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight}
+\title{Sell Prevention and Buy Cure: An Exploration of Indirect Conflicts}
 
@@
 \textit{indirect conflicts} in regards to prevent, detect, or resolve these conflicts as they arise
-through source code changes. While the techniques and tools for direct conflicts have had
-large success, tools either proposed or studied for indirect conflicts have had many common
-struggles and little developer interest. To better understand common indirect conflict tool issues, 
-we performed a grounded theory study with 97 developers involved in either
-interviews or a web survey. We found that: indirect conflicts are a real and frequent problem in
-software development, that developers tend to prefer to use detection and resolution processes or tools
-over those of prevention, that developers do not want awareness mechanisms which provide non actionable results,
-and that there exists a gap in software evolution analytical tools from the reliance on static analysis.
+from a result of source code changes. While the techniques and tools for direct conflicts have had
+large success, tools either proposed or studied for indirect conflicts have had common issues of
+information overload, false positives, and scalability. To better understand these issues, 
+we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
+cource projects, and a survey with 78 responses. Through grounded theory techniques, 
+we found that: indirect conflicts are a real and frequent problem in
+software development, developers prefer to use detection and resolution processes or tools
+over those of prevention, developers do not want awareness mechanisms which provide non actionable results,
+and that there exists a gap in software evolution analytical tools.
 
@@
 mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
-in practice we see these mitigated losses continue to appear quite frequently and can prove to be a significant
+empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
 and time-consuming chore for developers~\cite{Perry:2001:PCL}.
 
-Two types of conflicts have attracted the attention of academics, \textit{direct} and 
+Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
 \textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
-artifact, or finding expert knowledge about a particular file. Tools have been created and studied for direct conflicts
+artifact. Tools have been created and studied for direct conflicts
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. Indirect conflicts involve source code change influences such as making a change which
-causes an error in another location in the system, having code negatively affected by a library upgrade, or even
-breaking an integration test of a separate component than the one changed. Indirect conflict tools however, have
+positive developer feedback. Indirect conflicts involve source code change influences such as making a sour code change 
+which influences behavior negatively in another location in the system. Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 also compounded by the scale of many software development projects leading to further information overload.
-
-Human factors have also shown to cause issues as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+Lastly, social factors have also shown to cause issues as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
 congruence has been leveraged in several indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
-Factors again of information overload, human dependencies, and scalability lay at the root of these problems.
+Issues again of information overload, dependencies (of developer organizational structure), and scalability come up.
 
@@
 
-We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
-developers in order to answer the aforementioned questions. Our findings indicate that: indirect conflicts are very much a real 
+We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+developers, as well as 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
+indirect conflicts are very much a real 
 and frequent problem in software development, that developers tend to prefer to use detection and resolution processes or tools
@@
 
-\subsection{Evaluation}
+\subsection{Validation}
 
-Following our data collection and analysis, we re-interviewed some of our initial interview participants
+Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
 in order to validate our findings. We confirmed our findings as to whether or not they resonate with 
@@
 
-From the interviewed participants, 63\% of developers believe that a large contributing factor to the cause
+From the interviewed participants, 12 developers believe that a large contributing factor to the cause
 of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
@@
 to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
-In light of object contracts, 73\% of interviewed developers gave examples of indirect conflicts they had experienced
+In light of object contracts, 14 interviewed developers gave examples of indirect conflicts they had experienced
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
-towards the rest of the project. Of those 73\%, 21\% dealt
+towards the rest of the project. Of those 14, 3 dealt
 with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
@@
 library projects or are in test said that methods or functions were the root of their indirect conflict issues. 
-36\% of interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
+7 interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
 library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
@@
 
-From the interviewed participants, 58\% of developers explained that indirect conflicts occur ``all the time'' in
+From the interviewed participants, 11 explained that indirect conflicts occur ``all the time'' in
 their development life cycle with a minimum occurrence of once a week, with more serious issues tending
@@
 
-63\% of developers interviewed said that
+12 developers interviewed said that
 when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
@@
 use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
-Lastly, pure developer experience was mentioned with 37\% of developers mentioning that when planning code changes, 
+Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes, 
 either a very experienced member of the project was involved in the planning and has duties to foresee any 
@@
 
-In regards to catching indirect conflicts, 69\% of interviewed developers mentioned forms of testing (unit, and integration) 
+In regards to catching indirect conflicts, 13 interviewed developers mentioned forms of testing (unit, and integration) 
 as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
@@
 when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. 31\% of developers said build processes (either nightly builds or building the project 
+not all indirect conflicts should be caught. 31\% of surveyed developers said build processes (either nightly builds or building the project 
 themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
@@
 
-Once an indirect conflict has occurred and developers need to resolve it, 75\% of developers interviewed said
+Once an indirect conflict has occurred and developers need to resolve it, 14 developers interviewed said
 they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
@@
 \textbf{Unwanted Awareness} As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
-sporadically, and very greatly from case to case. These conditions pose very large issues for the creation of generalizable theories
+sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
 or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
-to proposed or implemented tools as discussed in Section~\ref{sec:related}. These complexities are what we believe to be the
+to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
 leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
-information being presented to them, rendering the system useless. These false positives are caused by a disjoint in what
+information being presented to them, rendering the system useless. These false positives are caused by a difference in what
 developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
-be seen, is cause by developers only considering events which cause some action to be taken on their part, to be true positives,
-where as awareness theory would state any event which is related to an individual's 
+be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
+where as current awareness understanding would state any event which is related to an individual's 
 work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
@@
 Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
-their project awareness. To be able to fully understand a developer's awareness intuition, we can look to the common statement of
-developers only wanting to know about an event if it affects them by forcing some sort of action on their part. In a sense, developers
+their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
+developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
 don't care about what they don't need to account for. This sense of unwanted awareness is crucial to understand why generalized
-awareness techniques of difficult to generalize problems, such as indirect conflicts, often encounter difficulties. 
+awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
+classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
+positives and information overload. 
 Developers simply do not want to know about events which effect them, but require no action on their part.
 
-This unwanted awareness, or limited awareness, seems counter intuitive to classical awareness theory which would state that
+This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
 being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
 but as it was found in practice, this is an incorrect assumption. Since software developers have limits on their time, awareness of
-all events surrounding a developer's work or project is not possible. Developers prefer to spend their limited time dealing with
+all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
 the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
-of events occurring around their work which pose no direct threat to the consumption of their time.
+of events occurring around their work which pose no direct threat to the consumption of their time.}
 
@@
 developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
-them initially, we must consider the pro and cons of adding this, for now, unwanted awareness. A pro of total, or near total
+them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
+influence of total, or near total
 awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
 with higher quality or understanding of the product. 
-A con would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
+A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
 producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
@@
 detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
-that developers tend to prefer workings on real issues that have already occurred as opposed to preventing issues of the future which
+that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
 may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
 is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
-
-Dromey~\cite{Dromey:2003} raises the debate of prevention versus cure in software development and how difficult a problem it is
-to measure. Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
+Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
 place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
@@
 The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
-Even with an experienced system architect, it was shown that identifying dependencies and notifying those involved is a daunting task 
+Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
 which is ignored more than dealt with. The second being the knowledge of when a dependency will fail also requires vast knowledge
@@
 
-The pros and cons of prevention versus cure are similar to those of unwanted awareness in that they result in a trade off 
+Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
+to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
+in that they result in a trade off 
 of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
@@
 A last interesting observation about current curative measures being taken in industry, is that developers view testing of software 
-are curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
+as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
 are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
@@
 %mature point, static analysis
-\textbf{Gap in Evolution Analysis} As was see in Section~\ref{sec:results}, we know that indirect conflicts are more likely to occur,
-from developer opinion, before a mature point is reached in the project's evolution. However, this context as well as any other
-potential contextual attributes are rarely identified or accounted for in regards to indirect conflicts. Developers have given
-us a clue into the idea that different change types may occur at different rates throughout a project's life time and that this
+\textbf{Gap in Software Evolution Analysis} Indirect conflicts are more likely to occur,
+from developer opinion, before a mature point is reached in the project's evolution. (This mature point cmay stem from a major release,
+end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
+mature point, as well as any other
+potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
+that different change types may occur at different rates throughout a project's life time and that this
 may drastically effect the outcomes of indirect conflict tools or processes.
 
+A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
+research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
+could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
+means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
 Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
-changes occur at a given point in time. This type of static analysis may be useful in regards to indirect conflicts in that we
-could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
-means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding. This sort
-of deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
-research than may have been previously thought.
+changes occur at a given point in time which could be used for determining such a context of mature points.
 
@@
 conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.
-These understandings may lead to the detection of indirect conflict issues quicker than traditional means.
 
-Aside from context, we have show that dependency identification and tracking is a key missing component of
+Aside from context, dependency identification and tracking is a key missing component of
 indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
@@
 
-Through our results, developers have said that relational database schemas are one of the highest sources of indirect
+Relational database schemas are one of the highest sources of indirect
 conflicts found in their projects, yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
@@
 The largest implication for future research found in this paper is the need for continued study of the open question
-``Which of prevention or cure is more effective for software development?''. This simple question will undoubtedly 
-produce extremely complex answers. With software developers focusing their current efforts on curative measures suggests
+``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly 
+produce extremely complex answers. With software developers focusing their current efforts on curative measures, for
+indirect conflicts, suggests
 that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
@@
 Tiwana~\cite{Tiwana:2008:ICD} in coordination tools. With these studies, we may find insight into the correct balance
-of prevention versus cure, thus being able to increase developer productivity as well as identify gaps in theory 
-where new discoveries may be made.
+of prevention versus cure, thus being able to increase developer productivity as well as identify more gaps in theory 
+versus practice which may lead to improved tools or the abandoning of existing ones as was shown with UML~\cite{Petre:2013:UP}.
 
-A second implication for research derived by the paper, which is more directly related to awareness, is the continued
-understanding of the disjoint between theoretical awareness and awareness as perceived by industry practitioners.
-As was identified in the previous discussion, while theory suggests that awareness of all events related to ones work
+A second implication for research more directly related to awareness, is the need to further characterize the
+mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
+by industry.
+As was identified in the previous discussion, while current understanding suggests that awareness of all events related to ones work
 produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
-to the point where they only want to be aware of events which require action on their part. This disjoint, of what 
-theoretically should be and what physically is, should be further understood to combat future potential failures in
+to the point where they only want to be aware of events which require action on their part. This difference, of what 
+should be and what is, should be further understood to combat future potential failures in
 tool development or theories attempted to be used in practice.
-
-These disjoints may be detected by conducting case 
-studies of tools built primarily on theoretical models, while having said tools tested in industry. We found developers
-to be quite honest when using a tool they dislike which should lend itself to the detection of the aforementioned
-disjoints.
 
@@
 developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
-said, if researchers wish for their tools to be largely adopted among current developers, a focus at these two stages
-is crucial for success. It should be noted here that detection must come with an almost zero rate of false positives as
+said, focus at these two stages for tool developer will lead to larger adoption among developers.
+It should be noted here that detection must come with an almost zero rate of false positives as
 current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
@@
 
-This research, in regards to tools, most importantly implies the further development of static analysis tools. Static 
+These tool implications themselves imply the need of further development of static analysis tools. Static 
 analysis lays at the heart of most if not all stages of indirect conflict research. We must be able to track and manage
@@
 10 of the 78 participants who were surveyed sent direct responses to us asking for any results of the research to be sent directly to
-them in order to improve their indirect conflict work flows. 7 of the 19 participants survey expressed interest concerning any possible
-tools or plans for tools to come out of this research as well. The combination of academia relatability and direct industry interest 
+them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
+tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest 
 in our results help us fulfill this criterion.
@@
 {\bfseries Variation.} ``Has variation been built into the findings?'' Variation shows that an event is complex and that any findings
-made accurately demonstrate that complexity. Since those participants interviewed came from such a diverse set of organizations, 
+made accurately demonstrate that complexity. Since interviewed participants came from such a diverse set of organizations, 
 software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
@@
 exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
-in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging can be used.
+in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging are presented.
 
@@
 
-Kim conducted several initial focus groups as well as web surveys to determine what developers are interested in with regards to software
-modifications~\cite{Kim:2011:ESA}. The top two interests found were that developers wanted to know whose recent code changes semantically
-interfere with their own code changes, and whether their code is impacted by a code change. These areas of interest resonate with what was
+For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
+recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
+Kim conducted several focus groups as well as web surveys to determine these results. These areas of interest resonate with what was
 found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
@@
 towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
-is the same mentality found in this paper as ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
+resonates with ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
 sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
@@
 While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
-can be, and should be, applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
+can be applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
 changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
@@
 
-We have brought to attention through discussion the disjoint of theoretical awareness needs versus the practical
-awareness needs found in industry, the debate of prevention versus cure and industries tendency towards curative measures,
-and the gap of analytical evolution tools needed for dependency identification and indirect conflicts.
-
-We hope that this study and its 
-findings will inspire future processes and tools for dealing with indirect conflicts that 
-will aid developers in industry as well as test and validate our theories put forth in this paper as well as spark a potential
-investigation into preventative versus curative approaches to software defects.
+We have shown the disjoint of current awareness understanding versus the practical
+awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
+ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure has been presented along with
+the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
+to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, we have shown
+the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
+future industrial adoption of tools produced by academia for lack of fit in industrial needs.
 
@@
 
-We would sincerely like to thank all participants who were willing to be interviewed or who participated
+We sincerely thank all participants who were willing to be interviewed or who participated
 in completing our survey. We thank these people for taking time out of their day to participate and for
-sharing their developer experience with us. Without these people our research could not have been possible.
+sharing their developer experience with us.
 

### 2013-11-21 iced-2013/paper.tex
ICED Draft 4 done.

@@
 
-Awareness techniques have been proposed and studied to aid in developer
+Awareness techniques have been proposed and studied to aid developer
 understanding, efficiency, and quality of software produced. Some of these techniques have focused 
 on either \textit{direct} or 
-\textit{indirect conflicts} in regards to prevent, detect, or resolve these conflicts as they arise
+\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
 from a result of source code changes. While the techniques and tools for direct conflicts have had
@@
 we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
-cource projects, and a survey with 78 responses. Through grounded theory techniques, 
-we found that: indirect conflicts are a real and frequent problem in
+source projects, and a survey with 78 responses. Through grounded theory techniques, 
+we found: events and conditions which indirect conflicts are likely to occur in
 software development, developers prefer to use detection and resolution processes or tools
@@
 also compounded by the scale of many software development projects leading to further information overload.
-Lastly, social factors have also shown to cause issues as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
-congruence has been leveraged in several indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
-Issues again of information overload, dependencies (of developer organizational structure), and scalability come up.
+Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
+However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
+come up.
 
-In the hopes of resolving some of the aforementioned issues surrounding indirect conflicts and their tools, we 
-are interested in exploring the following three research questions:
+In order to fully understand the root causes of information overload, false positives, and
+scalability issues in regards to indirect conflicts, we will take a step back and determine what events occur to
+cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
+We then set out to understand what mitigation strategies developers currently use as opposed to those created
+by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
+in both research and industry. To accomplish these goals, we ask the following:
 
 \begin{description}
-	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What is the nature of indirect conflicts?}
-	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What types of process or tools are being used by developers in regards to indirect conflicts?}
+	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
+	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
 	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
@@
 We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
-developers, as well as 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
-indirect conflicts are very much a real 
-and frequent problem in software development, that developers tend to prefer to use detection and resolution processes or tools
-over those of prevention, that developers do not want awareness mechanisms which provide non actionable results, 
-and that there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
+developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
+indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
+developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
+and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
 context of indirect conflicts.
-
-We continue this paper by first presenting our methodology used to perform this study in Section~\ref{sec:meth}. We
-then present and discuss the results of our study in regards to our 3 research questions in Section~\ref{sec:results}
-and Section~\ref{sec:disc}.
-Next, an evaluation of how our study was performed in terms of the credibility
-and quality of our results is described in Section~\ref{sec:eval}. We then cover related work in Section~\ref{sec:related}
-and conclude in Section~\ref{sec:conc}.
 
@@
 
-We performed a grounded theory study in two parts. First, a round of semi-structured interviews were conducted which 
+We performed a a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
 addressed the 3 research questions as mentioned above. Second, a survey was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
-larger developer opinion of the subject.
+larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
+interview participants to once again confirm and test what had been learned and the results derived by the paper.
+Grounded theory techniques were used to analyze the information provided from all three data gathering processes.
 
@@
 codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
-link our defined codes through categories of grounded theory such as context and consequences. The ultimate goal of grounded 
-theory is to produce a theory that is tied to the empirical data collected; our final theories can be seen in 
-Section~\ref{sec:disc}. In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+link our defined codes through categories of grounded theory such as context and consequences. 
+In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
 3 criteria that are commonly used in evaluating grounded theory studies.
@@
 
-\subsection{What is the nature of indirect conflicts?}
+\subsection{What events or conditions lead to indirect conflicts?}
 
@@
 
-\subsection{What types of process or tools are being used by developers in regards to indirect conflicts?}
+\subsection{What mitigation techniques are used by developers in regards to indirect conflicts?}
 
@@
 work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
+``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
+showing that not all changes around an object should be reported for awareness.
 
@@
 developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
-don't care about what they don't need to account for. This sense of unwanted awareness is crucial to understand why generalized
+don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
+that the developer only wants the awareness if the item will require action on their part to resolve the issue.
+This sense of unwanted or limited awareness is crucial to understand why generalized
 awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
@@
 being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
-but as it was found in practice, this is an incorrect assumption. Since software developers have limits on their time, awareness of
+but as it was found in practice, this is an incorrect assumption. 
+In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for 
+analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
 all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
@@
 conflicts occurred at either the detection or cure levels. Obviously, developers either prefer to spend their time in curative
-measures, or do not posses the proper tools to take better action in the preventative stages.
+measures, or do not posses the proper tools to take better action in the preventative stages. ``You're reaction time is much more crucial''
+said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
+matter what prevention was taken, the issue must simply be resolves as quick as possible now.
 
@@
 Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
-which is ignored more than dealt with. The second being the knowledge of when a dependency will fail also requires vast knowledge
-of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. The
+which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
+of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
+are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
 abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
@@
 when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
-measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative.
+measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
+``prevention is the goal'' was commonly said by developers.
 
@@
 \textbf{Gap in Software Evolution Analysis} Indirect conflicts are more likely to occur,
-from developer opinion, before a mature point is reached in the project's evolution. (This mature point cmay stem from a major release,
+from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
+better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
 end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
@@
 Relational database schemas are one of the highest sources of indirect
-conflicts found in their projects, yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
+conflicts found in their projects, ``it breaks stuff all over the place'',
+yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
 been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language

### 2013-12-02 msc-thesis/chapters/1/chapter-intro.tex
MSC Added thesis template files.

\startfirstchapter{Introduction}
\label{chapter:introduction}

The main goal of this document was to set up competently the Latex framework for a thesis document at UVic, with all the assorted details of front pages, numbering and so on. In order to present such a template to other users, the content has been filled from the guidelines of an experienced graduate advisor and supervisor on how to write a thesis in the first place.

The Latex framework given in these files should work in the Windows, Mac and Unix environments. In Windows, the development environment is MikTex, available as a free download from the web page at \textit{miktex.org}. Miktex includes all the tools for \LaTeX, without a convenient editor. I have been using
\textit{WinEdt} which is still the best as far as working in a well integrated manner with MikTex, even if one has to pay a small price. \textit{WinEdt} is available from \textit{winedt.com}. The corresponding package for the Mac is MacTex, available from \textit{www.tug.org$\backslash$mactex}.
Please note that in the actual file with the \LaTeX code there are extra local comments referring to small differences between platforms. This is especially crucial for the insertion of figures of any type (see also Chapter 3).

The package \textit{BibTex} is used for the automatic bibliography and it is incorporated in the MikTex package. The style of bibliography exemplified is this document is the "plain" one,
most often used in science theses. This is shown
by the entry \textit{plain} in the \textit{bibliographystyle}
command which can be found in the top file, namely
\textit{mainthesisUVIC.tex}. Substitute the
appropriate bibliography style for your needs by finding
the correct parameter. In order to help you there is a PDF
file entitled: "InformationOnBibliographyStyles.pdf" in the
same folder as the top file \textit{mainthesisUVIC.tex}.

The title pages are correctly formatted according to the guidelines. However the superset is given here as a template for a Ph.D. dissertation and adjustments must be made for a Master's thesis accordingly.

When presented with the task of having to write a large document reporting on the work done for the thesis portion of a graduate degree most people have anxiety attacks. It seems that the research work has been done, probably with great competence, yet one feels totally lost on how to approach the writing part - unless of course one is a writer to start with. While I am not at all the expert, I have experience both in writing documents and in supervising many graduate students. I have developed a simple enough structure which guides me and my students most of the times fairly well. We then customize it for different audiences and situations. The content of this pseudo mini thesis summarizes the structure I use.

This document, however, is \textit{not at all} a Latex manual. There are a few examples sprinkled here and there of various commands. However no explanations or full examples are given for how to use Latex itself. That is most certainly left to the user!

Starting at the very beginning, my first step is to set up the formatting framework before the content itself, just like the frame of a building is the sustaining structure. By having the content only in bullet points which can be moved and changed at will I can have a whole document which will always compile and be ready for presentation. I ask my student to prepare mock chapters which contains only these bulleted lists to be used for discussion. The bulleted list I had for this chapter is as follows:
\begin{enumerate}
        \item State the dual purpose of this document (just mention briefly);
        \item State what this document is not (just mention);
        \item Describe what I think should go in the Introduction in general;
        \item Describe what I think should not go in the "Introduction" chapter in general;
        \item Describe the structure of the document.
\end{enumerate}

When the writing for a chapter is particularly long, it might be better to have some, if not all, sections in separate files. It makes it a lot easier to find possible errors in Latex as well, since the command to input a file for a separate section can always be commented out. When looking at the \textit{.tex} file itself for this introduction chapter you will notice that there are 2 sections, namely "How to Start an Introduction" and "Is a Review of All previous Work Necessary Here?" which are included as separate files, immediately following this paragraph.
\input chapters/1/sec-intro
\input chapters/1/sec-review
%\pagebreak
\section{My Claims}
Something must be new in this work, no matter how small, since you are getting a graduate degree for it! Tell me about it clearly and succintly right now, just as you did in the abstract. Make an impact here. How about something like the following box:

I make \textit{four} claims which
my dissertation validates:
\\

\framebox{%
\parbox{5in}{
	My new algorithm to solve the problem of doing nothing include these important new features whose practical applicability can be proved both formally and empirically:
	\begin{enumerate}
	\item first feature;
	\item second feature;
	\item everything is much easier to understand, and therefore, easier to implement correctly.
	\end{enumerate}
}}
\\

\noindent Claim 1 and claim 2 are \textit{quantitative} - they will be proven by experiment.

\noindent Claim 3 is \textit{qualitative} - they will be demonstrated by argument.

\subsection{The Importance of My Claims}

Some very important positive consequences
arise from the validation of the above claims.
It is these consequences that comprise a significant
positive contribution to research in the field
of whatever the field is.
\\

\noindent Claim 1 implies that:
\begin{enumerate}
\item{Something profound which applies to:
	\begin{itemize}
	\item {something excellent;}
	\item {something important.}
	\end{itemize}}
\item{Something else just as profound.}
\end{enumerate}

\noindent Claim 2 implies that:
\begin{itemize}
\item{Repeat as above if necessary and useful.}
\end{itemize}

\noindent The consequence of claim 3 is that:
\begin{itemize}
\item{There must be something good coming out of all this work!}
\end{itemize}

\section{Agenda}

This section provides a map of the dissertation
to show the reader where and how it validates
the claims previously made. Here is where I am also presenting my own style of organization which may be totally different from what your supervisor thinks. However, trust me, this is a good solid beginning for a structure. Your supervisor may ask you to change it, but will still appreciate what you have! For each of the chapters below I also give a short summary of what the main focus should be and then I expand on it  a bit within the chapter itself.

\begin{description}
\item[\textbf{Chapter 1}] contains a statement of
the claims which will be proved by this dissertation followed by an overview of the structure of the document itself.
\item[\textbf{Chapter 2}] describes in details the open problem which is to be tackled together with its context, its impact and the overall motivation for the research overall.
\item[\textbf{Chapter 3}] gives the new research, its methodology, the algorithms involved, the new solution, the new work done. Formal proofs and arguments are made here. This is the first of the two contributions expected in a thesis for a graduate degree.
\item[\textbf{Chapter 4}] is where the experiments and the methodology for them is fully described. The first part includes all details of how the empirical side of the research has been conducted. Note that not every thesis has this empirical portion.
\item[\textbf{Chapter 5}] includes the evaluation of the data presented above and the comparisons with the work of others, to show how much better the new approach is. This is the second of the two contributions expected in a thesis for a graduate degree. Note that this part could be consolidated into the chapter above.
\item[\textbf{Chapter 6}] contains a restatement of the claims and results of the dissertation. It also enumerates avenues of future work for further development of the concept and its applications.
\end{description}

The list above is not complete. Chapter 3 actually includes a lot more, as I could not resist placing in it a few \LaTeX examples to help you along. This document is not a primer for \LaTeX, but there is no harm done in giving a little help.

### 2013-12-02 msc-thesis/chapters/3/chapter-new.tex
MSC Added thesis template files.

\startchapter{The New Approach and Solution}
\label{chapter:newsol}

This is where you go all out and tell us all about your new discovery and research related to the problem in the previous chapter. No arrogant sweeping statements which cannot be fully justified, but no false modesty either. You must impress your reader that you have accomplished something.

Simply summarized, this chapter should be comprised of at least two main sections, each with appropriate subsections. The first section should describe:
\begin{itemize}
\item {what the new approach is;}
\item {what is really totally new;}
\item {what is incrementally new;}
\item {what you built upon.}
\end{itemize}

The second part should describe fully how the new approach works, both with the overall theoretical exposition (e.g. an algorithm) and with as many examples as necessary for clarity. Remember that if the reader does not understand fully, you will get a lot of questions and doubts. Good examples, good figures, good diagrams with super clear tutorial explanations can be a joy to read and make even a small contribution appear to be more impressive. Are you afraid that if you are too tutorial your work will not seem as deep and difficult? Only shallow people will make such a superficial evaluation, have trust instead in the wisdom of your supervisory committee.

Use at least one good example throughout, and even better if this is one of the examples you used in Chapter 2 to describe the original problem.

By the way, this would be the first chapter I would write. This is what I know best right now, as I just finished working on it. It is clear to me and on the tip of my fingers. Start with your strengths! The second chapter I would write is the next one about the experiments, followed closely by chapter 2 describing the problem. It may not seem intuitive to you, but it works and it is the most productive way I ever found to finish a document.

\input chapters/3/sec-latexhelp

### 2013-12-03 msc-thesis/chapters/1/chapter-intro.tex
MSC Added in 2 motivational studies.

@@
 
-The main goal of this document was to set up competently the Latex framework for a thesis document at UVic, with all the assorted details of front pages, numbering and so on.
-In order to present such a template to other users, the content has been filled from the guidelines of an experienced graduate advisor and supervisor on how to write a thesis in the first place.
-
-The Latex framework given in these files should work in the Windows, Mac and Unix environments.
-In Windows, the development environment is MikTex, available as a free download from the web page at \textit{miktex.org}.
-Miktex includes all the tools for \LaTeX, without a convenient editor.
-I have been using
-\textit{WinEdt} which is still the best as far as working in a well integrated manner with MikTex, even if one has to pay a small price.
-\textit{WinEdt} is available from \textit{winedt.com}.
-The corresponding package for the Mac is MacTex, available from \textit{www.tug.org$\backslash$mactex}.
-Please note that in the actual file with the \LaTeX code there are extra local comments referring to small differences between platforms.
-This is especially crucial for the insertion of figures of any type (see also Chapter 3).
-
-The package \textit{BibTex} is used for the automatic bibliography and it is incorporated in the MikTex package. The style of bibliography exemplified is this document is the "plain" one,
-most often used in science theses. This is shown
-by the entry \textit{plain} in the \textit{bibliographystyle}
-command which can be found in the top file, namely
-\textit{mainthesisUVIC.tex}. Substitute the
-appropriate bibliography style for your needs by finding
-the correct parameter. In order to help you there is a PDF
-file entitled: "InformationOnBibliographyStyles.pdf" in the
-same folder as the top file \textit{mainthesisUVIC.tex}.
-
-The title pages are correctly formatted according to the guidelines.
-However the superset is given here as a template for a Ph.D.
-dissertation and adjustments must be made for a Master's thesis accordingly.
-
-When presented with the task of having to write a large document reporting on the work done for the thesis portion of a graduate degree most people have anxiety attacks.
-It seems that the research work has been done, probably with great competence, yet one feels totally lost on how to approach the writing part - unless of course one is a writer to start with.
-While I am not at all the expert, I have experience both in writing documents and in supervising many graduate students.
-I have developed a simple enough structure which guides me and my students most of the times fairly well.
-We then customize it for different audiences and situations.
-The content of this pseudo mini thesis summarizes the structure I use.
-
-This document, however, is \textit{not at all} a Latex manual.
-There are a few examples sprinkled here and there of various commands.
-However no explanations or full examples are given for how to use Latex itself.
-That is most certainly left to the user!
-
-Starting at the very beginning, my first step is to set up the formatting framework before the content itself, just like the frame of a building is the sustaining structure.
-By having the content only in bullet points which can be moved and changed at will I can have a whole document which will always compile and be ready for presentation.
-I ask my student to prepare mock chapters which contains only these bulleted lists to be used for discussion.
-The bulleted list I had for this chapter is as follows:
-\begin{enumerate}
-        \item State the dual purpose of this document (just mention briefly);
-        \item State what this document is not (just mention);
-        \item Describe what I think should go in the Introduction in general;
-        \item Describe what I think should not go in the "Introduction" chapter in general;
-        \item Describe the structure of the document.
-\end{enumerate}
-
-When the writing for a chapter is particularly long, it might be better to have some, if not all, sections in separate files.
-It makes it a lot easier to find possible errors in Latex as well, since the command to input a file for a separate section can always be commented out.
-When looking at the \textit{.tex} file itself for this introduction chapter you will notice that there are 2 sections, namely "How to Start an Introduction" and "Is a Review of All previous Work Necessary Here?" which are included as separate files, immediately following this paragraph.
 \input chapters/1/sec-intro
 \input chapters/1/sec-review
-%\pagebreak
-\section{My Claims}
-Something must be new in this work, no matter how small, since you are getting a graduate degree for it!
-Tell me about it clearly and succintly right now, just as you did in the abstract.
-Make an impact here.
-How about something like the following box:
-
-I make \textit{four} claims which
-my dissertation validates:
-\\
-
-\framebox{%
-\parbox{5in}{
-	My new algorithm to solve the problem of doing nothing include these important new features whose practical applicability can be proved both formally and empirically:
-	\begin{enumerate}
-	\item first feature;
-	\item second feature;
-	\item everything is much easier to understand, and therefore, easier to implement correctly.
-	\end{enumerate}
-}}
-\\
-
-\noindent Claim 1 and claim 2 are \textit{quantitative} - they will be proven by experiment.
-
-\noindent Claim 3 is \textit{qualitative} - they will be demonstrated by argument.
-
-\subsection{The Importance of My Claims}
-
-Some very important positive consequences
-arise from the validation of the above claims.
-It is these consequences that comprise a significant
-positive contribution to research in the field
-of whatever the field is.
-\\
-
-\noindent Claim 1 implies that:
-\begin{enumerate}
-\item{Something profound which applies to:
-	\begin{itemize}
-	\item {something excellent;}
-	\item {something important.}
-	\end{itemize}}
-\item{Something else just as profound.}
-\end{enumerate}
-
-\noindent Claim 2 implies that:
-\begin{itemize}
-\item{Repeat as above if necessary and useful.}
-\end{itemize}
-
-\noindent The consequence of claim 3 is that:
-\begin{itemize}
-\item{There must be something good coming out of all this work!}
-\end{itemize}
-
-\section{Agenda}
-
-This section provides a map of the dissertation
-to show the reader where and how it validates
-the claims previously made.
-Here is where I am also presenting my own style of organization which may be totally different from what your supervisor thinks.
-However, trust me, this is a good solid beginning for a structure.
-Your supervisor may ask you to change it, but will still appreciate what you have!
-For each of the chapters below I also give a short summary of what the main focus should be and then I expand on it  a bit within the chapter itself.
-
-\begin{description}
-\item[\textbf{Chapter 1}] contains a statement of
-the claims which will be proved by this dissertation followed by an overview of the structure of the document itself.
-\item[\textbf{Chapter 2}] describes in details the open problem which is to be tackled together with its context, its impact and the overall motivation for the research overall.
-\item[\textbf{Chapter 3}] gives the new research, its methodology, the algorithms involved, the new solution, the new work done.
-Formal proofs and arguments are made here.
-This is the first of the two contributions expected in a thesis for a graduate degree.
-\item[\textbf{Chapter 4}] is where the experiments and the methodology for them is fully described.
-The first part includes all details of how the empirical side of the research has been conducted.
-Note that not every thesis has this empirical portion.
-\item[\textbf{Chapter 5}] includes the evaluation of the data presented above and the comparisons with the work of others, to show how much better the new approach is.
-This is the second of the two contributions expected in a thesis for a graduate degree.
-Note that this part could be consolidated into the chapter above.
-\item[\textbf{Chapter 6}] contains a restatement of the claims and results of the dissertation. It also enumerates avenues of future work for further development of the concept and its applications.
-\end{description}
-
-The list above is not complete.
-Chapter 3 actually includes a lot more, as I could not resist placing in it a few \LaTeX examples to help you along.
-This document is not a primer for \LaTeX, but there is no harm done in giving a little help.

### 2013-12-03 msc-thesis/chapters/2/chapter-problem.tex
MSC Added in 2 motivational studies.

\startchapter{The Problem}
\label{chapter:problem}

\newlength{\savedunitlength}
\setlength{\unitlength}{2em}

Introduction paragraph about the problem and the various motivations.

\input chapters/2/sec-developer-pairs
\input chapters/2/sec-impact

\section{Problem Description}

Use the two case studies as motivating examples to show the full problem.

\setlength{\unitlength}{\savedunitlength}

### 2013-12-03 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC Added in 2 motivational studies.

--- 2013-03-15 icse-src-2013/paper.tex
@@
-% This will be the main document for the Technical Networks paper to
-% be written by the Eggnet team of Jordan Ell, Triet Huynh and Braden
-% Simpson in association with Adrian Schroeter and Daniela Damian.
-
-\documentclass[conference]{IEEEtran}
-
-% Use of outside images
-\usepackage{graphicx} 
-% Use text inside euqations
-\usepackage{amsmath}
-
-\usepackage{balance}
-\usepackage{float}
-\floatstyle{plaintop}
-\restylefloat{table}
-
-% Correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-% Begin the paper here
-\begin{document}
-
-% Paper title
-% Can use linebreaks \\ within to get better formatting as desired
-\title{Identifying Failure Inducing Developer Pairs within Developer Networks}
-
-% Authors names
-\author{\IEEEauthorblockN{Jordan Ell}
-\IEEEauthorblockA{University of Victoria,
-Victoria, British Columbia, Canada \\ jell@uvic.ca}
-}
-
-% Make the title area
-\maketitle
-
-\begin{abstract}
-Software systems have not only become larger over time, but the amount of
-technical contributors and dependencies have also increased. With these expansions also comes
-the increasing risk of introducing a software failure into a pre-existing system.
-Software failures are a multi-billion dollar problem in the industry today and while integration and
-other forms of testing are helping to ensure a minimal number of failures, research to understand
-full impacts of code changes and their social implications is still a major concern. This paper describes
-how analysis of code changes and the technical relationships they infer can be used to detect pairs
-of developers whose technical dependencies may induce software failures. These developer pairs may
-also be used to predict future software failures as well as provide recommendations to contributors
-to solve these failures caused by source code changes.
-\end{abstract}
-
-\section{Introduction}
-
-% Setup the problem.
-Large software projects are created using highly modular and 
-reusable code. This creates technical dependencies between methods or functions
-that can be used in a wide variety of locations throughout the project. 
-This causes changes to any given method to have a rippling 
-effect across the rest of the project~\cite{Acharya:2011:PCI}. The larger these effects are,
-the more likely they are to cause a software failure inside the system during the project's
-life span~\cite{Zimmermann:2008:PDU}. These observations of technical dependencies open
-the door to types of analysis on the developer networks they infer and preventing software
-failures by improving coordination amongst dependant developers.
+\section{Study 1: Failure Inducing Developer Pairs}
 
@@
 With the power of technical dependencies in predicting software failures, the question I
-investigated in this project is : 
+investigated in this study was : 
 ``\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
@@
 % Explain our brief intended approach
-This paper explains the approach used to locate these pairs of developers in developer networks.
+This study explains the approach used to locate these pairs of developers in developer networks.
 The process utilizes code changes and the call hierarchies effected  to find patterns of developer 
-relationships in successful and failed code changes.  These developer relationships can also be used
+relationships in successful and failed code changes. As it will be seen, we found 27 statistically
+significant failure inducing developer pairs. These developer relationships can also be used
 to promote the idea of leveraging socio-technical congruence, a measure of coordination compared
@@
 
-% Technical network figure
-\begin{figure*}[tb!]
-\centering
-\includegraphics[width=0.9\textwidth]{images/TecNetwork}
-\caption{A technical network for a code change. Carl has changed method getX() which is being
-called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
-\end{figure*}
+\subsection{Technical Approach}
 
-\section{Technical Approach}
-
-\subsection{Extracting Technical Networks}
+\subsubsection{Extracting Technical Networks}
 % End goal
@@
 repository although it can be used with any software repository.
+
+% Technical network figure
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/TecNetwork}
+\caption{A technical network for a code change. Carl has changed method getX() which is being
+called by Bret's method foo() as well as Daniel and Anne's method bar().\label{fig:network}}
+\end{figure}
 
@@
 
-\subsection{Identifying Failure Inducing Developer Pairs}
+\subsubsection{Identifying Failure Inducing Developer Pairs}
 To identify failure inducing developer pairs (edges) inside technical networks, 
@@
 
-\section{Results}
+\subsection{Results}
 To illustrate the use of the approach, I conducted a case study of
@@
 \hline
-(Daniel, Anne)	&	0&	14&	1.0000& 0.0001249		\\
-(Carl, Bret)	&	1&	12&	0.9190& 0.003468	\\
-(Emily, Frank)	&	1&	9&	0.8948& 0.02165      \\
+(Daniel, Anne) & 0 & 14 & 1.0000 & 0.0001249 \\
+(Carl, Bret)   & 1 & 12 & 0.9190 & 0.003468  \\
+(Emily, Frank) & 1 &  9 & 0.8948 & 0.02165   \\
 \hline
@@
 
-\section{Conclusion and Future Work}
+\subsection{Analysis}
 Technical dependencies are often used to predict software failures
 in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. 
-This paper has presented a method for detecting failure inducing pairs of developers inside
+This study has presented a method for detecting failure inducing pairs of developers inside
 of technical networks based on code changes. These developer pairs can be used in the prediction
@@
 
-In future work, the adding of communication networks on a per code change basis is planned. In this,
-the investigation of social and technical congruence on these networks and its effects on
-software quality will be studied.
-
-\bibliographystyle{IEEEtran}
-\balance
-\bibliography{paper}
-
-% End of the paper
-\end{document}
+This study however, did not consider the technical dependencies themselves to be the root cause of
+the software failures. This study focused purely on developer ownership of software methods and
+the dependencies between developers as the possible root cause of the failures. To study this root
+cause futher, a study of indirect conflicts and their relationship to developer code ownership
+will be conducted.

### 2013-12-03 msc-thesis/chapters/2/sec-impact.tex
MSC Added in 2 motivational studies.

--- 2013-06-05 impact-position-2013/paper.tex
@@
-% This paper is written for CHASE 2013 and features the concept and
-% creation of "Impact" the awareness tool.
+\section{Study 2: Awareness with Impact}
 
-\documentclass[conference]{IEEEtran}
+In response to Study 1, a second investigation was conducted. Study 1 revealed
+that pairs of developers can be used around technical dependencies in order
+to predict bugs. The natural follow up to these findings was to conduct a study
+of indirect conflicts surrounding source code changes that involved the source
+code being changed and the developers involved in those changes. These indirect
+conflicts were primarily studies through the notion of task awareness.
 
-\usepackage{graphicx}
-\usepackage{amsmath}
-
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-
-%Paper title
-\title{Supporting Awareness of Indirect Conflicts with Impact}
-
-%Author block
-\author{\IEEEauthorblockN{Jordan Ell}
-\IEEEauthorblockA{University of Victoria\\
-Victoria, British Columbia \\ jell@uvic.ca}
-\and
-\IEEEauthorblockN{Daniela Damian}
-\IEEEauthorblockA{University of Victoria\\
-Victoria, British Columbia \\ danielad@cs.uvic.ca}}
-
-\maketitle
-
-% Paper abstract
-\begin{abstract}
-Awareness has been largely studied in the field of CSCW research within
-software engineering. Many tools and techniques have been proposed and built
-in order to provide software developers and stakeholders with a greater
-sense of workspace and task awareness within their software projects. These
-techniques and tools have been largely focused on detecting \textit{direct} 
-conflicts which arise over a project's life time or have created 
-an exploratory ground for stakeholders to use as a means of resolving self
-discovered direct or indirect conflicts. However, detecting 
-and providing pertinent information regarding \textit{indirect} conflicts has been
-largely ignored partially due to their inherently larger
-complexity than direct conflicts. Indirect conflicts arise when changes
-in one software artifact affect another. In this paper, we present
-\textit{Impact}, a new task awareness tool directly aimed at both detecting
-and presenting indirect conflicts which arise inside of a software project.
-\textit{Impact} represents a first step towards the design and implementation
-of awareness tools that specifically address indirect conflicts in software
-projects. To describe \textit{Impact}, we introduce previous indirect 
-conflict awareness attempts, the existing problems in the field, our design 
-and implementation for a potential solution, as well as \textit{Impact's} future work
-on a proposed solution to information overload which arises with indirect
-conflict awareness.
-\end{abstract}
-
-\section{Introduction}
-Awareness is characterized as "an understanding of the activities of others
-which provides a context for one's own activities"~\cite{Dourish:1992:ACS}.
-The study of awareness and its tools has become an important topic of
-research in software engineering especially with the new importance of
-distributed work and collaboration. Awareness is generally associated with
-both technical and social dependencies that are created and evolve over
-a software project's life time. The study of these dependencies has become
-the primary focus of most awareness related research. Task awareness has
-become the most prevalent field of awareness research to understand 
-how developers cope with these technical and social dependencies.
-
+%Awareness tools
 Tools have been created to attempt to solve task awareness related issues
-with some success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
+with some success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. 
+These tools have been designed 
 to solve task awareness related issues at the direct conflict level. 
@@
 particular file, and knowing which developers are working on which files.
-Meanwhile, task awareness related issues at the indirect conflict level
-continue to be an issue which is largely unsolved by most coordination
-mechanisms. Examples of indirect conflict awareness
+On the other hand, task awareness related issues at the indirect conflict level
+have also been studied, with many tools being
+produced~\cite{Sarma:2007:TSA,Servant:2010:CPI,Begel:2010:CDE,Trainer:2005:BGT}. 
+Examples of indirect conflict awareness
 include having one's own code effected by another developer's source
@@
 
+% Limitations of indirect conflict tools
 Indirect conflicts arising in source code are inherently
@@
 created an exploratory environment which is used by developers to
-solve conflicts which may arise. These tools were not designed to detect
+solve conflicts which may arise~\cite{Servant:2010:CPI}. These tools were not designed to detect
 indirect conflicts that arise and alert developers to their presence 
@@
 
-While indirect conflicts remain to be a large problem area in task awareness,
-some attempts have been made to both detect and provide developer awareness
-of indirect conflicts. Servant et al.~\cite{Servant:2010:CPI}, have devised a tool
-which can both detect and alert developers to indirect conflicts arising in 
-their software projects. However, a main issue of information overload continues
-to arise from this and many other indirect conflict tools. As software systems
-grow larger and more complex, indirect dependencies tend to become more numerous.
-This causes small changes in source code to have a large impact on the software system
-which these tools often report exhaustively. Information overload must be addressed
-in any feasible attempt made on indirect conflicts.
-
-In this paper, we report on our research into supporting indirect conflicts
+%Outline of study
+In this study, I report on research into supporting indirect conflicts
 and present the design, implementation, and future evaluation of the tool \textit{Impact},
@@
 alert developers involved in such conflicts in task awareness while limiting information
-overload by using design by contract~\cite{Meyer:1988} solutions to method design. While this paper
+overload by using design by contract~\cite{Meyer:1988} solutions to method design. While this study
 outlines \textit{Impact's} specific implementation, its design is rather
 generic and can be implemented in similar indirect conflict awareness tools.
-\textit{Impact} represents a first step towards the design and implementation
-of awareness tools which address indirect conflicts in software development.
 
-The rest of this paper is organized as follows. First, we begin by discussing
-similar indirect conflict awareness tools which have partially solved the
-issues presented by this paper and how their designs can be applied to 
-\textit{Impact}. In the following section we describe a generic design and implementation
-of \textit{Impact} as an awareness tool. We then discuss future work in terms
-of evaluation and further understanding of the problems at hand.
-
-\section{Related Work}
-Although there is an abundance of awareness tools developed in research
-today, only a handful have made an attempt to examine indirect conflicts.
-Here, we will outline four of the forefront projects in indirect conflicts
-and how these projects have influenced the decision making process in
-the design and implementation of \textit{Impact}.
-
-We first start with both Codebook~\cite{Begel:2010:CDE} and 
-Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
-environment for developers to handle indirect conflicts. Exploratory
-pertains to the ability to solve self determined conflicts, meaning that
-once a developer discovers a conflict, they can use the tool as a type of
-lookup table to solve their issue. Codebook is a type of social developer
-network that relates developers to source code, issue repositories and
-other social media while Ariadne only looks at source code for developer
-to source code association. Through Codebook, developers become
-owners of source code artifacts. Both projects also use program 
-dependency graphs~\cite{Horwitz:1992:UPD}
-in order to relate technical artifacts to each other. These projects make 
-use of method call graphs in order to 
-determine which methods invoke others which forms the basis for 
-linking source code artifacts creating a directed graph. While these 
-projects can be great tools 
-for solving indirect conflicts which may arise, by querying such directed
-graphs to view impacts of conflict creating code, they lack the ability to
-detect potential conflicts on their own.
-
-A serious attempt at both detecting and informing developers of
-indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
-monitors developers activities in files with regards to class signatures.
-Once a developer changes the signature of a class, such as by modifying changes
-in name, parameters, or return values of public methods, any workspace
-of other developers which are using that class will be notified. Palantir utilizes
-a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
-a favored collection system among awareness tools. Sarma et al.
-~\cite{Sarma:2007:TSA} also
-developed a generic design for future indirect conflict awareness tools. 
-However, Palantir falls short in its collection and distribution
-mechanisms. First, Palantir only considers "outside" appearance of technical
-objects, being their return types, parameters, etc. Secondly, Palantir 
-only delivers
-detected conflicts to developers who are presently viewing or editing
-the indirect object while other developers who have used the modified 
-class previously are not notified.
-
-We will lastly examine the tool CASI~\cite{Servant:2010:CPI} which uses
-a sphere of influence for each developer to determine how source code
-changes are indirectly related to other components of a software project.
-CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
-call graphs in Ariadne~\cite{Trainer:2005:BGT} which gives dependencies among
-all source code entities. This provides a verbose output of dependencies when
-source code is changed. CASI also implements a visualization where
-a developer can see what parts of a software projects he or she may be effecting
-with the source code change. This allows the developer themselves to go and fix
-potential issues elsewhere in the project before the code change is committed
-to the software repository. While CASI covers great ground in its approach,
-it still leaves the issue of information overload, although attempts were made
-to solve this by having severity levels of indirect conflicts presented to
-the user.
-
-\textit{Impact} is designed to address these limitations by extending this work in
-two directions: (1) the detection of high importance indirect conflicts at an internal level
-of technical objects, (2) the dissemination of
-awareness information to all appropriate developers regardless of their current
-workspace activities, and (3) potential solutions to information overload. 
-\textit{Impact} is also designed around the successes these
-projects have had in the past with dependency graphs as 
-well as elements of collection, ownership, and distribution functionality.
-
-\section{Impact}
+\subsection{Impact}
 This section will proceed to give an outlined detail of \textit{Impact} in both its
@@
 
-\subsection{Design}
+\subsubsection{Design}
 Compared to tool design for direct conflicts, the major concern of 
@@
 \begin{figure}[t!]
-\includegraphics[width=\columnwidth]{images/CallGraph}
+\includegraphics[width=\columnwidth]{figures/CallGraph}
 \caption{Technical object directed graph with ownership\label{fig:graph}}
@@
 While technical object relationships form the basis of indirect conflicts,
-communication between developers is our ultimate goal of resolving such conflicts.
+communication between developers is my ultimate goal of resolving such conflicts
+(as was seen in Study 1).
 This being the case, developer ownership must be placed on the 
@@
 time~\cite{Fitzpatrick:2002:SPA}.
-We propose the use of commit time
+I propose the use of commit time
 information gathering as it avoids the issue of developers 
@@
 Once technical objects are found to be changed, appropriate owners
-of objects which use the changed object should be notified. However, to
-avoid information overload at this step, analyses of the changed object
-must be preformed to determine whether or not it is a ``notification worthy''
-change. Our proposed solution to this is explained in Section~\ref{sec:ns}.
+of objects which use the changed object should be notified. 
 In Figure~\ref{fig:graph}, Carl changes method (technical object) 1,
 which effects methods 2 and 3 resulting in the alerting of
-developers Bret, Daniel, and Anne. We have opted to alert the invoking
+developers Bret, Daniel, and Anne. I have opted to alert the invoking
 developers rather than the developer making the change to potential
-solutions as our conflicts are detected at commit time and this supports
+solutions as my conflicts are detected at commit time and this supports
 the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} 
@@
 objects, (ii) assigning ownership to those technical objects, and (iii)
-detecting ``notification worthy'' changes at commit time and the dissemination of conflict information
-to appropriate owners, we believe a wide variety of
-indirect conflict awareness tools can be created or extended. The
-on going implementation of \textit{Impact} described in the following 
-section follows these three design guidelines.
+detecting changes at commit time and the dissemination of conflict information
+to appropriate owners, I believe a wide variety of
+indirect conflict awareness tools can be created or extended.
 
-\subsection{Implementation}
-For \textit{Impact's} implementation, we decided to focus on methods as our
+\subsubsection{Implementation}
+For \textit{Impact's} implementation, I decided to focus on methods as my
 selected technical objects to infer a directed graph from. The ``uses'' 
 relationship described above for methods is method invocation.
-Thus, in our constructed dependency graph, methods represent nodes
-and method invocations represent our directed edges. In order to 
+Thus, in my constructed dependency graph, methods represent nodes
+and method invocations represent the directed edges. In order to 
 construct this directed graph, abstract syntax trees (ASTs) are 
@@
 
-Once the directed graph is constructed, we must now assign
-ownership to our technical objects (methods) as per our design.
-To do this, we simply query the source code repository. In our case
-we used Git as the source code repository, so the command \textit{git blame}
+Once the directed graph is constructed, I must now assign
+ownership to the technical objects (methods) as per the design.
+To do this, I simply query the source code repository. In this case
+I used Git as the source code repository, so the command \textit{git blame}
 is used for querying ownership information. (Most source code 
@@
 
-To detect changes to our technical objects (methods), we simply 
+To detect changes to technical objects (methods), I simply 
 use a commit's \textit{diff} which is a representation of all changes
-made inside a commit. We can use the lines changed in the \textit{diff} to 
+made inside a commit. I can use the lines changed in the \textit{diff} to 
 find methods that have been changed. This gives cause of potential
-indirect conflicts. Here is where information overload can occur and must
-be dealt with. Impact's current state does not address this as our potential 
-solution is still under investigation, but is the
-main focus of our next steps as detailed in Section~\ref{sec:ns}.
-We now find all methods in our directed graphs which invoke these changed methods. 
+indirect conflicts. 
+I now find all methods in the directed graphs which invoke these changed methods. 
 These are the final indirect conflicts.
 
-Once the indirect conflicts have been found, we use the
-ownership information of our technical objects to send notifications to
+Once the indirect conflicts have been found, I use the
+ownership information of technical objects to send notifications to
 those developers involved in the indirect conflict. All owners
@@
 invoking methods, and commit message are all displayed. 
-The weight provided will be an indication of importance based
-on the future work to be outline in Section~\ref{sec:ns}. This
-again helps solve the issue of information overload.
+The weight provided is the percent changed of changed method multiplied by 
+ownership of the invoking method. This allows developers to filter
+through high and low changes affecting their own source code.
 
 \begin{figure}[t!]
-\includegraphics[width=\columnwidth]{images/ImpactDemo}
+\includegraphics[width=\columnwidth]{figures/ImpactDemo}
 \caption{\textit{Impact's} RSS type information feed.\label{fig:impact}}
@@
 
-\section{Proposed Solution to Information Overload}
+\subsection{Evaluation}
+To fully evaluate both the generic design of detecting and resolving
+indirect conflicts as well as \textit{Impact}, extensive testing and evaluation
+must be performed. However, I felt that a simple evaluation is
+first needed to assess the foundation of \textit{Impact}'s design and claims
+about indirect conflicts at the method level.
+
+I performed a user case study where I gave \textit{Impact} to 
+two small development teams composed of three developers. Each team was
+free to use \textit{Impact} at their leisure during their development process,
+after which interviews were conducted with lead developers from 
+each development team. The interviews were conducted after each
+team had used \textit{Impact} for three weeks.
+
+I asked lead developers to address two main concerns: do indirect
+conflicts pose a threat
+at the method level (e.g. method 1 has a bug because it invokes method
+2 which has had its implementation changed recently), 
+and did \textit{Impact} help raise
+awareness and promote quicker conflict resolution for indirect
+conflicts. The two interviews largely supported the expectation of
+indirect conflicts posing a serious threat to developers, especially
+in medium to large teams or projects as opposed to the small
+teams which they were apart of. It was also pointed
+out that method use can be a particularly large area for indirect
+conflicts to arise. However, it was noted that
+any technical object which is used as an interface to some data
+construct or methodology, database access for instance, can be 
+a large potential issue for indirect conflicts.  Interview responses to
+\textit{Impact} were optimistically positive, as interviewees stated that \textit{Impact}
+had potential to raise awareness among their teams with what other developers
+are doing as well as the influence it has on their own work. However,
+\textit{Impact} was shown to have a major problem of information overload. 
+It was suggested
+that while all method changes were being detected,
+not all are notification worthy. One developer suggested to only notify
+developers to indirect conflicts if the internal structure of a method
+changes due to modification to input parameters or output parameters.
+In other words, the boundaries of the technical objects (changing
+how a parameter is used inside the method, modifying the return
+result inside the method) seem to be more of interest than other 
+internal workings. More complex inner workings of methods were also noted 
+to be of interest to developers
+such as cyclomatic complexity, or time and space requirements.
+
+These two studies have shown that my design and approach to
+detecting and alerting developers to indirect conflicts appear
+to be on the correct path. However, \textit{Impact} has clearly shown
+the achilles heel of indirect conflict tools, which is information
+overload because of an inability to detect ``notification worthy''
+changes.
+
+\subsection{Analysis}
 \label{sec:ns}
-As was previously stated, the main issue to be addressed with Impact is
-information overload. Our potential solution comes from the ideas of
+In this study, I have proposed a generic 
+design for the development of awareness tools in regards to
+handling indirect conflicts. I have presented a prototype 
+awareness tool, \textit{Impact}, which was designed around the generic 
+technical object approach. However, \textit{Impact} suffered from
+information overload, in that it had too many notifications sent to
+developers.
+
+A potential solution to information overload comes from the ideas of
 Meyer~\cite{Meyer:1988} on ``design by contract''. In this methodology, changes to method
@@
 postconditions have changed.
-We plan to bring these ideas into the realm of what makes a
-change ``notification worthy'', as to deal with information overload for
-developers using Impact. By only examining ``notification worthy'' changes,
-Impact would cut down the number of notifications 
-issued to just those which developers care about. To investigate this solution, we plan to
-interview and survey industry software developers at a wide range of
-experience to determine what makes a change in software important. We believe
-by only monitoring what developers and Meyer~\cite{Meyer:1988} find important to
-indirect conflict awareness, we will be able to eliminate information overload
-in Impact and other dependency awareness tools. We believe that these preconditions
-and postconditions will serve as the main tool for dealing with information overload.
 
-\section{Conclusion and Future Work}
-In this paper, we have presented the issues that arise from indirect 
-conflicts in present awareness tools. We have proposed a generic 
-design for the future development of awareness tools in regards to
-handling indirect conflicts. We have presented a prototype 
-awareness tool, \textit{Impact}, which was designed around our generic 
-technical object approach. We have also proposed a solution to information
-overload which occurs in most indirect conflict awareness tools. We plan
-on investigating our solution further through developer interview and
-surveys, and by finally implementing the solution into Impact.
-
-\bibliographystyle{IEEEtran}
-\bibliography{paper}
-
-%End of paper
-\end{document}
+Aside from better static analysis tools for examining source code changes,
+the results of this study potentially imply a lack of understanding into the
+root causes of indirect conflicts.  A theme of information overload to developers
+continues to crop up in indirect conflicts, of which the root cause should
+be examined in future studies.

### 2013-12-03 msc-thesis/chapters/2/chapter-problem.tex
MSC chapter 2 problem has initial draft.

@@
 
-Introduction paragraph about the problem and the various motivations.
+While the research problems have been briefly outlined in Chapter~\ref{chapter:introduction}, this chapter
+will focus on the underlying studies which motivated the research of this thesis as well as give a more full
+and rich description of the problem being solved.
+
+In this chapter, two studies will be presented that I conducted which motivated, and gave insights into, 
+the final research goals
+of this thesis. The first study entitled ``Failure Inducing Developer Pairs''
+(Section~\ref{study:pairs}), focuses on the prediction of
+software failures through identifying indirect conflicts of developers linked by their software modules. 
+This study found that certain pairs of developers when linked through indirect code changes are more prone
+to software failures than others. The ideas of developer pairs linked in indirect conflicts will be
+useful for the further development of indirect conflict tools as it shows that a human factor is present
+and may be used to help resolve such issues.
+
+The second study, ``Awareness with Impact'' ((Section~\ref{study:impact})), 
+takes the notion of developer
+pairs in indirect conflicts learned from Study 1, and adds in source code change detection in order to create
+an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to a developer
+to any source code changes preformed by another developers when the two are linked in a technical dependency through
+a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications for visualization. 
+While \textit{Impact} showed some promise through its user evaluation, it ultimately suffered the fate of information
+overload as was seen in other indirect conflict tools~\cite{Sarma:2007:TSA,Servant:2010:CPI,Trainer:2005:BGT}.
 
@@
 
-Use the two case studies as motivating examples to show the full problem.
+As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
+development has become the standard development process instead of the exception. With this parallel development
+comes the need for larger awareness among developers to have ``an understanding of the activities of others
+which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
+mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
+empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
+and time-consuming chore for developers~\cite{Perry:2001:PCL}.
+
+Through the two previous studies, I have shown that developers linked indirectly in source code changes can statistically
+related to software failures. In the attempts of mitigating these loses through added awareness, I implemented
+an indirect conflict tool called \textit{Impact}. However, \textit{Impact} ultimately suffered from information
+overload as seen in its evaluation which was caused by false positives and scalability of the tool.
+
+While other indirect conflict tools have shown potential from developer studies, some of the same problems continue
+to arise throughout most, if not all tools. The most prevalent issue is that of information
+overload and false positives. Through case studies, developers have noted that current indirect conflict tools provide too many 
+false positive results leading to information overload and the tool eventually being
+ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
+tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
+invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
+identification of failure inducing changes, other than those which are already identifiable by other means such
+as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
+also compounded by the scale of many software development projects leading to further information overload.
+
+Social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence, have also been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
+However, issues again of information overload, false positives, dependencies (in developer organizational structure), and scalability 
+come up.
+
+While these issues of information overload, false positives, dependencies, and scalability continue to come up
+in most indirect conflict tools, only a handful of attempts have been made at rectifying these issues or
+finding the root causes~\cite{Holmes:2010:CAR,Kim:2011:ESA}. 
+In order to fully understand the root causes of information overload, false positives, and
+scalability issues in regards to indirect conflicts, I will proceed by taking a step back and determine what events occur to
+cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
+causes of source code changes in indirect conflicts, we may be able to create indirect conflict tools which have filtered
+monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
+I then set out to understand what mitigation strategies developers currently use as opposed to those created
+by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
+the same time are not using the tools put forth by academia, I wish to find what they use to mitigate indirect
+conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
+the hopes of a higher adoption rate of tools produced by academia.
+Finally, I look to find what can be accomplished moving forward with indirect conflicts
+in both research and industry.
+
+I restate the research goals of this thesis for ease of the reader:
+
+\begin{description}
+  \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
+  \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
+  \item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
+\end{description}
 

### 2013-12-03 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC chapter 2 problem has initial draft.

@@
 \section{Study 1: Failure Inducing Developer Pairs}
+\label{study:pairs}
 
@@
 
-\subsection{Analysis}
+\subsection{Conclusions of Study}
 Technical dependencies are often used to predict software failures

### 2013-12-03 msc-thesis/chapters/2/sec-impact.tex
MSC chapter 2 problem has initial draft.

@@
 \section{Study 2: Awareness with Impact}
+\label{study:impact}
 
@@
 to predict bugs. The natural follow up to these findings was to conduct a study
-of indirect conflicts surrounding source code changes that involved the source
-code being changed and the developers involved in those changes. These indirect
+of indirect conflicts surrounding these developer pairs that are involved in 
+source code changes. These indirect
 conflicts were primarily studies through the notion of task awareness.
@@
 %Outline of study
-In this study, I report on research into supporting indirect conflicts
+In this study, I report on research into supporting developer pairs in indirect conflicts
 and present the design, implementation, and future evaluation of the tool \textit{Impact},
@@
 
-\subsection{Analysis}
+\subsection{Conclusions of Study}
 \label{sec:ns}

### 2013-12-03 msc-thesis/chapters/3/chapter-new.tex
MSC chapter 2 problem has initial draft.

@@
 
-This is where you go all out and tell us all about your new discovery and research related to the problem in the previous chapter.
-No arrogant sweeping statements which cannot be fully justified, but no false modesty either.
-You must impress your reader that you have accomplished something.
-
-Simply summarized, this chapter should be comprised of at least two main sections, each with appropriate subsections. The first section should describe:
-\begin{itemize}
-\item {what the new approach is;}
-\item {what is really totally new;}
-\item {what is incrementally new;}
-\item {what you built upon.}
-\end{itemize}
-
-The second part should describe fully how the new approach works, both with the overall theoretical exposition (e.g.
-an algorithm) and with as many examples as necessary for clarity.
-Remember that if the reader does not understand fully, you will get a lot of questions and doubts.
-Good examples, good figures, good diagrams with super clear tutorial explanations can be a joy to read and make even a small contribution appear to be more impressive.
-Are you afraid that if you are too tutorial your work will not seem as deep and difficult?
-Only shallow people will make such a superficial evaluation, have trust instead in the wisdom of your supervisory committee.
-
-Use at least one good example throughout, and even better if this is one of the examples you used in Chapter 2 to describe the original problem.
-
-By the way, this would be the first chapter I would write.
-This is what I know best right now, as I just finished working on it.
-It is clear to me and on the tip of my fingers.
-Start with your strengths!
-The second chapter I would write is the next one about the experiments, followed closely by chapter 2 describing the problem.
-It may not seem intuitive to you, but it works and it is the most productive way I ever found to finish a document.
-
 \input chapters/3/sec-latexhelp

### 2013-12-09 iced-2013/paper.tex
Added ic figure and highlighted implications section.

@@
 positive developer feedback. Indirect conflicts involve source code change influences such as making a sour code change 
-which influences behavior negatively in another location in the system. Indirect conflict tools however, have
+which influences behavior negatively in another location in the system (Figure~\ref{fig:ic}). Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 overload. Through case studies, developers have noted that current indirect conflict tools provide too many 
-false positive results leading to information overload and the tool eventually being
+false positive results, leading to information overload and the tool eventually being
 ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
@@
 \label{sec:meth}
+
+\begin{figure*}[tb!]
+\centering
+\includegraphics[width=0.9\textwidth]{images/indirect_conflict.png}
+\caption{An example of an indirect conflict. Bobby has changed method bar() which has negatively affected method
+func() through the chain of invocations.\label{fig:ic}}
+\end{figure*}
 
@@
 
+\textbf{Prevention versus Cure}
 The largest implication for future research found in this paper is the need for continued study of the open question
@@
 
+\textbf{Awareness Theory to Model Real World}
 A second implication for research more directly related to awareness, is the need to further characterize the
@@
 
+\textbf{Give Developers What They Want}
 The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
@@
 
+\textbf{Stronger Source Code Analysis}
 The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
@@
 We have shown through the findings of RQ1 why indirect conflicts occur, when indirect conflicts are more likely to
-occur, as well as what types of software objects are susceptible to these conflicts. Results of RQ 2 have shown
+occur, as well as what types of software objects are susceptible to these conflicts. Results of RQ2 have shown
 how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings of RQ3

### 2013-12-09 impact-uvic-2013/paper.tex
IMPACT completed uvic technical paper.

--- 2013-06-05 impact-position-2013/paper.tex
@@
 \usepackage{amsmath}
+\usepackage{fancyhdr}
 
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\pagestyle{fancy}
+\fancyhead{}
+\renewcommand{\headrulewidth}{0pt}
+\rfoot{University of Victoria Computer Science Technical Paper: DCS-351-IR}
+\fancyfoot[C]{}
+
+\fancypagestyle{firststyle}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[C]{}
+   \rfoot{University of Victoria Computer Science Technical Paper: DCS-351-IR}
+}
 
@@
 \section{Introduction}
+\thispagestyle{firststyle}
 Awareness is characterized as "an understanding of the activities of others
@@
 difficult to resolve as most of the time, source code analysis or
-program slicing~\cite{Weiser:1981:PS} must
+program slicing~\cite{Weiser:1982:PUS} must
 be performed in order to find relationships between technical objects

### 2013-12-11 iced-2013/paper.tex
ICED added reference to impact TR.
ICED added demographic table.

@@
 \usepackage{nameref}
+
+\usepackage{array} % for defining a new column type
+\usepackage{varwidth} %for the varwidth minipage environment
 
@@
 In addition to software development, some participants were also chosen based on their experience with project management
-at some capacity.
+at some capacity. See Table~\ref{tab:demo} for more demographic details.
+
+\begin{table*}[tb!]
+\begin{center}
+\begin{tabular}{| l | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
+\hline
+Company & \# of Participants & Software Development Experience (years) & Development Process & Software Access & Current Language Focuses \\
+\hline
+\hline
+Amazon & 2 & 5, 7 & Agile & Closed source & C++ \\ \hline
+Exporq Oy & 1 & 8 & Agile & Closed source & Ruby, JavaScript \\ \hline
+Fireworks Design & 1 & 6 & Agile & Closed source & JavaScript \\ \hline
+Frost Tree Games & 1 & 4 & Agile & Closed source & C\# \\ \hline
+GNOME & 1 & 13 & Agile & Open source & C \\ \hline
+James Evans and Associates & 5 & 3, 3, 3, 4, 13 & Waterfall & Closed source & Oracle Forms \\ \hline
+Kano Apps & 1 & 10 & Agile & Closed source & JavaScript, PHP \\ \hline
+IBM & 2 & 5, 18 & Agile & Open and closed source & Java, JavaScript \\ \hline
+Microsoft & 2 & 6, 10 & Agile & Closed source & C\# \\ \hline
+Mozilla & 1 & 25 & Agile & Open source & C++, JavaScript \\ \hline
+Ruboss & 1 & 5 & Agile & Closed source & JavaScript \\ \hline 
+Subnet Solutions & 1 & 5 & Agile & Closed source & C++ \\ \hline
+
+\end{tabular}
+\end{center}
+\caption{Demographic information of interview participants.\label{tab:demo}}
+\end{table*}
 
@@
 Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
-which were explored in this paper. Our own tool Impact!~\footnote{https://github.com/jordanell/Impact} also suffered these same fates.
+which were explored in this paper. Our own tool 
+Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
 

### 2013-12-13 iced-2013/paper.tex
ICED added italic blurb before each discussion and implications topic.

@@
 \usepackage{nameref}
-
-\usepackage{array} % for defining a new column type
-\usepackage{varwidth} %for the varwidth minipage environment
 
@@
 \IEEEauthorblockA{University of Victoria,
-Victoria, Canada \\ jell@uvic.ca, danielad@csc.uvic.ca}
+Victoria, Canada \\ jell@uvic.ca, danielad@cs.uvic.ca}
 }
@@
 we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
-source projects, and a survey with 78 responses. Through grounded theory techniques, 
-we found: events and conditions which indirect conflicts are likely to occur in
-software development, developers prefer to use detection and resolution processes or tools
-over those of prevention, developers do not want awareness mechanisms which provide non actionable results,
+source projects, and a survey with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
+we obtained insights about events and conditions in which indirect conflicts are likely to occur in
+software development, that developers prefer to use detection and resolution processes or tools
+over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
 and that there exists a gap in software evolution analytical tools.
@@
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. Indirect conflicts involve source code change influences such as making a sour code change 
-which influences behavior negatively in another location in the system (Figure~\ref{fig:ic}). Indirect conflict tools however, have
+positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
+which influences behavior negatively in another location in the system. Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 
-While indirect conflict tools have shown potential from developer studies, some of the same problems continue
+While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
 to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
-overload. Through case studies, developers have noted that current indirect conflict tools provide too many 
-false positive results, leading to information overload and the tool eventually being
+overload, tools eventually being
 ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
@@
 In order to fully understand the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, we will take a step back and determine what events occur to
+scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
 cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
@@
 
-\begin{figure*}[tb!]
-\centering
-\includegraphics[width=0.9\textwidth]{images/indirect_conflict.png}
-\caption{An example of an indirect conflict. Bobby has changed method bar() which has negatively affected method
-func() through the chain of invocations.\label{fig:ic}}
-\end{figure*}
-
-We performed a a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
-addressed the 3 research questions as mentioned above. Second, a survey was conducted
+We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
+addressed our 3 main research questions. Second, a survey was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
 larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
-interview participants to once again confirm and test what had been learned and the results derived by the paper.
-Grounded theory techniques were used to analyze the information provided from all three data gathering processes.
-
-\subsection{Grounded Theory}
-We approached our study based on grounded theory as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
-Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
-\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
-the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
-making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
-realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
-questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
-both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
-assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
-codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
-link our defined codes through categories of grounded theory such as context and consequences. 
-In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
-3 criteria that are commonly used in evaluating grounded theory studies.
+interview participants to once again confirm our insights.
+We used grounded theory techniques to analyze the information provided from all three data gathering stages.
 
@@
 
-We selected our interview participants from a large breadth of both open and closed source software development companies
-and projects as well as Agile and Waterfall based projects.
-The population which participated in our interview came from the following software groups: 
-IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
+Our interview participants came from a large breadth of both open and closed source software development companies
+and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation,
 Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
-Our participants were chosen based on their direct involvement in the actual writing of software for
+Our participants were invited based on their direct involvement in the actual writing of software for
 their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
 with an average of 8 years of experience.
-In addition to software development, some participants were also chosen based on their experience with project management
+In addition to software development, some participants were also invited based on their experience with project management
 at some capacity. See Table~\ref{tab:demo} for more demographic details.
@@
 Participants were invited to be interviewed by email and were sent a single reminder email one week
-after the initial invitation if no response was made. We directly emailed 22 participants and ended up conducting
+after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
 19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
 interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
-audio being stored for future use. We were pleased with the response rate to our initial invitation emails as
-our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
-and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.
+audio being stored for future use.
 
-Interview participants first answered a number of structured demographic items, followed by being  
-asked to describe various software development experiences regarding our three research questions.
-Our three questions were studied by having the interview participants talk about their experiences and opinions in
-\totalCategories{} semi-structured research topics.
+%We were pleased with the response rate to our initial invitation emails as
+%our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
+%and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.
 
-While each of the \totalCategories{} question categories had a number of starter questions, interviews 
+Interview participants first answered a number of  demographic questions. We then asked them to describe various software development experiences regarding our three research questions.
+Specifically, ten semi-structured  topics directly related to our research questions guided our interview:
+
+%Our three questions were studied by having the interview participants talk about their experiences and opinions in
+%\totalCategories{} .
+
+\begin{itemize}
+
+\item Software development tools for dependency tracking and awareness.
+
+\item Software development process for preventing indirect conflicts.
+\item Software artifact dependency levels and where conflicts can arise.
+
+\item How developers find internal or external software dependencies.
+
+\item Examples of indirect conflicts from real world experiences.
+\item How indirect conflicts are detected and found.
+
+\item How indirect conflict issues are solved or dealt with.
+\item Developer opinion of preemptive measures to prevent indirect conflicts.
+
+\item Developer opinion on what types of changes are worth a preemptive action.
+
+\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
+\end{itemize}
+
+While each of the \totalCategories{} topics had a number of starter questions, interviews 
 largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
@@
 either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
-being apart of any one software project for extended periods of time. We account for these non answers
-in our analysis and results as seen in Section~\ref{sec:results}.
+being apart of any one software project for extended periods of time.
+
+%We account for these non answers
+%in our analysis and results as seen in Section~\ref{sec:results}.
 
@@
 
-We selected our survey participants from a similar breadth of open and closed source software development 
-companies and projects as the interviews participants with two large exceptions. The software organizations
+Our survey respondents were different from our interviewees. We invited our survey participants from a similar breadth of open and closed source software development 
+companies and projects as the interviews participants with two main exceptions. The software organizations
 that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
 Subnet Solutions, and Amazon.
-Participants who took part in the round of interviews were asked to act as a contact point for other 
+Participants who took part in the round of interviews were not invited to the survey but were asked to act as a  point of contact for other 
 developers in their team, project, or organization who may be interested
-in completing the survey. Aside from this aforementioned list, two other groups of developers were asked to
+in completing the survey. Further, two other groups of developers were asked to
 participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
@@
 
-The created survey~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
-was based off of hypotheses created by the round of interviews. The survey
-was designed to test these hypotheses and to acquire a larger sample size of developers who may have similar
+The  survey we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
+was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
 or different opinions from those already acquired from the interviews. The survey went through two rounds of
@@
 with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
-the previously interviewed developers to examine what hypotheses were
-formed and what we would be moving forward with.
-Each non demographic question was made optional as it was shown 
-through the interviews that some questions require more experience from participants than may be provided.
+the previously interviewed developers to examine the insights we developed.
+%Each non demographic question was made optional as it was shown 
+%through the interviews that some questions require more experience from participants than may be provided.
+
+\subsection{Data Analysis}
+To analyze both the interview and survey data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
+Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
+\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
+the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
+making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
+realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
+questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
+both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
+assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
+codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
+link our defined codes through categories of grounded theory such as context and consequences. 
+In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+3 criteria that are commonly used in evaluating grounded theory studies.
 
@@
 
-\textbf{Unwanted Awareness} As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
+\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
+a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
+to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
+awareness about most, if not all actions.}
+
+As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
 sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
@@
 
-\textbf{Prevention versus Cure} We have seen through the results, that developers posses both tools and practices for the prevention,
+\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
+arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
+to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.}
+
+We have seen through the results, that developers posses both tools and practices for the prevention,
 detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
@@
 %mature point, static analysis
-\textbf{Gap in Software Evolution Analysis} Indirect conflicts are more likely to occur,
+\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
+and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
+some software project configurations and software languages to be quite prone to indirect conflicts.}
+
+Indirect conflicts are more likely to occur,
 from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
@@
 
-\textbf{Prevention versus Cure}
-The largest implication for future research found in this paper is the need for continued study of the open question
+\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
 ``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly 
-produce extremely complex answers. With software developers focusing their current efforts on curative measures, for
+produce extremely complex answers.}
+
+With software developers focusing their current efforts on curative measures, for
 indirect conflicts, suggests
@@
 
-\textbf{Awareness Theory to Model Real World}
-A second implication for research more directly related to awareness, is the need to further characterize the
+\textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
 mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
-by industry.
+by industry. We should push our awareness theories to model real development
+environments where interruptions, time management, and full development life cycles are often large factors.}
+
 As was identified in the previous discussion, while current understanding suggests that awareness of all events related to ones work
@@
 
-\textbf{Give Developers What They Want}
+\textbf{Give Developers What They Want} \textit{Developers have their own notions of what they want, and how ``good''
+that something is for their productivity. We should pay attention to what they want, but also be mindful that what
+developers want may not always be good for them.}
+
 The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
@@
 
-\textbf{Stronger Source Code Analysis}
+\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
+multiple languages, and often having a database schema, we should focus on better techniques for static analysis based
+on what industry standards are for languages and project structures.}
+
 The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing

### 2014-01-05 msc-thesis/chapters/2/chapter-motivating.tex
MSC Added all 4 studies to the thesis.

--- 2013-12-03 msc-thesis/chapters/2/chapter-problem.tex
@@
-\startchapter{The Problem}
-\label{chapter:problem}
+\startchapter{Motivating Studies}
+\label{chapter:motivating}
 
@@
 
-\section{Problem Description}
-
-Use the two case studies as motivating examples to show the full problem.
-
 \setlength{\unitlength}{\savedunitlength}

### 2014-01-05 msc-thesis/chapters/3/chapter-exploration.tex
MSC Added all 4 studies to the thesis.

--- 2013-12-03 msc-thesis/chapters/2/chapter-problem.tex
@@
-\startchapter{The Problem}
-\label{chapter:problem}
-
-\newlength{\savedunitlength}
-\setlength{\unitlength}{2em}
-
-While the research problems have been briefly outlined in Chapter~\ref{chapter:introduction}, this chapter
-will focus on the underlying studies which motivated the research of this thesis as well as give a more full
-and rich description of the problem being solved.
-
-In this chapter, two studies will be presented that I conducted which motivated, and gave insights into, 
-the final research goals
-of this thesis. The first study entitled ``Failure Inducing Developer Pairs''
-(Section~\ref{study:pairs}), focuses on the prediction of
-software failures through identifying indirect conflicts of developers linked by their software modules. 
-This study found that certain pairs of developers when linked through indirect code changes are more prone
-to software failures than others. The ideas of developer pairs linked in indirect conflicts will be
-useful for the further development of indirect conflict tools as it shows that a human factor is present
-and may be used to help resolve such issues.
-
-The second study, ``Awareness with Impact'' ((Section~\ref{study:impact})), 
-takes the notion of developer
-pairs in indirect conflicts learned from Study 1, and adds in source code change detection in order to create
-an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to a developer
-to any source code changes preformed by another developers when the two are linked in a technical dependency through
-a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications for visualization. 
-While \textit{Impact} showed some promise through its user evaluation, it ultimately suffered the fate of information
-overload as was seen in other indirect conflict tools~\cite{Sarma:2007:TSA,Servant:2010:CPI,Trainer:2005:BGT}.
-
-\input chapters/2/sec-developer-pairs
-\input chapters/2/sec-impact
-
-\section{Problem Description}
+\startchapter{Exploring Indirect Conflicts}
+\label{chapter:exploration}
 
@@
 \begin{description}
-  \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
-  \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
-  \item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
+  \item \textit{What events or conditions lead to indirect conflicts?}
+  \item \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
+  \item \textit{What do developers want from future indirect conflict tools?}
 \end{description}
 
-\setlength{\unitlength}{\savedunitlength}
+To answer these research questions, I performed a study (Section~\ref{study:exploration}) in which I interviewed 19 developers from across 12 commercial and open source projects, 
+followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews. Based on the findings (to be seen) to the aforementioned research questions, 
+I also performed a secondary complimentary study of software evolutionary trends (Section~\ref{study:apie}) to solidify
+developer opinion and provide a starting point for the future development of indirect conflict tools.
+
+\input chapters/3/sec-iced
+\input chapters/3/sec-apie

### 2014-01-05 msc-thesis/chapters/3/sec-apie.tex
MSC Added all 4 studies to the thesis.

\section{Study 4: Investigating Indirect Conflict Contextual Patterns}
\label{study:apie}

Release points are a vital milestone of software projects. From major releases of a Waterfall based project to minor iterations of an Agile development,
releases form an interesting single point of a project's development history.
Third party users (outside developers) of a system often only see a product at a release point either major or minor,
and expect the system to come with a sense of reliability and stability at this point. Developers often expect to be able to upgrade a library to
a newer version without having to make major revisions to their own projects to accommodate the upgrade (unless of course patch notes detailing major
changes are released). However, major and minor releases of a library or software resource can cause software quality to degrade in
a third party application as indirect conflicts may occur. The knowledge as to when a project is ready for public usage
as to its reliability, quality, stability, and thus indirect conflict proneness can be a difficult decision to make for most project owners or maintainers.

While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
the sub study of software stability and its implications on reliability and indirect conflict proneness remains a difficult subject to understand. 
The decision of what makes a project stable and ready
for a release often comes down to the release manager or maintainer of a project and is often a reflection of the open source community which surrounds 
the project~\cite{Conway:1968}. Code churn is an often looked to statistically for stability but can be grossly misleading
in terms of pre-release and post-release defects~\cite{Fenton:2000:QAF}, with some exceptions~\cite{Nagappan:2005:URC} as well as proneness to indirect conflicts
both internally and externally to third party applications. 
Creating an approach to determining software stability, release preparedness, and the proneness of indirect conflicts is still a large open area of
interest in software engineering research.

I turn my analysis to the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts, this study investigates the question:
``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
stability measure?}''.

In this study, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near and major release and more
frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.

\subsection{Methodology}
\label{sec:apie-meth}
In order to answer my research question, I decided to use the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows me to detect fine grained
source code changes in Java projects. This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
the smallest possible edit distance between the trees. This results in the source code change at a fine grained level performed in the commit.

I took ChangeDistiller and applied it across 10 open source Java projects. Java projects were necessary because ChangeDistiller only works for Java source code.
For each of the projects, I obtained the software configuration management (SCM) system
which is used to store all source code changes of a project. When it was necessary, I converted some forms of SCM system to Git in order to reduce implementation
burdens of using multiple SCMs. Once the SCM was obtained, I used ChangeDistiller and iterated over every commit found in a project's git master branch. I stored
34 of ChangeDistiller's built in source code change types for each commit. I noted how many of each change type was performed in each commit and stored that information
in a PostgreSQL database. In order to filter and protect my results, I manually inspected the 10 Java projects studied in order to identify code built for test
purposes.
I separated changes to this test code from all other code to ensure my results only focused on real implementation while allowing us to study changes to
test based code separately.

Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the project's I had selected. Since releases
have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
I decided that by studying the change types surrounding these releases, I could get a better understanding of
what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of our 10 project's 
web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
across my 10 studied projects.

Once the release dates were collected, I set about analyzing my data by creating average change ratios surrounding the release dates of each 
project as a way to measure the trend of a particular change type at a release type. This change ratio simply compares the number of change events (of a given
change type) before a release to after the release. Both of the before and after event totals are divided by the number of commits on their respective side
of the release to account for activity. I implemented this algorithm through Equation~\ref{eq:norm}.

Equation~\ref{eq:norm} works to create a change ratio by first creating a numerator by summing across all releases of
a given release type a sum of a particular change type in commits ($T_c$)
from the release date ($r$) to a given number of days after the release ($d$) divided by the number of commits in this date range ($|c|$). Next the denominator
is created by summing across all release of a given release type
a sum of that same particular change type in commits ($T_c$) from a given number of days ($d$) before the release date ($r$) to
the release date divided by the number of commits in this date range ($|c|$). This numerator and denominator form the final change ratio.
This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of our study, we set the
number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
formed much of the basis for the results to come in Section~\ref{sec:apie-results}

\begin{equation}
\text{ChangeRatio} = \frac{ \sum_{r_0}^{r_n}\sum_{c=r}^{r+d} T_c / |c|} { \sum_{r_0}\sum_{c=r}^{r-d} T_c / |c|}
\label{eq:norm}
\end{equation}

Aside from generating quantitative data, I also created a web application for the visualization of the data called API Evolution (APIE). This visualizer allowed
me to inspect a single project and a single change type metric at a time (see Figure~\ref{fig:apie}) for qualitative analysis of software evolution trends. I
used this tool to manually inspect 4 specific change type trends surround release dates. To do this, I aggregated change types across 50 commits, meaning that
each point in the graph represented the date of a commit and the sum of the particular change type's occurrences over the last 50 commits. This was used to smooth
out the curves presented by the tool to allow easier manual inspection. This method however does not take activity into account as seen in
Equation~\ref{eq:norm}, so it represents the true activity and change types occurring. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
trending, and local minimum. Since the graphs were quite turbulent, best estimations were conducted by two judges at each release point to fit the graph into the aforementioned
4 slope categories. The two judges used 1.5 months before and after the release date as start and end points for the graph trend line.

% APIE graph figure
\begin{figure}[tb!]
\centering
\includegraphics[width=0.9\textwidth]{figures/APIE.png}
\caption{A screen shot of the APIE visualizer showing project Eclipse.Mylyn.Context with change type PUBLIC\_ADDED\_METHODS being analyzed and showing
major releases as vertical yellow lines.\label{fig:apie}}
\end{figure}

I performed 1888 manual inspections across 10 projects, 472 release dates and 4 change types, and used this data to form the basis of my qualitative data.
Quantitative data was used to compliment the quantitative ratios found from the previous methodology.

\subsection{Results}
\label{sec:apie-results}

To answer my research question, I conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.

Due to time requirements, I focus my results on major releases of the 10 case study projects and select few of the calculated change ratios. 
There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
can be seen in Table~\ref{tab:apie-ratio}.

\begin{table}[ht]
\begin{center}
\tabcolsep=0.11cm
\begin{tabular}{| l | c | c | c |}
\hline
Object & Added & Changed & Removed\\
\hline
Public Classes & 1.14 & 0.86 & 1.16 \\
Public Methods (Signature) & 1.07 & 0.92 & 1.34 \\
Public Methods (Bodies) & - & 1.06 & - \\
Private Classes & 0.81 & 1.18 & 1.44 \\
Private Methods (Signatures) & 1.00 & 1.10 & 1.22 \\
Private Methods (Bodies) & - & 1.08 & - \\
Files & 1.12 & 0.96 & 1.14 \\
Documentation & - & 0.99 & - \\
\hline
\end{tabular}
\end{center}
\caption{Implementation oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:apie-ratio}}
\end{table}

To study the most prevalent change trends, I set a ratio threshold of greater than 1.2, or less than 0.83 (20\% greater trend of after the release date
or 20\% greater trend of before the release date) to indicate the greatest trends.

As it can be seen in Table~\ref{tab:apie-ratio}, there are few change type trends around major releases which pass my threshold. We can see that both public
and private methods
being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:apie-ratio} also shows significance in the changes to private
classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
All results in Table~\ref{tab:apie-ratio} could be used as identified trends of major software releases, while I have just highlighted the larger ratios
which meet my threshold criteria.

Another interesting trend that can been seen in Table~\ref{tab:apie-ratio} is that of changes to public objects. We can see for public classes and methods that
5 out of 7 ratios indicate changes occur to these objects after major release rather than before. I hypothesize that these changes to the public API
after a major release come from newly reported bugs from end users as well as having old features being deprecated while adding new features to the project
after the stable build had been released.

My complementary qualitative results from manual graph inspections can be seen in Table~\ref{tab:qual}. These results show that adding, changing signatures
and bodies of, and removing public methods tend to all be at a local minimum of change type trends at major releases when activity is not taken into
account. These results confirm previous results of low code churn as an indication of stability.

\begin{table*}[ht]
\begin{center}
\begin{tabular}{| p{2cm} | c | c | c | c |}
\hline
Change Type & Upward Trend & Local Maximum & Downward Trend & Local Minimum\\
\hline
Added Public Methods & 21.6\% & 17.2\% & 14.7\% & 33.6\% \\ \hline
Changed Public Methods (Signature) & 6.0\% & 19.8\% & 19.0\% & 39.7\% \\ \hline
Changed Public Methods (Bodies) & 9.2\% & 16.5\% & 26.6\% & 37.6\% \\ \hline
Removed Public Methods & 7.8\% & 16.4\% & 12.9\% & 41.4\% \\ \hline
\end{tabular}
\end{center}
\caption{Qualitative graph analysis results. \label{tab:qual}}
\end{table*}

Lastly, I found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet our threshold and are indicators of stability with regards to test based
changes are: the changing of test classes, the removal of test classes, the adding of methods, and the changing of method signatures, and test classes being changed.
Changes to documentation also meets my threshold and occurs more before a major release.

\begin{table}[ht]
\begin{center}
\begin{tabular}{| l | c | c | c |}
\hline
Object & Added & Changed & Removed\\
\hline
Classes & 1.07 & 1.21 & 0.76 \\
Methods (Signatures) & 1.23 & 0.83 & 1.01 \\
Methods (Bodies) & - & 0.90 & - \\
Documentation & - & 0.72 & - \\
\hline
\end{tabular}
\end{center}
\caption{Test oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:test}}
\end{table}

While all change ratios may need to be considered for continued analysis or a taxonomy of change trends, I have offered the strongest
change trends in these results as a suggestion for future focus.

\subsection{Conclusions of Study}

In this study, I have conducted a case study of 10 open source Java software projects in order to study their change trends surrounding
major releases as previous studies have shown that indirect conflicts occur less at a major release and more so after a major release
or as the beginning of a development cycle. I have presented here 9 major change trends which surround major releases in the open source
projects studied. The 4 change trends found in this study occurring before major releases are: added private classes, 
changed test method signatures, changed documentation, and removed test classes.
The 5 change trends found in this paper occurring after major release are: added test methods, changed test classes, removed public methods, removed
private classes, and removed private methods.

These 9 change trends can be used in future indirect conflict tools in order to identify a context for indirect conflicts.
For example. Any of the 9 change trends which occur more so after a major release may be used as a sign of high indirect conflict proneness
since after a major release a new development cycle is likely to begin. As per change trends which occur more so before a major release, indirect
conflicts may use this context in order to identify a low proneness to indirect conflicts in their results. These two contextual patterns
can be used throughout the life of a software project in order to help better inform indirect conflict tools as to the processes of developers
and provide better feedback to said developers of indirect conflicts.

Aside from contextual patterns for indirect conflicts, this study has also shown the beginnings of a visualization for source code change trends which may
be used as a visual cue towards project stability and potential areas of instability where action may need to be taken.

### 2014-01-05 msc-thesis/chapters/3/sec-iced.tex
MSC Added all 4 studies to the thesis.

--- 2013-12-13 iced-2013/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% Use of outside images
-\usepackage{graphicx} 
-% Use text inside euqations
-\usepackage{amsmath}
-
-%\usepackage{balance}
-\usepackage{float}
-
-\usepackage{hyperref}
-\usepackage{nameref}
-
-\floatstyle{plaintop}
-\restylefloat{table}
-
-% Correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\makeatletter
-\def\namedlabel#1#2{\begingroup
-   \def\@currentlabel{#2}%
-   \label{#1}\endgroup
-}
-\makeatother
-
-\newcommand{\totalCategories}{10}
-
-% Begin the paper here
-\begin{document}
-
-% Paper title
-% Can use linebreaks \\ within to get better formatting as desired
-\title{Sell Prevention and Buy Cure: An Exploration of Indirect Conflicts}
-
-\author{\IEEEauthorblockN{Jordan Ell and Daniela Damian}
-\IEEEauthorblockA{University of Victoria,
-Victoria, Canada \\ jell@uvic.ca, danielad@cs.uvic.ca}
-}
-
-% Make the title area
-\maketitle
-
-\begin{abstract}
-
-Awareness techniques have been proposed and studied to aid developer
-understanding, efficiency, and quality of software produced. Some of these techniques have focused 
-on either \textit{direct} or 
-\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
-from a result of source code changes. While the techniques and tools for direct conflicts have had
-large success, tools either proposed or studied for indirect conflicts have had common issues of
-information overload, false positives, and scalability. To better understand these issues, 
-we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
-source projects, and a survey with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
-we obtained insights about events and conditions in which indirect conflicts are likely to occur in
-software development, that developers prefer to use detection and resolution processes or tools
-over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
-and that there exists a gap in software evolution analytical tools.
-
-\end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
-development has become the standard development process instead of the exception. With this parallel development
-comes the need for larger awareness among developers to have ``an understanding of the activities of others
-which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
-mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
-empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
-and time-consuming chore for developers~\cite{Perry:2001:PCL}.
-
-Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
-\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
-artifact. Tools have been created and studied for direct conflicts
-~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
-which influences behavior negatively in another location in the system. Indirect conflict tools however, have
-not shared the same success as direct conflict 
-tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.
-
-While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
-to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
-overload, tools eventually being
-ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
-tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
-invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
-identification of failure inducing changes, other than those which are already identifiable by other means such
-as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
-also compounded by the scale of many software development projects leading to further information overload.
-Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
-congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
-However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
-come up.
+\section{Study 3: An Exploration of Indirect Conflicts}
+\label{study:exploration}
 
 In order to fully understand the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
+scalability issues in regards to indirect conflicts, I conducted an empirical study to determine what events occur to
 cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
-We then set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
-in both research and industry. To accomplish these goals, we ask the following:
+I then set out to understand what mitigation strategies developers currently use as opposed to those created
+by academia. Through this exploration, I look to find what can be accomplished moving forward with indirect conflicts
+in both research and industry.
 
-\begin{description}
-	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
-	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
-	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
-\end{description}
-
-We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
-developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
+I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
 indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
@@
 
-\section{Methodology}
-\label{sec:meth}
+\subsection{Methodology}
+\label{sec:exp-meth}
 
-We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
-addressed our 3 main research questions. Second, a survey was conducted
+I performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
+addressed my 3 main research questions. Second, a survey was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
 larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
-interview participants to once again confirm our insights.
-We used grounded theory techniques to analyze the information provided from all three data gathering stages.
+interview participants to once again confirm my insights.
+I used grounded theory techniques to analyze the information provided from all three data gathering stages.
 
-\subsection{Interview Participants}
+\subsubsection{Interview Participants}
 
-Our interview participants came from a large breadth of both open and closed source software development companies
+My interview participants came from a large breadth of both open and closed source software development companies
 and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation,
 Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
-Our participants were invited based on their direct involvement in the actual writing of software for
+My participants were invited based on their direct involvement in the actual writing of software for
 their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
@@
 \begin{center}
-\begin{tabular}{| l | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
+\begin{tabular}{| p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
 \hline
@@
 Subnet Solutions & 1 & 5 & Agile & Closed source & C++ \\ \hline
-
 \end{tabular}
@@
 
-\subsection{Interview Procedure}
+\subsubsection{Interview Procedure}
 
 Participants were invited to be interviewed by email and were sent a single reminder email one week
-after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
+after the initial invitation if no response was made. I directly emailed 22 participants and  conducted
 19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
@@
 
-%We were pleased with the response rate to our initial invitation emails as
-%our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
-%and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.
-
-Interview participants first answered a number of  demographic questions. We then asked them to describe various software development experiences regarding our three research questions.
+Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding our three research questions.
 Specifically, ten semi-structured  topics directly related to our research questions guided our interview:
-
-%Our three questions were studied by having the interview participants talk about their experiences and opinions in
-%\totalCategories{} .
 
@@
 
-While each of the \totalCategories{} topics had a number of starter questions, interviews 
+While each of the 10 topics had a number of starter questions, interviews 
 largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
 However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
-to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
+to the specific categories were not required or pressed upon. I attribute any non answer by a participant to
 either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
@@
 
-%We account for these non answers
-%in our analysis and results as seen in Section~\ref{sec:results}.
+\subsubsection{Survey Participants}
 
-\subsection{Survey Participants}
-
-Our survey respondents were different from our interviewees. We invited our survey participants from a similar breadth of open and closed source software development 
+My survey respondents were different from my interviewees. I invited our survey participants from a similar breadth of open and closed source software development 
 companies and projects as the interviews participants with two main exceptions. The software organizations
@@
 
-\subsection{Survey Procedure}
+\subsubsection{Survey Procedure}
 
@@
 survey responses were not connected with the invitation email addresses and thus participants who did respond
-could not be identified. We directly emailed 1300 participants and ended with 78 responses
-giving a response rate of 6\%. We attribute the low response rate with: the surveys
+could not be identified. I directly emailed 1300 participants and ended with 78 responses
+giving a response rate of 6\%. I attribute the low response rate with: the surveys
 were conducted during the months of July and August while many participants may be away from their regular positions.
-and our GitHub and Apache participants could not be verified as to whether or not they actively support the
+and my GitHub and Apache participants could not be verified as to whether or not they actively support the
 email addresses used in the invitations. In addition, the survey was considered by some to be long and require
@@
 
-The  survey we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
-was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
+The survey I designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
+was based on the insights I obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
 or different opinions from those already acquired from the interviews. The survey went through two rounds of
 piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
-with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
-the previously interviewed developers to examine the insights we developed.
-%Each non demographic question was made optional as it was shown 
-%through the interviews that some questions require more experience from participants than may be provided.
+with feedback given at the end. Not only did this allow me to create a more polished survey, but it also allowed 
+the previously interviewed developers to examine the insights I developed.
 
-\subsection{Data Analysis}
-To analyze both the interview and survey data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
+\subsubsection{Data Analysis}
+To analyze both the interview and survey data, I used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
 Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
 \textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
-the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
+mine, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
 making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
 realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
-questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
+questions that have been formulated in regards to forming hypotheses. In my study being presented, data collected from
 both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
 assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
-codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
-link our defined codes through categories of grounded theory such as context and consequences. 
-In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
+codes were defined as the study progressed and different hypotheses began to grow. I finally use axial coding it order to
+link my defined codes through categories of grounded theory such as context and consequences. 
+In Section~\ref{sec:exp-eval}, I give a brief evaluation of my studying using 
 3 criteria that are commonly used in evaluating grounded theory studies.
 
-\subsection{Validation}
+\subsubsection{Validation}
 
-Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
-in order to validate our findings. We confirmed our findings as to whether or not they resonate with 
+Following my data collection and analysis, I re-interviewed 5 of my initial interview participants
+in order to validate my findings. I confirmed my findings as to whether or not they resonate with 
 industry participants' opinions and experiences regarding indirect conflicts and as to their industrial 
-applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
+applicability. Due to limited time constraints of the interviewed participants, I could only re-interview
 five participants. Those that were re-interviewed came from the range of 5-10 years of software development
-experience. Re-interviewed participants were given our 3 research questions along with results and main
-discussion points, and asked open ended questions regarding their opinions and experiences to validate our 
-findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
-list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:eval}
+experience. Re-interviewed participants were given my 3 research questions along with results and main
+discussion points, and asked open ended questions regarding their opinions and experiences to validate my 
+findings. I also evaluated my grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
+list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:exp-eval}
 
-\section{Results}
-\label{sec:results}
+\subsection{Results}
+\label{sec:exp-results}
 
-We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
-outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
-results from which we draw our discussion to be seen in Section~\ref{sec:disc}.
+I now present my results of both the interviews and surveys conducted in regards to my 3 research questions
+outlined in this chapter and Chapter 1. I restate each research question, followed by my quantitative and qualitative
+results from which I draw my discussion to be seen in Chapter~\ref{chapter:discussion}.
 
-\subsection{What events or conditions lead to indirect conflicts?}
+\subsubsection{What events or conditions lead to indirect conflicts?}
 
@@
 
-\begin{table*}[tb!]
+\begin{table*}[ht]
 \begin{center}
-\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
+\begin{tabular}{| p{1.6cm} | c | c | c | c | c | c | c |}
 \hline
@@
 \hline
-Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
+Early stages of development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
 Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
@@
 
-\subsection{What mitigation techniques are used by developers in regards to indirect conflicts?}
+As per organizational structure, Table~\ref{tab:env} shows which development team environment developers believe to be the most
+prone to indirect conflicts.
+
+\begin{table*}[ht]
+\begin{center}
+\begin{tabular}{| p{3cm} | c | c | c | c | c |}
+\hline
+Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
+\hline
+\hline
+Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
+Developing in a team between 2 - 5 developers (Inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
+Developing in a multi team environment (Inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Results of survey questions to development environments in which indirect conflicts are likely to occur, in terms of percentage
+of developers surveyed.\label{tab:env}}
+\end{table*}
+
+\subsubsection{What mitigation techniques are used by developers in regards to indirect conflicts?}
 
@@
 
-\subsection{What do developers want from future indirect conflict tools?}
+\subsubsection{What do developers want from future indirect conflict tools?}
 
@@
 interface change. Other change types varied from never being notified to most times being notified, showcasing the
-complexity of change types which may or may not negatively affect a project.
+complexity of change types which may or may not negatively affect a project. A complete breakdown of change types and
+developer preference can be found in Table~\ref{tab:pre}.
+
+\begin{table*}[h]
+\begin{center}
+\begin{tabular}{| p{3cm} | c | c | c | c | c |}
+\hline
+Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
+\hline
+\hline
+Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
+Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
+Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
+User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
+Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
+Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Results of survey questions to source code changes that developers deem notification worthy, in terms of percentage
+of developers surveyed.\label{tab:pre}}
+\end{table*}
 
@@
 
-\section{Discussion}
-\label{sec:disc}
+\subsection{Evaluation}
+\label{sec:exp-eval}
 
-\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
-a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
-to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
-awareness about most, if not all actions.}
-
-As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
-sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
-or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
-to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
-leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
-information being presented to them, rendering the system useless. These false positives are caused by a difference in what
-developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
-be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
-where as current awareness understanding would state any event which is related to an individual's 
-work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
-``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
-showing that not all changes around an object should be reported for awareness.
-
-Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
-their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
-developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
-don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
-that the developer only wants the awareness if the item will require action on their part to resolve the issue.
-This sense of unwanted or limited awareness is crucial to understand why generalized
-awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
-classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
-positives and information overload. 
-Developers simply do not want to know about events which effect them, but require no action on their part.
-
-This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
-being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
-but as it was found in practice, this is an incorrect assumption. 
-In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for 
-analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
-all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
-the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
-of events occurring around their work which pose no direct threat to the consumption of their time.}
-
-Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
-developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
-them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
-influence of total, or near total
-awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
-with higher quality or understanding of the product. 
-A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
-producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
-however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
-be more productive.
-
-\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
-arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
-to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.}
-
-We have seen through the results, that developers posses both tools and practices for the prevention,
-detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
-that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
-may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
-is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
-Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
-place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
-them in the first place.
-
-Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
-accomplished essentially by
-not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
-developers are spending little to no time in prevention. Developers do however spend a large amount of time in detection and cure through the
-writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
-conflicts occurred at either the detection or cure levels. Obviously, developers either prefer to spend their time in curative
-measures, or do not posses the proper tools to take better action in the preventative stages. ``You're reaction time is much more crucial''
-said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
-matter what prevention was taken, the issue must simply be resolves as quick as possible now.
-
-The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
-Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
-which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
-of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
-are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
-abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
-real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
-this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
-curative measures are more productive than preventative measures.
-
-Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
-to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
-in that they result in a trade off 
-of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
-it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
-not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
-to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
-Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.
-
-A last interesting observation about current curative measures being taken in industry, is that developers view testing of software 
-as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
-are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
-thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
-However, if a test never fails, could it not be said that it is preventing changes from causing it to fail rather than detecting
-when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
-measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
-``prevention is the goal'' was commonly said by developers.
-
-This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
-designers, architects, or managerial stakeholders. 
-From the project manager's interviewed it was shown that they are heavily favored towards planning and 
-prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
-approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
-is being viewed in research.
-
-%mature point, static analysis
-\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
-and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
-some software project configurations and software languages to be quite prone to indirect conflicts.}
-
-Indirect conflicts are more likely to occur,
-from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
-better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
-end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
-mature point, as well as any other
-potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
-that different change types may occur at different rates throughout a project's life time and that this
-may drastically effect the outcomes of indirect conflict tools or processes.
-
-A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
-research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
-could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
-means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
-Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
-changes occur at a given point in time which could be used for determining such a context of mature points.
-
-This added level of context, through the notion of a mature point, 
-would add to the prevention versus cure debate as previously discussed. Prevention may be a better
-choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
-to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
-causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
-and the mature point in order to fully understand the nature of indirect conflicts and their context.
-
-While this added sense of understanding is important to indirect conflicts, contextual attributes of a project's current progress
-or measures of performance are often more synonymous with project management
-rather than software developers themselves. An understanding of a project's evolution, a mature point being reached as an example,
-may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
-conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.
-
-Aside from context, dependency identification and tracking is a key missing component of
-indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
-organizational structures of software teams. Between increased modularization (multiple sub projects or repositories), 
-cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
-become quite limited in identifying and tracking dependencies where they were once strong.
-
-Relational database schemas are one of the highest sources of indirect
-conflicts found in their projects, ``it breaks stuff all over the place'',
-yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
-been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
-dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.
-
-These increases of complexity in software products has left quite a gap in dependency identification and tracking which 
-has lead to some of the deficiencies of indirect conflict research.
-
-\subsection{Implication for Research}
-\label{sec:implr}
-
-\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
-``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly 
-produce extremely complex answers.}
-
-With software developers focusing their current efforts on curative measures, for
-indirect conflicts, suggests
-that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
-least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
-provide more insight into possibilities, as its answer may determine many future actions taken by the research community.
-
-Recommendations towards the study of prevention versus cure involve the examination of formal processes and tools used
-by industry professionals with measurements of efficiency and effectiveness, similar to the work of
-Tiwana~\cite{Tiwana:2008:ICD} in coordination tools. With these studies, we may find insight into the correct balance
-of prevention versus cure, thus being able to increase developer productivity as well as identify more gaps in theory 
-versus practice which may lead to improved tools or the abandoning of existing ones as was shown with UML~\cite{Petre:2013:UP}.
-
-\textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
-mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
-by industry. We should push our awareness theories to model real development
-environments where interruptions, time management, and full development life cycles are often large factors.}
-
-As was identified in the previous discussion, while current understanding suggests that awareness of all events related to ones work
-produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
-to the point where they only want to be aware of events which require action on their part. This difference, of what 
-should be and what is, should be further understood to combat future potential failures in
-tool development or theories attempted to be used in practice.
-
-\subsection{Implication for Tools}
-\label{sec:implt}
-
-\textbf{Give Developers What They Want} \textit{Developers have their own notions of what they want, and how ``good''
-that something is for their productivity. We should pay attention to what they want, but also be mindful that what
-developers want may not always be good for them.}
-
-The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
-developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
-said, focus at these two stages for tool developer will lead to larger adoption among developers.
-It should be noted here that detection must come with an almost zero rate of false positives as
-current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
-
-\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
-multiple languages, and often having a database schema, we should focus on better techniques for static analysis based
-on what industry standards are for languages and project structures.}
-
-The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
-tools are used for indirect conflicts alone (automatic debugging~\cite{Zeller:2005:WPF}, unit tests, etc), most of these tools
-have need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
-automatically written for a given software object's contract, the ability to find a change  in an external project
-which has broken a developer's own project, or any automation of the existing detection and resolution stages of indirect
-conflicts are what developers currently seek. But of course, most of these implications rely on the improvement of 
-static analysis tools.
-
-These tool implications themselves imply the need of further development of static analysis tools. Static 
-analysis lays at the heart of most if not all stages of indirect conflict research. We must be able to track and manage
-software dependencies across the new landscape of software development that is multiple projects, repositories, and cross
-language support. These improvements will allow the further development of both current and future indirect conflict
-tools.
-
-\section{Evaluation}
-\label{sec:eval}
-
-From the re-interviewed participants, we found extremely positive feedback regarding both our results and major discussion
+From the re-interviewed participants, I found extremely positive feedback regarding both my results and major discussion
 points. Participants often had new stories and experiences to share once they had heard the results of this paper which
@@
 As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.
-We have chosen three of these criteria and explain how we fulfill them.
+I have chosen three of these criteria and explain how I fulfill them.
 
 {\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
-criterion is used to verify the correctness of our finding and to ensure they resonate and fit with participant opinion. It is also
+criterion is used to verify the correctness of my findings and to ensure they resonate and fit with participant opinion. It is also
 required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
-after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
+after participants gave their own opinions on a topic, I presented them with previous participant opinions and asked them to comment
 on and potentially agree with what the majority has been on the topic. Often the developers own opinions already matched those of
@@
 
-As added insurance, we conducted 5 post results interviews with developers to once again confirm our results, and discussions. These
-procedures can be seen in Section~\ref{sec:meth}.
+As added insurance, I conducted 5 post results interviews with developers to once again confirm my results, and discussions. These
+procedures can be seen in Section~\ref{sec:exp-meth}.
 
-To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:results} to either a majority of agreeing
+To ensure the correctness of the results, I also linked all findings in Section~\ref{sec:exp-results} to either a majority of agreeing
 responses on a topic or to a large amount of direct quotes presented by participants.
@@
 {\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
-Although our results may not be entirely novel or even surprising, the combination of said results allowed us to discover the
+Although my results may not be entirely novel or even surprising, the combination of said results allowed me to discover a
 the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
-insight into the debate of prevention versus cure in software development. Given how many indirect conflict tools are left with the same common
-issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
+insight into the debate of prevention versus cure in software development (as to be seen in Chapter~\ref{chapter:discussion}). 
+Given how many indirect conflict tools are left with the same common
+issues, I believe that these findings will help researchers focus on what developers want and need moving into the future more than has
 been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.
 
-10 of the 78 participants who were surveyed sent direct responses to us asking for any results of the research to be sent directly to
+10 of the 78 participants who were surveyed sent direct responses to me asking for any results of the research to be sent directly to
 them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
 tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest 
-in our results help us fulfill this criterion.
+in my results help us fulfill this criterion.
 
@@
 software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
-expressed unique situations that did not fully meet our generalized findings or on going theories. In these cases, we worked in the specific
-cases which were presented as boundary cases and can be seen in Section~\ref{sec:results} as some unique findings or highly
+expressed unique situations that did not fully meet my generalized findings or on going theories. In these cases, I worked in the specific
+cases which were presented as boundary cases and can be seen in Section~\ref{sec:exp-results} as some unique findings or highly
 specialized cases. These cases add to the variation
@@
 
-\section{Related Work}
-\label{sec:related}
+\subsection{Conclusions of Study}
 
-Since this paper has covered a wide spectrum in regards to indirect conflicts (prevention, detection, resolution, process, and tools), there
-exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
-in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging are presented.
+In this study, I have explored indirect conflicts by examining their root causes, their current mitigation strategies, and how developers
+wish to handle indirect conflicts in the present in future. To achieve these results, I interviewed 19 developers from across 12 commercial 
+and open source projects, followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews.
 
-Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
-which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
-Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
-al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
-such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
-indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
-detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
-creates an environment where developers can see how source code changes will affect other areas of a project at the 
-method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
-been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
-needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
-instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
-Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
-which were explored in this paper. Our own tool 
-Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
-
-Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
-times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
-Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
-artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
-created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
-Socio-technical congruence however, is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
-
-For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
-recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
-Kim conducted several focus groups as well as web surveys to determine these results. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
-require action on the developer's half.  Kim
-also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
-were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
-noting ``I get a big laundry list... I see the email and I delete it''.
-
-In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
-preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
-development processes.
-
-In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
-ill fit of current static analysis to current development practices. Not being designed properly to fit current development work
-flow, and the over abundance of false positives were shown to be factors of developer non-use.
-
-Hattori et. al.~\cite{Hattori:2012:ICG} found, through qualitative user studies, that developers tend to use the bare amount of communication
-towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
-resonates with ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
-sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
-This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici:2009} give the 
-possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea
-that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
-Blincoe et al.~\cite{Blincoe:2013:UCC} also found, through data mining techniques, a subset of events related to awareness 
-which occur that developers will communicate about as opposed to all awareness events. This again confirms our understanding that developers
-only care about certain events which require some action on their part.
-
-While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
-can be applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
-changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
-with the idea that developers wait for something to go wrong, then wish to fix the issue as fast as possible. Program slicing~\cite{Xu:2005:BSP}
-has also been extensively studied in order to aid in debugging in manual or automated techniques.
-
-\section{Conclusions}
-\label{sec:conc}
-
-Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
-losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
-19 interviewed developers from 12 organizations as well as 78 surveyed developers, we have provided characterization of indirect conflicts,
-current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
-in their work flow with indirect conflicts.
-
-We have shown through the findings of RQ1 why indirect conflicts occur, when indirect conflicts are more likely to
-occur, as well as what types of software objects are susceptible to these conflicts. Results of RQ2 have shown
-how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings of RQ3
-have provided a foundation as to why past techniques and tools have had low adoption rates and where
-researchers should focus their current and future efforts in handling indirect conflicts.
-
-We have shown the disjoint of current awareness understanding versus the practical
-awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
-ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure has been presented along with
-the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
-to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, we have shown
-the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
-future industrial adoption of tools produced by academia for lack of fit in industrial needs.
-
-\section{Acknowledgments}
-
-We sincerely thank all participants who were willing to be interviewed or who participated
-in completing our survey. We thank these people for taking time out of their day to participate and for
-sharing their developer experience with us.
-
-\bibliographystyle{IEEEtran}
-%\balance
-\bibliography{paper}
-
-% End of the paper
-\end{document}
+My findings indicate that: 
+indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of change impacts,
+indirect conflicts occur quite frequently with a tendency to be more prevelant at the beginning of a development cycle, indirect conflicts
+occur more often as a software project grows in developer numbers, developers use design by contract, add and deprecate, and experience
+to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occured rather
+than tools which help prevent possible conflicts for the future.

### 2014-01-05 msc-thesis/chapters/4/chapter-discussion.tex
MSC Added all 4 studies to the thesis.

\startchapter{Discussion}
\label{chapter:discussion}

### 2014-01-05 msc-thesis/frontmatter/abs.tex
MSC Added all 4 studies to the thesis.

\newpage
\TOCadd{Abstract}

\noindent \textbf{Supervisory Committee}
\tpbreak
\panel

\begin{center}
\textbf{ABSTRACT}
\end{center}

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, scalability, information distribution and many others. 
To better understand these issues, this dissertation will focus on exploring the world of indirect conflicts
through 4 studies. The first two studies presented will focus on motivational circumstances which occur
during the software development life cycle and cause indirect conflicts. Developers interactions are
studied in order to create a tool which can aid in the workflows around indirect conflicts. The second
two studies present a deeper investigation into why most indirect conflict tools fail to attrack developer
inerest through exploring the root causes of indirect conflicts and how tools should be properly built
to support developer workflows.

### 2014-01-05 msc-thesis/frontmatter/tp.tex
MSC Added all 4 studies to the thesis.

\pagebreak
{
\centering
\thesistitle
\tpbreak
by
\tpbreak
\nameanddegrees
\tpbreak
% One needs to make adjustments for a Master's Thesis
A Dissertation Submitted in Partial Fulfillment of the \\
Requirements for the Degree of
\tpbreak
MASTER OF SCIENCE
\tpbreak
in the Department of Computer Science\\
\vfill
\begin{tabular}{cl}
& \copyright\ Jordan Ell, 2014\\
& \phantom{\copyright} University of Victoria
\end{tabular}
\tpbreak
All rights reserved. This dissertation may not be reproduced in whole or in part, by \\
\hfill photocopying or other means, without the permission of the author. 
\hfill
}
\pagebreak

### 2014-01-05 msc-thesis/chapters/2/chapter-motivating.tex
MSC starting discussion.

@@
 While the research problems have been briefly outlined in Chapter~\ref{chapter:introduction}, this chapter
-will focus on the underlying studies which motivated the research of this thesis as well as give a more full
-and rich description of the problem being solved.
+will focus on the underlying studies which motivated the research of this dissertation.
 

### 2014-01-05 msc-thesis/chapters/4/chapter-discussion.tex
MSC starting discussion.

@@
 \label{chapter:discussion}
+
+While each study presented in this dissertation is quite unique, this chapter will focus on the underlying themes and
+results of all 4 studies. This chapter will address the usefulness of the results presented towards academia and
+what those results mean for continued studies in future research as well as discuss how the results can be viewed
+for larger over arching outcomes than those presented in results sections.
+
+This chapter will proceed in 2 subsections. The first section will address the 2 motivational studies and the lessons
+learned from each as background information for the richer following studies. The second section will address the 2 large
+studies found in Chapter 3 in a combined manner. Since Study 4 was directly associated with developer opinion found in Study
+3, the two discussions will be intertwined to better support each other.
+
+\section{Motivating Studies Discussion}
+
+\section{Indirect Conflict Exploration Discussion}
+
+\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
+a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
+to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
+awareness about most, if not all actions.}
+
+As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
+sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
+or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
+to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what we believe to be the
+leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
+information being presented to them, rendering the system useless. These false positives are caused by a difference in what
+developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
+be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
+where as current awareness understanding would state any event which is related to an individual's 
+work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
+``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
+showing that not all changes around an object should be reported for awareness.
+
+Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
+their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
+developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
+don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
+that the developer only wants the awareness if the item will require action on their part to resolve the issue.
+This sense of unwanted or limited awareness is crucial to understand why generalized
+awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
+classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
+positives and information overload. 
+Developers simply do not want to know about events which effect them, but require no action on their part.
+
+This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
+being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
+but as it was found in practice, this is an incorrect assumption. 
+In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for 
+analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
+all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
+the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
+of events occurring around their work which pose no direct threat to the consumption of their time.}
+
+Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
+developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
+them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
+influence of total, or near total
+awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
+with higher quality or understanding of the product. 
+A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to 
+producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
+however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
+be more productive.
+
+\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
+arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
+to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.}
+
+We have seen through the results, that developers posses both tools and practices for the prevention,
+detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
+that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
+may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
+is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
+Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
+place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
+them in the first place.
+
+Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
+accomplished essentially by
+not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
+developers are spending little to no time in prevention. Developers do however spend a large amount of time in detection and cure through the
+writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
+conflicts occurred at either the detection or cure levels. Obviously, developers either prefer to spend their time in curative
+measures, or do not posses the proper tools to take better action in the preventative stages. ``You're reaction time is much more crucial''
+said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
+matter what prevention was taken, the issue must simply be resolves as quick as possible now.
+
+The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
+Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
+which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
+of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
+are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
+abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
+real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
+this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
+curative measures are more productive than preventative measures.
+
+Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
+to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
+in that they result in a trade off 
+of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
+it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
+not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
+to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
+Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.
+
+A last interesting observation about current curative measures being taken in industry, is that developers view testing of software 
+as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
+are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
+thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
+However, if a test never fails, could it not be said that it is preventing changes from causing it to fail rather than detecting
+when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
+measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
+``prevention is the goal'' was commonly said by developers.
+
+This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
+designers, architects, or managerial stakeholders. 
+From the project manager's interviewed it was shown that they are heavily favored towards planning and 
+prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
+approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
+is being viewed in research.
+
+%mature point, static analysis
+\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
+and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
+some software project configurations and software languages to be quite prone to indirect conflicts.}
+
+Indirect conflicts are more likely to occur,
+from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
+better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
+end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
+mature point, as well as any other
+potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
+that different change types may occur at different rates throughout a project's life time and that this
+may drastically effect the outcomes of indirect conflict tools or processes.
+
+A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
+research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
+could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
+means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
+Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
+changes occur at a given point in time which could be used for determining such a context of mature points.
+
+This added level of context, through the notion of a mature point, 
+would add to the prevention versus cure debate as previously discussed. Prevention may be a better
+choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
+to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
+causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
+and the mature point in order to fully understand the nature of indirect conflicts and their context.
+
+While this added sense of understanding is important to indirect conflicts, contextual attributes of a project's current progress
+or measures of performance are often more synonymous with project management
+rather than software developers themselves. An understanding of a project's evolution, a mature point being reached as an example,
+may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
+conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.
+
+Aside from context, dependency identification and tracking is a key missing component of
+indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
+organizational structures of software teams. Between increased modularization (multiple sub projects or repositories), 
+cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
+become quite limited in identifying and tracking dependencies where they were once strong.
+
+Relational database schemas are one of the highest sources of indirect
+conflicts found in their projects, ``it breaks stuff all over the place'',
+yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
+been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
+dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.
+
+These increases of complexity in software products has left quite a gap in dependency identification and tracking which 
+has lead to some of the deficiencies of indirect conflict research.
+
+\subsection{Implication for Research}
+\label{sec:implr}
+
+\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
+``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly 
+produce extremely complex answers.}
+
+With software developers focusing their current efforts on curative measures, for
+indirect conflicts, suggests
+that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
+least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
+provide more insight into possibilities, as its answer may determine many future actions taken by the research community.
+
+Recommendations towards the study of prevention versus cure involve the examination of formal processes and tools used
+by industry professionals with measurements of efficiency and effectiveness, similar to the work of
+Tiwana~\cite{Tiwana:2008:ICD} in coordination tools. With these studies, we may find insight into the correct balance
+of prevention versus cure, thus being able to increase developer productivity as well as identify more gaps in theory 
+versus practice which may lead to improved tools or the abandoning of existing ones as was shown with UML~\cite{Petre:2013:UP}.
+
+\textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
+mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
+by industry. We should push our awareness theories to model real development
+environments where interruptions, time management, and full development life cycles are often large factors.}
+
+As was identified in the previous discussion, while current understanding suggests that awareness of all events related to ones work
+produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
+to the point where they only want to be aware of events which require action on their part. This difference, of what 
+should be and what is, should be further understood to combat future potential failures in
+tool development or theories attempted to be used in practice.
+
+\subsection{Implication for Tools}
+\label{sec:implt}
+
+\textbf{Give Developers What They Want} \textit{Developers have their own notions of what they want, and how ``good''
+that something is for their productivity. We should pay attention to what they want, but also be mindful that what
+developers want may not always be good for them.}
+
+The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
+developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
+said, focus at these two stages for tool developer will lead to larger adoption among developers.
+It should be noted here that detection must come with an almost zero rate of false positives as
+current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
+
+\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
+multiple languages, and often having a database schema, we should focus on better techniques for static analysis based
+on what industry standards are for languages and project structures.}
+
+The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
+tools are used for indirect conflicts alone (automatic debugging~\cite{Zeller:2005:WPF}, unit tests, etc), most of these tools
+have need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
+automatically written for a given software object's contract, the ability to find a change  in an external project
+which has broken a developer's own project, or any automation of the existing detection and resolution stages of indirect
+conflicts are what developers currently seek. But of course, most of these implications rely on the improvement of 
+static analysis tools.
+
+These tool implications themselves imply the need of further development of static analysis tools. Static 
+analysis lays at the heart of most if not all stages of indirect conflict research. We must be able to track and manage
+software dependencies across the new landscape of software development that is multiple projects, repositories, and cross
+language support. These improvements will allow the further development of both current and future indirect conflict
+tools.

### 2014-01-07 msc-thesis/chapters/4/chapter-discussion.tex
MSC base discussions now in place.

@@
 
+\textbf{The Human Factor of Indirect Conflicts} \textit{An important component of indirect conflicts are the developers
+themselves and how their notions of other's work is perceived across a project.}
+
+As we have seen through Study 1, developers which are tied to source code objects can become a focal point of
+indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
+developers Daniel and Anne on Hibernate-ORM are always linked indirect conflicts when dependencies between
+their source code objects change. This is an important observation to make when moving forward with indirect
+conflicts.
+
+What is really happening between these two developers is the real issue to consider. Daniel could have an
+assumption about the way Anne's code works which causes Anne's changes to have negative impacts on Daniel's
+own work. For instance, Daniel may expect a method of Anne's to have a special return case, which may be correct,
+however when Anne changes that special case or removes it, Daniel's code can be negatively affected. In this extreme
+result found in Study 1, further analysis showed that one central method in the software project was being changed
+frequently and causing Daniel's code to be negatively affected in some way.
+
+The question of how to prevent this type of negative impact is ultimately the goal of indirect conflicts. Study 1 has
+shown that pairs of developers can often be a large component to that goal as well. (Impact from Study 2 was
+created the address just this issue.)
+
+\textbf{Information Overload} \textit{My tool Impact, as well as many other indirect conflict tools, suffer from
+information overload in their delivery to developers which is a key issue in preventing adopting and acceptance
+of indirect conflict tools.}
+
+As was previously stated, many indirect conflict tools end up suffering from information overload to developers
+and other end users. Impact was initially created to take the knowledge learned from Study 1 and attempt to create
+a new indirect conflict system based on developer interactions inside the code base which would potentially address
+information overload. However, we now know from the results of Study 2, that Impact once again suffered defeat to information
+overload based on the case study evaluation. Ultimately, this sense of information overload ends up causing
+adoption of indirect conflict tools to fail which in turn causes some research components to have failed as
+well.
+
+From previous works, as well as from Study 2, we know that information overload is a large issue in indirect
+conflicts. It has been found that a large number of dependencies in software caused by the nature large software
+projects is a root issue in information overload~\cite{Sarma:2007:TSA,Servant:2010:CPI}. These large sizes in
+dependencies are ultimately what cause information overload as dependencies cannot be evaluated as to their
+relevance in a certain source code change with ease. In other words, we cannot determine which of the numerous
+dependencies will fail (outside of compilation and testing failures) when source code is changed.
+
+A potential solution derived from the evaluation of Impact comes from the ideas of Meyer~\cite{Meyer:1988} on 
+``design by contract''. In this methodology, changes to method preconditions and postconditions are considered 
+to be the most harmful. This includes adding conditions that must be met by both input and output parameters 
+such as limitations to input and expected output. To achieve this level of source code analysis, the ideas of 
+Fluri et al.~\cite{Fluri:2007:CDT} can be used on the previously generated ASTs for high granularity source code 
+change extraction when determining if preconditions or postconditions have changed. While this solution does not 
+wholly address the problems of information overload as previously stated through failures in dependencies, it does 
+reduce the number of source code changes to  be analyzed which in turn will reduce the amount of information on 
+the whole which is put forth by the indirect
+conflict system. This solution however does also run the risk of missing more indirect conflicts as many conflicts
+can occur outside of changes to method contracts. This solution represents my first ideas of fixing information overload in regards to
+indirect conflicts and is again found in the later discussion of root causes of indirect conflicts found in
+the next section.
+
 \section{Indirect Conflict Exploration Discussion}
+
+While the previous section discussed the main motivations for taking a step back on indirect conflicts in order to understand
+better what can be done to improve developer work flow, this section will discuss exactly those steps back. From Studies 3 and
+4 we will notice 3 main trends that have been discovered for indirect conflicts in both industry and in academia. These 3 major
+trends to now be discussed are: unwanted awareness, prevention versus cure, and the gap in software evolution analysis. This
+section will also include a brief discussion regarding the implications, learned from studies 3 and 4, for both academia and
+for industry through tools.
 
@@
 or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
-to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what we believe to be the
+to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what I believe to be the
 leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 
@@
 Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
-to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
+to measure. The pros and cons of prevention versus cure I have identified, are similar to those of unwanted awareness,
 in that they result in a trade off 
@@
 \textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
-and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
+and analysis unfriendly languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
 some software project configurations and software languages to be quite prone to indirect conflicts.}
@@
 means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
-Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
-changes occur at a given point in time which could be used for determining such a context of mature points.
+These change trends are exactly what Study 4 set out to address.
+
+Through Study 4, I showed that 9 critical change trends exist at a major release of a variety of open source projects. This
+knowledge could help identify, as was previously stated, when indirect conflicts are more likely to occur in a project or not.
+For example, the trend of adding test methods was found to occur more after a major release than before. This trend along with
+the knowledge that indirect conflicts occur more at the start of a new development cycle can be used to predict times in development
+where indirect conflicts are likely to occur automatically. An automated system may see that more test methods are being added than
+before in the past 4 weeks or so and be able to alert developers to an instability in the code which in turn means an elevated
+risk of indirect conflicts.
+
+Similarly, the trend of adding private classes is seen to occur more so before a major release than after. This knowledge again
+can be used in an automated system along with the fact that indirect conflicts are less likely to occur towards the end of a development
+cycle. The automated system will be able to identify a stability in the project through private classes being added and alert developers
+that the code is stabilizing and that large changes to the project should be avoided until the start of the next development cycle
+in order to further reduce the risk of indirect conflicts.
 
 This added level of context, through the notion of a mature point, 
-would add to the prevention versus cure debate as previously discussed. Prevention may be a better
+would add to the prevention versus cure debate as previously discussed as well. Prevention may be a better
 choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
@@
 
-Aside from context, dependency identification and tracking is a key missing component of
+In regards to these contextual identifications in software projects, dependency identification and tracking is a key missing component of
 indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and 
@@
 cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have 
-become quite limited in identifying and tracking dependencies where they were once strong.
+become quite limited in identifying and tracking dependencies where they were once strong. As software has become more
+sophisticated over time, static analysis tools have gone from extremely useful to only occasionally useful. Since many
+projects involve several languages, sub projects, and database schemas, static analysis has become cripplingly obsolete
+in the industry of today. In order to move indirect conflicts and many other research areas forward, this gap of cross domain
+static analysis must be filled.
 
-Relational database schemas are one of the highest sources of indirect
+As an example, relational database schemas are one of the highest sources of indirect
 conflicts found in their projects, ``it breaks stuff all over the place'',
 yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have 
-been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
+been scarcely researched in terms of indirect conflicts. This falls in with my previous understanding of cross language
 dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.

### 2014-01-07 msc-thesis/chapters/2/sec-impact.tex
Added in study conclusions.

@@
 generic and can be implemented in similar indirect conflict awareness tools.
+
+\section{Related Work}
+Although there is an abundance of awareness tools developed in research
+today, only a handful have made an attempt to examine indirect conflicts.
+Here, I will outline four of the forefront projects in indirect conflicts
+and how these projects have influenced the decision making process in
+the design and implementation of \textit{Impact}.
+
+I first start with both Codebook~\cite{Begel:2010:CDE} and 
+Ariadne~\cite{Trainer:2005:BGT}. These projects produce an exploratory
+environment for developers to handle indirect conflicts. Exploratory
+pertains to the ability to solve self determined conflicts, meaning that
+once a developer discovers a conflict, they can use the tool as a type of
+lookup table to solve their issue. Codebook is a type of social developer
+network that relates developers to source code, issue repositories and
+other social media while Ariadne only looks at source code for developer
+to source code association. Through Codebook, developers become
+owners of source code artifacts. Both projects also use program 
+dependency graphs~\cite{Horwitz:1992:UPD}
+in order to relate technical artifacts to each other. These projects make 
+use of method call graphs in order to 
+determine which methods invoke others which forms the basis for 
+linking source code artifacts creating a directed graph. While these 
+projects can be great tools 
+for solving indirect conflicts which may arise, by querying such directed
+graphs to view impacts of conflict creating code, they lack the ability to
+detect potential conflicts on their own.
+
+A serious attempt at both detecting and informing developers of
+indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
+monitors developers activities in files with regards to class signatures.
+Once a developer changes the signature of a class, such as by modifying changes
+in name, parameters, or return values of public methods, any workspace
+of other developers which are using that class will be notified. Palantir utilizes
+a push-based event model~\cite{Fitzpatrick:2002:SPA} which seems to be
+a favored collection system among awareness tools. Sarma et al.
+~\cite{Sarma:2007:TSA} also
+developed a generic design for future indirect conflict awareness tools. 
+However, Palantir falls short in its collection and distribution
+mechanisms. First, Palantir only considers ``outside'' appearance of technical
+objects, being their return types, parameters, etc. Secondly, Palantir 
+only delivers
+detected conflicts to developers who are presently viewing or editing
+the indirect object while other developers who have used the modified 
+class previously are not notified.
+
+I will lastly examine the tool CASI~\cite{Servant:2010:CPI} which uses
+a sphere of influence for each developer to determine how source code
+changes are indirectly related to other components of a software project.
+CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
+call graphs in Ariadne~\cite{Trainer:2005:BGT} which gives dependencies among
+all source code entities. This provides a verbose output of dependencies when
+source code is changed. CASI also implements a visualization where
+a developer can see what parts of a software projects he or she may be effecting
+with the source code change. This allows the developer themselves to go and fix
+potential issues elsewhere in the project before the code change is committed
+to the software repository. While CASI covers great ground in its approach,
+it still leaves the issue of information overload, although attempts were made
+to solve this by having severity levels of indirect conflicts presented to
+the user.
 

### 2014-01-07 msc-thesis/chapters/3/sec-apie.tex
Added in study conclusions.

@@
 to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.
+
+\section{Related Work}
+\label{sec:rel}
+While very little has been published about release quality studies and stability (especially in regards to
+indirect conflicts), there have been a few studies which attempt to address the issues directly
+or indirectly. Wu et al.~\cite{Wu:2008:QAF} performed a case study of SoftPM, a widely adopted project management tool, to explore the relationships of
+pre-release and post-release failures at major releases. Wu et al. found that the ratio of post-release failures to pre-release failures is significantly low
+and can be used to show reliability and stability. Hindle et al.~\cite{Hindle:2007:RPD} performed a case study on MySQL which observed a project's behavior
+around major and minor release by monitoring artifact check-ins and changes. They found that there are temporary stoppages for source revisions around releases,
+indicating that a temporary freeze is taking place for developers and that last minute fixes and manual testing may be performed.
+Zaidman et al.~\cite{Zaidman:2011:SCP}, in comparison, studied the co-evolution of production and test code with inspections and analysis
+at major and minor releases, showing how test and production code can evolve at different rates and times. These results show that production
+and test code should be handled as different cases for a stability measure around major releases.
+
+The study of open source projects revolving around release points has become more accessible by the work of Tsay et al~\cite{Tsay:2011:EMO}. Tsay et al. created
+a resource of historical release dates for open source software projects to be used for future studies by other researchers.
+
+In terms of software stability, development techniques have been proposed to increase software stability. 
+Fayad~\cite{Fayad:2001:TOI}~\cite{Fayad:2002:ASS} suggests that ``business objects'' (BOs) do not change in nature and that they are inherently stable. These
+objects only need to change to accommodate external modules at the interface. Some studies such as Chow et al.~\cite{Chow:2011:SJI} have investigated
+the stability of changes to interfaces which are considered a good indication of stability.
+Mockus et al.~\cite{Mockus:2008:IQR} used major and minor release points to compare industry process quality to customer-perceived quality of the software
+project. Mockus et al. found that defect density is lowest at major releases but at the same time software quality is at its lowest all when compared to minor
+releases. The low software quality here relates to end-user errors of installations and configurations. Wermelinger et al.~\cite{Wermelinger:2008:AEE}
+showed that stable core architectures can be detected by using source code changes. Finally, Fayad et al.~\cite{Fayad:2010:SSM} have
+investigated the Software Stability Model (SSM) for Software Product Lines to show that the SSM's impact on architecture and design of a software product
+can help improve the life of the product line and make it more adaptable and applicable.
 
@@
 major releases as previous studies have shown that indirect conflicts occur less at a major release and more so after a major release
-or as the beginning of a development cycle. I have presented here 9 major change trends which surround major releases in the open source
+or at the beginning of a development cycle. I have presented here 9 major change trends which surround major releases in the open source
 projects studied. The 4 change trends found in this study occurring before major releases are: added private classes, 

### 2014-01-07 msc-thesis/chapters/3/sec-iced.tex
Added in study conclusions.

@@
 
+\section{Related Work}
+\label{sec:iced-related}
+
+Since this paper has covered a wide spectrum in regards to indirect conflicts (prevention, detection, resolution, process, and tools), there
+exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
+in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging are presented.
+
+Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
+which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
+Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
+al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
+such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
+indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
+detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
+creates an environment where developers can see how source code changes will affect other areas of a project at the 
+method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
+been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
+needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
+instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
+Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
+which were explored in this paper. Our own tool 
+Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
+
+Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
+Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
+artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
+created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
+Socio-technical congruence however, is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+
+For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
+recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
+Kim conducted several focus groups as well as web surveys to determine these results. These areas of interest resonate with what was
+found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
+require action on the developer's half.  Kim
+also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
+were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
+noting ``I get a big laundry list... I see the email and I delete it''.
+
+In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
+preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
+development processes.
+
+In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
+ill fit of current static analysis to current development practices. Not being designed properly to fit current development work
+flow, and the over abundance of false positives were shown to be factors of developer non-use.
+
+Hattori et. al.~\cite{Hattori:2012:ICG} found, through qualitative user studies, that developers tend to use the bare amount of communication
+towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
+resonates with ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
+sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
+This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici:2009} give the 
+possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea
+that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
+Blincoe et al.~\cite{Blincoe:2013:UCC} also found, through data mining techniques, a subset of events related to awareness 
+which occur that developers will communicate about as opposed to all awareness events. This again confirms our understanding that developers
+only care about certain events which require some action on their part.
+
+While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
+can be applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
+changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
+with the idea that developers wait for something to go wrong, then wish to fix the issue as fast as possible. Program slicing~\cite{Xu:2005:BSP}
+has also been extensively studied in order to aid in debugging in manual or automated techniques.
+
 \subsection{Conclusions of Study}

### 2014-01-07 msc-thesis/chapters/4/chapter-discussion.tex
Added in study conclusions.

@@
 or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
-to proposed or implemented tools as discussed in Section~\ref{sec:review}. This inability to generalize is what I believe to be the
+to proposed or implemented tools as discussed in Section~\ref{sec:iced-related}. This inability to generalize is what I believe to be the
 leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore 

### 2014-01-07 msc-thesis/chapters/5/chapter-conclusion.tex
Added in study conclusions.

\startchapter{Conclusions}
\label{concl}

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or \textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as 
they arise from a result of source code changes.

While the techniques and tools for direct conflicts have had large success, tools either proposed or studied for 
indirect conflicts have had common issues of information overload, false positives, scalability, information distribution 
and many others. These issues has ultimately lead to poor developer interest in indirect conflict tools as well
as a failed adoption rate and partial failures of research.

To better understand these issues, this dissertation has focused on exploring the world of indirect 
conflicts through 4 studies. The first two studies focused on motivational circumstances which occur during 
the software development life cycle and cause indirect conflicts. Developers interactions were studied in order to create 
a tool which can aid in the work-flows around indirect conflicts. The second two studies presented a deeper investigation 
as to why most indirect conflict tools fail to attract developer interest through exploring the root causes of indirect conflicts 
and how tools should be properly built to support developer work flows.

\section{Study 1}

Technical dependencies are often used to predict software failures
in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. However, human interactions as tied
to software objects can also be used to predict software failures.
Study 1 presented a method for detecting failure inducing pairs of developers inside
of technical networks based on code changes. The methodology used was to assigned developers to the source code artifacts
which they authored. Next, dependencies were identified between source code artifacts, implying dependencies between
developers. Lastly, failure inducing dependencies between developers were identified through using methods to identify
bugs in source code throughout the lifetime of a project. These developer pairs were used in the prediction
of future bugs as well as provide coordination recommendations for developers within a project. These bugs being predicted
directly correlate to indirect conflicts. Since the human dependencies being measured in Study 1 are directly tied to
software objects, we can say that the indirect conflicts being studied here are between developers themselves, while
the dependencies between the developers are found in the source code.

Study 1 however, did not consider the technical dependencies themselves to be the root cause of
the software failures. This study focused purely on developer ownership of software methods and
the dependencies between developers as the possible root cause of the failures.

Through the analysis of Study 1 as seen in Chapter 4, an important component of indirect conflicts are the developers
themselves and how their notions of other's work is perceived across a project. We can see from Study 1 results that
a developer Daniel may believe a source code artifact behaves in a particular way (from documentation or other forms
of communication for the artifact) so when that artifact changes its behavior, Daniel can become negatively impacted.
A human factor is present here as the person who makes the artifact change, say Anne, may not have communicated the
change to all other developers affected by the change. This is an important understanding that lead directly into the
creation of Impact (Study 2) as a way to mitigate the negative affects across human developers as a result of 
source code changes.

\section{Study 2}

As a direct response to the findings and analysis of Study 1, Study 2 set out to address the human factors presented
through an indirect conflict tool. In Study 2, I proposed a generic design for the development of an awareness tool 
in regards to handling indirect conflicts through human factors. I presented a prototype awareness tool, \textit{Impact}, 
which was designed around the generic technical object approach. Impact could track which developers were responsible
for which artifacts of code as well as their dependencies to other developer's code artifacts. When a dependency to
another developer's code was changed, the developer was notified of the change in order to avoid any bugs that
may arise from the change. The delivery system used for Impact was that of an RSS type feed where developers could view
their notifications in a stream of messages through a web application. However, \textit{Impact} suffered from information 
overload, in that it had too many notifications sent to developers.

This failure from information overload was ultimately equivalent to the various other indirect conflict tools from
previous research (even those not addressing human developer factors). This issue of information overload is the key issue 
in preventing adopting and acceptance of indirect conflict tools from developers and ultimately leads to some research
component failures to understand what developers truly want from indirect conflict tools.As a proposed solution to information overload, 
the ideas of Meyer~\cite{Meyer:1988} on ``design by contract'' were presented. This methodology looks at using the 
post and pre conditions of software objects in order to reduce the number of source code changes that are analyzed by 
Impact in order to reduce information overload.

While the previous proposed solution was designed to fix information overload in Impact and potentially other indirect
conflict tools, it was decided that further investigation into indirect conflicts was needed to truly attempt a solution.
The results of Study 2, combined with the knowledge of previous research having similar issues of information overload
in indirect conflict tools, sparked an interest in studying indirect conflicts at a deeper level (Study 3) which could
be used to better understand causes and developer strategies in solving indirect conflicts.

\section{Study 3}

Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
losses in this realm have been unsuccessful in attracting major support from developers (as was seen in Study 2). 
In Study 3, I conducted a qualitative study involving
19 interviewed developers from 12 organizations as well as 78 surveyed developers. Through these qualitative measures,
I provided characterization of indirect conflicts,
current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
in their work flow with indirect conflicts.

For the root causes of indirect conflicts, I found that indirect conflicts occur due to changing of a software object's 
contract and the lack of understanding of
the far reaching implications (through dependencies) of that change. I also found that indirect conflicts are more likely
to occur at the beginning of a development cycle when the code is quite unstable and that these scenarios can become
compounded in difficulty when more developers are present on a project.

As per current developer work-flows regarding indirect conflicts, I found that developers have 3 major mitigation strategies
to avoid indirect conflicts: ``Design by Contract'', add and deprecate, and personal experience. For catching indirect conflicts,
I found that use case coverage through proper testings (both unit and integration) are currently thought to be the best
developers can achieve. And finally, for the future of indirect conflict tools, developers made it clear that they would
prefer an improvement to resolution methods for indirect conflicts such as automatic or aided development techniques.

As per the analysis of Study 3, I have shown the disjoint of current awareness understanding versus the practical
awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure was presented along with
the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
future industrial adoption of tools produced by academia for lack of fit in industrial needs. This gap directly resulted in the final
study of this Dissertation to be completed (Study 4).

\section{Study 4}

As a result of the gap in software evolution analysis shown in Study 3, I conducted a case study of 10 open source Java software 
projects in order to study their change trends surrounding
major releases as Study 3 had shown that indirect conflicts occur less at a major release and more so after a major release
or at the beginning of a development cycle.

Through Study 4, I presented 9 major change trends which surround major releases in the open source
projects studied. The 4 change trends found in this study occurring before major releases are: added private classes, 
changed test method signatures, changed documentation, and removed test classes.
The 5 change trends found in this paper occurring after major release are: added test methods, changed test classes, removed public methods, removed
private classes, and removed private methods.

These 9 change trends can be used in indirect conflict tools in order to identify a context for indirect conflicts.
For example, the trend of adding test methods was found to occur more after a major release than before. This trend along with
the knowledge that indirect conflicts occur more at the start of a new development cycle can be used to predict times in development
where indirect conflicts are likely to occur automatically. An automated system may see that more test methods are being added than
before in the past 4 weeks or so and be able to alert developers to an instability in the code which in turn means an elevated
risk of indirect conflicts.

The results of Study 4 directly supplemented those from Study 3 in that it addressed the gap of software evolution techniques
needed for indirect conflicts. However, this study was also quite limited and the issues of cross domain analysis techniques
presented in Study 3 are still cause for future research.

\section{Final Conclusions}

### 2014-01-07 msc-thesis/chapters/5/chapter-conclusion.tex
MSC conslusions done.

@@
 \section{Final Conclusions}
+
+This dissertation has covered a wide range of interests all within indirect conflicts. I have shown how human factors can
+be an important part of indirect conflicts and how pairs of developers can be found to be statistically related to indirect
+conflicts bugs. I have shown how these human factors can be integrated into indirect conflict tools by using dependencies in 
+authored source code artifacts. Ultimately however, these new human factors added into indirect conflict tools resulted in
+similar failures of information overload as seen through many previous research tools in indirect conflicts. After these conclusions,
+I presented a study which found root causes of indirect conflicts to be contract changes and an unawareness of those changes
+implications, that developers use ``Design by Contract'', add and deprecate, and personal experience to avoid indirect conflicts,
+and some suggestions for future indirect conflict tool development while stating where gaps in source code analysis should
+be improved to improve the indirect conflict tools. Based on that gap in
+software evolution analysis in indirect conflicts, I presented a method for finding contextual patterns which can relate to
+indirect conflicts in order to aid future tool development for indirect conflicts.
+
+Through the analysis in Chapter 4, I have shown the disjoint of current awareness understanding versus the practical
+awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
+ultimately lead to tools with information overload and false positive issues. This was quite evident from my tool Impact as
+presented in Study 2. Even when the tool is based around the statistical results of Stud 1, the disjoint between what academics
+determin developers want versus what they themselves want is quite obvious. The debate of prevention versus cure has been presented along with
+the industrial tendency towards curative measures. However, while a curative approach may be favored by developers, further research is needed
+to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
+the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
+future industrial adoption of tools produced by academia for lack of fit in industrial needs. While, I have shown a potential solution to address
+a sub problem of the analysis gap through contextual patterns of indirect conflicts, many sub problems of the analysis gap remian problems
+for future research.

### 2014-01-08 msc-thesis/chapters/1/chapter-intro.tex
Added in study conclusions.
MSC intro now has big chunks in it.

@@
 
-\input chapters/1/sec-intro
-\input chapters/1/sec-review
+\section{Introduction}
+
+% Introduce awareness
+As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
+development has become the standard development process instead of the exception. With this parallel development
+comes the need for larger awareness among developers to have ``an understanding of the activities of others
+which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
+mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
+empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
+and time-consuming chore for developers~\cite{Perry:2001:PCL}.
+
+% Introduce dependencies
+Large software projects are created using highly modular and reusable code. This creates technical dependencies between
+methods or functions that can be used in a wide variety of locations throughout the project. This causes changes to any 
+given software object to have a rippling effect across the rest of the project~\cite{Acharya:2011:PCI}. The larger these 
+effects are, the more likely they are to cause a software failure inside the system during the project's life 
+span~\cite{Zimmermann:2008:PDU}. These observations of technical dependencies open the door to types of analysis on the 
+developer networks they infer and preventing software failures by improving coordination amongst dependent developers.
+
+% What has been done with technical dependencies
+Technical dependencies in a project can be used to predict success or failure in builds or code 
+changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most research in this area is based on identifying central 
+modules inside a large code base which are likely to cause software failures or detecting frequently changed code that 
+can be associated with previous failures~\cite{Kim:2006:AIB}. This module-based method also results in predictions at 
+the file or binary level of software development as opposed to a code change level and often lack the ability to provide 
+recommendations for improved coordination other than test focus.
+
+% Using technical dependencies for awareness tools
+With the power of technical dependencies in tracking unintended consequences from source code changes, several tools 
+have been created to attempt to solve task awareness related issues with some 
+success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
+to solve task awareness related issues at the direct conflict level.
+
+%Describe difference between direct and indirect
+Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
+\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
+artifact. Tools have been created and studied for direct conflicts
+~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
+positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
+which influences behavior negatively in another location in the system. Indirect conflict tools however, have
+not shared the same success as direct conflict 
+tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
+previous interviews and surveys conducted with software developers have shown a pattern that developers 
+of a software project view indirect conflict awareness  as a high priority issue in their 
+development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
+is required to address this developer concern.
+
+% Limitations of IC tools for indirect conflicts compared to Impact
+Indirect conflicts arising in source code are inherently difficult to resolve as most of the time, source code analysis must 
+be performed in order to find relationships between technical objects which are harmed by changes. While some awareness 
+tools have been created with these indirect conflicts primarily in mind~\cite{Begel:2010:CDE, Trainer:2005:BGT}, most have only 
+created an exploratory environment which is used by developers to solve conflicts which may arise. These tools were not designed 
+to detect indirect conflicts that arise and alert developers to their presence  inside the software system. 
+Sarma et al.~\cite{Sarma:2007:TSA} has started to work directly on solving indirect conflicts, however, these products are not 
+designed to handle internal structures of technical objects.
+
+% Issues with IC tools of information overload etc.
+While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
+to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
+overload, tools eventually being
+ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
+tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
+invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
+identification of failure inducing changes, other than those which are already identifiable by other means such
+as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
+also compounded by the scale of many software development projects leading to further information overload.
+Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
+However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
+come up.
+
+% There are 4 studies
+Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
+will be explored in this dissertation. In order to fully understand indirect conflicts, this dissertation is composed
+of 4 studies of which explore different components of indirect conflicts.
+
+% Study 1 introduction
+Study 1 focuses on the power of technical dependencies in software projects. The question I
+investigated was : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
+code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
+in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
+relationships in successful and failed code changes. These developers pairs can be seen as indirect conflicts occurring
+as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
+inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
+congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
+This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
+inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
+results of Study 1 directly influence Study 2.
+
+% Study 2 introduction
+Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool.
+In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
+evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
+notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
+projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationship 
+through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
+alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
+implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
+\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
+conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
+found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
+large problems found in many other indirect conflict
+tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to full understand 
+the causes of these indirect conflict tool issues, a third study was conducted.
+
+% Study 3 introduction
+
+In order to fully understand the root causes of information overload, false positives, and
+scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
+cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
+I also set out to understand what mitigation strategies developers currently use as opposed to those created
+by academia. Through this exploration, I looked to find what can be accomplished moving forward with indirect conflicts
+in both research and industry. I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
+indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
+developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
+and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
+context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
+a fourth and final study.
+
+% Study 4 introduction
+In order to begin to address the gap in software evolution analytical tools discovered in study 3, I turn my analysis to the notions 
+of software change trends, specifically those trends around major releases. Change trends are trends which indicate
+a likelihood for a change type to occur around a certain event. Change trends have been used to detect
+stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
+With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
+be leveraged to detect stability and the proneness of indirect conflicts (as will be seen in Study 3), this study investigates the question:
+``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
+stability measure?}''. I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
+throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
+prominently at major release points of the projects studied.
+
+% The rest of the thesis 
+The remainder of this dissertation is laid out as follows. Chapter two includes Study 1 and Study 2 as the motivational studies which
+ultimately led to the larger research en devours found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
+press upon indirect conflicts in a more in depth fashion than has been previously seen in research. Chapter 4 contains a lengthy discussion
+of what has been learned from all four studies of this dissertation as well as implications for further research and tool development
+in the field of indirect conflicts. Finally, Chapter 5 concludes this dissertation.

### 2014-01-08 msc-thesis/chapters/1/chapter-intro.tex
MSC draft 1 complete.

@@
 
-% There are 4 studies
+% What is being addressed in this thesis
 Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
-will be explored in this dissertation. In order to fully understand indirect conflicts, this dissertation is composed
-of 4 studies of which explore different components of indirect conflicts.
+will be explored in this dissertation. I will look to explore: technical dependencies, how developers are involved,
+socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
+what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
+are being used, what future steps should be taken in academia to better industry regarding indirect conflicts, and finally,
+how software evolution analysis can be used to better tools for indirect conflicts. I will address these issues by
+conducting the following four studies.
 
@@
 by academia. Through this exploration, I looked to find what can be accomplished moving forward with indirect conflicts
-in both research and industry. I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+in both research and industry. This study asked the following 3 research questions: 
+\textit{What events or conditions lead to indirect conflicts?}
+\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
+\textit{What do developers want from future indirect conflict tools?}
+
+ I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
 developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 

### 2014-01-08 msc-thesis/chapters/3/sec-iced.tex
MSC draft 1 complete.

@@
 by academia. Through this exploration, I look to find what can be accomplished moving forward with indirect conflicts
-in both research and industry.
+in both research and industry. My research questions for this particular study are as follows:
+
+\begin{description}
+        \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
+        \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
+        \item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
+\end{description}
 

### 2014-01-13 apie-2013/paper.tex
APIE next version done.

@@
 
-The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work of release study and stability. Sections~\ref{sec:meth}
-explains our methodology and tools used in solving our research problem.
-Section \ref{sec:results} gives a full representation of the results and trends found as well as the 9 change trends found
-to be most prominent. Finally, Sections~\ref{sec:fut} and~\ref{sec:con} give our future work to follow this study and final conclusions respectively.
+The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work, Sections~\ref{sec:meth}
+explains our methodology and tools used in solving our research problem,
+Section \ref{sec:results} gives a full representation of the results and trends found, and Sections~\ref{sec:fut} and~\ref{sec:con} give 
+our future work to follow this study and final conclusions respectively.
 

### 2014-01-13 iced-2013/paper.tex
ICED changed intro for new version.

@@
 
-In order to fully understand the root causes of information overload, false positives, and
+In order to investigate the root causes of information overload, false positives, and
 scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
@@
 by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
-in both research and industry. To accomplish these goals, we ask the following:
+in both research and industry and how we can get over the road block of these information overload related
+issues. To accomplish these goals, we ask the following:
 

### 2014-01-29 iced-2013/paper.tex
ICED peggy and casper changes.

@@
 we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
-source projects, and a survey with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
+source projects, and a questionnaire with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
 we obtained insights about events and conditions in which indirect conflicts are likely to occur in
@@
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
-which influences behavior negatively in another location in the system. Indirect conflict tools however, have
+positive developer feedback. Indirect conflicts (ICs) are caused by source code changes that negatively impact another 
+location in the software system such as when libraries are upgraded or when a method is changed
+and invoking methods are influence negatively. Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 
-We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78 
 developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
@@
 
+\section{Related Work}
+\label{sec:related}
+
+Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
+which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
+Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
+al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
+such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
+indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
+detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
+creates an environment where developers can see how source code changes will affect other areas of a project at the 
+method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
+been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
+needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
+instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
+Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
+which were explored in this paper. Our own tool 
+Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
+
+Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
+Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
+artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
+created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
+Socio-technical congruence however, is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+
+For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
+recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
+Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
+found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
+require action on the developer's half.  Kim
+also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
+were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
+noting ``I get a big laundry list... I see the email and I delete it''.
+
+In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
+preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
+development processes.
+
 \section{Methodology}
@@
 We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
-addressed our 3 main research questions. Second, a survey was conducted
+addressed our 3 main research questions. Second, a questionnaire was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
@@
 Our interview participants came from a large breadth of both open and closed source software development companies
-and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation,
-Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
+and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
 Our participants were invited based on their direct involvement in the actual writing of software for
-their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
-with an average of 8 years of experience.
+their respective companies or projects. These participants' software development experience had an average of 8 years of experience.
 In addition to software development, some participants were also invited based on their experience with project management
-at some capacity. See Table~\ref{tab:demo} for more demographic details.
+at some capacity.
 
@@
 after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
-19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
-interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
-audio being stored for future use.
+19 interviews. Interviews were conducted in person when possible, or over video chat when not possible, and recorded for audio content only.
 
@@
 
-Interview participants first answered a number of  demographic questions. We then asked them to describe various software development experiences regarding our three research questions.
-Specifically, ten semi-structured  topics directly related to our research questions guided our interview:
-
-%Our three questions were studied by having the interview participants talk about their experiences and opinions in
-%\totalCategories{} .
+Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences regarding our three research questions.
+Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:
 
 \begin{itemize}
-
-\item Software development tools for dependency tracking and awareness.
-
-\item Software development process for preventing indirect conflicts.
-\item Software artifact dependency levels and where conflicts can arise.
-
-\item How developers find internal or external software dependencies.
-
-\item Examples of indirect conflicts from real world experiences.
-\item How indirect conflicts are detected and found.
-
-\item How indirect conflict issues are solved or dealt with.
-\item Developer opinion of preemptive measures to prevent indirect conflicts.
-
-\item Developer opinion on what types of changes are worth a preemptive action.
-
-\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
+\item What tools are used for dependency tracking?
+\item What processes are used for preventing ICs?
+\item What artifact dependency levels are analyzed and where can conflicts arise?
+\item How are software dependencies found?
+\item Give examples of ICs from real world experiences.
+\item How are ICs detected or found?
+\item How are IC issues solved or dealt with?
+\item Opinion of preemptive measures to prevent ICs.
+\item What types of changes are worth a preemptive action?
+\item Who is responsible for fixing or preventing ICs?
 \end{itemize}
 
-While each of the \totalCategories{} topics had a number of starter questions, interviews 
-largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
-However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
-to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
+Not all participants had strong opinions or any experience on every category listed above. For these participants, answers 
+to the specific categories were not required. We attribute any non answer by a participant to
 either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
@@
 
-%We account for these non answers
-%in our analysis and results as seen in Section~\ref{sec:results}.
+\subsection{Questionnaire Participants}
 
-\subsection{Survey Participants}
-
-Our survey respondents were different from our interviewees. We invited our survey participants from a similar breadth of open and closed source software development 
-companies and projects as the interviews participants with two main exceptions. The software organizations
-that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
-Subnet Solutions, and Amazon.
-Participants who took part in the round of interviews were not invited to the survey but were asked to act as a  point of contact for other 
-developers in their team, project, or organization who may be interested
-in completing the survey. Further, two other groups of developers were asked to
-participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
+Our questionnaire respondents were different from our interviewees. We invited our questionnaire participants from a similar 
+breadth of open and closed source software development 
+companies and projects as the interviews participants with two main exceptions, the addition of GitHub users as well as Apache Software Foundation (Apache) developers.
+The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation, 
+Subnet Solutions, and Amazon. The GitHub
 users were selected based on large amounts of development activity on GitHub and the Apache developers
@@
 
-\subsection{Survey Procedure}
+\subsection{Questionnaire Procedure}
 
-Survey participants were invited to participate in the survey by email. No reminder email was sent as the
-survey responses were not connected with the invitation email addresses and thus participants who did respond
-could not be identified. We directly emailed 1300 participants and ended with 78 responses
-giving a response rate of 6\%. We attribute the low response rate with: the surveys
+Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and ended with 78 responses
+giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
 were conducted during the months of July and August while many participants may be away from their regular positions.
-and our GitHub and Apache participants could not be verified as to whether or not they actively support the
-email addresses used in the invitations. In addition, the survey was considered by some to be long and require
-more development experience than may have been typical of some of those invited to participate.
 
-The  survey we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
-was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
-or different opinions from those already acquired from the interviews. The survey went through two rounds of
-piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
-with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
-the previously interviewed developers to examine the insights we developed.
-%Each non demographic question was made optional as it was shown 
-%through the interviews that some questions require more experience from participants than may be provided.
+The questionnaire we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
+was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews.
+The question topics asked in the questionnaire were:
+
+\begin{itemize}
+\item What frequency do ICs occur at around different project milestones?
+\item How does team size affect IC frequency?
+\item What software change types do developers care about?
+\item What processes are used for preventing ICs?
+\item What tools are used for detecting ICs?
+\item What tools are used for debugging ICs?
+\end{itemize}
 
 \subsection{Data Analysis}
-To analyze both the interview and survey data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
+To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
 Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
-\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
+\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
+theory techniques for analysis and not for a final theory). For an exploratory study such as
 the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
@@
 questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
-both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
-assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
-codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
-link our defined codes through categories of grounded theory such as context and consequences. 
+both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding.
 In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
@@
 
-We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
+We now present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
 outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
-results from which we draw our discussion to be seen in Section~\ref{sec:disc}.
+results from which we draw our discussion to be seen in Section~\ref{sec:disc}. In each subsection, quantitative data
+given refers to interviews conducted unless explicitly said otherwise.
 
-\subsection{What events or conditions lead to indirect conflicts?}
+\phantomsection
+\addcontentsline{toc}{subsection}{RQ1}
+\subsection*{\normalfont{RQ1.} \it{What events or conditions lead to indirect conflicts?}}
 
-From the interviewed participants, 12 developers believe that a large contributing factor to the cause
+12 developers believe that a large contributing factor to the cause
 of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
@@
 to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
-In light of object contracts, 14 interviewed developers gave examples of indirect conflicts they had experienced
+In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
@@
 library projects or are in test said that methods or functions were the root of their indirect conflict issues. 
-7 interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
+7 developers mentioned that indirect conflicts occur when a major update to an external project, 
 library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
@@
 
-From the interviewed participants, 11 explained that indirect conflicts occur ``all the time'' in
+11 developers explained that indirect conflicts occur ``all the time'' in
 their development life cycle with a minimum occurrence of once a week, with more serious issues tending
-to occur once a month. 65\% of surveyed developers answered that indirect conflicts occur on minimum bi-weekly,
-with the majority of developers saying that weekly occurrences are most common.
+to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur on minimum bi-weekly,
+with the 25\% saying that weekly occurrences are most common (seen in ~\ref{tab:often}). 5 questionnaire participants
+also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
+differ.
 
-12 developers interviewed said that
+12 developers said that
 when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
@@
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
-changes a lot, especially with agile''. Surveyed developers also added ``indirect conflicts after a release 
+changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release 
 depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a 
 major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during 
-large revamps or the implementation of cross-cutting new features.''. Surveyed participants also answered that 
+large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results. 
+Questionnaire participants also answered that 
 indirect conflicts are more likely to occur before the first major release rather than after at the daily
@@
 \hline
+In general & 18\% & 25\% & 21\% & 16\% & 3\% & 5\% & 11\% \\ \hline
 Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
@@
 \end{center}
-\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
-of developers surveyed.\label{tab:often}}
+\caption{Results of questionnaire as to how often indirect conflicts occur, in terms of percentage
+of questionnaire participants.\label{tab:often}}
 \end{table*}
 
-In terms of organizational structure, surveyed participants answered that as a project becomes larger and more 
+In terms of organizational structure, questionnaire participants answered that as a project becomes larger and more 
 developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
 occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
-saying they are like to occur in single developer projects.
+saying they are like to occur in single developer projects. This can be seen in Table~\ref{tab:env}.
 
-\subsection{What mitigation techniques are used by developers in regards to indirect conflicts?}
+\begin{table*}[tb!]
+\begin{center}
+\begin{tabular}{| p{7cm} | c | c | c | c | c |}
+\hline
+Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
+\hline
+\hline
+Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
+Developing in a team between 2 - 5 developers (inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
+Developing in a multi team environment (inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Questionnaire results about development environments in which indirect conflicts are likely to occur, in terms of percentage
+of questionnaire participants.\label{tab:env}}
+\end{table*}
+
+\phantomsection
+\addcontentsline{toc}{subsection}{RQ2}
+\subsection*{\normalfont{RQ2.} \it{What mitigation techniques are used by developers in regards to indirect conflicts?}}
 
@@
 mean that problems don't still occur in practice.
-Second, 21\% of interview developers mentioned that the add and deprecate 
+Second, 21\% of developers mentioned that the add and deprecate 
 model is used to prevent indirect conflicts once
@@
 
-Of the the 37 developers surveyed who could give positive identification of preventative processes for indirect conflicts,
+Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
 by contract or the testing of a methods contract, and 14\% said that add and
-deprecate was used in their projects to avoid indirect conflicts.
+deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.
 
-In regards to catching indirect conflicts, 13 interviewed developers mentioned forms of testing (unit, and integration) 
+In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration) 
 as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
@@
 when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. 31\% of surveyed developers said build processes (either nightly builds or building the project 
+not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project 
 themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
-time testing. The surveyed developers confirmed these results with 49\% mentioning forms of testing as the major tool used to 
+time testing. The questionnaire participants confirmed these results with 49\% mentioning forms of testing as the major tool used to 
 catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
-Surveyed developers also mentioned review process and personal expertise as factors of catching indirect conflicts.
+Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.
 
-Once an indirect conflict has occurred and developers need to resolve it, 14 developers interviewed said
+Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
 they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
@@
 code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
-frames of when events occurred. Of the developers surveyed, 23\% said they used native IDE tools and 21\% said they use 
+frames of when events occurred. Of the questionnaire participants, 23\% said they used native IDE tools and 21\% said they use 
 features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers 
@@
 
-\subsection{What do developers want from future indirect conflict tools?}
+\begin{table*}[tb!]
+\begin{center}
+\begin{tabular}{| p{7cm} | c | c | c | c | c |}
+\hline
+Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
+\hline
+\hline
+Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
+Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
+Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
+User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
+Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
+Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
+\end{tabular}
+\end{center}
+\caption{Questionnaire results about source code changes that developers deem notification worthy, in terms of percentage
+of questionnaire participants.\label{tab:pre}}
+\end{table*}
+
+\phantomsection
+\addcontentsline{toc}{subsection}{RQ3}
+\subsection*{\normalfont{RQ3.} \it{What do developers want from future indirect conflict tools?}}
 
@@
 are already in place to catch potential issues such as testing, code review, 
-individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
-want to be notified about method signature changes as they have a high chance to break the code as opposed to only
-23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
-interface change. Other change types varied from never being notified to most times being notified, showcasing the
-complexity of change types which may or may not negatively affect a project.
+individual knowledge, or static language analysis tools. Further, developers said that having strict change type monitoring 
+for indirect conflicts does not work due to the inherent complexities of indirect conflicts.
+This can be seen confirmed in that questionnaire participants had very few cases in which they always wanted to be
+alerted to a change type as seen in Table~\ref{tab:pre}. Only method signature changes caused 68\% of questionnaire participants
+to want to be always notified as they have a high chance to break the code. This is opposed to other change types listed in
+Table~\ref{tab:pre} which have 27\% or lower of developers always wanting to be notified. This again showcases the
+complexity of indirect conflicts through change types which may or may not negatively affect a project.
 
 When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
-Surveyed developers suggested the following improvements to curative tools:
+Interview and questionnaire participants suggested the following improvements to curative tools:
 
@@
 \label{sec:disc}
+
+As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
+conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
+fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
+industry settings.
 
@@
 developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
-don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
-that the developer only wants the awareness if the item will require action on their part to resolve the issue.
+don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue''.
 This sense of unwanted or limited awareness is crucial to understand why generalized
@@
 analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
-all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
+all events surrounding a developer's work or project is not possible. 
+In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers 
+tend to use the bare amount of communication
+towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
+Bolici et. al~\cite{Bolici:2009} also give the 
+possibility of developer stigmergy as the reason that developers do not explicitly communicate with each other about issues opposed to
+time management. The idea
+that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication
+and unwanted awareness.
+\textit{Developers prefer to spend their limited time dealing with
 the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
@@
 writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
-conflicts occurred at either the detection or cure levels. Obviously, developers either prefer to spend their time in curative
-measures, or do not posses the proper tools to take better action in the preventative stages. ``You're reaction time is much more crucial''
+conflicts occurred at either the detection or cure levels. ``You're reaction time is much more crucial''
 said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
@@
 which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
-of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
+of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. 
+Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
+sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
+which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
 are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
 abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
-real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
-this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
-curative measures are more productive than preventative measures.
+real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately 
+presented in this area is if curative measures are more productive than preventative measures.
 
@@
 become quite limited in identifying and tracking dependencies where they were once strong.
+In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
+ill fit of current static analysis to current development practices..
 
@@
 
-10 of the 78 participants who were surveyed sent direct responses to us asking for any results of the research to be sent directly to
+10 of the 78 questionnaire participants who were sent direct responses to us asking for any results of the research to be sent directly to
 them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
@@
 made accurately demonstrate that complexity. Since interviewed participants came from such a diverse set of organizations, 
-software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
+software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and questionnaires, participants
 expressed unique situations that did not fully meet our generalized findings or on going theories. In these cases, we worked in the specific
@@
 
-\section{Related Work}
-\label{sec:related}
-
-Since this paper has covered a wide spectrum in regards to indirect conflicts (prevention, detection, resolution, process, and tools), there
-exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
-in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging are presented.
-
-Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
-which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
-Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
-al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
-such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
-indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
-detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
-creates an environment where developers can see how source code changes will affect other areas of a project at the 
-method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
-been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
-needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
-instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
-Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
-which were explored in this paper. Our own tool 
-Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
-
-Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
-times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
-Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
-artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
-created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
-Socio-technical congruence however, is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
-
-For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
-recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
-Kim conducted several focus groups as well as web surveys to determine these results. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
-require action on the developer's half.  Kim
-also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
-were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
-noting ``I get a big laundry list... I see the email and I delete it''.
-
-In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
-preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
-development processes.
-
-In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
-ill fit of current static analysis to current development practices. Not being designed properly to fit current development work
-flow, and the over abundance of false positives were shown to be factors of developer non-use.
-
-Hattori et. al.~\cite{Hattori:2012:ICG} found, through qualitative user studies, that developers tend to use the bare amount of communication
-towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
-resonates with ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
-sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
-This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici:2009} give the 
-possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea
-that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
-Blincoe et al.~\cite{Blincoe:2013:UCC} also found, through data mining techniques, a subset of events related to awareness 
-which occur that developers will communicate about as opposed to all awareness events. This again confirms our understanding that developers
-only care about certain events which require some action on their part.
-
-While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
-can be applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
-changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
-with the idea that developers wait for something to go wrong, then wish to fix the issue as fast as possible. Program slicing~\cite{Xu:2005:BSP}
-has also been extensively studied in order to aid in debugging in manual or automated techniques.
-
 \section{Conclusions}
@@
 losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
-19 interviewed developers from 12 organizations as well as 78 surveyed developers, we have provided characterization of indirect conflicts,
+19 interviewed developers from 12 organizations as well as 78 developers completing questionnaires, we have provided characterization of indirect conflicts,
 current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
@@
 We sincerely thank all participants who were willing to be interviewed or who participated
-in completing our survey. We thank these people for taking time out of their day to participate and for
+in completing our questionnaire. We thank these people for taking time out of their day to participate and for
 sharing their developer experience with us.

### 2014-01-29 apie-2013/paper.tex
APIE camera formatting.

@@
-\documentclass[conference]{IEEEtran}
+\documentclass{sig-alternate}
 
@@
 \usepackage{float}
-\floatstyle{plaintop}
-\restylefloat{table}
 
@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{API Evolution: A Study of Software Releases and Change Trends Leading to Stability}
+\title{API Evolution: A Study of Software Releases and Change Trends}
 
-% Authors names
-\author{\IEEEauthorblockN{Jordan Ell}
-\IEEEauthorblockA{University of Victoria\\
-British Columbia, Canada \\ jell@uvic.ca}
-\and
-\IEEEauthorblockN{Braden Simpson}
-\IEEEauthorblockA{University of Victoria\\
-British Columbia, Canada \\ braden@uvic.ca}
-\and
-\IEEEauthorblockN{Daniela Damian}
-\IEEEauthorblockA{University of Victoria\\
-British Columbia, Canada \\ danielad@csc.uvic.ca}
+\numberofauthors{3}
+\author{
+\alignauthor
+Jordan Ell\\
+       \affaddr{University of Victoria}\\
+       \affaddr{3800 Finnerty Rd}\\
+       \affaddr{Victoria, Canada}\\
+       \email{jell@uvic.ca}
+\alignauthor
+Braden Simpson\\
+       \affaddr{University of Victoria}\\
+       \affaddr{3800 Finnerty Rd}\\
+       \affaddr{Victoria, Canada}\\
+       \email{braden@uvic.ca}
+\alignauthor
+Daniela Damian\\
+       \affaddr{University of Victoria}\\
+       \affaddr{3800 Finnerty Rd}\\
+       \affaddr{Victoria, Canada}\\
+       \email{danielad@csc.uvic.ca}
 }
+
+\date{07 February 2014}
 
@@
 \end{abstract}
+
+\keywords{Change trends, software repositories, software releases}
 
@@
 
-The remainder of this paper is laid out as follows. Section~\ref{sec:rel} goes over the related work, Sections~\ref{sec:meth}
-explains our methodology and tools used in solving our research problem,
-Section \ref{sec:results} gives a full representation of the results and trends found, and Sections~\ref{sec:fut} and~\ref{sec:con} give 
-our future work to follow this study and final conclusions respectively.
-
 \section{Related Work}
@@
 \label{sec:meth}
-In order to answer our research question, we decided to use the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows us to detect fine grained
+To answer our research question, we used the tool ChangeDistiller created by Fluri et al.~\cite{Fluri:2007:CDT}. This tool allows us to detect fine grained
 source code changes in Java projects. This tool works by building an abstract syntax tree of a file before and after a code change, then it tries to determine
@@
 
-We took ChangeDistiller and applied it across 10 open source Java projects. Java projects were nesissary because ChangeDistiller only works for Java source code.
-For each of the projects, we obtained the software configuration management (SCM) system
-which is used to store all source code changes of a project. When it was necessary, we converted some forms of SCM system to Git in order to reduce implementation
-burdens of using multiple SCMs. Once the SCM was obtained, we used ChangeDistiller and iterated over every commit found in a project's git master branch. We stored
+We took ChangeDistiller and applied it across 10 open source Java projects. Java projects were necessary because ChangeDistiller only works for Java source code.
+For each of the projects, we obtained the software repository and used ChangeDistiller to iterate over every commit found in a project's master branch. We stored
 34 of ChangeDistiller's built in source code change types for each commit. We noted how many of each change type was performed in each commit and stored that information
@@
 
-Aside from generating quantitative data, we also created a web application for the visualization of the data called API Evolution (APIE). This visualizer allowed
+We also created a web application for the visualization of the data called API Evolution (APIE). This visualizer allowed
 us to inspect a single project and a single change type metric at a time (see Figure~\ref{fig:apie}) for qualitative analysis of software evolution trends. We
@@
 out the curves presented by the tool to allow easier manual inspection. This method however does not take activity into account as seen in
-Equation~\ref{eq:norm}, so it represents the true activity and change types occuring. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
+Equation~\ref{eq:norm}, so it represents the true activity and change types occurring. Manual inspections were labeled into 4 categories: upward trending, local maximum, downward
 trending, and local minimum. Since the graphs were quite turbulent, best estimations were conducted by two judges at each release point to fit the graph into the aforementioned
@@
 To answer our research question, we conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
-eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
+\\eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and \\eclipse.maven.surefire. These project were chosen
 because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.
 
-Due to space requirements of this paper, we focus our results on major releases of the 10 case study projects and select few of the calculated change ratios. 
+We focus our results on major releases of the 10 case study projects and select few of the calculated change ratios. 
 There were 109 major releases across the 10 studied projects. All of the major findings as per values computed from Equation~\ref{eq:norm} for non test metrics
@@
 \begin{center}
+\caption{Implementation oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:ratio}}
 \tabcolsep=0.11cm
@@
 \end{center}
-\caption{Implementation oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:ratio}}
 \end{table}
@@
 \begin{table*}[tb!]
-\begin{center}
+\centering
+\caption{Qualitative graph analysis results. \label{tab:qual}}
 \begin{tabular}{| l | c | c | c | c |}
@@
 \end{tabular}
-\end{center}
-\caption{Qualitative graph analysis results. \label{tab:qual}}
 \end{table*}
 
-Lastly we found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
+Lastly, we found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
 ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet our threshold and are indicators of stability with regards to test based
@@
 \begin{table}[h]
-\begin{center}
+\centering
+\caption{Test oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:test}}
 \begin{tabular}{| l | c | c | c |}
@@
 \end{tabular}
-\end{center}
-\caption{Test oriented change types and their normalized average change ratios at 60 days on each side of releases. \label{tab:test}}
 \end{table}
@@
 
-In future work of change trend analysis, we have planned further statistical tests of the change trends identified in this paper. We would like to
-be able to determine if these patterns are unique to major releases versus other types of releases (minor, alpha, beta, release candidate) and to
-be able to determine if these pattern are unique to releases alone or can be spotted throughout the history of a project.
+In future work, we have planned further statistical tests of the change trends identified in this paper. We would like to
+determine if these patterns are unique to major releases versus other types of releases (minor, alpha, beta, release candidate) and to
+determine if these pattern are unique to releases alone or can be spotted throughout the history of a project.
 
-If it is deemed that these change trends are present in no other releases, some, all, or in non release history of the project, we will
+If change trends are present in no other releases, some, all, or in non release history of the project, we will
 compare the history of the projects as to where these trends are found against preexisting software stability and quality metrics. We will use this data to
@@
 The source code change trends found in this paper, in combination with future work and analysis, will allow for the detection of stable code
-points throughout a software projects life time. This paper has also shown the beginnings of a visualization for source code change trends which may
-be used as a visual cue towards project stability and potential areas of instability where action may need to be taken.
+points throughout a software projects life time.
 
-\bibliographystyle{IEEEtran}
-%\balance
+\bibliographystyle{abbrv}
 \bibliography{paper}

### 2014-01-31 iced-2013/paper.tex
ICED intro fixes.

@@
 identification of failure inducing changes, other than those which are already identifiable by other means such
-as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
-also compounded by the scale of many software development projects leading to further information overload.
+as compilers, and unit tests, to these dependencies still remains an issue.
 Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
@@
 However, issues again of information overload, dependencies (in developer organizational structure), and scalability
-come up.
+come up. All the aforementioned tools have not been able to study their own problems as their evaluations have
+been small scale, involving a very limited number of industrial participants or student evaluators.
 
 In order to investigate the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
-cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
-We then set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
+scalability issues in regards to indirect conflicts, we conducted a large scale an empirical study involving industrial
+participants in order to determine: what events occur to
+cause indirect conflicts, when they occur, if compounding conditions exist, and what mitigation strategies developers currently
+use. Through this large exploration, we look to find what can be accomplished moving forward with indirect conflicts
 in both research and industry and how we can get over the road block of these information overload related

### 2014-02-04 apie-2013/paper.tex
APIE final edits.

@@
 \begin{abstract}
-As software evolves over the lifetime of a project, changes to software objects such as methods or classes occur, which can have ripple effects throughout
+As software evolves over the lifetime of a project, changes to software such as methods or classes occur, which can have ripple effects throughout
 the rest of the system causing relationships to be re-engineered and causing instability. However, projects are sometimes deemed stable at certain release
@@
 \section{Introduction}
-Release points are a vital milestone of software projects. From major releases of a Waterfall based project to minor iterations of an Agile development,
+Release points are a vital milestone of software projects. From major releases of a Waterfall based project to iterations of an Agile development,
 releases form an interesting single point of a project's development history.
@@
 
-We turn our analysis to the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
+In this paper, we perform analysis of software change trends, specifically those trends around major releases. Change trends are trends which indicate
 a likelihood for a change type to occur around a certain event. Change trends have been used to detect
@@
 ``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
-stability measure?}''. By answering this question, we hope to provide industry with the ability to determine a project's stability and perhaps
-reliability based on its current change trends. We also hope to provide researchers with the ability to perform further analysis of projects in
+stability measure?}''. By answering this question, we hope to provide the ability to determine a project's stability and perhaps
+reliability based on its current change trends. We also look to provide researchers with the ability to perform further analysis of projects in
 regards to software quality with the findings of this paper.
@@
 In this paper, we perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
-throughout their history. We studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
-prominently at major release points of the projects studied.
+throughout their history. We studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
+prominently at major release points of the projects studied. These 9 change trends can be used to determine if a project is likely near a major release point
+or can be deemed stable.
 
@@
 
-We took ChangeDistiller and applied it across 10 open source Java projects. Java projects were necessary because ChangeDistiller only works for Java source code.
+We conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.\\project, 
+eclipse.mylyn.commons, eclipse.mylyn.context, \\hibernate-orm, hibernate-ogm, hibernate-search, eclipse.\\maven.core, and eclipse.maven.surefire. These project were chosen
+because ChangeDistiller only works for Java source code and because of their high use amongst other Java projects and to study specific ecosystems 
+of projects and their evolution trends.
+
 For each of the projects, we obtained the software repository and used ChangeDistiller to iterate over every commit found in a project's master branch. We stored
@@
 
-Once the ChangeDistiller information was collected, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
+Next, we decided to examine software change trends surrounding releases of the project's we had selected. Since releases
 have preconceived notions of software stability, we decided that by studying the change types surrounding these releases, we could get a better understanding of
@@
 
-To answer our research question, we conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
-\\eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and \\eclipse.maven.surefire. These project were chosen
-because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.
-
 We focus our results on major releases of the 10 case study projects and select few of the calculated change ratios. 
@@
 
-As it can be seen in Table~\ref{tab:ratio}, there are few change type trends around major releases which pass our threshold. We can see that both public
+We found 4 major change trends occurring before major releases: adding of private classes, changing of test method signatures, changing of documentation,
+and removal of test classes.
+We also found 5 major change trends occurring after major release: adding of test methods, changing of test classes, removal of public methods, removal
+of private classes, and removal of private methods.
+
+Table~\ref{tab:ratio} shows there are few change type trends around major releases which pass our threshold. We can see that both public
 and private methods
@@
 
-Another interesting trend that can been seen in Table~\ref{tab:ratio} is that of changes to public objects. We can see for public classes and methods that
+Another interesting trend seen in Table~\ref{tab:ratio} is that of changes to public objects. We can see for public classes and methods that
 5 out of 7 ratios indicate changes occur to these objects after major release rather than before. We hypothesize that these changes to the public API
@@
 
-In future work, we have planned further statistical tests of the change trends identified in this paper. We would like to
+We have planned further statistical tests of the change trends identified in this paper. We would like to
 determine if these patterns are unique to major releases versus other types of releases (minor, alpha, beta, release candidate) and to
@@
 
-Software stability is an often used measure in determining some aspects of code quality. This paper has presented some initial indications as to how
+This paper has presented some initial indications as to how
 software stability may be found, based on major release stabilities, by analyzing source code change type trends both before and after major software

### 2014-02-11 msc-thesis/chapters/1/chapter-intro.tex
MSC Added changes to thesis.

--- 2014-01-08 msc-thesis/chapters/1/chapter-intro.tex
@@
 ~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
-positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
-which influences behavior negatively in another location in the system. Indirect conflict tools however, have
+positive developer feedback. Indirect conflicts are caused by source code changes that negatively impact another
+location in the software system such as when libraries are upgraded or when a method is changed
+and invoking methods are influence negatively. Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 also compounded by the scale of many software development projects leading to further information overload.
-Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+Lastly, social factors such as Cataldo et al's.~\cite{Cataldo:2006:ICR} notion of socio-technical
 congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
@@
 
-% There are 4 studies
+% What is being addressed in this thesis
 Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
-will be explored in this dissertation. In order to fully understand indirect conflicts, this dissertation is composed
-of 4 studies of which explore different components of indirect conflicts.
+will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
+with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
+in the are of indirect conflicts. To accomplish this goal, I have researcher the following sub topics of indirect
+conflicts: technical dependencies, how developers are involved in said dependencies,
+socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
+what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
+are being used, what future steps should be taken in academia to better industry regarding indirect conflicts, and finally,
+how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
+conducting the following four studies.
+
+\section{Research Methodology}
+
+In order to address the research goal as laid out in the previous section, I have conducted 4 studies which will
+now be briefly outlined. Each study builds off the previous one and has research questions informed from the
+findings of the previous study.
 
@@
 % Study 2 introduction
-Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool.
+Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool while
+answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
+developers who's changing technical dependencies statistically relate to bugs?}''.
 In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
@@
 large problems found in many other indirect conflict
-tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to full understand 
+tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to fully understand 
 the causes of these indirect conflict tool issues, a third study was conducted.
@@
 I also set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, I looked to find what can be accomplished moving forward with indirect conflicts
-in both research and industry. I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
-developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
+by academia. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
+in both research and industry. This study asked the following 3 research questions: 
+\textit{What events or conditions lead to indirect conflicts?}
+\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
+\textit{What do developers want from future indirect conflict tools?}
+
+I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. The study findings indicate that: 
 indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
@@
 stability measure?}''. I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
-throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
+throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
 prominently at major release points of the projects studied.

### 2014-02-11 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC Added all 4 studies to the thesis.
MSC intro now has big chunks in it.
MSC Added changes to thesis.

@@
 With the power of technical dependencies in predicting software failures, the question I
-investigated in this study was : 
-``\textit{Is it possible to identify pairs of developers whose technical dependencies in code changes
-statistically relate to bugs?}''
+investigated in this study was:
+
+\begin{description}
+        \item[RQ\namedlabel{itm:rq1-pairs}{RQ1-Pairs}] \textit{Is it possible to identify pairs 
+        of developers whose technical dependencies in code changes statistically relate to bugs?}
+\end{description}
 
@@
 The process utilizes code changes and the call hierarchies effected to find patterns of developer 
-relationships in successful and failed code changes. As it will be seen, we found 27 statistically
+relationships in successful and failed code changes. As it will be seen, I found 27 statistically
 significant failure inducing developer pairs. These developer relationships can also be used
@@
 % Technical network figure
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering

### 2014-02-11 msc-thesis/chapters/2/sec-impact.tex
MSC Added changes to thesis.

@@
 a web based tool that aims at detecting indirect conflicts among developers
-and notifying the appropriate members involved in these conflicts.
+and notifying the appropriate members involved in these conflicts. Through Impact and
+its evaluation I ask:
+
+\begin{description}
+        \item[RQ\namedlabel{itm:rq1-impact}{RQ1-Impact}] \textit{Can indirect conflicts be supported 
+        through an awareness mechanism which leverages pairs of developers who's changing technical 
+        dependencies statistically relate to bugs?}
+\end{description}
+
 By leveraging technical relationships inherent of 
@@
 
-\section{Related Work}
+After a brief evaluation of \textit{Impact} with two student software teams, it was
+found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
+large problems found in many other indirect conflict
+tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.
+
+\subsection{Related Work}
 Although there is an abundance of awareness tools developed in research
@@
 
+\subsection{Threats to Validity}
+% TODO
+
 \subsection{Conclusions of Study}

### 2014-02-11 msc-thesis/chapters/3/chapter-exploration.tex
MSC Added changes to thesis.

@@
 \label{chapter:exploration}
-
-As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
-development has become the standard development process instead of the exception. With this parallel development
-comes the need for larger awareness among developers to have ``an understanding of the activities of others
-which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
-mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
-empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
-and time-consuming chore for developers~\cite{Perry:2001:PCL}.
 
@@
 
-Social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+Social factors such as Cataldo et al's.~\cite{Cataldo:2006:ICR} notion of socio-technical
 congruence, have also been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
@@
 In order to fully understand the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, I will proceed by taking a step back and determine what events occur to
+scalability issues in regards to indirect conflicts in this study, I examine and determine what events occur to
 cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. By determining the root
@@
 monitoring in order to only detect those changes with a high likelihood of causing indirect conflicts.
-I then set out to understand what mitigation strategies developers currently use as opposed to those created
+I then determine what mitigation strategies developers currently use as opposed to those created
 by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
@@
 the hopes of a higher adoption rate of tools produced by academia.
-Finally, I look to find what can be accomplished moving forward with indirect conflicts
+Finally, I examine what can be accomplished moving forward with indirect conflicts
 in both research and industry.
 
-I restate the research goals of this thesis for ease of the reader:
+To explore and answer the research goals listed above, I performed a study (Section~\ref{study:exploration}) in which I 
+interviewed 19 developers from across 12 commercial and open source projects, 
+followed by a confirmatory questionnaire of 78 developers, and 5 confirmatory interviews.
 
-\begin{description}
-  \item \textit{What events or conditions lead to indirect conflicts?}
-  \item \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
-  \item \textit{What do developers want from future indirect conflict tools?}
-\end{description}
-
-To answer these research questions, I performed a study (Section~\ref{study:exploration}) in which I interviewed 19 developers from across 12 commercial and open source projects, 
-followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews. Based on the findings (to be seen) to the aforementioned research questions, 
-I also performed a secondary complimentary study of software evolutionary trends (Section~\ref{study:apie}) to solidify
-developer opinion and provide a starting point for the future development of indirect conflict tools.
+Based on some of the findings (to be seen in details in Section~\ref{study:exploration}) I performed a follow up study
+which did not relate directly to the themes of this dissertation but helped strengthen the results found in Study 3.
+Some results of Study 3 showed that indirect conflict tools should take into account a contextual setting of development
+progression in software projects to better inform developers of potential indirect conflicts. In other words, an indirect
+conflict tool should be able to tell what phase of development inside the development life cycle a project is currently
+active in. In order to better explore and support this finding, I performed a complimentary study of software evolutionary trends (Section~\ref{study:apie}).
+In Study 4, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
+throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
+prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
+to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
+frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
+to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.
 

### 2014-02-11 msc-thesis/chapters/3/sec-apie.tex
MSC Added changes to thesis.

@@
 
-Release points are a vital milestone of software projects. From major releases of a Waterfall based project to minor iterations of an Agile development,
+Release points are a vital milestone of software projects. From major releases of a Waterfall based project to iterations of an Agile development,
 releases form an interesting single point of a project's development history.
@@
 
-I turn my analysis to the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
+In this study, I examine the notions of software change trends, specifically those trends around major releases. Change trends are trends which indicate
 a likelihood for a change type to occur around a certain event. Change trends have been used to detect
@@
 be leveraged to detect stability and the proneness of indirect conflicts, this study investigates the question:
-``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
-stability measure?}''.
+
+\begin{description}
+        \item[RQ\namedlabel{itm:rq1-apie}{RQ1-APIE}] \textit{What trends exist in source code changes surrounding major releases of open 
+        source projects as a notion towards a project stability measure?}
+\end{description}
 
 In this study, I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
-throughout their history. I studied 26 quantitative and 16 qualitative change trends and identified a core group of 9 change trends which occur
+throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
 prominently at major release points of the projects studied. These change trends can provide context as to when indirect conflicts are more likely
-to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near and major release and more
+to occur based on the findings from Study 3 as I found that indirect conflicts tend to be become less frequent near major release and more
 frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
@@
 
-\section{Related Work}
+\subsection{Related Work}
 \label{sec:rel}
@@
 
-I took ChangeDistiller and applied it across 10 open source Java projects. Java projects were necessary because ChangeDistiller only works for Java source code.
+I conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
+eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
+because ChangeDistiller only works for Java source code and because of their high use amongst other Java projects and to study specific ecosystems 
+of projects and their evolution trends.
+
 For each of the projects, I obtained the software configuration management (SCM) system
@@
 
-To answer my research question, I conducted a case study of 10 open source Java projects. These projects are: eclipse.jdt.core, eclipse.jdt.ui, eclipse.jetty.project, 
-eclipse.mylyn.commons, eclipse.mylyn.context, hibernate-orm, hibernate-ogm, hibernate-search, eclipse.maven.core, and eclipse.maven.surefire. These project were chosen
-because of their high use amongst other Java projects and to study specific ecosystems of projects and their evolution trends.
-
 Due to time requirements, I focus my results on major releases of the 10 case study projects and select few of the calculated change ratios. 
@@
 or 20\% greater trend of before the release date) to indicate the greatest trends.
+
+I found 9 major change trends which surround major releases in the open source
+projects studied. 4 change trends found that occur before major releases are: added private classes, 
+changed test method signatures, changed documentation, and removed test classes.
+5 change trends found that occur after major release are: added test methods, changed test classes, removed public methods, removed
+private classes, and removed private methods.
 

### 2014-02-11 msc-thesis/chapters/3/sec-iced.tex
MSC Added changes to thesis.

@@
 In order to fully understand the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, I conducted an empirical study to determine what events occur to
-cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
-I then set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, I look to find what can be accomplished moving forward with indirect conflicts
-in both research and industry. My research questions for this particular study are as follows:
+scalability issues in regards to indirect conflicts, I conducted a large scale empirical study involving industrial
+participants in order to determine: what events occur to
+cause indirect conflicts, when they occur, if compounding conditions exist, and what mitigation strategies developers currently
+use. Through this large exploration, what can be accomplished moving forward with indirect conflicts
+in both research and industry and how we can get over the road block of these information overload related
+issues. My research questions for this particular study are as follows:
 
@@
 
-I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
+I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78 
 developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. My findings indicate that: 
@@
 
+\subsection{Related Work}
+\label{sec:iced-related}
+
+Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
+which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
+Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
+al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
+such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
+indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
+detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
+creates an environment where developers can see how source code changes will affect other areas of a project at the
+method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
+been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
+needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
+instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
+Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
+which were explored in this paper. Our own tool
+Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
+
+Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
+developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}.
+Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
+artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
+created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
+Socio-technical congruence however, is largely unproven in regards to its correlation to
+software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+
+For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
+recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
+Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
+found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
+require action on the developer's half.  Kim
+also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
+were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
+noting ``I get a big laundry list... I see the email and I delete it''.
+
+In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
+preventative measures I have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
+development processes.
+
 \subsection{Methodology}
@@
 I performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
-addressed my 3 main research questions. Second, a survey was conducted
+addressed my 3 main research questions. Second, a questionnaire was conducted
 which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
@@
 My interview participants came from a large breadth of both open and closed source software development companies
-and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation,
-Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
+and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
 My participants were invited based on their direct involvement in the actual writing of software for
@@
 \begin{itemize}
-
-\item Software development tools for dependency tracking and awareness.
-
-\item Software development process for preventing indirect conflicts.
-\item Software artifact dependency levels and where conflicts can arise.
-
-\item How developers find internal or external software dependencies.
-
-\item Examples of indirect conflicts from real world experiences.
-\item How indirect conflicts are detected and found.
-
-\item How indirect conflict issues are solved or dealt with.
-\item Developer opinion of preemptive measures to prevent indirect conflicts.
-
-\item Developer opinion on what types of changes are worth a preemptive action.
-
-\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
+\item What tools are used for dependency tracking?
+\item What processes are used for preventing indirect conflicts?
+\item What artifact dependency levels are analyzed and where can conflicts arise?
+\item How are software dependencies found?
+\item Give examples of indirect conflicts from real world experiences.
+\item How are indirect conflicts detected or found?
+\item How are indirect conflict issues solved or dealt with?
+\item Opinion of preemptive measures to prevent indirect conflicts.
+\item What types of changes are worth a preemptive action?
+\item Who is responsible for fixing or preventing indirect conflicts?
 \end{itemize}
@@
 
-\subsubsection{Survey Participants}
+\subsubsection{Questionnaire Participants}
 
-My survey respondents were different from my interviewees. I invited our survey participants from a similar breadth of open and closed source software development 
+My questionnaire respondents were different from my interviewees. I invited our questionnaire participants from a similar breadth of open and closed source software development 
 companies and projects as the interviews participants with two main exceptions. The software organizations
-that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
+that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation, 
 Subnet Solutions, and Amazon.
-Participants who took part in the round of interviews were not invited to the survey but were asked to act as a point of contact for other 
+Participants who took part in the round of interviews were not invited to the questionnaire but were asked to act as a point of contact for other 
 developers in their team, project, or organization who may be interested
-in completing the survey. Further, two other groups of developers were asked to
+in completing the questionnaire. Further, two other groups of developers were asked to
 participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
@@
 
-\subsubsection{Survey Procedure}
+\subsubsection{questionnaire Procedure}
 
-Survey participants were invited to participate in the survey by email. No reminder email was sent as the
-survey responses were not connected with the invitation email addresses and thus participants who did respond
+questionnaire participants were invited to participate in the questionnaire by email. No reminder email was sent as the
+questionnaire responses were not connected with the invitation email addresses and thus participants who did respond
 could not be identified. I directly emailed 1300 participants and ended with 78 responses
-giving a response rate of 6\%. I attribute the low response rate with: the surveys
+giving a response rate of 6\%. I attribute the low response rate with: the questionnaires
 were conducted during the months of July and August while many participants may be away from their regular positions.
 and my GitHub and Apache participants could not be verified as to whether or not they actively support the
-email addresses used in the invitations. In addition, the survey was considered by some to be long and require
+email addresses used in the invitations. In addition, the questionnaire was considered by some to be long and require
 more development experience than may have been typical of some of those invited to participate.
 
-The survey I designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
+The questionnaire I designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
 was based on the insights I obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
-or different opinions from those already acquired from the interviews. The survey went through two rounds of
-piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
-with feedback given at the end. Not only did this allow me to create a more polished survey, but it also allowed 
+or different opinions from those already acquired from the interviews. The questionnaire went through two rounds of
+piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the questionnaire
+with feedback given at the end. Not only did this allow me to create a more polished questionnaire, but it also allowed 
 the previously interviewed developers to examine the insights I developed.
 
+The question topics asked in the questionnaire were:
+
+\begin{itemize}
+\item What frequency do ICs occur at around different project milestones?
+\item How does team size affect IC frequency?
+\item What software change types do developers care about?
+\item What processes are used for preventing ICs?
+\item What tools are used for detecting ICs?
+\item What tools are used for debugging ICs?
+\end{itemize}
+
 \subsubsection{Data Analysis}
-To analyze both the interview and survey data, I used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
+To analyze both the interview and questionnaire data, I used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
 Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
@@
 questions that have been formulated in regards to forming hypotheses. In my study being presented, data collected from
-both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
+both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
 assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
@@
 
-I now present my results of both the interviews and surveys conducted in regards to my 3 research questions
+I now present my results of both the interviews and questionnaires conducted in regards to my 3 research questions
 outlined in this chapter and Chapter 1. I restate each research question, followed by my quantitative and qualitative
-results from which I draw my discussion to be seen in Chapter~\ref{chapter:discussion}.
+results from which I draw my discussion to be seen in Chapter~\ref{chapter:discussion}. In each subsection, quantitative data
+given refers to interviews conducted unless explicitly said otherwise.
 
@@
 
-From the interviewed participants, 12 developers believe that a large contributing factor to the cause
+12 developers believe that a large contributing factor to the cause
 of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
@@
 to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark.
-In light of object contracts, 14 interviewed developers gave examples of indirect conflicts they had experienced
+In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
@@
 library projects or are in test said that methods or functions were the root of their indirect conflict issues.
-7 interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
+7 developers mentioned that indirect conflicts occur when a major update to an external project,
 library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
@@
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
-changes a lot, especially with agile''. Surveyed developers also added ``indirect conflicts after a release 
+changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release
 depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
 major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
-large revamps or the implementation of cross-cutting new features.''. Surveyed participants also answered that 
+large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results.
+Questionnaire participants also answered that
 indirect conflicts are more likely to occur before the first major release rather than after at the daily
@@
 
-\begin{table*}[ht]
+\begin{table*}[tb!]
 \begin{center}
-\begin{tabular}{| p{1.6cm} | c | c | c | c | c | c | c |}
+\begin{tabular}{| p{2cm} | c | c | p{1.2cm} | c | p{1.3cm} | c | c |}
 \hline
@@
 \hline
-Early stages of development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
+In general & 18\% & 25\% & 21\% & 16\% & 3\% & 5\% & 11\% \\ \hline
+Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
 Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
@@
 \end{center}
-\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
-of developers surveyed.\label{tab:often}}
+\caption{Results of questionnaire as to how often indirect conflicts occur, in terms of percentage
+of questionnaire participants.\label{tab:often}}
 \end{table*}
 
-In terms of organizational structure, surveyed participants answered that as a project becomes larger and more 
+In terms of organizational structure, questionnaire participants answered that as a project becomes larger and more
 developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
 occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
-saying they are like to occur in single developer projects.
+saying they are like to occur in single developer projects. This can be seen in Table~\ref{tab:env}.
 
-As per organizational structure, Table~\ref{tab:env} shows which development team environment developers believe to be the most
-prone to indirect conflicts.
-
-\begin{table*}[ht]
+\begin{table*}[tb!]
 \begin{center}
@@
 Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
-Developing in a team between 2 - 5 developers (Inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
-Developing in a multi team environment (Inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
+Developing in a team between 2 - 5 developers (inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
+Developing in a multi team environment (inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
 \end{tabular}
 \end{center}
-\caption{Results of survey questions to development environments in which indirect conflicts are likely to occur, in terms of percentage
-of developers surveyed.\label{tab:env}}
+\caption{Questionnaire results about development environments in which indirect conflicts are likely to occur, in terms of percentage
+of questionnaire participants.\label{tab:env}}
 \end{table*}
@@
 mean that problems don't still occur in practice.
-Second, 21\% of interview developers mentioned that the add and deprecate 
+Second, 21\% of developers mentioned that the add and deprecate
 model is used to prevent indirect conflicts once
@@
 
-Of the the 37 developers surveyed who could give positive identification of preventative processes for indirect conflicts,
+Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
 27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
 by contract or the testing of a methods contract, and 14\% said that add and
-deprecate was used in their projects to avoid indirect conflicts.
+deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.
 
-In regards to catching indirect conflicts, 13 interviewed developers mentioned forms of testing (unit, and integration) 
+In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration)
 as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration
@@
 when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. 31\% of surveyed developers said build processes (either nightly builds or building the project 
+not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project
 themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run
-time testing. The surveyed developers confirmed these results with 49\% mentioning forms of testing as the major tool used to 
+time testing. The questionnaire participants confirmed these results with 49\% mentioning forms of testing as the major tool used to
 catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
-Surveyed developers also mentioned review process and personal expertise as factors of catching indirect conflicts.
+Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.
 
-Once an indirect conflict has occurred and developers need to resolve it, 14 developers interviewed said
+Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
 they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
@@
 code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time
-frames of when events occurred. Of the developers surveyed, 23\% said they used native IDE tools and 21\% said they use 
+frames of when events occurred. Of the questionnaire participants, 23\% said they used native IDE tools and 21\% said they use
 features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
@@
 
-When asked about preventative tools, interviewed developers had major concerns that the amount of false positives
+When asked about preventative tools, developers had major concerns that the amount of false positives
 provided by the tool which may render the tool useless. Developers said ``this would
@@
 
-In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
+In terms of catching indirect conflicts, developers suggested that proper software development processes
 are already in place to catch potential issues such as testing, code review,
-individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
-want to be notified about method signature changes as they have a high chance to break the code as opposed to only
-23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
-interface change. Other change types varied from never being notified to most times being notified, showcasing the
-complexity of change types which may or may not negatively affect a project. A complete breakdown of change types and
-developer preference can be found in Table~\ref{tab:pre}.
+individual knowledge, or static language analysis tools. Further, developers said that having strict change type monitoring
+for indirect conflicts does not work due to the inherent complexities of indirect conflicts.
+This can be seen confirmed in that questionnaire participants had very few cases in which they always wanted to be
+alerted to a change type as seen in Table~\ref{tab:pre}. Only method signature changes caused 68\% of questionnaire participants
+to want to be always notified as they have a high chance to break the code. This is opposed to other change types listed in
+Table~\ref{tab:pre} which have 27\% or lower of developers always wanting to be notified. This again showcases the
+complexity of indirect conflicts through change types which may or may not negatively affect a project.
 
-\begin{table*}[h]
+\begin{table*}[tb!]
 \begin{center}
@@
 \end{center}
-\caption{Results of survey questions to source code changes that developers deem notification worthy, in terms of percentage
-of developers surveyed.\label{tab:pre}}
+\caption{Questionnaire results about source code changes that developers deem notification worthy, in terms of percentage
+of questionnaire participants.\label{tab:pre}}
 \end{table*}
@@
 When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
-Surveyed developers suggested the following improvements to curative tools:
+Interview and questionnaire participants suggested the following improvements to curative tools:
 
@@
 
-10 of the 78 participants who were surveyed sent direct responses to me asking for any results of the research to be sent directly to
+10 of the 78 questionnaire participants sent direct responses to me asking for any results of the research to be sent directly to
 them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
@@
 made accurately demonstrate that complexity. Since interviewed participants came from such a diverse set of organizations, 
-software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and surveys, participants
+software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and questionnaires, participants
 expressed unique situations that did not fully meet my generalized findings or on going theories. In these cases, I worked in the specific
@@
 
-\section{Related Work}
-\label{sec:iced-related}
-
-Since this paper has covered a wide spectrum in regards to indirect conflicts (prevention, detection, resolution, process, and tools), there
-exists a large body of work in which to draw from regarding indirect conflicts. While some of the previous literature may not deal explicitly
-in the notion of indirect conflicts, lessons learned from topics in awareness, direct conflict detection, and debugging are presented.
-
-Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir, 
-which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts. 
-Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et 
-al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes, 
-such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause 
-indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives 
-detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT} 
-creates an environment where developers can see how source code changes will affect other areas of a project at the 
-method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has 
-been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
-needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS} 
-instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change. 
-Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload, 
-which were explored in this paper. Our own tool 
-Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
-
-Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
-times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
-developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}. 
-Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
-artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
-created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
-Socio-technical congruence however, is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
-
-For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
-recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
-Kim conducted several focus groups as well as web surveys to determine these results. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
-require action on the developer's half.  Kim
-also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
-were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
-noting ``I get a big laundry list... I see the email and I delete it''.
-
-In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
-preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
-development processes.
-
-In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
-ill fit of current static analysis to current development practices. Not being designed properly to fit current development work
-flow, and the over abundance of false positives were shown to be factors of developer non-use.
-
-Hattori et. al.~\cite{Hattori:2012:ICG} found, through qualitative user studies, that developers tend to use the bare amount of communication
-towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose. This
-resonates with ``I work until something breaks''. Hattori et. al. also show that with direct conflicts, the
-sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges.
-This is the situation many indirect conflict tools have strived for. In terms of communication, Bolici et. al~\cite{Bolici:2009} give the 
-possibility of developer stigmergy as the reason that developers do not need to explicitly communicate with each other about issues. The idea
-that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication.
-Blincoe et al.~\cite{Blincoe:2013:UCC} also found, through data mining techniques, a subset of events related to awareness 
-which occur that developers will communicate about as opposed to all awareness events. This again confirms our understanding that developers
-only care about certain events which require some action on their part.
-
-While not specifically developed with indirect conflicts in mind, Zeller's delta debugging techniques~\cite{Zeller:2005:WPF}, 
-can be applied to solving indirect conflict issues. Zeller gives automated techniques for identifying failure-inducing
-changes (as well as other failure causing components) of the software project in order to isolate and debug some issue. These practices fit
-with the idea that developers wait for something to go wrong, then wish to fix the issue as fast as possible. Program slicing~\cite{Xu:2005:BSP}
-has also been extensively studied in order to aid in debugging in manual or automated techniques.
+\subsection{Threats to Validity}
+% TODO
 
@@
 wish to handle indirect conflicts in the present in future. To achieve these results, I interviewed 19 developers from across 12 commercial 
-and open source projects, followed by a confirmatory survey of 78 developers, and 5 confirmatory interviews.
+and open source projects, followed by a confirmatory questionnaire of 78 developers, and 5 confirmatory interviews.
 
@@
 indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of change impacts,
-indirect conflicts occur quite frequently with a tendency to be more prevelant at the beginning of a development cycle, indirect conflicts
+indirect conflicts occur quite frequently with a tendency to be more prevalent at the beginning of a development cycle, indirect conflicts
 occur more often as a software project grows in developer numbers, developers use design by contract, add and deprecate, and experience
-to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occured rather
+to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occurred rather
 than tools which help prevent possible conflicts for the future.

### 2014-02-11 msc-thesis/chapters/4/chapter-discussion.tex
MSC Added changes to thesis.

@@
 While each study presented in this dissertation is quite unique, this chapter will focus on the underlying themes and
-results of all 4 studies. This chapter will address the usefulness of the results presented towards academia and
+results of all 4 studies. This chapter will address the usefulness of the results presented towards the
+research community and
 what those results mean for continued studies in future research as well as discuss how the results can be viewed
@@
 
-As we have seen through Study 1, developers which are tied to source code objects can become a focal point of
+As we have seen in Study 1, developers which are tied to source code objects can become a focal point of
 indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
@@
 shown that pairs of developers can often be a large component to that goal as well. (Impact from Study 2 was
-created the address just this issue.)
+created to use these pairs of developers in addressing the issue of indirect conflicts.)
 
 \textbf{Information Overload} \textit{My tool Impact, as well as many other indirect conflict tools, suffer from
-information overload in their delivery to developers which is a key issue in preventing adopting and acceptance
+information overload in their delivery to developers which is a key issue in preventing adoption and acceptance
 of indirect conflict tools.}
@@
 As was previously stated, many indirect conflict tools end up suffering from information overload to developers
-and other end users. Impact was initially created to take the knowledge learned from Study 1 and attempt to create
+and other end users. Impact was initially created to take the insights from Study 1 and attempt to create
 a new indirect conflict system based on developer interactions inside the code base which would potentially address
@@
 4 we will notice 3 main trends that have been discovered for indirect conflicts in both industry and in academia. These 3 major
-trends to now be discussed are: unwanted awareness, prevention versus cure, and the gap in software evolution analysis. This
+trends are: unwanted awareness, prevention versus cure, and the gap in software evolution analysis. This
 section will also include a brief discussion regarding the implications, learned from studies 3 and 4, for both academia and
@@
 
-\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
+\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this dissertation is the need for continued study of the open question
 ``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly 
@@
 \textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
-mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
+mismatch between awareness approaches and tools as developed in the research community, compared to awareness needs as perceived
 by industry. We should push our awareness theories to model real development
@@
 tools are used for indirect conflicts alone (automatic debugging~\cite{Zeller:2005:WPF}, unit tests, etc), most of these tools
-have need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
+have a need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
 automatically written for a given software object's contract, the ability to find a change  in an external project

### 2014-02-11 iced-2013/paper.tex
ICED danas last big edits.

--- 2014-01-29 iced-2013/paper.tex
@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{Sell Prevention and Buy Cure: An Exploration of Indirect Conflicts}
+\title{State of The Practice in Indirect Conflicts: Work Flows, Tools, and Future Direction}
 
@@
 to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
-overload, tools eventually being
+overload, which causes tools to eventually be
 ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
@@
 identification of failure inducing changes, other than those which are already identifiable by other means such
-as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
-also compounded by the scale of many software development projects leading to further information overload.
-Lastly, social factors such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
-congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
-However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
-come up.
+as compilers, and unit tests, to these dependencies still remains an issue.
+Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
+such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence to leverage developers in indirect conflict tools.
+However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
+problems as their evaluations have
+been small scale, involving a very limited number of industrial participants or student evaluators.
 
-In order to investigate the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, we conducted an empirical study to determine what events occur to
-cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
-We then set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, we look to find what can be accomplished moving forward with indirect conflicts
-in both research and industry and how we can get over the road block of these information overload related
-issues. To accomplish these goals, we ask the following:
+We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
+To do so, we take a step back and intend to obtain a broader view of indirect conflicts
+with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
+scalability issues in regards to indirect conflicts to better understand why indirect conflict tools fail to achieve
+the success of other domain tools. We determine: general events which 
+cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
+use, and what practitioners want from new tools.
+We ask:
 
 \begin{description}
-	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
+	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What are the types, factors, and frequencies of indirect conflicts?}
 	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
@@
 which were explored in this paper. Our own tool
-Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
+Impact!~\cite{Ell:2013} also suffered these same fates.
 
@@
 created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
-Socio-technical congruence however, is largely unproven in regards to its correlation to
-software quality~\cite{Kwan:2011:SCE} and again the problems of scalability and information overload become a factor.
+However, again the problems of scalability and information overload become a factor.
 
@@
 recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
-Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
-require action on the developer's half.  Kim
-also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
-were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
-noting ``I get a big laundry list... I see the email and I delete it''.
+Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same 
+issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
+Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
+developers which are not already detected by compilation or other static analysis tools. 
+For impact awareness, DeSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. However, DeSouza does not
+study which types of changes (types, frequencies, compounding factors) developers should concern themselves with more in terms of using their personal knowledge
+to stop indirect conflicts. DeSouza also does not study formal mitigation strategies, or resolutions of indirect conflicts directly.
 
-In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
-preventative measures we have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
-development processes.
+This paper is intended to fill the gap which has been left by aforementioned tools papers as well as the field study paper.
+We will not only study why information overload, false positives, and scalability are such difficult problems to tackle
+in indirect conflict tools, but we will also study how developers currently deal with indirect conflicts in practice through
+their mitigation strategies, their largest concerns, and their ideas for future suggestions in regards to indirect conflicts.
 
@@
 \begin{center}
-\begin{tabular}{| l | p{2cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
+\begin{tabular}{| l | p{1cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
 \hline
@@
 \addcontentsline{toc}{subsection}{RQ1}
-\subsection*{\normalfont{RQ1.} \it{What events or conditions lead to indirect conflicts?}}
+\subsection*{\normalfont{RQ1.} \it{What are the types, factors, and frequencies of indirect conflicts?}}
 
-12 developers believe that a large contributing factor to the cause
-of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
+\bf{The most common occurrence of indirect conflict is when a software object's contract changes. The frequency of
+indirect conflicts, while usually high, decreases as a stable point is reached in the development cycle. The frequency
+of indirect conflicts is compounded by the number of developers working on a project.}
+
+\normalfont{}
+
+A large contributing factor to the cause
+of indirect conflicts comes from the changing of a software object's contract said 12 developers. Object contracts are, in a sense,
 what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
@@
 their development life cycle with a minimum occurrence of once a week, with more serious issues tending
-to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur on minimum bi-weekly,
-with the 25\% saying that weekly occurrences are most common (seen in ~\ref{tab:often}). 5 questionnaire participants
+to occur once a month. The questionnaire showed that, 64\% of indirect conflicts occur bi-weekly
+or more frequent,
+with 25\% saying that weekly occurrences are most common (seen in Table~\ref{tab:often}). 5 questionnaire participants
 also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
@@
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
-changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release 
+changes a lot, especially with agile''. Questionnaire data also added ``indirect conflicts after a release
 depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
 major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
-large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results. 
-Questionnaire participants also answered that 
+large revamps or the implementation of cross-cutting new features.'' which corroborate our interview results.
+Questionnaire data also showed that
 indirect conflicts are more likely to occur before the first major release rather than after at the daily
@@
 
-In terms of organizational structure, questionnaire participants answered that as a project becomes larger and more 
+In terms of organizational structure, questionnaire data shows that as a project becomes larger and more
 developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
@@
 
+\bf{Three preventative techniques are used to mitigate indirect conflicts: design by contract, add and deprecate, and personal
+experience. For catching indirect conflicts, developers use: testing (unit and integration) with proper use case coverage,
+scheduled build processes, and static analysis tools built into IDEs. For resolving indirect conflicts, developers use: historical
+error and information logs, compiler and debugger tools, as well as static analysis tools built into IDEs.}
+
+\normalfont{}
+
 In terms of preventative processes used for indirect conflicts, 3 major components were found.
@@
 conflicts will occur while implementing.
-
-Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
-27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
-by contract or the testing of a methods contract, and 14\% said that add and
-deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.
-
-In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration) 
-as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
-tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
-when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project 
-themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
-time testing. The questionnaire participants confirmed these results with 49\% mentioning forms of testing as the major tool used to 
-catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
-Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.
-
-Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
-they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
-``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
-code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
-frames of when events occurred. Of the questionnaire participants, 23\% said they used native IDE tools and 21\% said they use 
-features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers 
-mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading 
-of formal documentation.
-
-%Very important
-Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
-most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
-approach. This means that while developers do have processes and tool for prevention, they would rather spend their time
-at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
-while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.
 
@@
 
+Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
+27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
+by contract or the testing of a methods contract, and 14\% said that add and
+deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.
+
+In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration)
+as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration
+tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
+when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
+not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project
+themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run
+time testing. The questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
+catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
+Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.
+
+Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
+they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
+``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
+code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time
+frames of when events occurred. Questionnaire data show that 23\% of developers use native IDE tools and 21\% said they use
+features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
+mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading
+of formal documentation.
+
+%Very important
+Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
+most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
+approach. This means that while developers do have processes and tool for prevention, they would rather spend their time
+at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
+while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.
+
 \phantomsection
@@
 
-When asked about preventative tools, interviewed developers had major concerns that the amount of false positives
+\bf{Being that developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts
+developers would rather see improvement to debugging processes. Developers are looking for automated debugging tools,
+and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}
+
+\normalfont{}
+
+When asked about preventative tools, developers had major concerns that the amount of false positives
 provided by the tool which may render the tool useless. Developers said ``this would
@@
 
-In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
+In terms of catching indirect conflicts, developers suggested that proper software development processes
 are already in place to catch potential issues such as testing, code review,
@@
 for indirect conflicts does not work due to the inherent complexities of indirect conflicts.
-This can be seen confirmed in that questionnaire participants had very few cases in which they always wanted to be
+This can be seen corroborated in that questionnaire participants had very few cases in which they always wanted to be
 alerted to a change type as seen in Table~\ref{tab:pre}. Only method signature changes caused 68\% of questionnaire participants
@@
 
+Having reported the aforementioned trends in industry (what, when, how, mitigation, and resolution), we now discuss these
+results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
+false positives, and scalability.
 As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
@@
 to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
-awareness about most, if not all actions.}
+awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
+false positives, as the changes being reported are not causing them to take action. These reported non actionable responses
+lead to information overload and false positives.}
 
@@
 arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
-to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.}
+to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.
+Focusing on cure is a direct response to the issue of scalability found in indirect conflict tools. Trying to prevent all conflicts 
+is too large a problem to handle, while curing existing conflicts is manageable by a developer.}
 
@@
 and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
-some software project configurations and software languages to be quite prone to indirect conflicts.}
+some software project configurations and software languages to be quite prone to indirect conflicts. This is yet another issue related
+to scalability of existing indirect conflict tools.}
 
@@
 
-\subsection{Implication for Research}
+\subsection{Implications for Future Research Work}
 \label{sec:implr}
@@
 
-\subsection{Implication for Tools}
+\subsection{Implications for Future Tool Work}
 \label{sec:implt}
@@
 
-We have shown through the findings of RQ1 why indirect conflicts occur, when indirect conflicts are more likely to
-occur, as well as what types of software objects are susceptible to these conflicts. Results of RQ2 have shown
-how developers in industry currently handle the prevention detection and solving of indirect conflicts. And Lastly, findings of RQ3
-have provided a foundation as to why past techniques and tools have had low adoption rates and where
-researchers should focus their current and future efforts in handling indirect conflicts.
+Our findings have indicated that:
+indirect conflicts occur frequently and are likely caused by software contract changes;
+while design by contract, add and deprecate, and personal experience help prevent indirect conflicts,
+developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, and developers want indirect conflict tools to focus on automatic debugging and better source code analysis.
 
-We have shown the disjoint of current awareness understanding versus the practical
-awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
-ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure has been presented along with
-the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
-to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, we have shown
-the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
-future industrial adoption of tools produced by academia for lack of fit in industrial needs.
-
-\section{Acknowledgments}
-
-We sincerely thank all participants who were willing to be interviewed or who participated
-in completing our questionnaire. We thank these people for taking time out of their day to participate and for
-sharing their developer experience with us.
+Our result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
+would rather focus on curing existing problems rather than preventing potential issues;
+and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.
 

### 2014-02-11 msc-thesis/chapters/1/chapter-intro.tex
MSC danas edits.

@@
 with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
-in the are of indirect conflicts. To accomplish this goal, I have researcher the following sub topics of indirect
+in the are of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
 conflicts: technical dependencies, how developers are involved in said dependencies,
@@
 what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
-are being used, what future steps should be taken in academia to better industry regarding indirect conflicts, and finally,
+are being used, what future steps should be taken by researchers to better industry regarding indirect conflicts, and finally,
 how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
@@
 Study 1 focuses on the power of technical dependencies in software projects. The question I
-investigated was : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
+investigated were : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
 code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
@@
 I also set out to understand what mitigation strategies developers currently use as opposed to those created
-by academia. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
+by researchers. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
 in both research and industry. This study asked the following 3 research questions: 
-\textit{What events or conditions lead to indirect conflicts?}
+\textit{What are the types, factors, and frequencies of indirect conflicts?}
 \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}

### 2014-02-11 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC danas edits.

@@
 to technical dependencies amongst stakeholders, to provide coordination recommendations.
+
+\subsection{Related Work}
+
+Research has shown multiple reasons for software failures in both technical dependencies as well
+as human or social dependencies in software development. On the technical side, studies have shown that
+technical dependencies in software are often powerful predictors or errors in software as well as
+in builds~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Hassan:4019574}. These technical dependencies
+are often accompanied by data mining algorithms in order to set apart failure inducing dependencies
+from non failing dependencies.
+
+On the other side with the human factor, researchers have examined predicting build outcomes of software
+using communication patterns from developers. Wolf et al.~\cite{5070503} used patterns of communication from between
+pre-existing build in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
+communication organization difference between developers who worked on the same software module had
+a negative influence on the quality of the software.
+
+Studies have also combined both technical and social dependencies into a notion of socio-technical congruence.
+Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that this notion of socio-technical congruence can be leveraged to predict
+and improve task completion times in a software projects.
+
+These studies have mostly focused at a very high and abstract level of software development (builds and task completion).
+Where they have fallen short is in fine grained analysis of software changes. This study is used to take the
+pre existing knowledge of both technical and social dependencies and apply it to a source code change level of
+granularity. Instead of large scale failures like long completion times or build failures, this study examines failures
+at the bug level induced by each code change.
 

### 2014-02-11 msc-thesis/chapters/2/sec-impact.tex
MSC danas edits.

@@
 network that relates developers to source code, issue repositories and
-other social media while Ariadne only looks at source code for developer
+other social media while Ariadne only examines there source code for developer
 to source code association. Through Codebook, developers become
@@
 This being the case, developer ownership must be placed on the 
-identified technical objects. With this ownership, we now infer
+identified technical objects. With this ownership, I now infer
 relationships among developers based on their technical objects
@@
 to be on the correct path. However, \textit{Impact} has clearly shown
-the achilles heel of indirect conflict tools, which is information
+the Achilles heel of indirect conflict tools, which is information
 overload because of an inability to detect ``notification worthy''
@@
 \subsection{Threats to Validity}
-% TODO
+Because of the tool validation nature of this work, I chose participant interviews as my research validation method, which has some implications regarding the
+limitations and threats to validity of this study. While I did some positive results regarding the potential of Impact in this study
+populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
+from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.
+
+As per my interview participants, I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
+may not have had enough real world experience to validate this study at an industry level. The students were also consumed with their regular course
+work which could have limited the time spend using Impact and the enthusiasm put forward while conducting this study.
+
+To counter act this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
+on the study or provide appropriate feedback where needed.
 

### 2014-02-11 msc-thesis/chapters/3/chapter-exploration.tex
MSC danas edits.

@@
 I then determine what mitigation strategies developers currently use as opposed to those created
-by academia. Since developers have identified indirect conflicts as a major concern for themselves, but at
-the same time are not using the tools put forth by academia, I wish to find what they use to mitigate indirect
+by researchers. Since developers have identified indirect conflicts as a major concern for themselves, but at
+the same time are not using the tools put forth by researchers, I wish to find what they use to mitigate indirect
 conflicts. Through these findings, we can create tools which are similar to those already in use by developers in
-the hopes of a higher adoption rate of tools produced by academia.
+the hopes of a higher adoption rate of tools produced by researchers.
 Finally, I examine what can be accomplished moving forward with indirect conflicts

### 2014-02-11 msc-thesis/chapters/3/sec-apie.tex
MSC danas edits.

@@
 I decided that by studying the change types surrounding these releases, I could get a better understanding of
-what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of our 10 project's 
+what types of source code changes or trends constitute software stability or maturity. In order to study the release points, I went to each of my 10 project's 
 web pages and looked through their release histories for major, minor, alpha, beta, and release candidate type releases. In total I identified 472 releases
@@
 This equation gives us a ratio of a particular change type happening before and after a particular release. If the ratio is above 1 then that particular change
-type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of our study, we set the
+type occurs more frequently after the release and if it is below 1 then it occurs more frequently before the release. For the purposes of my study, I set the
 number of days before and after the release ($d$) to 60 as the projects studies had many months in between their major releases. This quantitative data
@@
 Lastly, I found that software changes related to testing can be an indicator of a major release points within the projects studied. The change
-ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet our threshold and are indicators of stability with regards to test based
+ratios found can be seen in Table~\ref{tab:test}. As it can be seen, the four ratios which meet my threshold and are indicators of stability with regards to test based
 changes are: the changing of test classes, the removal of test classes, the adding of methods, and the changing of method signatures, and test classes being changed.
@@
 changed test method signatures, changed documentation, and removed test classes.
-The 5 change trends found in this paper occurring after major release are: added test methods, changed test classes, removed public methods, removed
+The 5 change trends found in this study occurring after major release are: added test methods, changed test classes, removed public methods, removed
 private classes, and removed private methods.

### 2014-02-11 msc-thesis/chapters/3/sec-iced.tex
MSC danas edits.

@@
 
-In order to fully understand the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts, I conducted a large scale empirical study involving industrial
-participants in order to determine: what events occur to
-cause indirect conflicts, when they occur, if compounding conditions exist, and what mitigation strategies developers currently
-use. Through this large exploration, what can be accomplished moving forward with indirect conflicts
-in both research and industry and how we can get over the road block of these information overload related
-issues. My research questions for this particular study are as follows:
+I want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
+To do so, I take a step back and intend to obtain a broader view of indirect conflicts
+with a large field study of practitioners. I investigate the root causes of information overload, false positives, and
+scalability issues in regards to indirect conflicts to better understand why indirect conflict tools fail to achieve
+the success of other domain tools. I determine: general events which 
+cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
+use, and what practitioners want from new tools. My research questions for this particular study are as follows:
 
 \begin{description}
-        \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What events or conditions lead to indirect conflicts?}
+        \item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What are the types, factors, and frequencies of indirect conflicts?}
         \item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
@@
 Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
-which were explored in this paper. Our own tool
-Impact!~\footnote{University of Victoria Computer Science Technical Paper: DCS-351-IR} also suffered these same fates.
+which were explored in this study. My own tool
+Impact!~\cite{Ell:2013} also suffered these same fates.
 
@@
 recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
-Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
-found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
-require action on the developer's half.  Kim
-also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
-were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
-noting ``I get a big laundry list... I see the email and I delete it''.
+Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same 
+issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
+Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
+developers which are not already detected by compilation or other static analysis tools. 
+For impact awareness, DeSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. However, DeSouza does not
+study which types of changes (types, frequencies, compounding factors) developers should concern themselves with more in terms of using their personal knowledge
+to stop indirect conflicts. DeSouza also does not study formal mitigation strategies, or resolutions of indirect conflicts directly.
 
-In a case study of impact awareness, de Souza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. This corresponds with the findings of
-preventative measures I have found in that some human interaction or knowledge is required in preventing indirect conflicts along with
-development processes.
+This study is intended to fill the gap which has been left by aforementioned tools papers as well as the field study paper.
+I will not only study why information overload, false positives, and scalability are such difficult problems to tackle
+in indirect conflict tools, but I will also study how developers currently deal with indirect conflicts in practice through
+their mitigation strategies, their largest concerns, and their ideas for future suggestions in regards to indirect conflicts.
 
@@
 
-Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding our three research questions.
-Specifically, ten semi-structured  topics directly related to our research questions guided our interview:
+Interview participants first answered a number of demographic questions. I then asked them to describe various software development experiences regarding my three research questions.
+Specifically, ten semi-structured topics directly related to my research questions guided my interview:
 
@@
 
-My questionnaire respondents were different from my interviewees. I invited our questionnaire participants from a similar breadth of open and closed source software development 
+My questionnaire respondents were different from my interviewees. I invited my questionnaire participants from a similar breadth of open and closed source software development 
 companies and projects as the interviews participants with two main exceptions. The software organizations
@@
 
-\subsubsection{What events or conditions lead to indirect conflicts?}
+\subsubsection{What are the types, factors, and frequencies of indirect conflicts?}
+
+\bf{The most common occurrence of indirect conflict is when a software object's contract changes. The frequency of
+indirect conflicts, while usually high, decreases as a stable point is reached in the development cycle. The frequency
+of indirect conflicts is compounded by the number of developers working on a project.}
+
+\normalfont{}
 
@@
 their development life cycle with a minimum occurrence of once a week, with more serious issues tending
-to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur on minimum bi-weekly,
-with the 25\% saying that weekly occurrences are most common (seen in ~\ref{tab:often}). 5 questionnaire participants
+to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur bi-weekly or more
+frequent,
+with the 25\% saying that weekly occurrences are most common (seen in Table~\ref{tab:often}). 5 questionnaire participants
 also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
@@
 major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
-large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results.
+large revamps or the implementation of cross-cutting new features.'' in order to confirm mu interview results.
 Questionnaire participants also answered that
@@
 \subsubsection{What mitigation techniques are used by developers in regards to indirect conflicts?}
+
+\bf{Three preventative techniques are used to mitigate indirect conflicts: design by contract, add and deprecate, and personal
+experience. For catching indirect conflicts, developers use: testing (unit and integration) with proper use case coverage,
+scheduled build processes, and static analysis tools built into IDEs. For resolving indirect conflicts, developers use: historical
+error and information logs, compiler and debugger tools, as well as static analysis tools built into IDEs.}
+
+\normalfont{}
 
@@
 \subsubsection{What do developers want from future indirect conflict tools?}
+
+\bf{Being that developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts
+developers would rather see improvement to debugging processes. Developers are looking for automated debugging tools,
+and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}
+
+\normalfont{}
 
@@
 From the re-interviewed participants, I found extremely positive feedback regarding both my results and major discussion
-points. Participants often had new stories and experiences to share once they had heard the results of this paper which
+points. Participants often had new stories and experiences to share once they had heard the results of this study which
 confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
@@
 issues, I believe that these findings will help researchers focus on what developers want and need moving into the future more than has
-been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.
+been possible in the past. These finding set a course of action for where effort should be spent in research to better benefit industry.
 
@@
 them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
-tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest 
+tools or plans for tools inspired by this research as well. The combination of research relatability and direct industry interest 
 in my results help us fulfill this criterion.
@@
 \subsection{Threats to Validity}
-% TODO
+Because of the exploratory nature of this work, I chose Grounded Theory as my research method, which has some implications regarding the
+limitations and threats to validity of this study. While I achieved high saturation regarding the topics focused on in this study, other
+populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
+from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.
+
+As per my interview participants, I attempted to cover as great a breadth as possible of software developers in their practices, companies,
+project, languages, etc. However, it would be extremely difficult to conduct a completely thorough study of indirect conflicts in practice
+as there exist too many demographic options for participants. This same applies for questionnaire participants. This being said, questionnaire 
+or interview results could become heavily biased towards a developers good or bad experiences with any given project.
+
+To counter act this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
+I have uncovered valuable insight regarding indirect conflicts in practice.
 
@@
 
-My findings indicate that: 
-indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of change impacts,
-indirect conflicts occur quite frequently with a tendency to be more prevalent at the beginning of a development cycle, indirect conflicts
-occur more often as a software project grows in developer numbers, developers use design by contract, add and deprecate, and experience
-to mitigate indirect conflicts, and finally, developer prefer to have tools which help them debug quicker once a conflict has occurred rather
-than tools which help prevent possible conflicts for the future.
+My findings have indicated that:
+indirect conflicts occur frequently and are likely caused by software contract changes;
+while design by contract, add and deprecate, and personal experience help prevent indirect conflicts,
+developers tend to prefer to use detection and resolution processes or tools
+over those of prevention, and developers want indirect conflict tools to focus on automatic debugging and better source code analysis.
+
+My result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
+would rather focus on curing existing problems rather than preventing potential issues;
+and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.

### 2014-02-11 msc-thesis/chapters/4/chapter-discussion.tex
MSC danas edits.

@@
 learned from each as background information for the richer following studies. The second section will address the 2 large
-studies found in Chapter 3 in a combined manner. Since Study 4 was directly associated with developer opinion found in Study
+studies found in Chapter 3 in a combined manner. Having reported the trends in industry (what, when, how, mitigation, and resolution)
+from Study 3, the Study 3 discussion will focus these
+results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
+false positives, and scalability. Since Study 4 was directly associated with developer opinion found in Study
 3, the two discussions will be intertwined to better support each other.
@@
 better what can be done to improve developer work flow, this section will discuss exactly those steps back. From Studies 3 and
-4 we will notice 3 main trends that have been discovered for indirect conflicts in both industry and in academia. These 3 major
+4 we will notice 3 main trends that have been discovered for indirect conflicts in both industry and in research. These 3 major
 trends are: unwanted awareness, prevention versus cure, and the gap in software evolution analysis. This
-section will also include a brief discussion regarding the implications, learned from studies 3 and 4, for both academia and
+section will also include a brief discussion regarding the implications, learned from studies 3 and 4, for both research and
 for industry through tools.
@@
 Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
-their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
+their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this study that
 developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers

### 2014-02-11 msc-thesis/chapters/5/chapter-conclusion.tex
MSC danas edits.

@@
 component failures to understand what developers truly want from indirect conflict tools.As a proposed solution to information overload, 
-the ideas of Meyer~\cite{Meyer:1988} on ``design by contract'' were presented. This methodology looks at using the 
+the ideas of Meyer~\cite{Meyer:1988} on ``design by contract'' were presented. This methodology examines the 
 post and pre conditions of software objects in order to reduce the number of source code changes that are analyzed by 
@@
 the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
-future industrial adoption of tools produced by academia for lack of fit in industrial needs. This gap directly resulted in the final
+future industrial adoption of tools produced by researchers for lack of fit in industrial needs. This gap directly resulted in the final
 study of this Dissertation to be completed (Study 4).
@@
 the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
-future industrial adoption of tools produced by academia for lack of fit in industrial needs. While, I have shown a potential solution to address
+future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While, I have shown a potential solution to address
 a sub problem of the analysis gap through contextual patterns of indirect conflicts, many sub problems of the analysis gap remian problems

### 2014-02-11 msc-thesis/chapters/4/chapter-discussion.tex
MSC transfering changes from iced to msc.

@@
 and analysis unfriendly languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
-some software project configurations and software languages to be quite prone to indirect conflicts.}
+some software project configurations and software languages to be quite prone to indirect conflicts. This is yet another issue related
+to scalability of existing indirect conflict tools.}
 

### 2014-03-10 iced-2013/paper.tex
ICED bird's edits.

@@
 
-As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel
+As Software Configuration Management (SCM) has grown, the maturity and norm of parallel
 development has become the standard development process instead of the exception. With this parallel development
@@
 location in the software system such as when libraries are upgraded or when a method is changed
-and invoking methods are influence negatively. Indirect conflict tools however, have
+and invoking methods are influenced negatively. Indirect conflict tools however, have
 not shared the same success as direct conflict
@@
 Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
-such as Cataldo et al.~\cite{Cataldo:2006:ICR} notion of socio-technical
-congruence to leverage developers in indirect conflict tools.
+such as Cataldo et al.'s~\cite{Cataldo:2006:ICR} notion of socio-technical
+congruence.
 However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
 problems as their evaluations have
-been small scale, involving a very limited number of industrial participants or student evaluators.
+been small scale, involving a limited number of industrial participants or student evaluators.
 
 We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
-To do so, we take a step back and intend to obtain a broader view of indirect conflicts
+To do so, step back and intend to obtain a broader view of indirect conflicts
 with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
-scalability issues in regards to indirect conflicts to better understand why indirect conflict tools fail to achieve
+scalability issues to better understand why indirect conflict tools fail to achieve
 the success of other domain tools. We determine: general events which 
@@
 developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that:
-indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
-developers tend to prefer to use detection and resolution processes or tools
-over those of prevention, developers do not want awareness mechanisms which provide non actionable results,
-and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
-context of indirect conflicts.
+\begin{itemize}
+	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding
+	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
+	\item developers do not want awareness mechanisms which provide non actionable results
+	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
+				context of indirect conflicts
+\end{itemize}
 
@@
 either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
-being apart of any one software project for extended periods of time.
+being part of any one software project for extended periods of time.
 
@@
 questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
-both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding.
+both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding (relating
+code and concepts to each other).
 In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
@@
 are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
-object contacts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
+object contracts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
 Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
@@
 when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
-not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project
-themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run
-time testing. The questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
+not all indirect conflicts should be caught. In corroboration, 
+questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
 catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
@@
 features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
-mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading
+mentioned a form of communication with other developers in aid to resolving these conflicts and only 4\% mentioned the reading
 of formal documentation.
@@
 most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
-approach. This means that while developers do have processes and tool for prevention, they would rather spend their time
+approach. This means that while developers do have processes and tools for prevention, they would rather spend their time
 at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
@@
 
-Having reported the aforementioned trends in industry (what, when, how, mitigation, and resolution), we now discuss these
+We now discuss the aforementioned
 results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
@@
 
-\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
+\subsection{Unwanted Awareness}
+
+\textit{Developers tend to only care about the awareness of other's activities, if it causes
 a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
@@
 awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
-false positives, as the changes being reported are not causing them to take action. These reported non actionable responses
-lead to information overload and false positives.}
+false positives, as the changes being reported are not causing them to take action.}
 
@@
 leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore
-information being presented to them, rendering the system useless. These false positives are caused by a difference in what
-developers consider to be false positives versus what awareness literature considers they are. This disjoint, as will
+information being presented to them, rendering the system useless.
+
+\begin{itemize}
+	\item These false positives are caused by a difference in what
+				developers consider to be false positives versus what awareness literature considers they are.
+\end{itemize}
+
+This disjoint, as will
 be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
@@
 their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
-developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
-don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue''.
+developers only want awareness of an event if the event forces the developer to take some action.
+
+\begin{itemize}
+	\item Developers don't care about what they don't need to account for
+\end{itemize}
+
+``We would want a high probability of the [reported] change being an issue''.
 This sense of unwanted or limited awareness is crucial to understand why generalized
@@
 towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
-Bolici et. al~\cite{Bolici:2009} also give the
-possibility of developer stigmergy as the reason that developers do not explicitly communicate with each other about issues opposed to
-time management. The idea
-that developers use artifacts left behind by other developers to solve their issues could play a large role in a lack of formal communication
-and unwanted awareness.
 \textit{Developers prefer to spend their limited time dealing with
@@
 producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
-however, when given the choice, developers tend to, as previously stated, lean towards less awareness in order to, in their eyes,
-be more productive.
+however, when given the choice, developers tend to lean towards less awareness in order to be more productive in their eyes.
 
-\textbf{Prevention versus Cure}  \textit{Developers would rather spend their time on curative measures (fixing problems after they
-arise) rather than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
+\subsection{Prevention versus Cure}
+
+\textit{Developers would rather spend their time on curative measures (fixing problems after they
+arise) than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
 to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.
@@
 
-We have seen through the results, that developers posses both tools and practices for the prevention,
-detection, and resolution of indirect conflicts. We have also seen that through unwanted awareness, and the use of developer time,
-that developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
-may never arise. This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
+We have seen through the results, that developers possess both tools and practices for the prevention,
+detection, and resolution of indirect conflicts.
+
+\begin{itemize}
+	\item Through unwanted awareness, and the use of developer time,
+				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
+				may never arise.
+\end{itemize}
+
+This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
 is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
-Prevention here refers to taking precautionary steps to stop issues, indirect conflicts, form occurring in the first
-place while cure refers to fixing issues as they arise which includes not attempting, or putting little attempt, into preventing
+Prevention here refers to taking precautionary steps to stop potential or some probability of issues, or indirect conflicts, form occurring in the first
+place, while cure refers to fixing non probable (have happened) issues as they arise which includes not attempting, or putting little attempt, into preventing
 them in the first place.
@@
 not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
-developers are spending little to no time in prevention. Developers do however spend a large amount of time in detection and cure through the
-writing of tests and the debugging of issues. In fact, most improvements mentioned by developers in regard to dealing with indirect
+developers are spending little to no time in prevention (or what the perceive as prevention). Developers do however spend a 
+large amount of time in detection and cure through the
+writing of tests and the debugging of issues (although writing tests can be viewed as prevention while their utility is only see at detection). 
+In fact, most improvements mentioned by developers in regard to dealing with indirect
 conflicts occurred at either the detection or cure levels. ``You're reaction time is much more crucial''
@@
 
-The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
+The lack of prevention process and tools being used is believed to be due to 2 factors as discovered in the interview process. 
+The first being the identification of dependencies.
 Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
@@
 it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
-not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
-to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
+not be thinking of how to move developers into a more preventative mindset. If the inverse is true, should we not be putting more time into automatic debugging, such as
 Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.
@@
 thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
-However, if a test never fails, could it not be said that it is preventing changes from causing it to fail rather than detecting
-when failures or indirect conflicts do occur? This may suggest that if, given the write tools, developers may no longer view preventative
+This may suggest that if, given the right tools, developers may no longer view preventative
 measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
@@
 prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
-approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
-is being viewed in research.
+approaches to their developers.
 
 %mature point, static analysis
-\textbf{Gap in Software Evolution Analysis} \textit{Due to a lack of productive software analysis tools, caused by project infrastructures
+\subsection{Gap in Software Evolution Analysis}
+
+\textit{Due to a lack of productive software analysis tools, caused by project infrastructures
 and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
@@
 
-This added level of context, through the notion of a mature point,
+This added level of context,
 would add to the prevention versus cure debate as previously discussed. Prevention may be a better
@@
 to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
-causes more bugs than can be prevented. These possibilities existing, it may be imperative to discover more of project stability
+causes more bugs than can be prevented. It may be imperative to discover more on project stability
 and the mature point in order to fully understand the nature of indirect conflicts and their context.
 
-While this added sense of understanding is important to indirect conflicts, contextual attributes of a project's current progress
-or measures of performance are often more synonymous with project management
-rather than software developers themselves. An understanding of a project's evolution, a mature point being reached as an example,
+While this added sense of understanding is important to indirect conflicts, an understanding of a project's evolution, 
+a mature point being reached as an example,
 may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
@@
 organizational structures of software teams. Between increased modularization (multiple sub projects or repositories),
-cross language dependencies, and languages which do not lend themselves to static analysis, static analysis tools have
+cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
 become quite limited in identifying and tracking dependencies where they were once strong.
@@
 
-\textbf{Prevention versus Cure} \textit{The largest implication for future research found in this paper is the need for continued study of the open question
+\textbf{Prevention versus Cure} \textit{The need for continued study of the open question
 ``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly
@@
 
-\textbf{Give Developers What They Want} \textit{Developers have their own notions of what they want, and how ``good''
+\textbf{Give Developers What They Want} \textit{Developers have their own notions of ``good''
 that something is for their productivity. We should pay attention to what they want, but also be mindful that what
@@
 \textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
-multiple languages, and often having a database schema, we should focus on better techniques for static analysis based
-on what industry standards are for languages and project structures.}
+multiple languages, and more modular structures, we should focus on better techniques for static analysis based
+on what industry standards.}
 
@@
 
-\section{Evaluation}
+\section{Evaluation and Threats to Validity}
 \label{sec:eval}
@@
 As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.
-We have chosen three of these criteria and explain how we fulfill them.
+We have chosen three of these criteria and explain how we fulfill them. The three evaluation criteria selected are more relevant
+to a qualitative study than the usual threats to validity associated with qualitative work. These selected criteria can also
+be viewed as the threats to validity of this paper.
 
@@
 
-10 of the 78 questionnaire participants who were sent direct responses to us asking for any results of the research to be sent directly to
+10 of the 78 questionnaire participants sent direct responses to us asking for any results of the research to be sent directly to
 them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible

### 2014-03-10 iced-2013/paper.tex
ICED quick edits.

@@
 
-\begin{itemize}
-	\item These false positives are caused by a difference in what
-				developers consider to be false positives versus what awareness literature considers they are.
-\end{itemize}
+\textbf{These false positives are caused by a difference in what
+developers consider to be false positives versus what awareness literature considers they are.}
 
@@
 
-\begin{itemize}
-	\item Developers don't care about what they don't need to account for
-\end{itemize}
+\textbf{Developers don't care about what they don't need to account for}
 
@@
 towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
-\textit{Developers prefer to spend their limited time dealing with
+\textbf{Developers prefer to spend their limited time dealing with
 the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
@@
 
-\begin{itemize}
-	\item Through unwanted awareness, and the use of developer time,
+\textbf{Through unwanted awareness, and the use of developer time,
 				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
-				may never arise.
-\end{itemize}
+				may never arise.}
 
@@
 which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
-are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
-abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
+are what ultimately have led to the amount of false positives and information overload seen in previous tools. \textbf{The
+abundance of detection and curative process and tools shows once again developers willingness to debug
 real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
-presented in this area is if curative measures are more productive than preventative measures.
+presented in this area is if curative measures are more productive than preventative measures.}
 
@@
 
-A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
-research than may have been previously thought. Static analysis may be useful in regards to indirect conflict context in that we
+\textbf{A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
+research than may have been previously thought.} Static analysis may be useful in regards to indirect conflict context in that we
 could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
@@
 indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
-organizational structures of software teams. Between increased modularization (multiple sub projects or repositories),
+organizational structures of software teams. \textbf{Between increased modularization (multiple sub projects or repositories),
 cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
-become quite limited in identifying and tracking dependencies where they were once strong.
+become quite limited in identifying and tracking dependencies where they were once strong.}
 In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the

### 2014-03-16 iced-2013/paper.tex
ICED final commit.

@@
 % Can use linebreaks \\ within to get better formatting as desired
-\title{State of The Practice in Indirect Conflicts: Work Flows, Tools, and Future Direction}
+\title{State of The Practice in Indirect Conflicts: Causes, Mitigation Strategies, and Future Directions}
 
@@
 mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
-empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
-and time-consuming chore for developers~\cite{Perry:2001:PCL}.
+empirical evidence shows that despite of recent research has proposed mitigation strategies, these losses continue to appear quite frequently and can prove to be a significant
+and time-consuming chore for developers~\cite{Perry:2001:PCL, Sarma:2009:TIV}.
 
@@
 overload, which causes tools to eventually be
-ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
+ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of code dependency identification and
 tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
-invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the
+invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success; however, the
 identification of failure inducing changes, other than those which are already identifiable by other means such
-as compilers, and unit tests, to these dependencies still remains an issue.
-Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also incorporated social factors 
-such as Cataldo et al.'s~\cite{Cataldo:2006:ICR} notion of socio-technical
-congruence.
-However, the same aforementioned problems still arise. These mentioned tools have not been able to study their own 
-problems as their evaluations have
-been small scale, involving a limited number of industrial participants or student evaluators.
+as compilers or unit tests, to these dependencies still remains an issue.
+Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also leveraged the notion of socio-technical congruence ~\cite{Cataldo:2006:ICR} to help developers pursue dependencies through social means.
+However, the same aforementioned problems still arise.
+Evaluations of these tools or approaches have largely been small in scale, and involved a limited number of industrial participants or student evaluators.
+As a result, the problems of scalability, information overload and false positives in tackling indirect conflicts remains an outstanding research problem.
 
-We want to understand why it is so hard to tackle indirect conflicts, specifically through tool based solutions.
-To do so, step back and intend to obtain a broader view of indirect conflicts
-with a large field study of practitioners. We investigate the root causes of information overload, false positives, and
-scalability issues to better understand why indirect conflict tools fail to achieve
-the success of other domain tools. We determine: general events which 
-cause indirect conflicts, when they occur, if compounding conditions exist, mitigation strategies developers
-use, and what practitioners want from new tools.
-We ask:
+In this paper we report from a field study of software practitioners that sought to characterize the state-of-the-practice in indirect conflicts, particularly their causes and mitigation strategies.
+Our intention has been to uncover the root causes of information overload, false positives, and
+scalability issues, in order to better understand why indirect conflict tools and methods fail to fully achieve
+their intended success. Three research questions that guided our study:
 
@@
 
-We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78
-developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that:
+To answer these questions, we interviewed 19 developers from across 12 commercial and open source projects, sought confirmation of the themes we identified in our interviews through a questionnaire with 78
+developers, as well as 5 follow-up interviews. Our findings indicate that:
 \begin{itemize}
-	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding
+	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of such changes
 	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
-	\item developers do not want awareness mechanisms which provide non actionable results
+	\item Developers do not want awareness mechanisms which provide non actionable results
 	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
@@
 
-Many indirect conflict tools have been created, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
+Many indirect conflict tools have been developed, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
 which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
@@
 instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
-Most of these tools have all shown to have the same common difficulties of scalability, false positives, and information overload,
-which were explored in this paper. Our own tool
+The evaluations of these tools indicate that they have the same common difficulties of scalability, false positives, and information overload. Our own tool
 Impact!~\cite{Ell:2013} also suffered these same fates.
@@
 developers which are not already detected by compilation or other static analysis tools.
-For impact awareness, DeSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
-code base to determine the impact of their code changes on fellow developers, teams, and projects. However, DeSouza does not
-study which types of changes (types, frequencies, compounding factors) developers should concern themselves with more in terms of using their personal knowledge
-to stop indirect conflicts. DeSouza also does not study formal mitigation strategies, or resolutions of indirect conflicts directly.
+In studying awareness of change impact, deSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
+code base to determine the impact of their code changes on fellow developers, teams, and projects. However, their work does not
+study which types of changes (types, frequencies, compounding factors) developers encounter in their work, nor provides concrete information on formal mitigation strategies or resolutions of indirect conflicts.
 
-This paper is intended to fill the gap which has been left by aforementioned tools papers as well as the field study paper.
-We will not only study why information overload, false positives, and scalability are such difficult problems to tackle
-in indirect conflict tools, but we will also study how developers currently deal with indirect conflicts in practice through
-their mitigation strategies, their largest concerns, and their ideas for future suggestions in regards to indirect conflicts.
+The insights of our study intends to fill the gap. In our field study we seek to understand not only  why information overload, false positives, and scalability are such difficult problems to tackle
+in indirect conflict tools, but also how developers currently deal with indirect conflicts in practice through
+their mitigation strategies, what are their major concerns as well as suggestions for future research and tools to tackle these concerns.
 
@@
 
-Our interview participants came from a large breadth of both open and closed source software development companies
-and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
-Our participants were invited based on their direct involvement in the actual writing of software for
-their respective companies or projects. These participants' software development experience had an average of 8 years of experience.
-In addition to software development, some participants were also invited based on their experience with project management
-at some capacity.
+Our interview participants came from a wide range of both open and closed source software development companies
+and projects, using both  agile and waterfall based methodologies, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
+The participants had software development experience in average of 8 years, as well as some project management experience.
 
@@
 
-Participants were invited to be interviewed by email and were sent a single reminder email one week
+Participants were invited based on their software engineering and project management experience.
+We recruited participants through an email message to a number of software organizations in our working business network, and followed up in a single reminder email one week
 after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
@@
 
-Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences regarding our three research questions.
+Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences and practices.
 Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:
@@
 
-Our questionnaire respondents were different from our interviewees. We invited our questionnaire participants from a similar
+Our questionnaire respondents were different from our interviewees. We recruited the respondents to our questionnaire from a similar
 breadth of open and closed source software development
-companies and projects as the interviews participants with two main exceptions, the addition of GitHub users as well as Apache Software Foundation (Apache) developers.
+companies and projects as the interviews participants with two main exceptions.
+The GitHub users were recruited from the Google BigQuery database; the Apache respondents were recruited from the list of tops contributors to the Apache Software Foundation.
 The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation,
@@
 
-Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and ended with 78 responses
+Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and obtained 78 responses
 giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
@@
 
-The questionnaire we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
-was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews.
+The questionnaire we designed ~\footnote{http://thesegalgroup.org/projects/indirect-conflicts/}
+was based on the insights we obtained from the round of interviews, and was intended to confirm these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews.
 The question topics asked in the questionnaire were:
@@
 Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
-\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
-theory techniques for analysis and not for a final theory). For an exploratory study such as
-the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
+\textit{open coding} to create a theory ``grounded'' in the empirical data. In this study we used grounded
+theory techniques for analysis and not to develop a theory per say.
+For an exploratory study such as ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
 making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
@@
 findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
-list of criteria to evaluate quality and credibility. This evaluation can be seen in Section~\ref{sec:eval}
+list of criteria to evaluate quality and credibility, described in Section~\ref{sec:eval}.
 
@@
 
-We now present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
-outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
-results from which we draw our discussion to be seen in Section~\ref{sec:disc}. In each subsection, quantitative data
+In this section we present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
+outlined in Section~\ref{sec:intro}. For each research question we summarize the quantitative and qualitative
+evidence we obtained. In each subsection, quantitative data
 given refers to interviews conducted unless explicitly said otherwise.
@@
 
-A large contributing factor to the cause
-of indirect conflicts comes from the changing of a software object's contract said 12 developers. Object contracts are, in a sense,
-what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
-to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark.
-In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
+We found that a large contributing factor to the cause
+of indirect conflicts comes from the changing of a software object's contract (reported by 12 developers directly).
+Object contracts, or ``Design by Contract''\texttrademark, a concept made well-known through the work on Eiffel Software's~\footnote{http://www.eiffel.com/} are, in a sense,
+what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed.
+In our interviews, 14 developers gave examples of indirect conflicts they had experienced
 which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
@@
 7 developers mentioned that indirect conflicts occur when a major update to an external project,
-library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
+library, or service occurs with one developer noting that ``their build never breaks, but it breaks ours''. Some
 other notable indirect conflict artifacts were user interfaces in web development and full components in component
@@
 
-12 developers said that
+12 developers reported that
 when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
-than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
+than once a stable point is reached. For example, ``At a stable point we decided we are not going to change
 [this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
-changes a lot, especially with agile''. Questionnaire data also added ``indirect conflicts after a release
+changes a lot, especially with agile''. From the survey respondents we also learned that ``indirect conflicts after a release
 depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
@@
 developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
-occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
-saying they are like to occur in single developer projects. This can be seen in Table~\ref{tab:env}.
+occur. However, as shown in Table~\ref{tab:env}, indirect conflicts still occur in small teams, with  43\% of respondents indicating
+ they are likely to occur in single developer projects.
 
@@
 
-In terms of preventative processes used for indirect conflicts, 3 major components were found.
-First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or understand
+Our analysis indicates indicates three major preventative processes used for indirect conflicts.
+First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or to understand
 when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
@@
 
-\bf{Being that developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts
-developers would rather see improvement to debugging processes. Developers are looking for automated debugging tools,
+\bf{Because developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts, we identified a strong preference for improvements to debugging processes.
+Developers prefer automated debugging tools
 and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}
@@
 
-We now discuss the aforementioned
-results in relation to the outstanding issues of tool based solutions regarding indirect conflicts which are information overload,
+We now discuss these results in relation to the outstanding issues of tool-based solutions regarding indirect conflicts and which are information overload,
 false positives, and scalability.
-As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
-conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
-fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
+As per our findings, three main discussion points emerge: (1) What developers really want and don't want from indirect
+conflict awareness tools, (2) Whether indirect conflicts are better handled by extensively trying to prevent them or by simply
+fixing them as they occur with less prevention, and (3) A gap in analytical tools for software dependencies found in current
 industry settings.
@@
 
-As we have seen, indirect conflicts are found to be quite a serious problem that occur frequently,
+According to our study participants, indirect conflicts represent a serious problem that occur frequently,
 sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
@@
 to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
-leading cause of information overload and false positives in the awareness system, causing developers to eventually ignore
+leading cause of information overload and false positives  awareness systems, causing developers to eventually ignore
 information being presented to them, rendering the system useless.
@@
 \textbf{These false positives are caused by a difference in what
-developers consider to be false positives versus what awareness literature considers they are.}
-
-This disjoint, as will
-be seen, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
+developers consider to be false positives versus what awareness literature considers they are.} This disjoint, we argue, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
 where as current awareness understanding would state any event which is related to an individual's
@@
 
-\textbf{Developers don't care about what they don't need to account for}
-
-``We would want a high probability of the [reported] change being an issue''.
+\textbf{Developers don't care about what they don't need to account for}. ``We would want a high probability of the [reported] change being an issue''.
 This sense of unwanted or limited awareness is crucial to understand why generalized
@@
 
-We have seen through the results, that developers possess both tools and practices for the prevention,
+Our findings suggest that developers possess both tools and practices for the prevention,
 detection, and resolution of indirect conflicts.
@@
 				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
-				may never arise.}
-
-This is neatly explained by the popular adage ``I work until something breaks'' taken by most developers. This mindset
+				may never arise.} Related to the popular adage ``I work until something breaks'' taken by most developers, this mindset
 is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
@@
 
-\textbf{Prevention versus Cure} \textit{The need for continued study of the open question
-``Which of prevention or cure is more effective for software development and indirect conflicts?''. This simple question will undoubtedly
-produce extremely complex answers.}
+\textbf{Prevention versus Cure} \textit{There is a need for continued study of the open question
+``Which of prevention or cure is more effective for software development and indirect conflicts?''. We believe that this simple question will undoubtedly
+produce extremely complex answers. Developers are currently favoring cure, but we should be mindful that what developers want
+may not always be the best choice.}
 
-With software developers focusing their current efforts on curative measures, for
+The fact that software developers focus their current efforts on curative measures, for
 indirect conflicts, suggests
@@
 
-\textbf{Awareness Theory to Model Real World} \textit{the need to further characterize the
+\textbf{Theories of Awareness to Model Real World} \textit{There is a need to further characterize the
 mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
-by industry. We should push our awareness theories to model real development
+by industry. Theories of awareness should consider modeling real development
 environments where interruptions, time management, and full development life cycles are often large factors.}
@@
 
-\textbf{Give Developers What They Want} \textit{Developers have their own notions of ``good''
-that something is for their productivity. We should pay attention to what they want, but also be mindful that what
-developers want may not always be good for them.}
+\textbf{Give Developers What They Want} \textit{Developers have their own notions of how ``good''
+something is for their productivity. We should pay attention to what they want, in order to shape our
+models and better address their needs.}
 
-The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
+The direct implication this research has on tool development is that of current adoption among developers. As was stated,
 developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being
@@
 after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
-on and potentially agree with what the majority has been on the topic. Often the developers own opinions already matched those of
+on and potentially agree with what the majority has been on the topic. Often the developers' own opinions already matched those of
 the majority before them and did not necessarily have to directly verify it themselves.
 
-As added insurance, we conducted 5 post results interviews with developers to once again confirm our results, and discussions. These
-procedures can be seen in Section~\ref{sec:meth}.
+As added insurance, we conducted 5 confirmatory interviews after we analyzed our data and drafted themes in our findings. We describe this
+procedure in Section~\ref{sec:meth}.
 
@@
 {\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
-Although our results may not be entirely novel or even surprising, the combination of said results allowed us to discover the
-the disjoint between theoretical awareness and practical awareness regarding indirect conflict tools as well as provide more
-insight into the debate of prevention versus cure in software development. Given how many indirect conflict tools are left with the same common
+Although our results may not be entirely novel or even surprising, our findings do (1) suggest
+a disjoint between the current research understanding of developers' awareness needs and the these awareness needs related to indirect conflict tools in practice, and (2) provide some
+insight into the debate of prevention versus cure in software development. Given that many indirect conflict tools still suffer from these same common
 issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
@@
 
-10 of the 78 questionnaire participants sent direct responses to us asking for any results of the research to be sent directly to
-them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
-tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest
-in our results help us fulfill this criterion.
+%10 of the 78 questionnaire participants sent direct responses to us asking for any results of the research to be sent directly to
+%them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
+%tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest
+%in our results help us fulfill this criterion.
 
@@
 
-Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
-losses in this realm have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
-19 interviewed developers from 12 organizations as well as 78 developers completing questionnaires, we have provided characterization of indirect conflicts,
-current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
-in their work flow with indirect conflicts.
+Indirect conflicts are a significant issue with real world development, however, many of our research proposed techniques and tools have been unsuccessful in attracting major support from developers.
+Based on our qualitative study involving
+19 interviewed developers from 12 organizations as well as 78 questionnaires respondents, we sought to characterize indirect conflicts,
+current developer work flows surrounding indirect conflicts, and frame directions for future tool development that support developers
+in their tackling of indirect conflicts.
 
-Our findings have indicated that:
-indirect conflicts occur frequently and are likely caused by software contract changes;
+Our findings  indicate that indirect conflicts occur frequently and are likely caused by software contract changes;
 while design by contract, add and deprecate, and personal experience help prevent indirect conflicts,
@@
 
-Our result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
+Our study of practitioners intends to fill a gap in the software engineering research with respect to the outstanding problems in current tool-based solutions to indirect conflicts, namely the issues of information overload, false positives and scalability.
+Overall, we described a disjoint between the current research understanding of developers' awareness needs and these awareness needs related to indirect conflict tools in practice.
+More specifically, our findings suggest that developers do not want awareness mechanisms which provide non actionable results; developers
 would rather focus on curing existing problems rather than preventing potential issues;
 and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.
+We hope that these insights will be found valuable in future research endeavours and development of tools to assist developers in tackling indirect conflicts.
 

### 2014-03-16 iced-2013/paper.tex
ICED done.

@@
 instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
-The evaluations of these tools indicate that they have the same common difficulties of scalability, false positives, and information overload. Our own tool
-Impact!~\cite{Ell:2013} also suffered these same fates.
+The evaluations of these tools (including our own prototype Impact! ~\cite{Ell:2013}) indicate that they have the same common difficulties of scalability, false positives, and information overload.
 
-Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
+Following Cataldo et al.~\cite{Cataldo:2006:ICR}'s work that indicated that socio-technical congruence can be leveraged to improve task completion
 times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
@@
 Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and obtained 78 responses
-giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
-were conducted during the months of July and August while many participants may be away from their regular positions.
+giving a response rate of 6\%. We attribute the low response rate with the questionnaires
+being sent out during the months of July and August when many participants may be away from their regular positions.
 
-The questionnaire we designed ~\footnote{http://thesegalgroup.org/projects/indirect-conflicts/}
-was based on the insights we obtained from the round of interviews, and was intended to confirm these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews.
+We designed the questionnaire~\footnote{http://thesegalgroup.org/projects/indirect-conflicts/} to follow the interviews.
+We intended to confirm some of the insights we obtained from the interviews with a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews.
 The question topics asked in the questionnaire were:
@@
 The direct implication this research has on tool development is that of current adoption among developers. As was stated,
-developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being
-said, focus at these two stages for tool developer will lead to larger adoption among developers.
-It should be noted here that detection must come with an almost zero rate of false positives as
-current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
+developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts.
+We believe that focus at these two stages for tool developer will lead to larger adoption among developers.
+Similarly, detection must come with an almost zero rate of false positives as current tools (such as unit and integration tests) only report real problems with almost 100\% precision (even though their recall may not be 100\%).
+
+%It should be noted here that detection must come with an almost zero rate of false positives as
+%current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.
 

### 2014-03-17 msc-thesis/chapters/1/chapter-intro.tex
MSC final melanie edits.

@@
 location in the software system such as when libraries are upgraded or when a method is changed
-and invoking methods are influence negatively. Indirect conflict tools however, have
+and invoking methods are influenced negatively. Indirect conflict tools however, have
 not shared the same success as direct conflict 
@@
 will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
-with supporting indirect conflicts through awareness techniques as well as to explore possible solution for industry practice 
-in the are of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
+with supporting indirect conflicts through awareness techniques as well as to explore possible solutions for industry practice 
+in the area of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
 conflicts: technical dependencies, how developers are involved in said dependencies,
@@
 how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
-conducting the following four studies.
+conducting four studies.
 
@@
 in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
-relationships in successful and failed code changes. These developers pairs can be seen as indirect conflicts occurring
+relationships in successful and failed code changes. These developer pairs can be seen as indirect conflicts occurring
 as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
@@
 answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
-developers who's changing technical dependencies statistically relate to bugs?}''.
+developers whose changing technical dependencies statistically relate to bugs?}''.
 In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
@@
 notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
-projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationship 
+projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationships 
 through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
@@
 over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
-and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
+and that there exists a gap in software evolution analytical tools arising from the reliance on static analysis resulting in missed
 context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
@@
 The remainder of this dissertation is laid out as follows. Chapter two includes Study 1 and Study 2 as the motivational studies which
-ultimately led to the larger research en devours found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
+ultimately led to the larger research studies found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
 press upon indirect conflicts in a more in depth fashion than has been previously seen in research. Chapter 4 contains a lengthy discussion

### 2014-03-17 msc-thesis/chapters/2/chapter-motivating.tex
MSC final melanie edits.

@@
 pairs in indirect conflicts learned from Study 1, and adds in source code change detection in order to create
-an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to a developer
-to any source code changes preformed by another developers when the two are linked in a technical dependency through
-a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications for visualization. 
+an awareness notification system for developers called \textit{Impact}. \textit{Impact} was designed to alert a developer
+to any source code changes preformed by another developer when the two are linked in a technical dependency through
+a developer pair. \textit{Impact} utilized a non-obtrusive RSS style feed for notifications. 
 While \textit{Impact} showed some promise through its user evaluation, it ultimately suffered the fate of information

### 2014-03-17 msc-thesis/chapters/2/sec-developer-pairs.tex
MSC final melanie edits.

@@
 ~\cite{Kim:2006:AIB}. 
-This module-based method also results in predictions 
+These module-based methods also result in predictions 
 at the file or binary level of software development as opposed to a code change level
@@
 using communication patterns from developers. Wolf et al.~\cite{5070503} used patterns of communication from between
-pre-existing build in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
-communication organization difference between developers who worked on the same software module had
+pre-existing builds in order to predict later build outcomes. Naggappan et al.~\cite{nagappan08} showed that having a large
+communication organizational difference between developers who worked on the same software module had
 a negative influence on the quality of the software.
@@
 The code change effect, if any, to the call hierarchy is now found. The Git
-software repository is used to determine what changes were made to any give file inside a 
+software repository is used to determine what changes were made to any given file inside a 
 code change. Specifically, methods modified by a code change are searched for. The call graph 

### 2014-03-17 msc-thesis/chapters/2/sec-impact.tex
MSC final melanie edits.

@@
 In this study, I report on research into supporting developer pairs in indirect conflicts
-and present the design, implementation, and future evaluation of the tool \textit{Impact},
+and present the design, and implementation of the tool \textit{Impact},
 a web based tool that aims at detecting indirect conflicts among developers
@@
         \item[RQ\namedlabel{itm:rq1-impact}{RQ1-Impact}] \textit{Can indirect conflicts be supported 
-        through an awareness mechanism which leverages pairs of developers who's changing technical 
+        through an awareness mechanism which leverages pairs of developers whose changing technical 
         dependencies statistically relate to bugs?}
@@
 network that relates developers to source code, issue repositories and
-other social media while Ariadne only examines there source code for developer
+other social media while Ariadne only examines their source code for developer
 to source code association. Through Codebook, developers become
@@
 indirect conflicts is the tool Palantir~\cite{Sarma:2007:TSA}. Palantir
-monitors developers activities in files with regards to class signatures.
+monitors developer's activities in files with regards to class signatures.
 Once a developer changes the signature of a class, such as by modifying changes
@@
 a sphere of influence for each developer to determine how source code
-changes are indirectly related to other components of a software project.
+changes are indirectly related to other components of the software.
 CASI uses dependency slicing~\cite{Bajracharya:2009:SIS} instead of the 
@@
 a developer can see what parts of a software projects he or she may be effecting
-with the source code change. This allows the developer themselves to go and fix
+with the source code change. This allows the developers themselves to go and fix
 potential issues elsewhere in the project before the code change is committed
@@
 trade off is that indirect conflicts must be committed before detected,
-which results in conflicts being apart of the system before being able
+which results in conflicts being integrated into the system before being able
 to be dealt with as opposed to catching conflicts before they happen.
@@
 in medium to large teams or projects as opposed to the small
-teams which they were apart of. It was also pointed
+teams which they were a part of. It was also pointed
 out that method use can be a particularly large area for indirect
@@
 not all are notification worthy. One developer suggested to only notify
-developers to indirect conflicts if the internal structure of a method
+developers if the internal structure of a method
 changes due to modification to input parameters or output parameters.
@@
 Because of the tool validation nature of this work, I chose participant interviews as my research validation method, which has some implications regarding the
-limitations and threats to validity of this study. While I did some positive results regarding the potential of Impact in this study
-populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
+limitations and threats to validity of this study. While I did have some positive results regarding the potential of Impact in this study,
+populations studied from outside of this study's participants may add new insights into the pool of findings. As a result of this, findings
 from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.
 
-As per my interview participants, I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
+I conducted this case study with undergraduate students at the University of Victoria. This being said, participants
 may not have had enough real world experience to validate this study at an industry level. The students were also consumed with their regular course
-work which could have limited the time spend using Impact and the enthusiasm put forward while conducting this study.
+work which could have limited the time spent using Impact and the enthusiasm put forward while conducting this study.
 
-To counter act this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
+To counter this, my study was conducted purely on a volunteer basis to eliminate those participants which may have been to busy to focus
 on the study or provide appropriate feedback where needed.

### 2014-03-17 msc-thesis/chapters/3/chapter-exploration.tex
MSC final melanie edits.

@@
 
-Through the two previous studies, I have shown that developers linked indirectly in source code changes can statistically
+Through the two previous studies, I have shown that developers linked indirectly in source code changes can be statistically
 related to software failures. In the attempts of mitigating these loses through added awareness, I implemented
@@
 frequent after a release or at the start of a new development cycle. The findings of this study can be applied over the lifetime of a project
-to determine the proneness of indirect conflicts and thus aiding developers in dealing with indirect conflicts in their projects.
+to determine the probability of indirect conflicts occurring and thus aiding developers in dealing with indirect conflicts in their projects.
 

### 2014-03-17 msc-thesis/chapters/3/sec-apie.tex
MSC final melanie edits.

@@
 While measuring software quality has had a major focus in software engineering research for many years~\cite{Bowen:1978:CAS}~\cite{Grady:1993:PRM}~\cite{ISOIEC9126},
-the sub study of software stability and its implications on reliability and indirect conflict proneness remains a difficult subject to understand. 
+the study of software stability and its implications on reliability and indirect conflict proneness remains a difficult subject to understand. 
 The decision of what makes a project stable and ready
@@
 
-Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the project's I had selected. Since releases
+Once the ChangeDistiller information was collected, I decided to examine software change trends surrounding releases of the projects I had selected. Since releases
 have preconceived notions of software stability and a lack of proneness to indirect conflicts, 
@@
 and private methods
-being removed from a project is more likely to occur after a major release than after. Table~\ref{tab:apie-ratio} also shows significance in the changes to private
+being removed from a project is more likely to occur after a major release than before. Table~\ref{tab:apie-ratio} also shows significance in the changes to private
 classes. We see that private classes are added more (24\%) before a major release and removed more after (44\%) the release. 
@@
 since after a major release a new development cycle is likely to begin. As per change trends which occur more so before a major release, indirect
-conflicts may use this context in order to identify a low proneness to indirect conflicts in their results. These two contextual patterns
+conflict tools may use this context in order to identify a low proneness to indirect conflicts in their results. These two contextual patterns
 can be used throughout the life of a software project in order to help better inform indirect conflict tools as to the processes of developers

### 2014-03-17 msc-thesis/chapters/3/sec-iced.tex
MSC final melanie edits.

@@
 with an average of 8 years of experience.
-In addition to software development, some participants were also invited based on their experience with project management
-at some capacity. See Table~\ref{tab:demo} for more demographic details.
+In addition to software development, some participants were also invited based on their experience with project management. 
+See Table~\ref{tab:demo} for more demographic details.
 
@@
 either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
-being apart of any one software project for extended periods of time.
+being a part of any one software project for extended periods of time.
 
@@
 
-\subsubsection{questionnaire Procedure}
+\subsubsection{Questionnaire Procedure}
 
-questionnaire participants were invited to participate in the questionnaire by email. No reminder email was sent as the
+Questionnaire participants were invited to participate in the questionnaire by email. No reminder email was sent as the
 questionnaire responses were not connected with the invitation email addresses and thus participants who did respond
@@
 were conducted during the months of July and August while many participants may be away from their regular positions.
-and my GitHub and Apache participants could not be verified as to whether or not they actively support the
-email addresses used in the invitations. In addition, the questionnaire was considered by some to be long and require
+Also, my GitHub and Apache participants could not be verified as to whether or not they actively support the
+email addresses used in the invitations. In addition, the questionnaire was considered by some to be long and required
 more development experience than may have been typical of some of those invited to participate.
@@
 
-As per my interview participants, I attempted to cover as great a breadth as possible of software developers in their practices, companies,
+I attempted to cover as great a breadth as possible of software developers in their practices, companies,
 project, languages, etc. However, it would be extremely difficult to conduct a completely thorough study of indirect conflicts in practice
 as there exist too many demographic options for participants. This same applies for questionnaire participants. This being said, questionnaire 
-or interview results could become heavily biased towards a developers good or bad experiences with any given project.
+or interview results could become heavily biased towards a developer's good or bad experiences with any given project.
 
-To counter act this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
+To counter this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
 I have uncovered valuable insight regarding indirect conflicts in practice.
@@
 would rather focus on curing existing problems rather than preventing potential issues;
-and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.
+and there exists a gap in software evolution analytical tools between what is available and what practitioners need.

### 2014-03-17 msc-thesis/chapters/4/chapter-discussion.tex
MSC final melanie edits.

@@
 
-As we have seen in Study 1, developers which are tied to source code objects can become a focal point of
+As we have seen in Study 1, developers that are tied to source code objects can become a focal point of
 indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
@@
 a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
-to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
+to take some action because of it. This limited awareness is quite different from what literature suggests, which is larger
 awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
-false positives, as the changes being reported are not causing them to take action. These reported non actionable responses
+false positives, as the changes being reported are not causing them to take action. These reported non-actionable responses
 lead to information overload and false positives.}
@@
 are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
-abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
+abundance of detection and curative process and tools on the other hand shows once again developer's willingness to debug
 real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
@@
 it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
-not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
+not be moving developers into a more preventative mindset; if curative can be shown to be more effective in that it takes less time
 to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
@@
 designers, architects, or managerial stakeholders. 
-From the project manager's interviewed it was shown that they are heavily favored towards planning and 
+From the project managers interviewed it was shown that they heavily favored planning and 
 prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
@@
 better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
-end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
+end of an iteration, a new feature being released or any point considered stable and reliable.) However, this context of a 
 mature point, as well as any other
@@
 developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
-said, focus at these two stages for tool developer will lead to larger adoption among developers.
+said, focus at these two stages for tool development will lead to larger adoption among developers.
 It should be noted here that detection must come with an almost zero rate of false positives as

### 2014-03-17 msc-thesis/chapters/5/chapter-conclusion.tex
MSC final melanie edits.

@@
 indirect conflicts have had common issues of information overload, false positives, scalability, information distribution 
-and many others. These issues has ultimately lead to poor developer interest in indirect conflict tools as well
+and many others. These issues have ultimately led to poor developer interest in indirect conflict tools as well
 as a failed adoption rate and partial failures of research.
@@
 conflicts through 4 studies. The first two studies focused on motivational circumstances which occur during 
-the software development life cycle and cause indirect conflicts. Developers interactions were studied in order to create 
+the software development life cycle and cause indirect conflicts. Developer's interactions were studied in order to create 
 a tool which can aid in the work-flows around indirect conflicts. The second two studies presented a deeper investigation 
@@
 bugs in source code throughout the lifetime of a project. These developer pairs were used in the prediction
-of future bugs as well as provide coordination recommendations for developers within a project. These bugs being predicted
+of future bugs and to provide coordination recommendations for developers within a project. These bugs being predicted
 directly correlate to indirect conflicts. Since the human dependencies being measured in Study 1 are directly tied to
@@
 A human factor is present here as the person who makes the artifact change, say Anne, may not have communicated the
-change to all other developers affected by the change. This is an important understanding that lead directly into the
+change to all other developers affected by the change. This is an important understanding that led directly into the
 creation of Impact (Study 2) as a way to mitigate the negative affects across human developers as a result of 
@@
 In Study 3, I conducted a qualitative study involving
-19 interviewed developers from 12 organizations as well as 78 surveyed developers. Through these qualitative measures,
-I provided characterization of indirect conflicts,
+19 interviewed developers from 12 organizations as well as 78 surveyed developers. I provided characterization of indirect conflicts,
 current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
@@
 to avoid indirect conflicts: ``Design by Contract'', add and deprecate, and personal experience. For catching indirect conflicts,
-I found that use case coverage through proper testings (both unit and integration) are currently thought to be the best
+I found that use case coverage through proper testing (both unit and integration) are currently thought to be the best
 developers can achieve. And finally, for the future of indirect conflict tools, developers made it clear that they would
@@
 the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
-to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
+to fully assess the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
 the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
@@
 awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
-ultimately lead to tools with information overload and false positive issues. This was quite evident from my tool Impact as
-presented in Study 2. Even when the tool is based around the statistical results of Stud 1, the disjoint between what academics
-determin developers want versus what they themselves want is quite obvious. The debate of prevention versus cure has been presented along with
+ultimately led to tools with information overload and false positive issues. This was quite evident from my tool Impact as
+presented in Study 2. Even when the tool is based around the statistical results of Study 1, the disjoint between what academics
+determine developers want versus what they themselves want is quite obvious. The debate of prevention versus cure has been presented along with
 the industrial tendency towards curative measures. However, while a curative approach may be favored by developers, further research is needed
-to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
+to fully assess the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
 the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
 future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While, I have shown a potential solution to address
-a sub problem of the analysis gap through contextual patterns of indirect conflicts, many sub problems of the analysis gap remian problems
+a fraction of the analysis gap through contextual patterns of indirect conflicts, many other problems of the analysis gap remain problems
 for future research.

### 2014-03-17 msc-thesis/frontmatter/abs.tex
MSC final melanie edits.

@@
 two studies present a deeper investigation into why most indirect conflict tools fail to attrack developer
-inerest through exploring the root causes of indirect conflicts and how tools should be properly built
+interest through exploring the root causes of indirect conflicts and how tools should be properly built
 to support developer workflows.

### 2014-03-19 msc-thesis/chapters/3/sec-iced.tex
MSC final edits

@@
 populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
-from this study may not relate to everyone or generalize to outside of the group of participants involved in this study.
+from this study may only relate to developers or managers from the projects I studied.
 
@@
 
-My findings have indicated that:
+In addresses the three issues of scalability, false positives, and information overload, my findings indicate:
 indirect conflicts occur frequently and are likely caused by software contract changes;

### 2014-03-19 msc-thesis/chapters/5/chapter-conclusion.tex
MSC final edits

@@
 the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
-future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While, I have shown a potential solution to address
+future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While I have shown a potential solution to address
 a fraction of the analysis gap through contextual patterns of indirect conflicts, many other problems of the analysis gap remain problems

### 2014-04-24 msc-thesis/chapters/3/sec-apie.tex
MSC final version.

@@
 stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
+This study was conducted in order to address the results of Study 3 which identified that indirect conflicts are likely to occur more
+before a major release (i.e. stable point in the code). I look to find the context in which trends occur in order to support
+indirect conflict tools in their context for when indirect conflicts are more likely to occur.
 With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can

### 2014-04-24 msc-thesis/chapters/3/sec-iced.tex
MSC final version.

@@
 Because of the exploratory nature of this work, I chose Grounded Theory as my research method, which has some implications regarding the
-limitations and threats to validity of this study. While I achieved high saturation regarding the topics focused on in this study, other
+limitations and threats to validity of this study. Inter rater reliability was not used due to a time constraint which means
+that the code from open and axial coding have not been confirmed by a separate source.
+While I achieved high saturation regarding the topics focused on in this study, other
 populations studied from outside of this studies participants may add new insights into the pool of findings. As a result of this, findings
-from this study may only relate to developers or managers from the projects I studied.
+from this study may only relate to developers or managers from the projects I studied. The saturation indicated here regards the majority
+results reached in this study. Since only 19 developers were interviewed, saturation was reached when majority agreed upon
+trends emerged in the results, thus further interviews did not have to be conducted. This is different from saturation of edge or
+unique cases often found in other qualitative research where the goal is to find all cases of a phenomenon where this study
+was only looking for majority cases.
 
